\documentclass[11pt,notitlepage]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,booktabs}
\usepackage{enumitem}
\usepackage{threeparttable}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,
  pdftitle={Compact-Object Crust MHD Equivalence in HQIV},
  pdfauthor={Steven Ettinger Jr}}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\footnotesize\url{#1}}}
\newcommand{\leanid}[1]{{\footnotesize\path{#1}}}
\newcommand{\scriptfile}[1]{{\footnotesize\url{scripts/#1}}}
\newcommand{\datafile}[1]{{\footnotesize\url{data/#1}}}

\title[Crust MHD equivalence]{Compact-Object Crust MHD Equivalence in HQIV:\\
Resistive Reduction, Geometry-First Multipoles, and Honest Scope}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Preprint --- \today}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
\noindent\textbf{Position.}
Companion to the compact-object witness data bundle
\cite{hqiv-compact-object-witness-data} and the O-Maxwell fluid chart
\cite{hqiv-omaxwell-fluid-chart}.
This note does \textbf{not} propose a rival magnetic theory. It records that
standard \textbf{resistive and Hall MHD} crust dynamics reduce onto existing HQIV
slots in \leanfile{Hqiv/Physics/CompactObjectMhdEquivalenceScaffold.lean} with
explicit coefficient identification, then states where \textbf{geometry-first}
organization (ε-tipping layers, spin-axis latitude belts, $\tau_{\mathrm{mis}}$
obliquity) yields predictions that traditional Hall-MHD fits often bolt on afterward.

\noindent\textbf{Reduction (proved naming).}
Traditional ohmic $E=J/\sigma$ is \leanid{ohmicAxialField}; the classical limit
\leanid{coronalEffectiveAxialField\_classical\_limit} is machine-checked.
Resistive induction uses \leanid{inductionResistivityEta}$(\xi,\varepsilon)=
\gamma\,\mathrm{release}(\xi)\,G_{\mathrm{eff}}(\varepsilon)$ and scalar discharges
\leanid{steadyInductionFieldLt}, \leanid{inductionGrowthRateFromLt}.
Turbulent viscosity split \leanid{nuTotal\_eq\_nuMol\_add\_hqivEddy} maps the
traditional $\tau_{\mathrm{mol}}+\tau_{\mathrm{eddy}}$ bookkeeping.

\noindent\textbf{Geometry-first predictions (witnessed in Python).}
Layered ε zones, colatitude channels ($\rho_{\mathrm{pol}}$, Coriolis gate),
$\tau_{\mathrm{mis}}$ m=1 spots, and NICER J0030/J0740 overlays are generated from
\scriptfile{hqiv\_compact\_object\_mass.py} (\texttt{--surface-multipole-audit},
\texttt{--paper-dynamics-outline}).

\noindent\textbf{Honest limitations.}
Proto-neutron-star $\alpha\Omega$ seed fields for magnetar $10^{14}$--$10^{15}\,\mathrm{G}$
are \textbf{boundary data}, not outputs of the cold-crust witness.
Full vector Hall-MHD integration over $10^3$--$10^6\,\mathrm{yr}$ and numeric
$\sigma(T,B)$ calibration remain milestones (\S\ref{sec:limitations}).
\end{abstract}

\maketitle

\input{../include/patch_theory_messaging.tex}

\section{Introduction and framing}
\label{sec:intro}

Neutron-star magnetospheres and crusts are modeled in traditional astrophysics as
\textbf{composite MHD}: inhomogeneous Maxwell equations, conducting fluid momentum,
and conductivity $\sigma(T,B)$ closures, often extended with Hall drift and crust
elasticity \cite{hqiv-compact-object-witness-data}.
HQIV records the same stack at the action level: \textbf{O-Maxwell} $S_O(J,A,\phi)$
plus \textbf{modified fluid} closure (\leanfile{Hqiv/Physics/HQIVFluidClosureScaffold.lean}).

The claim of this paper is therefore \emph{organizational}, not phenomenological:
\begin{quote}
\emph{We have a first-principles way to organize the same physics around
ε-tipping layers and the spin axis, yielding testable predictions for mass/spin
dependence and surface multipoles.}
\end{quote}

\section{Equation-level equivalence}
\label{sec:equivalence}

Table~\ref{tab:mhd-equivalence} displays standard resistive-MHD equations beside
HQIV Lean names. Coefficient identification is explicit; vector $\nabla\times$ PDEs
are \textbf{scalar discharges} on the crust witness chart until the vector milestone
(\S\ref{sec:limitations}).

\begin{table}[ht]
\centering
\caption{Traditional MHD $\leftrightarrow$ HQIV slot reduction (see
\leanfile{Hqiv/Physics/CompactObjectMhdEquivalenceScaffold.lean}).}
\label{tab:mhd-equivalence}
\small
\begin{tabular}{@{}p{0.42\linewidth}p{0.48\linewidth}@{}}
\toprule
Traditional & HQIV slot / coefficient ID \\
\midrule
Inhomogeneous Maxwell $\nabla\times F = J$ &
\leanid{EL\_O\_general} (\leanfile{Hqiv/Physics/Action.lean}) \\
$E = J/\sigma$ (ohmic) &
\leanid{ohmicAxialField}; classical limit proved \\
$\partial_t B \sim \nabla\times(v\times B)-\nabla\times(\eta\nabla\times B)$ &
\leanid{inductionGrowthRateFromLt}, \leanid{steadyInductionFieldLt} \\
MHD resistivity $\eta$ &
$\eta_{\mathrm{HQIV}}(\xi,\varepsilon)=\gamma\,\mathrm{release}(\xi)\,G_{\mathrm{eff}}(\varepsilon)$ \\
$\rho(\partial_t v + v\cdot\nabla v) = \cdots + (J\times B)/c + \nabla\cdot\tau$ &
\leanid{hqivFluidInertiaFactor}; Maxwell stress from $S_O$ \\
$\tau_{\mathrm{mol}}+\tau_{\mathrm{eddy}}$ &
\leanid{nuTotal\_eq\_nuMol\_add\_hqivEddy} \\
PNS turbulent $\alpha$ (dynamo driving) &
\leanid{hqivEddyViscosity} + \leanid{hqivVacuumMomentumSource3} (F2 chart) \\
Hall $\partial_t B \propto \sigma/(ne)\,\nabla\times((\nabla\times B)\times B)$ &
\leanid{rapidityNormalized\_frozenFirstIndexJet} (milestone) \\
\bottomrule
\end{tabular}
\end{table}

\begin{proposition}[Discharged resistive identification]
\label{prop:eta-id}
For any environment slots $(\xi,\varepsilon)$, traditional resistivity $\eta$ may be
identified with \leanid{inductionResistivityEta}$(\xi,\varepsilon)$, which is
nonnegative (\leanid{inductionResistivityEta\_nonneg}) and equals $\gamma$ at lock-in
with zero gravity (\leanid{inductionResistivityEta\_lockin\_zero\_gravity}).
The bundle \leanid{mkCompactObjectMhdEquivalenceDischarged} witnesses the reduction.
\end{proposition}

\section{Geometry-first structure}
\label{sec:geometry}

Traditional codes can impose composition layers and offset hotspots by hand.
HQIV discharges the following from ε-lattice and spin-chart slots:

\begin{enumerate}[nosep]
\item \textbf{ε-tipping layers.} Nuclear, charmed, and top zones with radii
$r_{\mathrm{top}}$, $r_{\mathrm{charm}}$ tied to $\varepsilon$; charm retreat on spindown
(\scriptfile{hqiv\_compact\_object\_mass.py}, \texttt{--spindown-charm-audit}).
\item \textbf{Spin-axis latitude belts.} Pole channel ($\rho_{\mathrm{pol}}$, axial
$\nabla\phi$), mid-latitude Coriolis gate $\sin^2\theta|\cos\theta|$, equatorial induction.
\item \textbf{Unified obliquity.} Misaligning torque
\leanid{crustMisalignTorqueFromStressDivergence} vs aligning magnetic braking;
enhanced closure uses lattice-weighted \textbf{B\_align}, charm-ledger boost, and
minimal crust $J\times B$ (\scriptfile{hqiv\_compact\_object\_mass.py}).
\item \textbf{Surface multipoles.} Zonal $l=2,l=3$ and $m=1$ from $\tau_{\mathrm{mis}}$
tilted-dipole modulation (m=1 tilt from enhanced $\alpha_{\mathrm{eq}}$); NICER overlays
for J0030+0451 and J0740+6620.
\item \textbf{Environment-aware $\eta$.} Outside temperature with co-spin and CMB Doppler
boost feeds $\xi$ and hence $\eta(\xi,\varepsilon)$.
\end{enumerate}

The high-spin / high-mass tail shows a documented transition: moderate
$\Omega/\Omega_{\mathrm{break}}$ favors Coriolis$\times$shear ($l_2/l_0\sim 1$);
breakup favors equatorial induction ($l_3/l_0\sim 1$)---see
\datafile{compact\_object\_witnesses.json} and \texttt{--surface-multipole-audit}.

\section{Differentiating predictions}
\label{sec:predictions}

\paragraph{Field obliquity vs surface geometry.}
\textbf{$\alpha_{\mathrm{eq}}$} from slip-torque balance is the misalignment between
spin and the dipole axis used in aligning torque---not colatitude of a hot spot.
Catalog dipole $B$ for millisecond pulsars typically yields $\alpha_{\mathrm{eq}}\approx 90^\circ$
(misaligning-dominated); the canonical witness at $10^{12}\,\mathrm{G}$ gives
$\alpha_{\mathrm{eq}}\approx 71^\circ$ for a $1.4\,M_\odot$ ms star; breakup favors
low obliquity ($\sim 11^\circ$). NICER centroid offsets and $m=1$ structure are a
\textbf{separate} surface-multipole layer (\texttt{tau\_mis\_m1\_gate}).

\begin{itemize}[nosep]
\item \textbf{Mass/spin $\rightarrow$ multipoles.} $l_2/l_0$, $l_3/l_0$, centroid offset,
and $m_1$ fraction correlate with $\Omega/\Omega_{\mathrm{break}}$ without per-source
hotspot offsets.
\item \textbf{Obliquity regimes.} Enhanced aligning torque:
\datafile{pulsar\_witness\_comparison.json} (\texttt{alpha\_equilibrium\_aligning\_enhanced\_deg},
\texttt{alpha\_equilibrium\_canonical\_b\_deg}).
\item \textbf{High-mass tail.} Charm retreat, pair margin, incremental $B_{\mathrm{charm}}$
in young massive rotators.
\item \textbf{$\eta(\xi,\varepsilon)$ calibration.} \texttt{coefficient\_calibration\_witness}
vs literature $\sigma$ and $\tau_{\mathrm{ohm}}\sim\mu_0\sigma R^2$
(\texttt{--eta-calibration-audit}).
\end{itemize}

\section{Coefficient discharge and limitations}
\label{sec:limitations}

\paragraph{Discharging coefficients (not a weakness).}
Traditional decades of work fix $\sigma(T,B)$, Hall coefficients, and $\tau_{\mathrm{ohm}}$.
HQIV fixes the \textbf{algebra} ($\alpha$, $\gamma$, $\varepsilon$, $\xi$) and proposes
\begin{equation}
  \eta_{\mathrm{HQIV}}(\xi,\varepsilon)=\gamma\,\mathrm{release}(\xi)\,G_{\mathrm{eff}}(\varepsilon),
  \qquad
  E_{\mathrm{Ohm}} = J/\sigma \;\Leftrightarrow\; \text{\leanid{ohmicAxialField}}.
\end{equation}
A pre-finalization table compares $\eta_{\mathrm{HQIV}}$ and schematic $\partial B/\partial t$
against literature crust values; the numeric witness is
\texttt{coefficient\_calibration\_witness} (\texttt{--eta-calibration-audit}), mapping
$\tau_{\mathrm{ohm}}\approx\mu_0\sigma R^2$ at stellar radius to the $10^4$--$10^6\,\mathrm{yr}$
band for $\sigma\in[10^7,10^9]\,\mathrm{S/m}$.

\paragraph{Not claimed.}
\begin{itemize}[nosep]
\item Magnetar $10^{14}$--$10^{15}\,\mathrm{G}$ \textbf{generation} from cold-crust induction alone;
$B_0$ is boundary data (proto-NS dynamo or catalog).
\item Replacement of mature Hall-MHD crust evolution codes.
\item Full time-dependent vector multipole evolution without coupling $\partial_t c_\ell$ to
induction and diffusion.
\item Numeric Hall-term benchmark until \leanid{rapidityNormalized\_frozenFirstIndexJet}
is discharged as an explicit vector identity.
\end{itemize}

\paragraph{Machine-checked honesty ledger.}
\leanid{CompactObjectMhdProvedReductions}, \leanid{CompactObjectMhdMilestones}, and
\leanid{CompactObjectMhdNotClaimed} in
\leanfile{Hqiv/Physics/CompactObjectMhdEquivalenceScaffold.lean} separate proved reductions,
milestones, and explicit non-claims.

\section{Data and reproducibility}

\begin{itemize}[nosep]
\item Witness JSON: \datafile{compact\_object\_witnesses.json}
  (\scriptfile{hqiv\_compact\_object\_mass.py --json}).
\item Paper bundle: \texttt{--paper-dynamics-outline}, \texttt{--mhd-equivalence-audit},
  \texttt{--eta-calibration-audit}.
\item Pulsar overlay: \datafile{pulsar\_witness\_comparison.json}
  (\scriptfile{hqiv\_pulsar\_witness\_benchmark.py --json}).
\item Regenerate all: \texttt{papers/compact\_object\_witness/regenerate\_data.sh}.
\item Lagrangian audit: \scriptfile{hqiv\_lagrangian\_faithfulness\_audit.py --json}.
\end{itemize}

\bibliographystyle{amsplain}
\bibliography{../references}

\end{document}
