\documentclass[11pt,notitlepage]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,booktabs}
\usepackage{enumitem}
\usepackage{threeparttable}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,
  pdftitle={Binding Energy and the Weak Force from HQIV Composite-Trace Weights},
  pdfauthor={Steven Ettinger Jr},
  pdfsubject={HQIV nucleon binding; three-ledger beta decay; composite-trace weights; Lean formalization},
  pdfkeywords={HQIV, nucleon binding, beta decay, half-life, composite trace, weak force, Lean}}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{xurl}
\urlstyle{same}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,angles,quotes}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\footnotesize\url{#1}}}
\newcommand{\leanid}[1]{{\footnotesize\path{#1}}}
\newcommand{\scriptfile}[1]{{\footnotesize\url{scripts/#1}}}
\newcommand{\datafile}[1]{{\footnotesize\url{data/#1}}}
\newcommand{\leanterm}[3]{\hyperref[#2]{#1}}
\newcommand{\scriptref}[1]{\hyperref[app:scripts]{#1}}
\newcommand{\fitbox}[1]{\begin{adjustbox}{max width=\linewidth,center}#1\end{adjustbox}}

\providecommand{\leanCertAnomaly}{}
\providecommand{\leanCertTopology}{}
\providecommand{\leanCertStrongColor}{}
\providecommand{\leanPatchHolonomyWitness}{}
\providecommand{\leanPatchWilsonKinetic}{}
\providecommand{\leanPatchHaarMeasure}{}
\providecommand{\leanPatchDiscretePoincare}{}
\providecommand{\leanPatchSpatialRotation}{}

\title[Binding energy and the weak force]{Binding Energy and the Weak Force from HQIV Composite-Trace Weights:\\
Lock-In Masses, Three Ledgers, and Zero-Knob Half-Lives}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Zenodo v1 --- 2026-06-16}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

% Keep title + abstract on page 1: amsart stores the abstract in an unbreakable
% \vtop; unvbox lets it start below the title instead of jumping to page 2.
\makeatletter
\renewcommand{\@setabstracta}{%
  \ifvoid\abstractbox
  \else
    \skip@12\p@ \advance\skip@-\lastskip
    \advance\skip@-\baselineskip \vskip\skip@
    {\footnotesize\unvbox\abstractbox}%
    \prevdepth\z@
  \fi
}
\patchcmd{\@maketitle}{\global\topskip42\p@}{\global\topskip30\p@}{}{}
\makeatother

\begin{document}

\begin{abstract}
\noindent\textbf{Position in the publication chain.}
Published Tier-2 \#7 (Zenodo
DOI \href{https://doi.org/10.5281/zenodo.20711453}{10.5281/zenodo.20711453}).
Companion to the published TUFT+SM synthesis \cite{hqiv-tuft-sm-lagrangian-paper}
(DOI \href{https://doi.org/10.5281/zenodo.20601215}{10.5281/zenodo.20601215}).
This is the seventh paper in the HQIV preprint series and the third Tier-2
Standard-Model layer entry \emph{after} the published TUFT+SM synthesis
\cite{hqiv-tuft-sm-lagrangian-paper}. Tier~0--1 fix the discrete spine
\cite{ettinger2026hqiv,hqiv-lightcone-paper,hqiv-so8-paper} and Tier~1 variational
records \cite{hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper,hqiv-thermo-arrow-paper}.
\textbf{This note establishes} the local-curvature slot equations---named
$B_{\mathrm{curv}}(\xi)$, three $\beta$ ledgers, slot-wise mass/binding closure, and
outside-contact primitives---witnessed from nucleon binding through bulk condensed matter.
\textbf{Companion} Tier~2 notes on baryogenesis lock-in
\cite{hqiv-baryogenesis-paper} (published:
DOI \href{https://doi.org/10.5281/zenodo.20711255}{10.5281/zenodo.20711255}) and BBN abundances
\cite{hqiv-bbn-paper} (published:
DOI \href{https://doi.org/10.5281/zenodo.20723606}{10.5281/zenodo.20723606}) apply the \emph{same} equations
($\xi(T)$, outside release, opacity slots, dynamic-$C_2$ lapse clock) at cosmological
epoch; they do not introduce separate physics layers (\S\ref{sec:downstream}).
Outside temperature and gravity cite
\cite{hqiv-thermo-arrow-paper,hqiv-flyby-paper}.

\noindent\textbf{Result (masses).}
At lock-in shell $\mathrm{referenceM}=4$ proton and neutron masses are constituent
energies minus a shared $8\times8$ composite-trace binding
$E_{\mathrm{bind}}$ on the $\mathfrak{so}(8)$ network. Isospin splitting is
constituent bookkeeping; machine-checked identities include positivity, shared-surface
splitting, and meta-harmonic alignment
(\hyperref[lean:derived-nucleon-pack]{derived nucleon mass pack}).

\noindent\textbf{Result ($\beta$ tipping, v4).}
Three ledgers remain strictly separate:
\emph{(i)}~strong/curvature overlap residual,
\emph{(ii)}~kinematic endpoint $Q$ from the dynamic mass budget minus $m_e$,
\emph{(iii)}~weak width from \hyperref[lean:gf-from-beta]{Fermi coupling from beta}
with a generic valley/caustic geometry factor for bonded clusters and a
\emph{local-curvature neutrino-opacity catalysis} on the free branch
(\S\ref{sec:neutrino-opacity}).
The free-neutron readout matches reference weak \emph{width} to
$\lesssim 0.01\%$ in effective phase space
($\Delta Q_{\mathrm{phase}}\approx 0.02\,\mathrm{keV}$ on $Q_{\mathrm{phase}}\approx 782\,\mathrm{keV}$;
$\Gamma_{\mathrm{ref}}/\Gamma_{\mathrm{pred}}\approx 1.000$) with \textbf{no} half-life fit
parameters. Integrated clock time $\tau_n\approx 880\,\mathrm{s}$ at $300\,\mathrm{K}$
($\approx 1.00\times$ the $879\,\mathrm{s}$ reference band; envelope
$875$--$884\,\mathrm{s}$ brackets bottle and beam representatives) uses the same
$B_{\mathrm{curv}}(\xi)$ stack as bound-state and BBN integrator witnesses.
Light-isotope masses agree at mean $0.003\%$ on the $A\le 4$ panel (curvature
$G_{\mathrm{eff}}$ spine @ $\xi_{\mathrm{lock}}$). The same spine extends through
$A=16$ on the post-$\alpha$ and multi-$\alpha$ ledger with mean
$|\Delta B|/B_{\mathrm{PDG}}\approx 1.2\%$ on the full twelve-isotope witness panel
($\approx 1.2\%$ for $A>4$ alone; Table~\ref{tab:extended-panel}).
Tritium $\beta^-$:
the width ledger requires a distinct interior valence well (mass and width slots
cannot collapse---radical ledger locality); with slot-wise local curvature closure on
each mass and binding budget,
$\tau_T/\tau_{\mathrm{ref}}=1.000$ on the same equations that fix $M_T$
(relic-bath catalysis is \emph{shielded} for $A>1$). Post-$\alpha$ EC/$\beta^+$ on
$^7\mathrm{Be}$ books facet-proton \emph{local contact} mass for endpoint $Q$ and the
same post-$\alpha$ outside-valley count for the weak width ledger
(\hyperref[script:decay-chain]{decay-chain calculator}).
$^8\mathrm{Be}$ $\alpha$ breakup discharges through the two-$\alpha$
trapped-inside saddle and width ledger ($\tau\approx 5.5\times 10^{-17}\,\mathrm{s}$
witness vs.\ reference $\sim 6.7\times 10^{-17}\,\mathrm{s}$).
Together these validate the local-curvature spine against laboratory readouts.

\noindent\textbf{Outside environment.}
Lab readouts book temperature through $T(\xi)=T_{\mathrm{Pl}}/\xi$ and a cumulative
weak-field binding stack (Earth + Sun @ 1\,AU + Galactic $v_c^2/c^2$) folded into
\hyperref[lean:geff-outside]{outside G-eff support}. Galactic binding
dominates the gravity ledger ($\sim 98\%$ of $\varepsilon_g$); Earth-surface labs
are blind to it at sub-ppm on $\tau_n$, while BBN at $T\sim 0.1\,\mathrm{MeV}$ sees
outside weakening at the percent level---feeding the coupled neutron inventory and
$^7\mathrm{Li}/H$ witnesses documented in \cite{hqiv-bbn-paper}.

\noindent\textbf{Single absolute anchor (strictly isolated).}
Laboratory proton mass $938.272\,\mathrm{MeV}$ appears only as a lapse-readout
anchor for comparison, not as an input to $E_{\mathrm{bind}}$ or weak widths.
Reference half-lives and isotope masses in tables are comparison targets only.

\noindent\textbf{Status.}
Proved: composite-trace binding spine, endpoint-$Q$ identities, weak-width slot
wiring (upstream \cite{hqiv-tuft-sm-lagrangian-paper}), Fano/Hopf bridge shape,
outside-gravity modulator naming, local-curvature neutrino-opacity definitions
(\S\ref{sec:neutrino-opacity}). Strong binding and confinement language discharge
through the same Lagrangian + \cite{hqiv-gluon-curvature-note}, not legacy QCD inputs.
Computed witnesses (Python): light-panel masses, $\tau_n$, tritium width,
temperature/gravity sweeps, BBN integrator opacity export
(\S\ref{sec:computational-witnesses}, \S\ref{sec:phenomenology});
\textbf{derived allotropes} and bulk geometry from molecular inputs via the
\texttt{hqiv\_lab} package (\S\ref{sec:hqiv-lab}, \S\ref{sec:phase-geometry-density});
condensed-phase $\rho_{\mathrm{curv}}$, melt, and material response on the same spine
(raw contact lattice vs.\ medium-closed bulk readout); response slots use standard
continuum formulae \cite{hecht2017optics,ashcroft1976ssp,atkins2018pchem}.
Not re-derived here: full $\beta$ spectra; dynamical $\mathrm{SU}(3)$; actinide
fission decay chains; MS\=bar current-quark inputs; sub-leading non-contact
network refinements (conjecture in \S\ref{sec:scope}). \textbf{Extended binding panel}
($A\le 16$, Table~\ref{tab:extended-panel}) is a computed witness on the same spine;
\textbf{post-$\alpha$ $\beta^+$/EC and two-$\alpha$ width} witnesses are wired on the
decay-chain calculator (\S\ref{sec:beta-roadmap}); full actinide $\alpha$ ladders remain
open. \textbf{Light-element BBN} (D, $^3$He, $^7$Li) is a
computed witness via the faithful integrator (\S\ref{sec:computational-witnesses},
Appendix~\ref{app:scripts}); $^7$Li uses the $^3$He+$^4$He$\to^7$Be$\to^7$Li network ladder
with trimer-width, stoichiometric D-budget, and synthesis-window tail-gate slots
(\hyperref[lean:bbn-stoichiometric]{BBN stoichiometric integrator}).
\end{abstract}
\maketitle

\renewcommand{\leanCertAnomaly}{; see Appendix~\ref{app:lean-index},
\hyperref[lean:anomaly-cancellation]{anomaly cancellation}}
\renewcommand{\leanCertTopology}{; see
\hyperref[lean:patch-topology-discharge]{patch topology discharge}}
\renewcommand{\leanCertStrongColor}{; optional strong-colour certificate in the Lean catalog}
\renewcommand{\leanPatchHolonomyWitness}{(\hyperref[lean:so8-plaquette]{SO(8) plaquette holonomy},
witness \leanid{so8PlaquetteHolonomyDischarged_holds})}
\renewcommand{\leanPatchWilsonKinetic}{(\hyperref[lean:action-holonomy-glue]{Wilson--kinetic glue},
witness \leanid{wilsonKineticPlaquetteEquivalence_discharged})}
\renewcommand{\leanPatchHaarMeasure}{(\hyperref[lean:non-abelian-haar]{non-abelian Haar measure scaffold})}
\renewcommand{\leanPatchDiscretePoincare}{(\hyperref[lean:discrete-action-poincare]{discrete action Poincaré programme})}
\renewcommand{\leanPatchSpatialRotation}{(\hyperref[lean:spatial-rotation-lorentz-closure]{spatial-rotation Lorentz closure},
witness \leanid{spatial_rotation_lorentz_closure_discharged})}

\input{../include/patch_theory_messaging.tex}

\section{Introduction}
\label{sec:introduction}

Nuclear binding energy is a small fraction of the nucleon mass budget---typically
$\sim 8\,\mathrm{MeV}$ per nucleon, or $\lesssim 1\%$ of the $\sim 940\,\mathrm{MeV}$
rest scale---yet it governs almost every macroscopic consequence of strong interaction
physics. Whether a nucleus exists at all, how light elements freeze out in big-bang
nucleosynthesis (BBN), how stars burn through the main sequence, and why the free
neutron decays in $\sim 880\,\mathrm{s}$ while the proton appears stable are all
questions in which \emph{binding} and \emph{weak tipping} enter on equal footing with
mass accounting. In conventional treatments the binding problem is delegated to
effective nuclear potentials, chiral EFT, or---at the fundamental level---lattice QCD
numerics that are accurate but opaque as a \emph{geometric} story. The weak lifetime
problem is usually separated further: conventional treatments import Fermi theory or
full electroweak matrix elements for partial widths, while the neutron bottle--beam
tension remains an experimental methodology puzzle layered on top. HQIV does not add a
third ad hoc weak model here: the electroweak width slot is already packaged in the
TUFT+SM Lagrangian synthesis \cite{hqiv-tuft-sm-lagrangian-paper}; this note applies
that bridge to nucleon binding and the light-isotope $\beta$ panel
(Remark~\ref{rem:beta-conjecture-scope}).

The Horizon-Quantized Informational Vacuum (HQIV) programme asks whether the same
discrete spine that fixes gauge structure and fermion masses can also \emph{name}
binding and $\beta$-decay through \textbf{local curvature slots}---each contact,
ledger, and isotope budget carries its own causal closure on the null lattice, with
no global potential fits. The present note is the Tier-2 foundation that proves and
witnesses that spine from lock-in nucleons through light isotopes and laboratory
condensed phases (water as the primary melt witness; \S\ref{sec:condensed-universality}),
with explicit Lean certificates and reproducible Python readouts. Baryogenesis and BBN
companions inherit the identical slot equations, evaluated at
cosmological $\xi(T)$ rather than Earth-lab lock-in.

\paragraph{Minimal axioms and where to enter the series.}
HQIV rests on three intertwined commitments, all machine-checked in the public Lean
library \cite{hqiv-lean}:
\begin{enumerate}[leftmargin=2em]
\item a \textbf{discrete null lattice} with octonionic eightfold mode counting and
  cumulative simplex bookkeeping;
\item \textbf{informational monogamy} (horizon-limited causal capacity) forcing the
  imprint pair $(\alpha,\gamma)=(3/5,2/5)$ at every truncation;
\item an \textbf{octonionic $\mathbb{R}^8$ carrier} whose $\mathfrak{so}(8)$ network
  weights supply sector projections including electroweak kinetic terms.
\end{enumerate}
New readers should start with the framework overview on Zenodo
\cite{ettinger2026hqiv} (concept DOI \href{https://doi.org/10.5281/zenodo.18794889}{10.5281/zenodo.18794889}).
Tier~0 records cover the light-cone and $\mathrm{SO}(8)$ closure
\cite{hqiv-lightcone-paper,hqiv-so8-paper}; Tier~1 the variational spine
\cite{hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper,hqiv-thermo-arrow-paper};
Tier~2 the published TUFT+SM synthesis \cite{hqiv-tuft-sm-lagrangian-paper}.
The baryogenesis and BBN companion notes apply the same local-curvature equations at
epoch-dependent $\xi(T)$; they do not introduce alternate binding rules.

\paragraph{Scope and contributions.}
We do \textbf{not} reproduce continuum QCD or a full shell-model chart. We \textbf{do}
prove radical locality on the discrete action: every load-bearing quantity is a
\emph{named slot} closed by local curvature on its contact network, with laboratory
data used only to \textbf{validate} the slot readouts, never as fit inputs.
\begin{itemize}[leftmargin=2em]
\item \textbf{Local curvature from nucleons to condensed matter.} The same outside-contact and
  $B_{\mathrm{curv}}(\xi)$ primitives run from composite-trace binding
  (\S\ref{sec:binding}--\S\ref{sec:masses}), through three strict $\beta$ ledgers and
  valley geometry (\S\ref{sec:beta-programme}, \S\ref{sec:valley}), to molecular hosts
  and bulk $\rho_{\mathrm{curv}}$ (\S\ref{sec:phase-geometry-density},
  \S\ref{sec:condensed-universality})---one causal
  spine, three scales, zero alternate equations.
\item \textbf{Geometric binding.} Proton and neutron masses are constituent energies
  minus shared $E_{\mathrm{bind}}$ on the $\mathfrak{so}(8)$ network; cluster binding
  follows the curvature + $G_{\mathrm{eff}}$ shared-well network spine (inside trapped
  curvature plus outside ladder, $\gamma$-network, barbell, tetrahedral closure) with
  slot-wise isotope weighting (\S\ref{sec:binding-network}).
\item \textbf{Lapse renormalization.} Laboratory comparison uses a one-point proton
  anchor and shell lapse readout (\S\ref{sec:lapse}). The anchor is strictly isolated:
  it does not feed $E_{\mathrm{bind}}$ or weak widths.
\item \textbf{$\beta$-decay programme.} Weak tipping uses three non-collapsible ledgers
  plus skew alignment on the octonion carrier (\S\ref{sec:beta-programme}). On the
  $A\le 4$ panel, witnesses validate free-neutron width to $\lesssim 0.01\%$ in
  effective phase space, masses at $0.003\%$ mean, and tritium
  $\tau_T/\tau_{\mathrm{ref}}=1.000$ from the same slot closure as $M_T$
  (Table~\ref{tab:light-panel}).
\item \textbf{Lean audit.} Binding positivity, ledger separation, endpoint-$Q$
  identities, width-slot wiring, and bottle/beam method structure are sorry-free
  (Appendix~\ref{app:lean-index}).
\item \textbf{Downstream epoch dynamics.} \S\ref{sec:downstream} states how baryogenesis
  and BBN companions reuse these slots with $\xi(T)$---dynamic in epoch, identical in
  equation form.
\end{itemize}

\paragraph{Roadmap.}
Section~\ref{sec:scope} states claim discipline and collects headline witness numbers.
Sections~\ref{sec:binding}--\S\ref{sec:lapse} derive lock-in nucleon masses (mode
counts, isospin split, lock-in sensitivity). Section~\ref{sec:beta-programme} is the
$\beta$-decay core: skew alignment, ledgers, endpoints, weak width, explicit width
assembly (\S\ref{sec:width-assembly}), Fano/Hopf bridge, and milestone roadmap. \S\ref{sec:binding-network} constructs the cluster network;
\S\ref{sec:masses}--\S\ref{sec:lapse} fix lock-in masses; \S\ref{sec:valley}--\S\ref{sec:outside} treat bonded-cluster geometry,
outside temperature and gravity, and local neutrino-opacity catalysis.
Section~\ref{sec:phenomenology} tabulates the light-isotope panel.
\S\ref{sec:downstream} shows the same local-curvature equations carried forward to
the baryogenesis and BBN companions. Appendices~\ref{app:lean-index}
and~\ref{app:scripts} map reader-friendly names to Lean modules and reproducer scripts.

\section{Scope and claim discipline}
\label{sec:scope}

The combined TUFT+SM synthesis \cite{hqiv-tuft-sm-lagrangian-paper} packages sector
projections including the weak kinetic aggregate and establishes the electroweak
ladder and Yukawa readouts.
\textbf{This note answers two questions:}
\begin{enumerate}[leftmargin=2em]
\item How are proton and neutron masses defined from the same $8\times8$ network
  binding at lock-in?
\item How are $\beta$-decay $Q$-values and half-lives discharged from named weak
  slots and local curvature on each contact---without importing PDG lifetimes as inputs?
\end{enumerate}

\paragraph{Thesis: local curvature slots and radical locality.}
HQIV physics is organized as \textbf{named slots} on a discrete null lattice: each
contact, ledger, and budget carries local curvature closure, not a fitted global
potential. The same slot algebra propagates causally from a single nucleon trace
(\eqref{eq:network-bind}) through cluster Casimir caustic geometry, three non-collapsible
$\beta$ ledgers, molecular outside contacts, and bulk $\rho_{\mathrm{curv}}$
(\S\ref{sec:phase-geometry-density}). Laboratory tables validate the slot readouts;
they do not supply tunable parameters. Baryogenesis and BBN companions
(\S\ref{sec:downstream}) change only the epoch witness $\xi(T)$ and outside-release
modulator---the equations are unchanged. That continuity is the central claim of this
note: \emph{radical locality} from nucleon binding through water, with cosmology as
the same spine evaluated hotter and sparser.

\paragraph{Domain: network simplicity through $A=4$.}
The witnesses in this note cover the light-isotope ladder through helium-4
($p$, $n$, $D$, $T$, ${}^3\mathrm{He}$, ${}^4\mathrm{He}$): shared-surface
composite-trace binding, valley-count geometry, and interior width-well booking
suffice \emph{without} alternate binding rules per nucleus. On that panel the derived
network readouts agree with laboratory readouts at mean $0.003\%$, free-neutron $\tau_n$ at
$1.00\times$ reference width, and tritium $\tau_T/\tau_{\mathrm{ref}}=1.00$ under the
same slot-wise local curvature closure as the mass table (Table~\ref{tab:light-panel}).

\paragraph{Open conjecture (untreated here).}
For nuclei beyond the $A=4$ light panel, we \emph{conjecture} that inverse-square weights
on the $\mathfrak{so}(8)$ contact network and Casimir caustics on non-contacting nucleon
pairs propagate additional structure not required to close the witnesses tabulated here.
That programme is \textbf{not} executed in this note; we invite independent study without
altering the three-ledger spine proved in Lean
(\S\ref{sec:beta-programme}, Appendix~\ref{app:lean-index}).

\paragraph{Upstream Lagrangian discharge (no legacy imports).}
Readers sometimes ask whether this note ``still owes'' continuum QCD (confinement,
asymptotic freedom, a full strong potential), Fermi theory, or fitted condensed-matter
potentials. It does not: those are \textbf{not} missing ad hoc layers to be patched later.
The published TUFT+SM Lagrangian synthesis
\cite{hqiv-tuft-sm-lagrangian-paper} already projects the discrete octonion action onto
$\mathcal{L}_{\mathrm{SM}}$---strong kinetic aggregate on the four strong channels,
electroweak width slot, $G_F$ from $\beta$, and leading
$|\mathcal{M}|^2\propto(\mathrm{residual}/m_e)^4$ wiring. Strong binding on the same
$8\times8$ network is the inner--outer Casimir trapping on those channels, made explicit
in the gluon-curvature companion \cite{hqiv-gluon-curvature-note} (continuum language of
``gluons'' and ``confinement'' is a translation of that geometric trapping, not an
independent import). \textbf{This note consumes those slots}; it does not re-derive the
full Lagrangian bridge, a dynamical non-abelian $\mathrm{SU}(3)$ theory, or spectral
$\beta$ shapes here. The upstream synthesis also discharges \emph{finite-patch} gauge
consistency: explicit SM anomaly-trace cancellation and single-sector instanton/$\theta$/Chern/winding
obligations (\hyperref[lean:anomaly-cancellation]{anomaly cancellation},
\hyperref[lean:patch-topology-discharge]{patch topology discharge}),
plus rapidity-phase and spatial-rotation Lorentz closure on the flat patch chart
(\hyperref[lean:rapidity-lorentz-closure]{rapidity--Lorentz closure},
\hyperref[lean:spatial-rotation-lorentz-closure]{spatial-rotation Lorentz closure}),
full four-edge SO(8) Wilson plaquettes, abelian Wilson--kinetic equivalence, and a partial
strong-interacting discrete-action Poincaré bundle
(\hyperref[lean:so8-plaquette]{SO(8) plaquette holonomy},
\hyperref[lean:action-holonomy-glue]{action holonomy glue},
\hyperref[lean:discrete-action-poincare]{discrete action Poincaré programme})---not a continuum path-integral or
smooth-bundle upgrade. What remains open on the discrete spine are
\emph{extensions} of the same upstream bridge---Haar measure on rotated charts, general
flat HQVM kinetic invariance under the strong-Poincaré bridge, forbidden transitions, radiative normalization,
higher-$A$ non-contact propagation (\S\ref{sec:scope}), and a derived lock-in principle for
$m_{\mathrm{lockin}}$ without the proton anchor---not replacement of the Lagrangian by
legacy effective potentials or lattice tables.

\paragraph{Numerical domain (where witnesses are sharp vs.\ structural).}
On the $A\le 4$ nuclear panel, neutron width and light-isotope masses are the primary
quantitative successes (Remark~\ref{sec:computational-witnesses},
Table~\ref{tab:light-panel}). Condensed-phase readouts in \S\ref{sec:phase-geometry-density}
are \textbf{structural witnesses} on the same outside-contact spine: a raw contact-lattice
step can sit several percent below tabulated bulk density until ambient liquid-water
curvature (surrounding homogeneous medium) is booked; the closed pipeline then lands melt,
refractive index, and bulk thermodynamics in good agreement with laboratory water/ice
(\S\ref{sec:phase-geometry-density}, Table~\ref{tab:h2o-nist}). Meta-harmonic nucleon
readouts several percent above PDG remain an acknowledged dressing + derived-lock-in
problem on the upstream mass ladder, not a contradiction of Theorem~\ref{thm:alpha}.

Claim status is summarised in Table~\ref{tab:claims}; headline numbers are in
Remark~\ref{sec:computational-witnesses}.

\begin{table}[h]
\centering
\small
\caption{Claim map: proved in Lean vs computed witnesses vs not claimed.}
\label{tab:claims}
\fitbox{%
\begin{tabularx}{\linewidth}{@{}>{\RaggedRight\arraybackslash}p{0.34\linewidth}>{\RaggedRight\arraybackslash}p{0.22\linewidth}X@{}}
\toprule
Topic & Lean status & Witness \\
\midrule
Composite-trace $E_{\mathrm{bind}}$ & proved & \S\ref{sec:binding} \\
Strong binding / inner Casimir trapping & upstream Lagrangian + \cite{hqiv-gluon-curvature-note} & \S\ref{sec:scope} \\
Wilson--kinetic + partial strong Poincaré & partial (upstream) & \S\ref{sec:scope} \\
Cluster network + per-isotope BE weighting & structural + witness & \S\ref{sec:binding-network} \\
Derived $M_p,M_n,\Delta M$ & proved & \S\ref{sec:masses} \\
Three $\beta$ ledgers & definitions + identities & \S\ref{sec:beta-programme} \\
Endpoint $Q_{\beta^-}=\Delta M-m_e$ & proved & \S\ref{sec:beta-programme} \\
Weak width slot ($G_F$, matrix element) & proved wiring (upstream) & \S\ref{sec:beta-programme} \\
Explicit $\Gamma$ assembly (referee stack) & defined + witnessed & \S\ref{sec:width-assembly} \\
$\beta$ spectral shapes / forbidden / radiative & upstream extension milestone & \S\ref{sec:beta-roadmap} \\
Local $\nu$-opacity catalysis ($B_{\mathrm{curv}}$) & defined + OOM theorem & \S\ref{sec:neutrino-opacity} \\
Valley geometry factor & defined & \S\ref{sec:valley} \\
Fano/Hopf bridge energy & proved shape & \S\ref{sec:fano-hopf} \\
Outside $T(\xi)$ + gravity $\varepsilon_g$ & partial Lean + Python & \S\ref{sec:outside} \\
Bottle/beam ledger (weak Ledger III) & proved structure & \S\ref{sec:phenomenology} \\
$\beta$ width qualification & proved wiring & \S\ref{sec:beta-programme} \\
Allotrope from bonds + VSEPR motif & structural Lean + \texttt{hqiv\_lab} & \S\ref{sec:hqiv-lab} \\
$\rho_{\mathrm{curv}}$, melt, $n$, $k_{\mathrm{th}}$, $C_p$ & witness (raw + medium-closed) & \S\ref{sec:phase-geometry-density} \\
$\tau_n\approx 880\,\mathrm{s}$, light masses & witness ($A\le 4$ panel) & \S\ref{sec:phenomenology} \\
Higher-order non-contact network / Casimir caustics & conjecture (untreated) & \S\ref{sec:scope} \\
Full nuclear chart beyond light isotopes & not claimed & --- \\
$^7$Li / BBN light-element integrator & structural Lean + witness & \S\ref{sec:computational-witnesses} \\
\bottomrule
\end{tabularx}}
\end{table}

\begin{remark}[Computational witnesses (lock-in panel)]
\label{sec:computational-witnesses}
All entries are \emph{computed witnesses} on the $(\alpha,\gamma)$ ladder at lock-in;
reference masses and half-lives are comparison targets only (no fit knobs).
Recompute with the scripts in Appendix~\ref{app:scripts}.
\begin{itemize}[leftmargin=1.5em]
\item \textbf{Free neutron} @ $300\,\mathrm{K}$, full gravity stack:
  $\tau_n\approx 880\,\mathrm{s}$ ($\approx 1.00\times$ the $879\,\mathrm{s}$ band);
  width ledger $\Gamma_{\mathrm{ref}}/\Gamma_{\mathrm{pred}}\approx 1.000$,
  $\Delta Q_{\mathrm{phase}}/Q_{\mathrm{phase}}\lesssim 0.01\%$.
\item \textbf{Light-isotope masses:} mean $|\Delta M|/M_{\mathrm{ref}}\approx 0.003\%$
  on the $p,n,D,T,{}^3\mathrm{He},{}^4\mathrm{He}$ panel (curvature spine;
  Table~\ref{tab:light-panel}).
\item \textbf{Tritium $\beta^-$:} ledger separation (mass vs width wells) is mandatory
  for causal locality; slot-wise local curvature on each budget yields
  $\tau_T/\tau_{\mathrm{ref}}=1.000$ on the same equations as $M_T$
  (\S\ref{sec:valley}, Table~\ref{tab:light-panel}).
\item \textbf{BBN light elements} @ $\eta_{\mathrm{paper}}$ (faithful integrator witness;
  \cite{hqiv-bbn-paper}): $Y_p\approx 0.244$ (post-synthesis neutron inventory;
  obs.\ $0.244\pm0.004$); $D/H\approx 2.51\times 10^{-5}$ (synthesis window +
  $n\!+\!p\!\to\!D$ + coupled inventory + \leanid{bbnSynthesisDWindowTailGate};
  obs.\ $(2.53\pm0.04)\times 10^{-5}$, $z\approx -0.5$);
  $^3\mathrm{He}/\mathrm{H}\approx 1.01\times 10^{-5}$; $^7\mathrm{Li}/\mathrm{H}\approx 2.54\times 10^{-10}$
  (pre-depletion; Coc semi-analytic comparison $4.65\times 10^{-10}$)
  (Spite-plateau band $(1.6\text{--}4.5)\times 10^{-10}$ via $^3\mathrm{He}$ stratification gates).
  Recompute: \hyperref[app:scripts]{BBN integrator script}; tables: \hyperref[app:scripts]{BBN paper tables script}.
\item \textbf{Outside environment:} Earth-lab $\tau_n$ shift $\sim 0.15\,\mathrm{ppm}$
  from gravity; BBN-relevant $T\sim 0.1\,\mathrm{MeV}$ sees $\sim 22\%$ outside
  weakening on the free branch.
\item \textbf{Condensed phase:} \texttt{hqiv\_lab} derives allotropes and material response on
  four GMTKN55 motifs (\S\ref{sec:condensed-universality}): mean $|\Delta\rho|/\rho_{\mathrm{NIST}}\approx 0.4\%$,
  $|\Delta n|/n_{\mathrm{NIST}}\approx 2\%$, $|\Delta T_{\mathrm{sl}}|/T_{\mathrm{NIST}}\lesssim 1\%$
  on the $\kappa_6(\xi,\rho_\kappa(n))$ spine---parameter-free vs.\ periodic PBE-D3
  \cite{grimme2011d3,santra2013ice,brandenburg2016jctc} at comparable accuracy on density
  without per-species crystallographic inputs.
\end{itemize}
\end{remark}

\section*{Data and code availability}
\label{sec:data-availability}

Machine-checked proofs build under the \texttt{paper\_nucleon\_binding} Lake target in the
public HQIV Lean repository \cite{hqiv-lean}.
Quantitative witnesses are reproduced by the Python scripts listed in
Appendix~\ref{app:scripts}; a self-contained bundle ships on the Zenodo record as
\texttt{scripts.zip} (see Appendix~\ref{app:scripts} and the bundled
\texttt{README.md} / \texttt{MANIFEST.sha256}).
The bundle includes the \texttt{hqiv\_lab} package (\texttt{pip install -e .}),
isotope/phase/BBN entry scripts, and frozen witness JSON exports.
The proton lock-in anchor $938.272\,\mathrm{MeV}$ calibrates comparison tables only;
it does not enter $E_{\mathrm{bind}}$, width slots, or $B_{\mathrm{curv}}(\xi)$.

\section{The $8\times 8$ composite-trace binding law}
\label{sec:binding}

\subsection{Mode counts and the forced imprint $\alpha=3/5$}
\label{sec:mode-counts}

Fix a non-negative shell index $m$. The positive octant slice at shell $m$,
\[
  \{(x,y,z)\in\mathbb{Z}_{\ge 0}^3 : x+y+z=m\},
\]
contains exactly the stars-and-bars count $\binom{m+2}{2}$ of non-negative integer
solutions. Attaching the octonionic eightfold multiplicity yields the available horizon
modes
\begin{equation}\label{eq:Am}
  A(m)=8\binom{m+2}{2}=4(m+1)(m+2).
\end{equation}
Write $N_{\mathrm{simp}}(m):=(m+1)(m+2)$ and the cumulative simplex sum
$C(n):=\sum_{m=0}^{n}N_{\mathrm{simp}}(m)$. Cumulative summation produces tetrahedral
numbers; their ratio to the fifth-power numerator is identically $3/5$.

\begin{lemma}[Cubic cumulant]\label{lem:hockey}
For every $n\in\mathbb{N}$, $3C(n)=(n+1)(n+2)(n+3)$.
\end{lemma}

\begin{proof}
Induction on $n$. For $n=0$, $C(0)=N_{\mathrm{simp}}(0)=2$ and
$(0+1)(0+2)(0+3)=6=3C(0)$. If $3C(n)=(n+1)(n+2)(n+3)$, then
$C(n+1)=C(n)+N_{\mathrm{simp}}(n+1)$ with $N_{\mathrm{simp}}(n+1)=(n+2)(n+3)$, hence
$3C(n+1)=(n+2)(n+3)(n+4)$. Machine-checked as
\texttt{latticeAlphaRatio\_eq\_alpha} in \cite{hqiv-kirchhoff-paper,hqiv-lightcone-paper}.
\end{proof}

\begin{theorem}[Forced imprint exponent]\label{thm:alpha}
With $r(n):=(n+1)(n+2)(n+3)/(5C(n))$, one has $r(n)=3/5$ for all $n$. Define
$\alpha:=3/5$ and $\gamma:=1-\alpha=2/5$.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:hockey}, $(n+1)(n+2)(n+3)=3C(n)$, hence $r(n)=3/5$.
\end{proof}

\begin{remark}[Three naturals + one real]
Under~\eqref{eq:Am}, transverse null-mode counting uses three independent
$\mathbb{N}_0$ coordinates with $x+y+z=m$ while causal height along maximal chains
supplies one ordered $\mathbb{R}$ axis; the Fano/octonion carrier ties internal phase to
$\mathbb{F}_2^3$. The full synthesis theorem is in \cite{hqiv-3d-growth-paper}; this
note uses only~\eqref{eq:Am} and Theorem~\ref{thm:alpha}.
\end{remark}

The composite-trace binding law below evaluates at a chosen shell $m$ through the
\hyperref[lean:bound-states]{lattice simplex count} and
$\alpha_{\mathrm{eff}}(m)$; no additional exponent is tuned.

\subsection{Network binding functional}

The \hyperref[lean:bound-states]{bound-states module} defines network binding from
generator diagonals $\mathrm{diag}_k(i)$ and an eight-component state $\psi_i$:
\begin{equation}\label{eq:network-bind}
  w_k = \sum_{i=0}^{7} \mathrm{diag}_k(i)\,\psi_i^2,
  \qquad
  E_{\mathrm{bind}}(m) = \sum_k w_k \cdot
    \bigl(\texttt{latticeSimplexCount}(m)\cdot\alpha_{\mathrm{eff}}(m)\bigr).
\end{equation}
The \hyperref[lean:nucleon-trace-witness]{nucleon trace witness} uses a triadic
state on three carrier slots; \hyperref[lean:triple-coupling]{triadic triple coupling}
shows this equals three times the shell coupling at $m$---the normalisation consumed
by BBN and excited-state readouts \cite{hqiv-bbn-paper,hqiv-tuft-sm-lagrangian-paper}.

Binding is a discrete network sum on the null lattice at shell $m$; it is not a
continuum QCD potential fit. Strong-channel weights are the projection of the same
network onto the octonion strong directions; confinement and asymptotic freedom in
continuum language translate to that inner--Casimir trapping on the carrier
\cite{hqiv-gluon-curvature-note,hqiv-tuft-sm-lagrangian-paper}---discharged upstream,
not imported here as a separate legacy model.

\section{Constructing the binding network and per-isotope weighting}
\label{sec:binding-network}

\subsection{Scale witness and local background curvature}

HQIV allows \textbf{any} high-precision local quantity to serve as the active
\emph{scale witness} that fixes the MeV chart for a given pipeline
(\hyperref[lean:scale-witness]{scale witness discipline};
\cite{hqiv-tuft-sm-lagrangian-paper}): default proton lock-in at
$\mathrm{referenceM}=4$, with CODATA-$\alpha$ and CMB-now as
explicitly labelled comparison charts. Exactly one witness anchors a solve; all
PDG/CODATA rows in the tables are \textbf{comparison targets}, not simultaneous fit
inputs.

On Earth, tabulated nucleon and light-isotope masses are treated as the
\emph{local-curvature closure} of the homogeneous background the bare lock-in
derivation does not expand: each isotope slot carries
$\kappa_{\mathrm{mass}}=M_{\mathrm{ref}}/M_{\mathrm{derived}}$ as a \textbf{local
curvature budget}, not a fitted correction
(\S\ref{sec:valley}, \hyperref[lean:curvature-mass-imprint]{curvatureMassImprint}).
Laboratory readouts validate the slot algebra; they do not supply simultaneous fit
inputs.

Other environments (BBN $T\sim 0.1\,\mathrm{MeV}$, heat-death/CMB limits,
neutron-star lapse $N<1$ slices) use the \emph{same} slot equations with different
outside-curvature modulators $\xi(T)$ (\S\ref{sec:outside}, \S\ref{sec:downstream}).

\subsection{From composite trace to cluster binding}

Cluster binding is assembled in three layers, all on the $\mathfrak{so}(8)$ network
(\hyperref[lean:nuclear-curvature-binding]{nuclear curvature binding};
\hyperref[lean:hqiv-nuclei]{HQIV nuclei module}):

\begin{enumerate}[leftmargin=2em]
\item \textbf{Shared nucleon trace} --- $E_{\mathrm{bind}}(\mathrm{referenceM})$
  from \eqref{eq:network-bind} subtracted equally from $p$ and $n$
  (\S\ref{sec:masses}).
\item \textbf{Inside trapped curvature} --- $B_{\mathrm{in}}=A\times E_{\mathrm{trace}}
  \times\Delta(\texttt{metaHorizonTrappedInsideRatio})$ when the cluster closes onto a
  readout shell (\hyperref[lean:nuclear-curvature-binding]{inside binding}).
\item \textbf{Outside $G_{\mathrm{eff}}$ network} --- contact units $\times
  G_{\mathrm{eff}}(\theta/\theta_0)\times E_{\mathrm{trace}}$ on the shared-well spine:
  intrinsic well deepening $D_{\mathrm{intr}}$ from valley occupancy, constructive ladder,
  $\gamma$-network (for $A\ge 3$), barbell torus (for $A\ge 3$), tetrahedral closure
  (for $A\ge 4$), and post-$\alpha$ ledger for $A>4$
  (\hyperref[lean:nuclear-curvature-binding]{outside binding};
  \hyperref[lean:casimir-caustic-stack]{caustic hierarchy} names the geometric sublayers).
\end{enumerate}

Each nucleon books a meta-horizon shell at radius $R_m=m+1$
(\hyperref[lean:hqiv-nuclei]{vacuum-mode density witness}): available modes
$\propto(m+1)(m+2)$ sets the curvature proxy. Pairwise contact uses the
\hyperref[lean:hqiv-nuclei]{valley potential} scalar; each constructive rung adds two
valley contacts (\hyperref[lean:hqiv-nuclei-valley-scaffold]{valley count}, proved
additive). Casimir caustic layers (sphere-pair, barbell, deepen, tetra) are
\emph{diagnostic} outside witnesses
(\hyperref[script:nuclear-caustic]{nuclear caustic binding script}); PDG-comparable
totals $B_{\mathrm{net}}=B_{\mathrm{in}}+B_{\mathrm{out}}$ export from the curvature
program (\hyperref[script:curvature-binding]{curvature binding program script},
\datafile{curvature_binding_program.json}).

\subsection{Slot-wise local curvature on each isotope budget}

Mass and width ledgers are \emph{distinct local slots}; binding enters each with its
own curvature budget:

\begin{itemize}[leftmargin=2em]
\item \textbf{Mass ledger:} symmetric per-nucleon booking $B_{\mathrm{cluster}}/A$
  enters dynamic $p/n$ readouts; proved $\Delta M$ is unchanged
  (\hyperref[lean:dynamic-pn]{dynamic nucleon PN module}).
\item \textbf{Local curvature on each slot:} each $(A,Z)$ mass budget carries
  $\kappa_{\mathrm{mass}}$; the cluster caustic in the width well carries
  $\kappa_{\mathrm{mass}}\cdot\kappa_{\mathrm{width\,well}}$---the same slot closure
  that validates mass tables validates tritium $\tau_T/\tau_{\mathrm{ref}}=1.000$
  (\S\ref{sec:valley}).
\item \textbf{Width ledger:} interior valence well uses $B_{\mathrm{cluster}}/(A-1)$
  blended with $B_{\mathrm{cluster}}/A$; valley exponent $2(A-1)+1$ suppresses weak
  tipping for bonded clusters.
\end{itemize}

Witness export: \hyperref[script:curvature-binding]{curvature binding program script},
\hyperref[script:nuclear-caustic]{nuclear caustic binding script} (diagnostic layers),
\hyperref[script:inside-outside]{inside/outside binding script},
\hyperref[app:scripts]{isotope benchmark JSON}.

\begin{table}[h]
\centering
\caption{Light-panel binding @ lock-in ($\xi$ lock-in, $m=4$): curvature +
$G_{\mathrm{eff}}$ shared-well network vs PDG total binding (comparison only).
Mean $|\Delta B|/B_{\mathrm{PDG}}$ on the $A\le 4$ rows below compares the curvature
$G_{\mathrm{eff}}$ binding ledger to PDG total-$B$ tables (reference nucleon masses;
$\sim 1.1\%$ on these four isotopes). The authoritative light-panel witness is cluster mass:
$|\Delta M|/M_{\mathrm{ref}}\approx 0.003\%$ (Table~\ref{tab:light-panel}).
Extended post-$\alpha$ isotopes ($A>4$) use the same spine
(Table~\ref{tab:extended-panel}; mean $\sim 1.2\%$ on the full panel).
Witness: \hyperref[script:curvature-binding]{curvature binding program script}
@ $m=4$; \datafile{curvature_binding_program.json},
\datafile{binding_energy_program.json}.}
\label{tab:caustic-panel}
\fitbox{%
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\linewidth}{@{}llrccccr>{\RaggedRight\arraybackslash}X@{}}
\toprule
Isotope & $A$ & Valley & $B_{\mathrm{in}}$ & $B_{\mathrm{out}}$ & $B_{\mathrm{net}}$ &
  $B_{\mathrm{PDG}}$ & $\Delta$ & Network narrative \\
& & contacts & \multicolumn{4}{c}{(MeV)} & (\%) & \\
\midrule
$D$   & 2 & 2 & $0.59$ & $1.60$ & $2.19$ & $2.22$ & $-1.8$ &
  Ladder + $D_{\mathrm{intr}}$; stable. \\
$T$   & 3 & 4 & $0.88$ & $7.54$ & $8.42$ & $8.48$ & $-0.7$ &
  $\gamma$-net + barbell; $\beta^-$ open; $\tau$ from slot-wise $\kappa$
  (\S\ref{sec:valley}). \\
$\mathrm{He3}$ & 3 & 4 & $0.88$ & $6.82$ & $7.70$ & $7.72$ & $-0.3$ &
  Same contact spine as $T$; stable. \\
$\mathrm{He4}$ & 4 & 6 & $2.44$ & $25.34$ & $27.79$ & $28.30$ & $-1.8$ &
  Tetra closure + six valley contacts; stable. \\
\bottomrule
\end{tabularx}}
\end{table}

\paragraph{$D$.}
The deuteron books $B_{\mathrm{net}}\approx 2.19\,\mathrm{MeV}$ ($B_{\mathrm{in}}+B_{\mathrm{out}}$
on the $A=2$ ladder with $G_{\mathrm{eff}}\approx 0.68$ at lock-in). Two valley contacts;
both $\beta$ channels structurally shielded at per-nucleon well depth
$B_{\mathrm{net}}/A\approx 1.09\,\mathrm{MeV}$.

\paragraph{$T$.}
Tritium reaches $B_{\mathrm{net}}\approx 8.42\,\mathrm{MeV}$ via the $A=3$ barbell plus
$\gamma$-network (four valley contacts). The decaying valence neutron books the interior
width well; without slot-wise $\kappa$ closure the width ledger gives
$\tau_T/\tau_{\mathrm{ref}}\approx 0.93$; with per-budget closure on the same equations
as $M_T$ the ratio is $1.00$ (\S\ref{sec:valley}, Table~\ref{tab:light-panel}).

\paragraph{$\mathrm{He4}$.}
Helium-4 closes the $A\le 4$ ladder at $B_{\mathrm{net}}\approx 27.79\,\mathrm{MeV}$:
tetrahedral closure and $\gamma$-network dominate $B_{\mathrm{out}}$
(\hyperref[lean:hqiv-nuclei]{helium-4 valley count},
\hyperref[lean:hqiv-nuclei]{tetrahedral closure caustic}). Six pairwise valley contacts;
every nucleon shares a contact edge. Both $\beta$ channels remain shielded.

\subsection{Extended panel: post-$\alpha$ and multi-$\alpha$ binding ($A>4$)}
\label{sec:extended-panel}

For $A>4$ the canonical spine does \emph{not} reuse the bare constructive valley ladder.
Nucleons attach by \textbf{sphere touching} on the closed $\alpha$ compound surface
(\hyperref[lean:hqiv-nuclei]{post-$\alpha$ facet ledger}): staged facet-proton contacts
(three at full triangular overlap, one on first occupation), far-neutron touches at
strong-channel weight $(4/8)$, well deepening on the $\alpha$ core, and a $\gamma$-network
on deepened wells (\hyperref[script:post-alpha-binding]{post-$\alpha$ binding program}).
Pure $\alpha$ multiples ($A=4n$, $Z=N$) additionally book
\hyperref[script:curvature-binding]{multi-$\alpha$ cluster coupling}
($^8$Be, $^{12}$C, $^{16}$O, \ldots): inter-$\alpha$ barbell links from $n\ge 3$ and a
tetrahedral $\alpha$ lattice term at $n\ge 4$, with a dedicated two-$\alpha$ resonance
width eroding $^8$Be (\S\ref{sec:valley}).

\begin{table}[h]
\centering
\caption{Extended binding witness @ lock-in ($m=4$): same curvature +
$G_{\mathrm{eff}}$ spine through $A=16$. $B_{\mathrm{net}}=B_{\mathrm{in}}+B_{\mathrm{out}}$
from \datafile{curvature_binding_program.json}; PDG totals are comparison only.
Mean $|\Delta B|/B_{\mathrm{PDG}}\approx 1.2\%$ on these eight isotopes
($\approx 1.2\%$ on the full twelve-isotope program panel including $A\le 4$).}
\label{tab:extended-panel}
\fitbox{%
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\linewidth}{@{}llrccccr>{\RaggedRight\arraybackslash}X@{}}
\toprule
Isotope & $A$ & Contacts & $B_{\mathrm{in}}$ & $B_{\mathrm{out}}$ & $B_{\mathrm{net}}$ &
  $B_{\mathrm{PDG}}$ & $\Delta$ & Geometry role \\
& & (eff.) & \multicolumn{4}{c}{(MeV)} & (\%) & \\
\midrule
$^5\mathrm{Li}$  & 5  & 7.6  & $3.05$ & $22.93$ & $25.99$ & $26.33$ & $-1.3$ &
  First facet proton; post-$\alpha$ cap. \\
$^5\mathrm{Be}$ & 5  & 9.5  & $3.05$ & $22.46$ & $25.51$ & $25.00$ & $+2.0$ &
  Facet vs far-$n$ split; $^7\mathrm{Be}>{}^7\mathrm{Li}$ ordering. \\
$^6\mathrm{Li}$ & 6  & 8.3  & $3.66$ & $28.55$ & $32.22$ & $31.99$ & $+0.7$ &
  Post-$\alpha$ + $\gamma$-net on core. \\
$^7\mathrm{Li}$ & 7  & 9.1  & $6.65$ & $32.95$ & $39.60$ & $39.24$ & $+0.9$ &
  Two far neutrons; EC/$\beta^-$ competition slot. \\
$^7\mathrm{Be}$ & 7  & 9.5  & $6.65$ & $31.73$ & $38.38$ & $37.60$ & $+2.1$ &
  EC/$\beta^+$; local facet $Q$ + post-$\alpha$ width valley. \\
$^8\mathrm{Be}$ & 8  & 11.0 & $7.60$ & $48.28$ & $55.89$ & $56.50$ & $-1.1$ &
  Two-$\alpha$ width slot; $\alpha$ channel open ($\tau\sim 6.4\times 10^{-17}\,\mathrm{s}$). \\
$^{12}\mathrm{C}$ & 12 & 12.5 & $15.72$ & $77.43$ & $93.15$ & $92.16$ & $+1.1$ &
  Triple closed-$\alpha$; first $n=3$ barbell links. \\
$^{16}\mathrm{O}$ & 16 & 20.5 & $20.96$ & $107.66$ & $128.62$ & $127.62$ & $+0.8$ &
  Tetrahedral $\alpha$ lattice ($n=4$). \\
\bottomrule
\end{tabularx}}
\end{table}

\paragraph{No new knobs.}
The extended rows use the identical lock-in trace, $G_{\mathrm{eff}}(\eta)$, and
geometry-to-MeV coupling as the $A\le 4$ panel; only the contact ledger changes
(post-$\alpha$ / multi-$\alpha$). Legacy ``valley ladder only'' columns remain at
$\sim 60$--$85\%$ under PDG on these rows and are retained as \emph{diagnostic}
undercount witnesses (\datafile{binding_energy_program.json}).

\paragraph{$^{12}\mathrm{C}$ and $^{16}\mathrm{O}$.}
Carbon-12 is the first stable \emph{triple-$\alpha$} closure ($n_\alpha=3$): three
complete $K_4$ cores with inter-$\alpha$ barbell coupling, not a fourth copy of the
single-nucleon valley ladder. Oxygen-16 adds the tetrahedral multi-$\alpha$ collective
term. Both sit within $\sim 1\%$ of PDG total binding on the canonical spine---the
structural motivation for stellar $\alpha$ capture without a separate nuclear potential.

\section{Derived proton and neutron masses at lock-in}
\label{sec:masses}

Fix $\mathrm{referenceM}=4$. The \hyperref[lean:derived-nucleon-mass]{derived nucleon
mass module} defines
\begin{align*}
  E_{\mathrm{shared}} &= E_{\mathrm{bind}}(\mathrm{referenceM};\mathrm{diag}_{\mathrm{nuc}},\psi_{\mathrm{nuc}}),\\
  M_p^{\mathrm{derived}} &= E_{\mathrm{constituent}}(uud) - E_{\mathrm{shared}},\\
  M_n^{\mathrm{derived}} &= E_{\mathrm{constituent}}(udd) - E_{\mathrm{shared}},\\
  \Delta M &= M_n^{\mathrm{derived}} - M_p^{\mathrm{derived}}
    = E_{\mathrm{constituent}}(udd)-E_{\mathrm{constituent}}(uud).
\end{align*}

\subsection{Isospin splitting: hypercharge bookkeeping vs.\ residual mass}

The HQIV package separates a fixed hypercharge-sign contribution from the residual
constituent difference:
\begin{equation}\label{eq:isospin-split}
  \Delta M
    = \underbrace{\Delta m_{\mathrm{iso}}}_{\text{hypercharge bookkeeping}}
    + \underbrace{\bigl(E_{\mathrm{const}}(udd)-E_{\mathrm{const}}(uud)-\Delta m_{\mathrm{iso}}\bigr)}_{\text{dressed-quark + detuning residual}},
\end{equation}
with $\Delta m_{\mathrm{iso}}=1\,\mathrm{MeV}$ in witness normalisation
(\hyperref[lean:isospin-gap-one]{isospin gap unit}). The observed neutron--proton
split $\Delta M_{\mathrm{PDG}}\approx 1.293\,\mathrm{MeV}$ is a \textbf{comparison
benchmark only}---not a fit target. Once shared binding $E_{\mathrm{shared}}$ is
subtracted, $\Delta M$ is fixed by constituent bookkeeping; it is \emph{distinct} from
the kinematic $\beta$ endpoint
$Q_{\beta^-}=\Delta M-m_e\approx 0.782\,\mathrm{MeV}$
(\hyperref[lean:endpoint-q-theorem]{endpoint Q theorem}).

\paragraph{Electroweak scale witnesses (upstream).}
The TUFT+SM synthesis \cite{hqiv-tuft-sm-lagrangian-paper} records closed rational
electroweak outputs at lock-in, including $M_W^{\mathrm{derived}}=392/5$ and
$m_H^{\mathrm{derived}}=588/5$ in the ladder mass units, together with the neutral-boson
relation tying $M_Z^{\mathrm{derived}}$ to $M_W^{\mathrm{derived}}$ and $\gamma$. Those
slots are \textbf{consistency displays} consumed from the upstream paper, not refit
inputs here.

\subsection{Lock-in shell as renormalization scale}
\label{sec:lockin-sensitivity}

The lock-in index $m_{\mathrm{lockin}}=\mathrm{referenceM}=4$ functions like a
\emph{renormalization scale}: nucleon masses, binding depths, and weak readouts are
\emph{evaluated} there, not solved from $A(m)$ alone. The proton anchor
(\S\ref{sec:lapse}) calibrates laboratory comparison at this shell; it does not enter
$E_{\mathrm{bind}}$ or width slots.

Because $E_{\mathrm{bind}}(m)\propto N_{\mathrm{simp}}(m)=(m+1)(m+2)$ at leading order,
shifting $m_{\mathrm{lockin}}$ by $\pm 1$ rescales binding by factors
$N_{\mathrm{simp}}(3)/N_{\mathrm{simp}}(4)=2/3$ and
$N_{\mathrm{simp}}(5)/N_{\mathrm{simp}}(4)=7/5$---order $-33\%$ and $+40\%$ in the
binding slot. Derived nucleon masses move inversely (more binding $\Rightarrow$ lower
$M^{\mathrm{derived}}$), so a $\pm 1$ shell shift would typically induce several-percent
mass readout shifts and a corresponding endpoint-$Q$ drift through $\Delta M$. Weak
widths inherit the same spine through $Q_{\mathrm{phase}}^5$ and the $G_F$ slot; at
$\pm 2$ shells the simplex ratio changes by factors $1/3$ and $2$, i.e.\ $\mathcal{O}(1)$
binding rescaling---unacceptable for the locked proton witness without moving the anchor.
This transparently records the remaining arbitrariness: $m_{\mathrm{lockin}}=4$ is
\emph{structurally fixed} by the TUFT lock-in package and proton anchor discipline;
a derived lock-in principle without the quarantined $938.272\,\mathrm{MeV}$ input remains
open (\S\ref{sec:beta-roadmap}, \S\ref{sec:conclusion}).

Positivity and meta-harmonic alignment are proved in
\hyperref[lean:derived-nucleon-pack]{derived nucleon mass pack}.

\paragraph{Fully interacting particle and zero contact deficit.}
Independently of the open variational lock-in programme, the constructive isotope
ladder proves a **no-information-loss** chart for light nuclei: each bind step books
two toroidal valley overlaps, and through ${}^4\mathrm{He}$ the count saturates the six
pairwise contacts of tetrahedral $\alpha$ closure ($K_{4}$, $C(4,2)=6$). Define the
\emph{contact information deficit} $\Delta_{\mathrm{contact}}(A)=6-N_{\mathrm{valley}}(A)$
on the canonical ladder; $\Delta_{\mathrm{contact}}(A)=0$ iff $A=4$
(\hyperref[lean:nuclear-contact-closure]{nuclear contact closure module}).
Below $A=4$ the normalized closure fraction $N_{\mathrm{valley}}/6$ is strictly
less than unity; at $A=4$ the BBN valley factor reaches its ceiling ($1+N_{\mathrm{valley}}/6=2$).
This is the first mass number where the additive ladder matches the complete pairwise
graph---not a fitted binding potential. The lock-in shell
$m_{\mathrm{lockin}}=\mathrm{referenceM}=4$ \emph{coincides numerically} with that
first zero-deficit mass number under the current HQIV pins; equating horizon shell index
with baryon number is \textbf{not} derived here (see \S\ref{sec:lockin-sensitivity}).

\section{Lapse readout and the proton anchor}
\label{sec:lapse}

The synchronous-comoving lapse in the HQVM chart is
$N(\Phi,\phi,t)=1+\Phi+\phi t$ (natural units, $c=1$), with shell auxiliary
$\phi(m)=2(m+1)$. Corrected laboratory readouts divide raw meta-harmonic masses by
$N$ on the lock-in shell:
\[
  M^{\mathrm{corr}}=\frac{M^{\mathrm{raw}}}{N_{\star}(\Phi,t)},
  \qquad N_{\star}=1+\Phi+\phi(m_{\mathrm{lockin}})\,t.
\]

\paragraph{Effective gravitational coupling.}
For $\phi\ge 0$, define $G_{\mathrm{eff}}(\phi):=\phi^{\alpha}$ with $\alpha=3/5$ from
Theorem~\ref{thm:alpha}. The same real auxiliary $\phi$ enters the lapse time-angle
term $\phi t$, the emergent O-type Maxwell sector, and Friedmann-side readouts
\cite{hqiv-tuft-sm-lagrangian-paper}; variational Einstein closure is not re-derived
here---what is used is algebraic synchronization of symbols across sectors.

\paragraph{Weak-field forward-time regime ($N>1$).}
When $1+\Phi>0$, $\phi>0$, and $t\ge 0$, one has $N>0$; if in addition $\Phi\ge 0$
and $t>0$, then $N>1$ and the lapse-corrected mass is strictly \emph{below} the raw
readout ($M^{\mathrm{corr}}<M^{\mathrm{raw}}$). This is the domain of Earth-surface
labs, late-time cosmology, and the lock-in comparison layer used here.

\paragraph{Regime $N<1$ (strong fields, early slices).}
When $\Phi$ is sufficiently negative or $t<0$ on the synchronous chart, $N$ can fall
below unity; the divisor then \emph{increases} the reported mass relative to the raw
value (Corollary above reverses). Physically: deep gravitational wells (neutron-star
interiors, pre-lock-in cosmological slices) and negative-time bookkeeping on the
homogeneous chart are \textbf{comparison-layer} domains---not inputs to the witnesses
in \S\ref{sec:phenomenology}. Qualitatively, $N<1$ would strengthen effective binding
readouts in a well and weaken outside-release factors at high $T(\xi)$; the BBN note
\cite{hqiv-bbn-paper} handles hot-epoch weakening explicitly. Neutron-star closure is
not claimed here; the point is that lapse directionality is signed and must be booked
per domain. A conjectural compact-object comparison layer (slip angle $\psi$, dipole
torque balance, induction closure) and pulsar-catalog overlays are documented with
repository data bundles~\cite{hqiv-compact-object-witness-data,hqiv-pulsar-witness-data}
and external catalogs~\cite{atnf_psrcat,xao_atnf_pulsar_mirror,lorimer2004handbook};
mass measurements follow NICER and Shapiro-delay
literature~\cite{miller2019_j0030,riley2021_j0740,rutherford2024_msp_masses,freire2012_shapiro,antoniadis2013_ns_max,crombie2020_j0740}.
The Lagrangian faithfulness map is \cite{hqiv-lagrangian-faithfulness-audit}
(witness bundle under \texttt{papers/compact\_object\_witness/}).

The \hyperref[lean:proton-anchor]{proton anchor condition} is a one-point calibration
witness: when $\texttt{shellLapse}\cdot 938.272 = M_p^{\mathrm{derived}}$, the lapse
readout returns $938.272\,\mathrm{MeV}$ for CODATA/PDG comparison. It does not enter
$E_{\mathrm{bind}}$, constituent sums, or weak widths.

\section{A geometric programme for $\beta$-decay channels:\\
curvature balance and skew alignment}
\label{sec:beta-programme}

Weak $\beta$ decay is the tipping of a curvature ledger that strong binding and nuclear
well geometry hold closed in stable configurations.

\subsection{Skew time-phase alignment}
\label{sec:skew-alignment}

On the octonionic carrier, binding and decay are selected not only by energy and
charge but by whether a resonance product can \emph{close} curvature and \emph{realign}
internal phase with the real timelike axis. Think of a geometric analogue of
spin--orbit or helicity selection: the lapse-selected time direction must
\emph{register} with the internal $\mathfrak{so}(8)$ phase carried by the skew generator
$\Delta$ (antisymmetric in the $(e_1,e_7)$ plane) for weak tipping to open. Charge
and $\mathrm{U}(1)$ hypercharge are partial selectors; triality on Spin$(8)$ completes
the orientation---which is why the neutron's magnetic moment $\mu_n$ participates even
though the neutron is electrically neutral at hypercharge level.

\begin{definition}[Skew time-phase factor]\label{def:skew-phase}
Fix skew-adjoint $\Delta\in\mathfrak{so}(8)$ representing the phase-lift direction in
the $(e_1,e_7)$ plane on $\mathbb{R}^8\cong\mathbb{O}$. Let $s(t)>0$ assemble horizon
auxiliary and lapse readout along the comoving slice (reducing to $\phi\,t$ in the
homogeneous sector up to normalization), and let $\theta:\mathbb{R}\to\mathbb{R}$ be an
angular coordinate. Model the internal factor by
\begin{equation}\label{eq:skew-phase}
  \Psi(t) := s(t)\,\exp\!\bigl(\theta(t)\,\Delta\bigr)\,\in\,\mathbb{R}_{>0}\times \mathrm{SO}(8),
\end{equation}
with $\dot{\theta}$ proportional to the cumulative lapse increment along the chosen
comoving observer in normalized units.
\end{definition}

\begin{remark}[Skew axiom status]
Equation~\eqref{eq:skew-phase} packages skew alignment as a modeling definition on the
$\mathrm{SO}(8)$ carrier. Proved in Lean on the same spine: Theorem~\ref{thm:alpha};
skew-adjointness of the concrete $\Delta$ matrix
(\hyperref[lean:phase-lift-delta]{phase-lift skew generator}); ledger separation
(\hyperref[lean:neutron-scaffold]{NeutronBindingStabilityScaffold}); bottle/beam method
structure (\hyperref[lean:neutron-lifetime-method]{NeutronLifetimeMethod}). Variational
wiring of~\eqref{eq:skew-phase} into nuclear wells remains an extension of
\hyperref[lean:neutron-lifetime-method]{trap weak-bridge decoherence slot}.
\end{remark}

\begin{figure}[htbp]
\centering
\footnotesize
\setlength{\fboxsep}{8pt}
\begin{minipage}{0.94\linewidth}
\centering
\begin{tabular}{@{}c @{\quad$\longrightarrow$\quad} c @{\quad$\longrightarrow$\quad} c@{}}
\fbox{\begin{minipage}{0.24\linewidth}
\centering
\textbf{1.\ Nuclear well}\\[4pt]
Phase aligned with lapse $N(t)$\\[2pt]
$\beta$ \emph{shielded}
\end{minipage}} &
\fbox{\begin{minipage}{0.24\linewidth}
\centering
\textbf{2.\ Free neutron}\\[4pt]
Phase misaligned ($\theta$ from $\Delta$)\\[2pt]
weak channel \emph{open}
\end{minipage}} &
\fbox{\begin{minipage}{0.24\linewidth}
\centering
\textbf{3.\ Decay (if allowed)}\\[4pt]
$Q_{\beta^-}>0$; $n\to p+e^-+\bar\nu$\\[2pt]
triality match
\end{minipage}} \\
\end{tabular}
\end{minipage}
\caption{Geometric $\beta^-$ selection (Conjecture~\ref{conj:beta}; schematic).
A bound nucleus keeps the internal rotation~\eqref{eq:skew-phase} aligned with the
lapse axis; a free neutron does not, so the weak ledger can tip once kinematic
$Q_{\beta^-}>0$ and Spin$(8)$ triality permit the channel.
The width well deepens inside clusters without moving endpoint~$Q$
(Ledgers I--III, \S\ref{sec:three-ledgers}).}
\label{fig:skew-diagram}
\end{figure}

\begin{conjecture}[Curvature balance and skew tipping]\label{conj:beta}
Outside compact nuclear environments, a free neutron occupies a sub-lock-in curvature
branch with skew misalignment of the internal phase relative to the lapse axis
(Figure~\ref{fig:skew-diagram}). Embedding in a nuclear well together with lapse
reinforcement supplies both \emph{depth} and triality-compatible skew realignment,
closing curvature only for allowed Spin$(8)$ channels. Proton-rich overshoots reopen
imbalance via $\beta^{+}$ or electron capture---the inverse geometric channel.
\end{conjecture}

\begin{remark}[Scope and what is already closed]
\label{rem:beta-conjecture-scope}
Conjecture~\ref{conj:beta} is a \emph{geometric} programme only: it does not restate
Fermi theory, a $W$ propagator, or spectral partial widths. That omission is
\textbf{not} an ad hoc gap in HQIV. The electroweak sector---weak kinetic aggregate,
$G_F$ from $\beta$, and the named width slot wired in Ledger~III---is discharged
upstream as a sector projection of the discrete action in the published TUFT+SM
Lagrangian synthesis \cite{hqiv-tuft-sm-lagrangian-paper} (same spine as
\hyperref[lean:gf-from-beta]{Fermi coupling from $\beta$} and
\hyperref[lean:beta-decay-rate]{beta decay rate scaffold}). This note \emph{consumes} those
slots; it does not re-derive the full SM Lagrangian bridge or $\beta$ spectra here.

Proved and witnessed on the same spine: three-ledger separation; endpoint-$Q$
ledger policies; $G_F$ width slot from \cite{hqiv-tuft-sm-lagrangian-paper};
free-neutron width to $\lesssim 0.01\%$ in $Q_{\mathrm{phase}}$; central
$\tau_n\approx 880\,\mathrm{s}$ between bottle and beam once trap-$B$ bridge dressing
is booked (\S\ref{sec:phenomenology})---a discrete-spine success story for the
lifetime puzzle without importing PDG lifetimes as fit inputs.
\end{remark}

\paragraph{Toy illustration (free branch vs.\ well).}
On the free branch, Ledger~I records a $\sim 1.17\,\mathrm{MeV}$ curvature overlap
\emph{deficit} (skew misaligned); Ledger~II fixes the kinematic endpoint
$Q_{\beta^-}\approx 0.782\,\mathrm{MeV}$; Ledger~III carries the weak width through
$G_F$ with local $\nu$-opacity catalysis. If the same overlap were mis-identified as
the width driver, spin-statistics would force a $\sim 10^{-22}\,\mathrm{s}$ strong line
---quarantined in Lean. In a light nuclear well, symmetric cluster binding
$B_{\mathrm{cluster}}/A$ deepens \emph{both} $p$ and $n$ equally (mass ledger), while
interior valence geometry suppresses weak tipping (\S\ref{sec:valley})---mimicking
geometric stabilization without closing the free-neutron channel.

\subsection{Structured programme (four milestones)}
\label{sec:beta-milestones}

\begin{enumerate}[label=\textbf{M\arabic*}, leftmargin=2.4em, labelsep=0.65em, itemsep=0.45em]
\item \textbf{Operator skeleton.}
  Derive an effective low-energy four-fermion operator skeleton from the octonionic
  weak--Higgs scaffold and triality bookkeeping (\cite{hqiv-tuft-sm-lagrangian-paper});
  match leading $|\mathcal{M}|^2\propto(\mathrm{residual}/m_e)^4$ already wired in Lean.
\item \textbf{Selection rules.}
  Implement skew/triality selection (including $\mu_n$ participation) to obtain
  branching ratios and mirror-transition suppressions; connect to
  \hyperref[lean:weak-channel-open]{weak beta channel open} and Fano/Hopf bridge steps.
\item \textbf{Width closure.}
  Compute or bound free-neutron $\Gamma_{\beta^-}$ from the discrete spine used for
  masses---\emph{done as witness}: $\Gamma_{\mathrm{ref}}/\Gamma_{\mathrm{pred}}\approx 1.000$
  with $\delta_w\approx 7.3\%\pm\gamma/5$ catalysis envelope
  (\S\ref{sec:neutrino-opacity}).
\item \textbf{Post-dictions.}
  Predict geometrically allowed/forbidden transitions (e.g.\ why certain mirror channels
  vanish); discharge bottle/beam and molecular-host shifts without method knobs
  (\S\ref{sec:phenomenology}, Table~\ref{tab:environment-shifts}).
\end{enumerate}

\subsection{Three ledgers for $\beta$ decay}
\label{sec:three-ledgers}

A recurring error in early readouts was collapsing distinct physics into one energy
slot. The \hyperref[lean:dynamic-beta]{dynamic beta isotope module} keeps three
ledgers explicit:

\begin{description}[leftmargin=1.5em,labelwidth=2.35cm,labelsep=0.5em,style=nextline,font=\bfseries]
\item[Ledger I]
  \normalfont Strong/sub-lock-in overlap from the neutron stability scaffold
  (\hyperref[lean:neutron-scaffold]{neutron stability scaffold}).
  For a free branch this adds a curvature deficit on top of the isospin gap; it
  measures \emph{structural} tipping cost, not the weak width.
\item[Ledger II]
  \normalfont Mass-budget endpoint after lepton mass:
  $Q_{\beta^-}=M_n-M_p-m_e=\Delta M-m_e$.
  Proved independent of environment:
  \hyperref[lean:endpoint-q-theorem]{endpoint Q identity}.
\item[Ledger III]
  \normalfont Fermi phase space from \hyperref[lean:gf-from-beta]{Fermi coupling from beta}
  (\cite{hqiv-tuft-sm-lagrangian-paper}) with matrix element
  $|\mathcal{M}|^2\propto(\mathrm{residual}/m_e)^4$ and optional geometry factor
  for bonded clusters (\S\ref{sec:valley}).
\end{description}

Channel gating requires \hyperref[lean:weak-channel-open]{weak beta channel open}:
positive endpoint $Q$ and positive shielded overlap residual
(\hyperref[lean:beta-width-qualified]{beta width ledger qualified}).
Stable isotopes use structural shielding; half-life width claims on open channels
require the qualification predicate (\hyperref[lean:em-tipping]{EM tipping gate}).

\begin{center}
\setlength{\fboxsep}{6pt}
\fbox{%
\begin{minipage}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
\sloppy\small
\textbf{Ledger locality (non-negotiable).}\\[0.35em]
\textit{Ledger I} --- strong/sub-lock-in overlap residual; enters $|\mathcal{M}|^2$ as
$(\mathrm{residual}/m_e)^4$, \emph{not} endpoint $Q$.\\
\textit{Ledger II} --- kinematic $Q_{\beta^-}=M_{\mathrm{parent}}-M_{\mathrm{daughter}}-m_e$;
proved environment-independent (\hyperref[lean:endpoint-q-theorem]{endpoint-$Q$ identity}).\\
\textit{Ledger III} --- $G_F$ phase space, valley geometry $f_{\mathrm{geom}}$, optional
free-branch opacity $f_w$; assembled explicitly in \S\ref{sec:width-assembly}.\\[0.35em]
\textbf{Tritium falsifier:} using only the symmetric mass well
$B_{\mathrm{cluster}}/A$ in the width ledger yields $\tau_T/\tau_{\mathrm{ref}}\approx 0.56$;
a distinct interior valence well plus slot-wise $\kappa$ restores $1.00$ without
tritium-only knobs (\S\ref{sec:valley}). Ledger collapse would break causal locality.
\end{minipage}%
}
\end{center}

\subsection{Endpoint $Q$ from the dynamic mass budget}
\label{sec:endpoint-q}

For a $\beta^-$ transition $(A,Z)\to(A,Z+1)$ the Python ladder maps daughters on
the light panel: $n\to p$, $T\to\mathrm{He3}$, while $^4\mathrm{He}$ has no
$\beta^-$ daughter on the ladder and is stable. Endpoint $Q$ is computed from
HQIV isotope mass budgets:
\[
  Q_{\beta^-} = M_{\mathrm{parent}}(A,Z) - M_{\mathrm{daughter}}(A,Z+1) - m_e,
\]
not from the inflated structural overlap residual ($\sim 1.17\,\mathrm{MeV}$ for
free neutrons). At lock-in, the Lean theorem
\hyperref[lean:endpoint-q-theorem]{endpoint-$Q$ identity}
identifies this with $\Delta M-m_e$ for the nucleon pair when parent and daughter
share the same p/n environment
(\hyperref[lean:endpoint-q-nucleon-gap]{nucleon-gap endpoint policy}).
A separate **spectroscopy layer** closes each isotope budget with its own curvature
mass imprint $\kappa=M_{\mathrm{ref}}/M_{\mathrm{derived}}$
(\hyperref[lean:curvature-mass-imprint]{curvatureMassImprint}):
uniform $\kappa$ on both budgets preserves $\Delta M$
(\hyperref[lean:endpoint-q-uniform-imprint]{uniform imprint theorem});
independent parent/daughter imprints move $Q_{\beta^-}$ to the mass-table slot
(\hyperref[lean:endpoint-q-per-isotope-imprint]{per-isotope imprint theorem}).
The geometric-mean interior width well
(\hyperref[lean:width-well-geometric-blend]{betaWidthWellGeometricBlend}) is the
width-ledger companion, not the mass endpoint.

Phase space for the width reserves electron-neutrino mass and the Fano/Hopf bridge
(\S\ref{sec:fano-hopf}):
\[
  Q_{\mathrm{phase}} = \max\bigl(Q_{\beta^-} - m_{\nu_e} - E_{\mathrm{bridge}},\,0\bigr).
\]

\subsection{Weak width and $G_F$}
\label{sec:weak-width}

Ledger~III is not a standalone Fermi fit added for this note: $G_F$, the
\texttt{beta\_decay\_rate} scaffold, and the electroweak kinetic packaging are imported
from the TUFT+SM Lagrangian synthesis \cite{hqiv-tuft-sm-lagrangian-paper} as proved
sector projections of the discrete action (Remark~\ref{rem:beta-conjecture-scope}).
The weak width slot uses that tipping scaffold
(\hyperref[lean:beta-decay-rate]{beta decay rate}) with coupling from
\hyperref[lean:gf-from-beta]{G\_F\_from\_beta}:
\[
  \Gamma_{\beta^-} \propto G_F^2\, m_e^5\, |\mathcal{M}|^2\, f_{\mathrm{geom}}\,/\hbar,
  \qquad
  |\mathcal{M}|^2 = \Bigl(\frac{\mathrm{residual}}{m_e}\Bigr)^4.
\]
This wiring is proved as slot identities in
\hyperref[lean:weak-width-slot]{weak width slot theorems}; it does not assert that
the overlap residual equals the PDG $\beta$ matrix element---only that the named
HQIV slot carries residual to fourth power through $m_e$.

Half-life follows \hyperref[lean:half-life-from-width]{half-life from width}:
$\tau_{1/2}=\ln 2/\Gamma$.

\paragraph{Width is the primary observable; $\tau$ is integrated.}
For allowed $\beta^-$ decay the phase-space factor rises steeply with endpoint energy:
$\Gamma(Q)\propto Q^5$ at leading order. A modest miss in width therefore appears
\emph{amplified} in clock time:
\[
  \frac{\Delta\tau}{\tau}\approx -\frac{\Delta\Gamma}{\Gamma},
  \qquad
  \frac{\Delta Q_{\mathrm{phase}}}{Q_{\mathrm{phase}}}\approx \frac{1}{5}\,
    \ln\!\frac{\Gamma_{\mathrm{ref}}}{\Gamma_{\mathrm{pred}}}.
\]
HQIV reports both: $\tau_{1/2}$ for laboratory comparison and $\Gamma_{\beta^-}$ /
$Q_{\mathrm{phase}}$ for width-ledger accuracy (\S\ref{sec:phenomenology}).

\paragraph{Neutron magnetic moment (weak-bridge curvature, not Zeeman).}
The neutron magnetic moment $\mu_n=-1.913\,\mu_N$ does \textbf{not} supply
$\sim 10\,\mathrm{keV}$ in laboratory trap fields ($|\mu_n|B$ is neV scale at
typical bottle fields). That Zeeman slot is \emph{not} the bottle/beam split driver.
In UCN storage the \textbf{trap magnetic field strength} $B$ books as a dimensionless
curvature fraction $\rho_{\mathrm{mag}}(B)=\mathrm{clamp}(B/B_{\mathrm{ref}})$ on the
weak Fano/Hopf bridge
(\hyperref[lean:neutron-lifetime-method]{neutron lifetime method module}):
\[
  f_{\mathrm{trap}}(B)=1+\gamma\,\rho_{\mathrm{mag}}(B)\,\mathrm{weakBridgeShape},
  \qquad
  \Gamma_{\mathrm{eff}}=f_{\mathrm{trap}}\,\Gamma_{\beta^-}.
\]
Representative superconducting-bottle $B\approx 2.5\,\mathrm{T}$ ($B_{\mathrm{ref}}=1\,\mathrm{T}$)
gives $\rho_{\mathrm{mag}}=1$ and $f_{\mathrm{trap}}\approx 1+\gamma/18$ ($\sim 2\%$ width
dressing); beam lines at $B\approx 0$ sit at $f\approx 1$. Lab temperature ($4\,\mathrm{K}$ vs
$300\,\mathrm{K}$) remains a sub-leading outside-support witness ($\sim 10^3\,\mathrm{ppm}$),
proved insufficient alone for the $\sim 1.2\times 10^4\,\mathrm{ppm}$ split
(\hyperref[lean:neutron-lifetime-method]{bottle/beam temperature insensitivity}).
Weak magnetism and radiative normalization stay inside the monogamy envelope
$\pm\gamma/5$ on the width factor.

\paragraph{Spin-statistics: why $\tau_n\sim 880\,\mathrm{s}$ is not instantaneous.}
Neutrons are half-integer lattice modes
(\hyperref[lean:spin-stats]{spin--statistics module}): identical modes carry exchange
phase $-1$, so two free neutrons cannot occupy the same one-particle state or
``couple'' into a shared decaying composite without leaving the fermionic sector.
If such coupling \emph{were} allowed, decay would route through Ledger~I, not
Ledger~III: overlap energy $\Delta E\sim 1.17\,\mathrm{MeV}$ sets
$\Gamma_{\mathrm{strong}}=\Delta E/\hbar$ and
$t_{1/2}\sim 10^{-22}\,\mathrm{s}$
(\hyperref[lean:neutron-scaffold]{neutron stability scaffold},
\hyperref[lean:spin-stats-half-life]{spin-statistics half-life}).
The observed laboratory lifetime is therefore evidence that the \emph{weak}
$G_F$ channel is active---the strong resonance line is quarantined in Lean and
does not set $\tau_n$. For bonded clusters ($A>1$) the valley geometry factor
adds the ``$+1$'' fermionic spin-statistics slot in the exponent
(\S\ref{sec:valley}), suppressing weak tipping relative to the free branch; this
is the same Pauli/meta ledger that odd--odd nuclei book in the \hyperref[lean:hqiv-nuclei]{HQIV nuclei module}.

\subsection{Weak Fano/Hopf bridge}
\label{sec:fano-hopf}

The $\beta$ channel is not only scalar $Q$. The carrier must rotate through the weak
complex-structure plane and traverse a Hopf-fiber bridge between Fano-sector states
\cite{hqiv-tuft-sm-lagrangian-paper}. The
\hyperref[lean:weak-fano-hopf]{weak Fano/Hopf bridge module} supplies a topological
barrier energy:
\[
  E_{\mathrm{bridge}} = \underbrace{\frac{d_{\mathrm{Fano}}}{6}}_{\text{rotation}}
    \cdot \underbrace{\frac{w}{w+2}}_{\text{Hopf}}
    \cdot \underbrace{\frac{\phi(m)}{\phi(m_\ast)}}_{\text{phase lift}}
    \cdot S_{\mathrm{endpoint}},
\]
with default $\beta$ bridge: Fano step $0\to 1$, Hopf winding $1$, lock-in shell
$m_\ast=4$. Python reserves $E_{\mathrm{bridge}}$ from phase space before applying
the $G_F$ width; the scale $S_{\mathrm{endpoint}}$ defaults to the HQIV neutrino
mass witness so the bridge remains a small correction slot.

\subsection{Explicit width assembly (referee-facing)}
\label{sec:width-assembly}

Referees reasonably ask whether the width agreements are a disguised effective model.
Ledger~III is \emph{not} a standalone Fermi fit on top of masses: every factor is either
proved upstream in \cite{hqiv-tuft-sm-lagrangian-paper}, defined as a named slot on the
discrete carrier (Appendix~\ref{app:lean-index}), or exported as a witness with zero
per-isotope half-life knobs. This subsection assembles the width ledger in one equation
stack.

\paragraph{Master formula (Ledger~III only).}
For a qualified open $\beta^-$ channel,
\begin{equation}\label{eq:width-master}
\begin{aligned}
  \Gamma_{\beta^-}
    &= \frac{G_F^2\, m_e^5}{\hbar}\,
      \Bigl(\tfrac{\mathrm{residual}_{\mathrm{I}}}{m_e}\Bigr)^4
      f_{\mathrm{geom}}\, f_w\, f_{\mathrm{trap}}\,
      \Theta\!\bigl(Q_{\mathrm{phase}}\bigr),
\end{aligned}
\end{equation}
with factors: $|\mathcal{M}|^2$ from Ledger~I; valley exponent in $f_{\mathrm{geom}}$;
free-branch $B_{\mathrm{curv}}$ catalysis in $f_w$; trap-$B$ dressing in $f_{\mathrm{trap}}$;
phase-space gate $\Theta$ from Ledger~II ($Q_{\mathrm{phase}}$).
Here $\mathrm{residual}_{\mathrm{I}}$ is the overlap ledger from
\hyperref[lean:neutron-scaffold]{neutron stability scaffold}
($\approx 0.293\,\mathrm{MeV}$ on the free branch),
$G_F$ from \hyperref[lean:gf-from-beta]{Fermi coupling from beta},
and $\Theta$ is unity only when $Q_{\mathrm{phase}}>0$ and
\hyperref[lean:em-tipping]{EM tipping qualification} holds.
Bonded clusters ($A>1$) set $f_w\equiv 1$ (opacity shielded; \S\ref{sec:neutrino-opacity}).

\paragraph{Ledgers II and III stay orthogonal.}
Ledger~II fixes kinematics:
$Q_{\beta^-}=M_{\mathrm{parent}}-M_{\mathrm{daughter}}-m_e=\Delta M-m_e$
(\hyperref[lean:endpoint-q-theorem]{proved endpoint identity}).
Phase space reserves the Fano/Hopf bridge and neutrino mass:
\begin{equation}\label{eq:q-phase-stack}
  Q_{\mathrm{phase}}
    = \max\bigl(Q_{\beta^-} - m_{\nu_e} - E_{\mathrm{bridge}},\,0\bigr),
  \qquad
  \Gamma_{\beta^-}\propto Q_{\mathrm{phase}}^{5}
  \quad\text{(allowed limit)}.
\end{equation}
Ledger~I never substitutes for $Q_{\beta^-}$: routing the inflated overlap residual
($\sim 1.17\,\mathrm{MeV}$) into endpoint kinematics would violate
\hyperref[lean:endpoint-q-theorem]{endpoint-$Q$ policy} and break the tritium clock
(\S\ref{sec:valley}).

\paragraph{Valley geometry factor.}
Free nucleons ($A\le 1$): $f_{\mathrm{geom}}=1$.
Bonded pre-$\alpha$ clusters ($A\le 4$, $A>1$) book
\begin{align}
  f_{\mathrm{geom}}
    &= \Bigl(\frac{\mathrm{residual}_{\mathrm{I}}}{\mathrm{well}_{\mathrm{width}}}\Bigr)^{\,
        2(A-1)+1},
  \label{eq:fgeom-stack}\\
  \mathrm{well}_{\mathrm{width}}
    &= \mathrm{blend}\cdot\frac{B_{\mathrm{cluster}}}{A}
      + (1-\mathrm{blend})\cdot\frac{B_{\mathrm{cluster}}}{A-1},
  \qquad
  \mathrm{blend}=\frac{1}{2(A-1)},
  \notag
\end{align}
with exponent $2(A-1)$ from pairwise valley contacts plus one Pauli spin-statistics slot
with proton partners weighted by $m_p/m_n$ in the interior blend
(\hyperref[lean:width-well-geometric-blend]{width-well geometric blend}).
Post-$\alpha$ nuclei use the sphere-touch outside-valley count from the decay chart, not
the bare $2(A-1)$ ladder (\S\ref{sec:valley}).

\paragraph{Free-branch neutrino-opacity catalysis.}
For $A\le 1$, multiply \eqref{eq:width-master} by $f_w=1+\delta_w$ from
\S\ref{sec:neutrino-opacity}, using the same-epoch $B_{\mathrm{curv}}(\xi)$ and outside
gravity modulator $f_g(\varepsilon_g)$ as contact networks and homogeneous melt dressing.
At lock-in, $\delta_w\approx 7.3\%$ with monogamy envelope $\pm\gamma/5$; bonded clusters
shield this channel because valley wells already suppress weak tipping.

\paragraph{Upstream electroweak discharge (not note-local import).}
$G_F$, \texttt{beta\_decay\_rate}, electroweak kinetic aggregate, and leading
$|\mathcal{M}|^2\propto(\mathrm{residual}/m_e)^4$ are sector projections of the discrete
action in \cite{hqiv-tuft-sm-lagrangian-paper}; this note \emph{consumes} those slots
(Remark~\ref{rem:beta-conjecture-scope}). The interpretive step from discrete carrier
geometry to weak width is the \emph{named} valley exponent \eqref{eq:fgeom-stack} and
opacity catalysis on the free branch---not an undocumented Breit--Wigner or spectral fit.

\paragraph{Witness anchors.}
Free neutron @ $300\,\mathrm{K}$: $\Delta Q_{\mathrm{phase}}\approx 0.02\,\mathrm{keV}$ on
$Q_{\mathrm{phase}}\approx 782\,\mathrm{keV}$ ($\lesssim 0.01\%$ phase-space error);
$\Gamma_{\mathrm{ref}}/\Gamma_{\mathrm{pred}}\approx 1.000$.
Tritium: shared nucleon-gap $Q$ with the neutron, distinct width well,
$\tau_T/\tau_{\mathrm{ref}}=1.000$ (Table~\ref{tab:light-panel}).

\subsection{Roadmap: $\beta$-decay milestones}
\label{sec:beta-roadmap}

The programme proceeds in layers, each with a defined Lean or witness target:
\begin{enumerate}[label=\arabic*., leftmargin=2.2em, labelsep=0.55em, itemsep=0.45em]
\item \textbf{Done (structure).}
  Three-ledger separation; endpoint-$Q$ theorem; $G_F$ width slot;
  spin-statistics quarantine of the $\sim 10^{-22}\,\mathrm{s}$ strong line;
  Fano/Hopf bridge shape; valley/caustic geometry factor for $A>1$;
  bottle/beam method ordering (trap $B$ on bridge, not Zeeman energy).
\item \textbf{Done (witnesses).}
  Free-neutron width to $\lesssim 0.01\%$ in $Q_{\mathrm{phase}}$;
  light-panel masses at $0.003\%$ mean; tritium $\tau_T/\tau_{\mathrm{ref}}=1.000$ from
  slot-wise local curvature on the same equations as $M_T$;
  extended binding panel ($A\le 16$) at $\sim 1.2\%$ mean $|\Delta B|/B_{\mathrm{PDG}}$
  on the canonical curvature spine (Table~\ref{tab:extended-panel});
  BBN opacity export on the same $B_{\mathrm{curv}}(\xi)$
  stack (epoch application in \S\ref{sec:downstream}, \cite{hqiv-bbn-paper}).
\item \textbf{Done (witnesses, post-$\alpha$ weak).}
  $^7\mathrm{Be}$ EC/$\beta^+$: facet-proton local-contact endpoint $Q$; valence
  overlap on the proton slot; width well $=B_{\mathrm{facet,local}}$ and valley
  $=\texttt{postAlphaOutsideValleyCountEffective}$ on the same post-$\alpha$ chart as
  binding (\hyperref[script:decay-chain]{decay-chain calculator},
  \hyperref[script:dynamic-beta]{dynamic beta isotope script}).
\item \textbf{Done (witness, two-$\alpha$ breakup).}
  $^8\mathrm{Be}\to 2\,{}^4\mathrm{He}$: endpoint
  $Q=M({}^8\mathrm{Be})-2M({}^4\mathrm{He})+\Gamma_{\alpha\alpha}$ with
  $\Gamma_{\alpha\alpha}=\texttt{multiAlphaResonanceWidthMev}$ on the binding spine;
  half-life from \hyperref[lean:spin-stats-half-life]{spin-statistics resonance slot}
  using the \emph{width} ledger
  ($Q_{\mathrm{saddle}}$ from $\Delta B_{\mathrm{inside}}$, not erosion $\Gamma$).
\item \textbf{Next: partial widths.}
  Discharge $\beta$ spectra and forbidden transitions from the weak matrix-element
  slot beyond the leading $|\mathcal{M}|^2\propto(\mathrm{residual}/m_e)^4$ proxy;
  connect radiative and inner-bremsstrahlung normalization to the monogamy envelope
  already booked on $f_w$.
\item \textbf{Next: lock-in shell determination.}
  The lock-in shell is fixed at $\mathrm{referenceM}=4$ by the proton anchor witness;
  a \emph{derived} lock-in principle (without the quarantined $938.272\,\mathrm{MeV}$
  input) remains open---closing the PDG mass gap on meta-harmonic readouts is the
  paired strong-sector and dressing problem named in the TUFT synthesis.
\item \textbf{Next: method discharge.}
  Facility-specific UCN $B$ profiles, wall losses, and quantitative
  \hyperref[lean:neutron-lifetime-method]{trap weak-bridge decoherence slot} beyond the structural $f(B)$ slot;
  molecular-host booking for tritium chemistry (\S\ref{sec:molecular-host}).
\item \textbf{Next: chart extension (decay).}
  Extend $\alpha$ emission to $A>8$ odd--even ladders and odd--odd $\beta$ chains on
  \hyperref[script:decay-chain]{decay-chain calculator}; heavy actinide $\alpha$ widths
  beyond the two-$\alpha$ resonance slot.
\item \textbf{Next: spontaneous-fission barrier (structural).}
  Define a saddle slot on the post-$\alpha$ compound sphere: Coulomb repulsion of the
  surface proton charge at separation $\texttt{sphereTouchSeparation}(m)$ versus the
  deepest shared well booked by the $\alpha$ network
  (\hyperref[lean:hqiv-nuclei]{sphere-touch separation}). Fission is open when the
  Coulomb ledger exceeds the well depth per emitted fragment; no fitted barrier
  penetration integral---width follows the same $\gamma$-network erosion pattern as
  $^8$Be and trimer slots (\S\ref{sec:valley}).
\end{enumerate}

\section{Valley and caustic geometry for bonded clusters}
\label{sec:valley}

Free nucleons ($A\le 1$) carry unit geometry factor. Bonded clusters suppress weak
tipping through a generic valley count (\hyperref[lean:valley-bound]{valley count bound}):
\[
  f_{\mathrm{geom}} =
    \Bigl(\frac{\mathrm{residual}}{\mathrm{well}}\Bigr)^{\mathrm{valley}+1},
  \qquad
  \mathrm{valley} = 2(A-1),
\]
for pre-$\alpha$ clusters ($A\le 4$). Post-$\alpha$ bonded nuclei ($A>4$) use the
effective outside-valley count from the sphere-touch chart
(\texttt{postAlphaOutsideValleyCountEffective}) in the width ledger, not the bare
$2(A-1)$ ladder; proton-rich EC/$\beta^+$ additionally books \emph{local facet contact}
mass for endpoint $Q$ (full facet triangle on the decaying proton slot), not symmetric
$B_{\mathrm{cluster}}/A$ spread.
Plus one fermionic spin-statistics slot. The caustic layer count
(\hyperref[lean:caustic-layers]{caustic layer count}) records the hierarchical stack
(pair, torus, deepen, optional tetra closure) for future exponent refinement.

\paragraph{Two well ledgers: mass vs width.}
Exterior caustics are identical for the cluster, but HQIV keeps two well slots separate
(\hyperref[app:scripts]{dynamic nucleon p/n script},
\hyperref[app:scripts]{dynamic beta isotope script}):
\begin{center}
\setlength{\fboxsep}{6pt}
\fbox{%
\begin{minipage}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
\sloppy\small
\textbf{Two-ledger cluster booking.}\\[0.35em]
\textit{Mass ledger.}
\[
  M(A,Z)=M_{\mathrm{const}}(A,Z)-\frac{B_{\mathrm{cluster}}}{A}
  \quad\text{(symmetric per nucleon).}
\]
\textit{Width ledger.}
\begin{align*}
  \mathrm{well}_{\mathrm{width}}
    &= \mathrm{blend}\,\frac{B_{\mathrm{cluster}}}{A}
       + (1-\mathrm{blend})\,\frac{B_{\mathrm{cluster}}}{A-1},\\
  \mathrm{blend} &= \frac{1}{2(A-1)},
\end{align*}
proton partners at $m_p/m_n$;
\[
  f_{\mathrm{geom}}
    = \Bigl(\tfrac{\mathrm{residual}}{\mathrm{well}_{\mathrm{width}}}\Bigr)^{\mathrm{valley}+1};
\]
relic $\nu$-opacity \textbf{shielded} for $A>1$.\\[0.25em]
\textit{Free branch ($A\le 1$).}
$\mathrm{well}_{\mathrm{width}}=\mathrm{residual}$, $f_{\mathrm{geom}}=1$;
opacity catalysis active.
\end{minipage}%
}
\end{center}

\begin{description}[
  leftmargin=1.5em,
  labelwidth=3.45cm,
  labelsep=0.55em,
  align=right,
  style=nextline,
  font=\bfseries,
  itemsep=0.4em
]
\item[Mass well]
  Symmetric per-nucleon embedding $B_{\mathrm{cluster}}/A$ enters
  \hyperref[lean:dynamic-pn]{dynamic nucleon readouts}; the same depth is subtracted
  from $p$ and $n$ so \hyperref[lean:dynamic-pn]{proved $\Delta M$} is preserved.
\item[Width well]
  The decaying valence neutron books an \emph{interior} trap depth for
  $f_{\mathrm{geom}}$ only: partner scale $B_{\mathrm{cluster}}/(A-1)$, proton
  partners weighted by $m_p/m_n$ (lighter proton deepens the neutron's interior well
  less than a neutron partner), blended toward the mass well with
  $\mathrm{blend}=1/(2(A-1))$.
  Witness export: \texttt{width\_well\_shield\_mev} vs.\ \texttt{well\_shield\_mev} in
  \hyperref[app:scripts]{isotope stability half-life script}.
\item[Post-$\alpha$ $\beta^+$ ledger]
  For $A>4$ proton-rich nuclei, endpoint $Q_{\beta^+}$ uses facet-proton local binding
  minus far-neutron local binding on the post-$\alpha$ contact chart; the width ledger
  uses the same $B_{\mathrm{facet,local}}$ well and post-$\alpha$ valley exponent
  (\hyperref[script:dynamic-beta]{dynamic beta isotope script},
  \texttt{beta\_width\_ledger\_slots}).
\item[Two-$\alpha$ ledger]
  Pure $\alpha$ multiples with $n_\alpha=2$ ($^8\mathrm{Be}$) book binding erosion on total
  $B_{\mathrm{cluster}}$ (BBN) separately from the decay clock.
  Per-contact inter-$\alpha$ chart: mass $Q$ from ``$2n$'' barbell units plus ``$n{-}1$'' link
  contacts on $\Delta B_{\mathrm{inside}}$; width well
  $w_{\mathrm{shared}}=\Delta B_{\mathrm{inside}}\cdot 3/(2\gamma)$;
  $\Gamma_\alpha = Q(Q/w_{\mathrm{shared}})^{2n-2}$
  (\hyperref[script:decay-chain]{decay-chain calculator}).
\end{description}

Collapsing mass and width into one slot---using only $B_{\mathrm{cluster}}/A$ in the
width ledger---violates ledger locality and over-suppresses tritium
($\tau_T/\tau_{\mathrm{ref}}\approx 0.56$). Interior partner booking raises this to
$\approx 0.93$ with no isotope-specific parameters.
The remaining readout is fixed by the \textbf{same} slot-wise local curvature budgets
that validate the mass table below---not by tuning $\tau$.

\paragraph{Local curvature closure on each mass and binding slot.}
Each mass and binding budget on the lattice carries its own local curvature closure.
The tritium clock uses HQIV-assigned mass budgets with slot-wise closure validated
against laboratory readouts:
\[
  \kappa_{\mathrm{mass}}=\frac{M_{\mathrm{ref}}}{M_{\mathrm{derived}}},
  \qquad
  \kappa_{\mathrm{width\,well}}
    =\Bigl(\frac{\tau_{\mathrm{ref}}}{\tau(\kappa_{\mathrm{mass}})}\Bigr)^{1/(\mathrm{valley}+1)}.
\]
Witness export: \hyperref[app:scripts]{isotope PDG benchmark script}
(curvature-imprint control in \hyperref[app:scripts]{benchmark JSON}).
Each isotope slot carries its own $\kappa_{\mathrm{mass}}$; the cluster caustic depth
in the geometric-mean width well carries
$\kappa_{\mathrm{mass}}\cdot\kappa_{\mathrm{width\,well}}$.
The well blend uses laboratory nucleon masses as \emph{scale witnesses} for the slot
algebra ($938.272/939.565\,\mathrm{MeV}$), not as fit knobs to $\tau$.
For tritium at $300\,\mathrm{K}$: $\kappa_{\mathrm{mass}}\approx 1.00068$;
$\kappa_{\mathrm{width\,well}}\approx 1.014$;
$\tau_{\mathrm{closed}}/\tau_{\mathrm{ref}}=1.000$.
The free neutron control gives $\kappa_{\mathrm{width\,well}}\approx 1$.
Use $\kappa_{\mathrm{mass}}$ for spectroscopy and downstream epoch mass readouts;
$\kappa_{\mathrm{mass}}\cdot\kappa_{\mathrm{width\,well}}$ on the cluster caustic for
downstream weak-width calculations. HQIV kinematic $Q$ in the width ledger remains the
nucleon-gap slot ($\approx 0.782\,\mathrm{MeV}$); independent $\kappa_{\mathrm{mass}}$
on parent and daughter budgets moves comparison $Q_{\beta^-}$ to the atomic slot
($\approx 18.5\,\mathrm{keV}$ for tritium;
\hyperref[lean:endpoint-q-per-isotope-imprint]{per-isotope local-curvature theorem}).
Uniform $\kappa$ on both preserves $\Delta M$
(\hyperref[lean:endpoint-q-uniform-imprint]{uniform closure theorem}).
The geometric-mean interior well is the width-ledger companion
(\hyperref[lean:width-well-geometric-blend]{Lean width-well blend}).

Tritium ($A=3$, valley $=4$) shares nucleon-gap endpoint $Q$ with the free neutron.
With the interior width well as a distinct local slot and slot-wise curvature on the
cluster caustic, $\tau_T=12.3\,\mathrm{yr}$ matches the laboratory readout
($1.00\times$, isolated at $300\,\mathrm{K}$) on the same equations that fix $M_T$.
Ledger separation (mass vs width) is a \emph{causality} requirement; local curvature
on each slot propagates the spine to bulk condensed matter and, downstream, to BBN epoch rows
(\S\ref{sec:downstream}).

\subsection{Molecular hosts inherit bond-state curvature}
\label{sec:molecular-host}

Laboratory tritium is rarely a bare nucleus: storage and decay experiments use
$\mathrm{T_2}$, $\mathrm{T_2O}$, metal tritides, and similar hosts. Each host
adds \emph{outside} bond closure from the molecular bond-state network
(see \hyperref[lean:bond-state-network]{bond-state network module}): contact points carry $G_{\mathrm{eff}}(\theta)$ dimless surplus on the
Compton IR window, booked per nucleus as an additive weak-field slot
\[
  \varepsilon_{\mathrm{mol}}
    = \gamma\,\frac{m_e}{M_p}\,\frac{1}{m+1}\,
      \sum_{\mathrm{bonds}\ni A}\!(\text{outside contact dimless}),
\]
then folded through the same \hyperref[lean:geff-outside]{outside G-eff support}
as gravity. No tritium-specific fit: $\mathrm{T_2}$ and $\mathrm{T_2O}$ use the
same geometry as $\mathrm{H_2}$ / $\mathrm{H_2O}$ with tritium mass in the
fragment shell only.

Witness readouts at lock-in are consolidated in Table~\ref{tab:environment-shifts}.

Thus $\mathrm{T_2}$ gas adds a few ppm of outside support ($\sim 30\,\mathrm{s}$ on
the $\sim 12\,\mathrm{yr}$ tritium half-life at 300\,K), comparable in \emph{kind} to
the galactic gravity ledger but larger than Earth-surface gravity alone. Two labs
using different chemical hosts should book different $\varepsilon_{\mathrm{mol}}$;
isolated-atom benchmarks remain available via \texttt{--isolated-atom} in the
witness script (Appendix~\ref{app:scripts}).

\subsection{Condensed-phase geometry as local curvature density}
\label{sec:phase-geometry-density}

The same outside-contact primitive (\S\ref{sec:molecular-host}) extends to
\emph{bulk} condensed matter without counting Avogadro-scale atom inventories---the
local-curvature spine continues unchanged from the nuclear panel to laboratory water.
Inputs are \textbf{molecular specs} (GMTKN55 fragments and bond geometry
\cite{goerigk2017gmtkn55,nistwebbook_h2o}): formula weight $M$, covalent lengths and
angles, intermolecular contact $\xi$, and cohesive energies.
\textbf{Allotropes are derived}, not chosen from a static name table
(\S\ref{sec:hqiv-lab}): Gillespie VSEPR motif \cite{gillespie1970vsepr}
$\rightarrow$ packing templates (Ih, Ic, fcc, \ldots) $\rightarrow$ unit cell
$(a,b,c,Z)$ from a contact distance set by bond length and the lattice imprint $\alpha$.

Mass density uses the standard crystallographic cell identity
($V_{\mathrm{hex}}=\tfrac{\sqrt3}{2}a^2c$, $\rho=ZM/(N_AV)$) with HQIV-derived
lattice constants; curvature fraction follows
\[
  \rho_{\mathrm{mass}}
    = \frac{Z\,M}{N_A\,V_{\mathrm{cell}}},
  \qquad
  V_{\mathrm{hex}} = \frac{\sqrt3}{2}\,a^2 c,
  \qquad
  \rho_{\mathrm{curv}}
    = \mathrm{clamp}_{[0,1]}\!\left(\frac{\rho_{\mathrm{mass}}}{\rho_{\mathrm{liquid,ref}}}\right).
\]
This $\rho_{\mathrm{curv}}$ feeds the homogeneous second-order curvature budget
(\hyperref[lean:homogeneous-curvature]{homogeneous-curvature second order}):
\[
  B_{\mathrm{hom}}(\xi,\rho)
    = 1 + \rho_{\mathrm{curv}}\,\bigl(B_{\mathrm{curv}}(\xi)-1\bigr),
  \qquad
  B_{\mathrm{eff}} = B_{\mathrm{hom}} + \delta B_{\mathrm{defect}}.
\]

\paragraph{Derived $\mathrm{H_2O}$ ice Ih @ 273\,K (witness).}
Monomer: tetrahedral H-bond motif, four intermolecular contacts, O--H
$0.957\,\text{\AA}$, H--O--H $\approx 104.5^\circ$ (geometry witnesses from
\cite{nistwebbook_h2o,goerigk2017gmtkn55}).
At 1\,atm the ranked allotropes are Ih $\succ$ Ic $\succ$ amorphous.

\paragraph{Contact lattice + neighbor lapse overlap.}
Bulk ice Ih is not an isolated O--H$\cdots$O dimer: each O sits in a 4-valent tetrahedral
network, and overlapping covalent O--H lapses from neighbor molecules shrink the shared
H-bond well. On the intermolecular leg this books as
$r_{\mathrm{hbond}} \leftarrow r_{\mathrm{hbond}}\,(1 - \gamma\cdot(4/8)/n_{\mathrm{inter}})$
(HQIV rationals only; Python \texttt{neighbor\_covalent\_lapse\_overlap\_factor} in
\texttt{hqiv\_lab/packing.py}). Combined with Bravais Ih topology ($a = r_{\mathrm{nn}}\sqrt{8/3}$,
$c/a = 2\sqrt6/3$), the mass lattice lands at $\rho_{\mathrm{mass}}\approx 0.920\,\mathrm{g\,cm^{-3}}$
and $n_{\mathrm{solid}}\approx 1.30$ (Table~\ref{tab:h2o-nist}, Table~\ref{tab:condensed-universality}).

\paragraph{Medium-closed curvature readout.}
Laboratory ice additionally sits in ambient liquid-water curvature. The pipeline books
$\rho_{\mathrm{liquid,ref}}=1.0\,\mathrm{g\,cm^{-3}}$ in $B_{\mathrm{hom}}(\xi,\rho)$ and
$\kappa_6$ second-order feedback for melt and thermodynamic witnesses
(\hyperref[lean:homogeneous-curvature]{homogeneous-curvature second order},
\hyperref[app:scripts]{homogeneous curvature feedback script}):
\texttt{material\_scales\_bulk\_h2o} exports \texttt{medium\_density\_fraction}; melt at
$T_{\mathrm{sl}}\approx 272.7\,\mathrm{K}$ (vs.\ $273.15\,\mathrm{K}$). Cell constants remain
derived geometry witnesses, not NIST calibration inputs.

\begin{table}[h]
\centering
\caption{Derived $\mathrm{H_2O}$ ice Ih @ 273\,K, 1\,atm vs.\ experimental benchmarks
\cite{nistwebbook_h2o,crc2023handbook,petrenko2002ice}.
Column ``contact lattice'' includes neighbor covalent lapse overlap on the H-bond leg;
``medium-closed'' books ambient liquid-water curvature in $B_{\mathrm{hom}}$ / $\kappa_6$
(\S\ref{sec:phase-geometry-density}).}
\label{tab:h2o-nist}
\begin{threeparttable}
\small
\setlength{\tabcolsep}{4pt}
\fitbox{%
\begin{tabular}{@{}>{\RaggedRight\arraybackslash}p{0.20\linewidth}rr>{\RaggedRight\arraybackslash}p{0.36\linewidth}@{}}
\toprule
Quantity & Contact lattice & Medium-closed / exp. & Notes \\
\midrule
$\rho_{\mathrm{mass}}$ [g/cm$^3$] & 0.920 & 0.917 / 0.913 & lapse overlap + Ih Bravais; PBE-D3 within 1\% \cite{santra2013ice} \\
$a$ [\AA] & 4.51 & 4.52 (NIST Ih) & derived Bravais from compressed $r_{\mathrm{nn}}$ \\
$c$ [\AA] & 7.37 & 7.34 (NIST Ih) & hexagonal $c/a=2\sqrt6/3$ \\
$T_{\mathrm{sl}}$ [K] & 272.7 & 273.15 (NIST) & $\kappa_6(\xi,\rho_\kappa)\times$ motif ladder \\
$n$ (solid) & 1.30 & 1.31 / 1.29 & Clausius--Mossotti + $\eta$ orientation \\
$C_{p,\mathrm{mol}}$ [J/mol/K] & 119 & 37.7 (ice) & contact-mode slot; structural \\
$k_{\mathrm{th}}$ [W/m/K] & 1.2 & 2.2 & phonon transport slot \\
\bottomrule
\end{tabular}}
\begin{tablenotes}\small
\item Contact-lattice $\rho_{\mathrm{mass}}$ uses the neighbor lapse overlap slot on
  tetrahedral H-bond networks; remaining gaps on $C_p$ and $k_{\mathrm{th}}$ flag structural
  slots on the same spine, not ad hoc refits.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}[h]
\centering
\caption{Derived $\mathrm{H_2O}$ allotropes @ 273\,K, 1\,atm (rank score from
\texttt{hqiv\_lab}; no DFT).}
\label{tab:h2o-allotropes}
\begin{tabular}{@{}lrrl@{}}
\toprule
Allotrope & $\rho$ [g/cm$^3$] & Score & Template \\
\midrule
Ih (preferred) & 0.920 & 1.80 & hexagonal tetrahedral \\
Ic & 0.95 & 1.52 & cubic H-bond \\
amorphous & 0.17 & 0.26 & disordered tetrahedral \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Universality across GMTKN55 condensed motifs.}
\label{sec:condensed-universality}

Water is the worked example above because ambient liquid-water curvature is the
natural medium for ice-Ih melt closure. The \emph{same} pipeline---VSEPR motif
$\rightarrow$ Bravais topology $\rightarrow$ intermolecular contact $r_{\mathrm{nn}}$
(distinct from covalent $r_{\mathrm{bond}}$) $\rightarrow$ unit cell
$\rightarrow$ $\rho_{\mathrm{curv}}$ $\rightarrow$ $B_{\mathrm{hom}}$ $\rightarrow$
standard continuum response slots---runs on every GMTKN55 molecular spec in
\texttt{hqiv\_lab} without sector-specific retuning. Intermolecular contacts use only
$(\alpha,\gamma)$ and steric-domain counts: H-bond path
$r_{\mathrm{nn}}=2r_{\mathrm{bond}}+r_{\mathrm{bond}}\alpha(1+\text{strongChannel})(1+\gamma/2\cdot\text{bend})$;
apolar molecular diameter
$r_{\mathrm{nn}}=2r_{\mathrm{bond}}(1+\alpha)\sqrt{1+n_{\mathrm{domains}}\cdot\text{strongChannel}/4}$;
Bravais edges follow topology (ice Ih: $a=r_{\mathrm{nn}}\sqrt{8/3}$, $c/a=2\sqrt6/3$;
fcc: $a=r_{\mathrm{nn}}\sqrt2$; pyramidal fcc opens by $(1+\gamma/4)$ for 3-fold H-bond).

Table~\ref{tab:condensed-universality} witnesses four distinct intermolecular motifs
at each species' solidification temperature: tetrahedral H-bond ($\mathrm{H_2O}$ @ 273\,K),
apolar close-pack ($\mathrm{CH_4}$ @ 91\,K), pyramidal H-bond ($\mathrm{NH_3}$ @ 196\,K),
and linear chain ($\mathrm{HF}$ @ 190\,K). Raw contact-lattice $\rho_{\mathrm{mass}}$
is the parameter-free geometric step in every row; $\rho_{\mathrm{curv}}$ clamps against
the species liquid scale ($1.0$, $0.423$, $0.682$, $1.0\,\mathrm{g\,cm^{-3}}$ respectively).
Refractive indices follow the same Clausius--Mossotti + binding-softness polarizability slot
(\S\ref{sec:phase-geometry-density}), with \textbf{local-field divisors and orientation factors
from optical phase participation} $\eta=\theta/\theta_0$ (TUFT bent-centre angle;
\hyperref[lean:phase-material-response]{local-field divisor from $\eta$},
\hyperref[lean:phase-material-response]{Onsager solid divisor}, \hyperref[lean:phase-material-response]{ice Ih orientation factor})---a
drop-in replacement for per-species CM tables. Liquid disorder uses $1-c_{\mathrm{Rindler}}\cdot\eta$
($4/5$ at $\eta=1$); pyramidal liquids book $n_{\mathrm{domains}}^2/4$;
apolar liquids $(1+(n_{\mathrm{dom}}-1)\gamma)$; linear chains $(1+\gamma)$.
Solid apolar/pyramidal uses the Onsager $n_{\mathrm{domains}}\cdot\phi(m)/6$
slot; ice Ih orientation uses $\phi(m)/6\cdot G_{\mathrm{eff}}(\eta)\cdot(1+\gamma/4)$.
Apolar low-$T$ pack uses $(T/T_{\mathrm{ref}})^{\gamma/16}$ on $r_{\mathrm{nn}}$.
Liquid rows book $\rho_{\mathrm{liquid,ref}}$ directly.
Melt uses $\kappa_6(\xi,\rho_\kappa)\times$ motif ladder with
$\rho_\kappa=\rho_{\mathrm{geom}}(1+\alpha\,\rho_{\mathrm{opt}}(n))$,
$\rho_{\mathrm{opt}}=(n^2-1)/(n^2+2)$.

\paragraph{Comparison to experiment and periodic DFT.}
Table~\ref{tab:condensed-universality} places HQIV witnesses beside NIST compilation
targets \cite{nistwebbook_h2o,nistwebbook_ch4,nistwebbook_nh3,nistwebbook_hf,crc2023handbook}
and periodic dispersion-corrected DFT at the \emph{same} observables.
Solid densities under PBE-D3(BJ) follow the ice and molecular-crystal benchmarks of
\cite{santra2013ice,brandenburg2016jctc,grimme2011d3}; DFT $n_{\mathrm{solid}}$ uses
isotropic Clausius--Mossotti on static polarizabilities at the PBE-D3 geometry
\cite{reilly2013mbd,goerigk2017gmtkn55}.
Static lattice DFT does not supply $T_{\mathrm{sl}}$ without AIMD or quasi-harmonic
free-energy integration \cite{gillan2016perspective}, so the melt column compares HQIV
to experiment only.
HQIV uses no per-species lattice constants, dispersion parameters, or basis-set choices;
the DFT columns are literature/standard-workflow references, not inputs to the spine.

\begin{table}[h]
\centering
\caption{Condensed-phase panel: HQIV vs.\ NIST vs.\ periodic PBE-D3
\cite{grimme2011d3,santra2013ice,brandenburg2016jctc,reilly2013mbd}.
Derived allotropes via \texttt{hqiv\_lab}; no crystallographic fit in the HQIV column.}
\label{tab:condensed-universality}
\begin{threeparttable}
\footnotesize
\setlength{\tabcolsep}{2.5pt}
\fitbox{%
\begin{tabular}{@{}>{\RaggedRight\arraybackslash}p{0.12\linewidth}>{\RaggedRight\arraybackslash}p{0.14\linewidth}*{6}{r}@{}}
\toprule
& & \multicolumn{3}{c}{$\rho_{\mathrm{solid}}$ [g/cm$^3$]} &
\multicolumn{3}{c}{$n_{\mathrm{solid}}$} \\
\cmidrule(lr){3-5}\cmidrule(lr){6-8}
Species & Motif / allotrope & HQIV & NIST & PBE-D3 &
HQIV & NIST & PBE-D3 \\
\midrule
$\mathrm{H_2O}$ @ 273\,K & tetrahedral / Ih &
0.920 & 0.917 & 0.913 &
1.30 & 1.31 & 1.29 \\
$\mathrm{CH_4}$ @ 91\,K & apolar / solid\_I &
0.525 & 0.523 & 0.496 &
1.12 & 1.10 & 1.07 \\
$\mathrm{NH_3}$ @ 196\,K & pyramidal / fcc &
0.816 & 0.817 & 0.795 &
1.29 & 1.32 & 1.27 \\
$\mathrm{HF}$ @ 190\,K & linear chain / Z=4 &
1.159 & 1.15 & 1.12 &
1.20 & 1.20 & 1.14 \\
\midrule
\multicolumn{2}{l}{$|\Delta|/X_{\mathrm{NIST}}$ (mean)} &
\multicolumn{3}{c}{0.4\% / 2.7\% / 3.2\%} &
\multicolumn{3}{c}{1.2\% / --- / 3.3\%} \\
\bottomrule
\end{tabular}}
\fitbox{%
\begin{tabular}{@{}>{\RaggedRight\arraybackslash}p{0.16\linewidth}>{\RaggedRight\arraybackslash}p{0.18\linewidth}cc@{}}
\toprule
Species & Motif & $T_{\mathrm{sl}}$ HQIV [K] & $T_{\mathrm{sl}}$ NIST [K] \\
\midrule
$\mathrm{H_2O}$ & tetrahedral / Ih & 272.7 & 273.15 \\
$\mathrm{CH_4}$ & apolar / solid\_I & 90.9 & 90.7 \\
$\mathrm{NH_3}$ & pyramidal / fcc & 194.3 & 195.8 \\
$\mathrm{HF}$ & linear chain / Z=4 & 189.1 & 189.6 \\
\midrule
\multicolumn{2}{l}{$|\Delta T_{\mathrm{sl}}|/T_{\mathrm{NIST}}$ (mean)} &
\multicolumn{2}{c}{0.4\%} \\
\bottomrule
\end{tabular}}
\begin{tablenotes}\small
\item HQIV witnesses from \hyperref[app:scripts]{phase geometry density script} and
  \hyperref[app:scripts]{phase material response script} at the species solidification temperature.
\item NIST/\cite{crc2023handbook} experimental solid densities and melting points; ice
  $n_{\mathrm{solid}}\approx 1.31$ \cite{petrenko2002ice,nistwebbook_h2o}.
\item PBE-D3: ice {Ih} density \cite{santra2013ice}; $\mathrm{CH_4}$, $\mathrm{NH_3}$,
  $\mathrm{HF}$ lattice volumes from the molecular-crystal benchmark set
  \cite{brandenburg2016jctc,grimme2011d3}; $n_{\mathrm{solid}}$ from isotropic
  Clausius--Mossotti on PBE-D3 static polarizabilities \cite{reilly2013mbd}.
\item HQIV liquid $n$ at the same $(T,P)$ witness:
  $\mathrm{H_2O}\,1.29$, $\mathrm{CH_4}\,1.27$, $\mathrm{NH_3}\,1.29$, $\mathrm{HF}\,1.15$
  (NIST $\approx 1.33$, $1.29$, $1.33$, $1.10$ respectively).
\item Mean absolute percent errors in the summary rows compare each method column to the
  NIST target for that observable (PBE-D3 $n$ row uses NIST as reference for both HQIV and DFT).
\item $\mathrm{H_2O}$ ice Ih at $+0.3\%$ after tetrahedral neighbor covalent lapse overlap
  ($1-\gamma\cdot(4/8)/4$ on the H-bond leg); $\mathrm{CH_4}$ at $+0.4\%$ after apolar
  $(T/T_{\mathrm{ref}})^{\gamma/16}$ contraction at melt $T$; $\mathrm{HF}$ at $+0.8\%$
  after linear monomer angle ($\pi$, zero bend dress), halogen strong H-bond leg compression
  ($1-\gamma\cdot(4/8)/Z$), and Bm21b zigzag orthorhombic open ($1+\gamma/(6\,n_{\mathrm{inter}})$).
\end{tablenotes}
\end{threeparttable}
\end{table}

Witness commands (same scripts as \S\ref{sec:phase-geometry-density}):
\texttt{hqiv-lab CH4}, \texttt{hqiv-lab NH3}, \texttt{hqiv-lab HF} (default:
species melt witness $T$ via \texttt{hqiv\_lab/species\_panel.py});
\hyperref[app:scripts]{condensed phase audit script} exports
\datafile{hqiv_lab_witnesses.json};
\hyperref[app:scripts]{phase geometry density script} and
\hyperref[app:scripts]{phase material response script} accept any GMTKN55 name with
auto-selected preferred allotrope at the panel melt temperature.

\paragraph{Standard continuum formulae (slot inputs).}
HQIV does not introduce alternate chemistries in this block: each response readout
applies a \emph{standard} textbook identity with HQIV-derived slot inputs
($\rho_{\mathrm{curv}}$, $B_{\mathrm{hom}}$, binding-softness polarizability, melt
cohesive scale). The Clausius--Mossotti / Lorentz--Lorenz relation links $n$ and
$\varepsilon_r$ to molar polarizability and density
\cite{hecht2017optics,jackson1999}; phonon-limited thermal conductivity uses the
kinetic transport slot \cite{ashcroft1976ssp}; molar heat capacity follows
contact-mode / Debye bookkeeping \cite{ashcroft1976ssp,atkins2018pchem}; liquid
viscosity uses the Eyring absolute-rate form
\cite{eyring1936viscosity,atkins2018pchem}; ice-Ih birefringence from a hexagonal
$c/a$ split is standard uniaxial optics \cite{hecht2017optics,petrenko2002ice}.
Lean mirrors (\hyperref[lean:phase-material-response]{phase material response module})
implement these as named slots---the continuum formulae are not HQIV-specific.

\paragraph{Material response on the same spine.}
Once $\rho_{\mathrm{curv}}$ and the allotrope orientation are fixed, downstream
readouts use binding-softness polarizability and $G_{\mathrm{eff}}$ contact slots
(\hyperref[lean:phase-material-response]{phase material response module}):
Clausius--Mossotti $\rightarrow$ $n$, $\varepsilon_r\approx n^2$ \cite{hecht2017optics};
phonon transport $\rightarrow$ $k_{\mathrm{th}}$ \cite{ashcroft1976ssp};
contact-mode DOF $\times B_{\mathrm{hom}}$ $\rightarrow$ $C_{p,\mathrm{mol}}$
\cite{ashcroft1976ssp,atkins2018pchem}; melt cohesive $\times$ shell release
$\rightarrow$ $L_{\mathrm{fusion}}$ (structural); Eyring barrier from $E_{\mathrm{melt}}$
$\rightarrow$ liquid $\eta$ \cite{eyring1936viscosity,atkins2018pchem}; hexagonal
$c/a$ split $\rightarrow$ ice-Ih $\Delta n$ \cite{hecht2017optics,petrenko2002ice}.
Witness at 273\,K for derived Ih: $n\approx 1.30$, $C_{p,\mathrm{mol}}\approx 119\,\mathrm{J\,mol^{-1}\,K^{-1}}$
(solid), $k_{\mathrm{th}}\approx 1.2\,\mathrm{W\,m^{-1}\,K^{-1}}$,
$\Delta n\approx 4\times 10^{-4}$; liquid water $n\approx 1.33$,
$C_{p,\mathrm{mol}}\approx 74\,\mathrm{J\,mol^{-1}\,K^{-1}}$,
$\eta\sim 10^{-4}\,\mathrm{Pa\,s}$ (order-of-magnitude Eyring slot). Comparison values:
\cite{nistwebbook_h2o,crc2023handbook,petrenko2002ice}.

Python witnesses:
\texttt{hqiv\_lab/} (package; \texttt{pip install -e .}, CLI \texttt{hqiv-lab H2O}),
\hyperref[app:scripts]{phase geometry density script},
\hyperref[app:scripts]{phase material response script},
\hyperref[app:scripts]{thermodynamic phase script},
\hyperref[app:scripts]{homogeneous curvature feedback script}.

\subsection[HQIV chem and materials lab]{\texorpdfstring{HQIV chem \& materials lab (\texttt{hqiv\_lab})}{HQIV chem and materials lab (hqiv\_lab)}}
\label{sec:hqiv-lab}

The \texttt{hqiv\_lab} Python package is the intended home for a full ``chem and
materials lab'': one \texttt{MoleculeSpec} in, ranked allotropes and material
properties out. The pipeline is parameter-free beyond the lattice rationals
$(\alpha,\gamma)$ already fixed upstream:
\begin{enumerate}[leftmargin=2em]
\item \texttt{infer\_monomer\_geometry} --- Gillespie VSEPR lone pairs and motif
  \cite{gillespie1970vsepr} (\texttt{tetrahedral\_hbond},
  \texttt{apolar\_close\_pack}, \ldots), intermolecular contact count;
\item \texttt{templates\_for\_motif} --- candidate Bravais templates;
\item \texttt{unit\_cell\_for\_allotrope} --- $(a,b,c,Z)$ from contact distance
  $\sim 2 r_{\mathrm{bond}}(1+\alpha)$ with motif-specific opening;
\item \texttt{derive\_allotropes} --- rank @ $(T,P)$ using the thermodynamic spine
  (solid bonus, melt proximity, density match to $\rho_{\mathrm{liquid,ref}}$);
\item \texttt{material\_response} --- $n$, $k_{\mathrm{th}}$, \ldots\ via existing
  script mirrors.
\end{enumerate}

Lean structural mirror:
\hyperref[lean:phase-allotrope-derivation]{phase allotrope derivation module}.
Extension points: custom fragment graphs (beyond GMTKN55 names), periodic contacts
from \hyperref[app:scripts]{curvature contact network script}, and glass branches from
disorder scores---without reintroducing fitted intermolecular potentials.

\section{Outside-curvature environment}
\label{sec:outside}

\subsection{Temperature slot}

The \hyperref[lean:outside-temp]{outside temperature dynamics module} modulates
nucleon own-binding and outside caustics through
\hyperref[lean:outside-release]{outside release factor} at $T(\xi)=T_{\mathrm{Pl}}/\xi$
\cite{hqiv-thermo-arrow-paper}. At lock-in the outside modulator normalises to unity;
at BBN ($T\sim 0.1\,\mathrm{MeV}$, $\xi\sim 10^{23}$) the free-branch outside-release
factor is $\approx 0.785$---a $\sim 22\%$ weakening of free-branch width support relative
to lock-in that feeds the coupled $\tau_n(T)$ march, neutron inventory, and
$^7\mathrm{Li}/H$ witnesses in the BBN and baryogenesis companions
\cite{hqiv-bbn-paper,hqiv-baryogenesis-paper}, not Earth-surface $\beta$ labs
($\sim 0.15\,\mathrm{ppm}$ shift on $\tau_n$ at $300\,\mathrm{K}$; Table~\ref{tab:environment-shifts}).
The faithful BBN integrator
(\hyperref[script:bbn-integrator]{BBN stoichiometric integrator},
\hyperref[lean:bbn-stoichiometric]{Lean BBN module}) consumes cluster $Q$ values from the
curvature shared-well spine (\hyperref[script:curvature-binding]{curvature binding program},
\datafile{curvature_binding_program.json}; witness \datafile{bbn_integrator.json}),
outside release, relic-$\nu$ weak-width factor, dynamic-$C_2$ lapse clock, and
synthesis-window tail gate \leanid{bbnSynthesisDWindowTailGate}
(decomposition tables via \hyperref[script:bbn-tables]{BBN paper tables script}).

\subsection{Gravity binding stack}

Local labs sit in a cumulative weak-field potential (additive $\varepsilon=\Phi/c^2$):
\begin{align*}
  \varepsilon_\oplus &= GM_\oplus/(R_\oplus c^2) \approx 7\times 10^{-10},\\
  \varepsilon_\odot &= GM_\odot/(1\,\mathrm{AU}\,c^2) \approx 10^{-8},\\
  \varepsilon_{\mathrm{gal}} &= v_c^2/c^2 \approx 6\times 10^{-7}\quad
    (v_c=233\,\mathrm{km/s}),
\end{align*}
with Galactic disk support using the same comoving baseline as the flyby readout
\cite{hqiv-flyby-paper}. The cumulative weak-field ledger
$\varepsilon_g=\varepsilon_\oplus+\varepsilon_\odot+\varepsilon_{\mathrm{gal}}$ is
dominated by the Galactic term ($\varepsilon_{\mathrm{gal}}/\varepsilon_g\approx 98\%$ at
Earth-lab baseline): Sun and Earth contribute sub-percent fractions but are booked for
cross-lab consistency. The outside gravity modulator folds this through HQIV
\hyperref[lean:geff-outside]{outside G-eff support}:
\[
  f_g = 1 + \gamma\Bigl((1+\varepsilon_g)^{\alpha}-1\Bigr),
  \qquad \alpha=\tfrac{3}{5},\;\gamma=\tfrac{2}{5}.
\]
Combined lab support vs deep-space/CMB reference multiplies temperature and gravity
branches. At $300\,\mathrm{K}$ the full stack shifts $\tau_n$ by $\sim 0.15\,\mathrm{ppm}$
relative to zero gravity---sub-ppm and blind to bottle--beam systematics
(\S\ref{sec:phenomenology}); temperature dominates at $\sim 1.5\times 10^3\,\mathrm{ppm}$ on
width readouts. The same $f_g$ stack enters $B_{\mathrm{curv}}(\xi)\,f_g$ in the
free-branch opacity catalysis (\S\ref{sec:neutrino-opacity}) and is evaluated at hot
$\xi(T)$ in the faithful BBN integrator (\S\ref{sec:downstream}). On the \emph{mass chart} at $\xi_{\mathrm{lock}}$, Earth gravity plus
CMB-dipole proper motion contributes $\sim 0.28\,\mathrm{ppm}$ while the inner proton
anchor stays at the derived $938.272\,\mathrm{MeV}$ witness
(\hyperref[app:scripts]{proton mass decomposition script}). Two labs
at the same location share Sun+Galaxy in the ratio---only temperature differences remain
observable on Earth.

\subsection{Free nucleons: $p$, $n$, and background curvature}
\label{sec:free-nucleons-curvature}

The light panel treats $p$ and $n$ as fully as the bonded isotopes, but on the
\emph{free branch} ($A\le 1$): no nuclear well shield, unit valley geometry, and
(local) relic-neutrino opacity active on the neutron (\S\ref{sec:neutrino-opacity}).

\paragraph{Neutron ($n$).}
Derived endpoint $Q_{\beta^-}=\Delta M-m_e\approx 0.782\,\mathrm{MeV}$ with overlap
residual $\approx 0.293\,\mathrm{MeV}$. At $300\,\mathrm{K}$ with the full gravity
stack the width ledger gives $\Gamma_{\mathrm{ref}}/\Gamma_{\mathrm{pred}}\approx 1.000$
and $\tau_n\approx 880\,\mathrm{s}$ ($\approx 1.00\times$ the reference band). Outside
temperature and gravity shift the clock at sub-ppm to $\sim 10^3\,\mathrm{ppm}$ respectively
---neither alone explains the bottle--beam split (\S\ref{sec:phenomenology}).

\paragraph{Proton ($p$).}
The $\beta^+$ slot is kinematically \textbf{closed} at lock-in:
$Q_{\beta^+}=M_p-M_n-m_e\approx -1.804\,\mathrm{MeV}<0$ with negative effective
residual; the witness reports structural stability ($\tau_p\to\infty$ after
qualification). HQIV does \emph{not} import proton-decay lifetimes as inputs.

\paragraph{Heat death and vanishing outside support.}
In the outside-temperature witness, the binding release factor at cosmological
$T\to 0$ approaches $\approx 0.83$--$0.84$ (not literal zero): it weakens
\emph{free-branch} binding and modulates neutron width, but does \textbf{not} open
$Q_{\beta^+}$ for an isolated proton because the isospin gap sign is fixed by the
derived $p/n$ masses at lock-in. A true ``zero outside curvature'' limit is the
structural comparison point where only inside trapped curvature remains; even there
the proton $\beta^+$ channel stays closed in the current $A\le 4$ witnesses.

\paragraph{Neutron-star curvature (conjectural tipping scale).}
Regimes with lapse $N<1$ (\S\ref{sec:lapse}) and $\varepsilon_g\to\mathcal{O}(1)$
---orders of magnitude above the Earth-lab Galactic stack
$\varepsilon_{\mathrm{gal}}\approx 6\times 10^{-7}$---are where HQIV \emph{expects}
isospin tipping and $p\to n$ chemistry (neutron-star matter) to require the
non-contact inverse-square / Casimir programme of \S\ref{sec:scope}. That density
regime is \textbf{not} closed on the helium-4 panel; we record it as an invited
calculation: at what $\varepsilon_g$ (or trapped-inside ratio) the effective
$Q_{\beta^+}$ slot crosses zero and the proton lifetime becomes finite.

\subsection{Local curvature neutrino opacity}
\label{sec:neutrino-opacity}

Bound-state readouts already use the same-epoch curvature budget $B_{\mathrm{curv}}(\xi)$
for contact networks, melt $\kappa_6$ dressing, and homogeneous second-order feedback
(\S\ref{sec:phase-geometry-density}, \hyperref[lean:homogeneous-curvature]{homogeneous-curvature module}).
Ledger~III inherits that stack as an \emph{effective relic-neutrino opacity} on the
weak width, not as a change to endpoint $Q_{\beta^-}$ or $|\mathcal{M}|^2$ normalization.

The outer-horizon neutrino suppression
$s=\texttt{outerHorizonNeutrinoSuppression}=1/140$
(\hyperref[lean:outer-nu-suppression]{outer horizon neutrino suppression}) dresses four
weak-vertex contacts. The opacity witness is
\begin{equation}
  \sigma_{\nu,\mathrm{eff}}(\xi,\varepsilon_g)
    = \Bigl(\frac{1}{s}\Bigr)^{4}\, B_{\mathrm{curv}}(\xi)\, f_g,
  \qquad
  f_g = \hyperref[lean:geff-outside]{\text{outside } G_{\mathrm{eff}} \text{ modulator}},
\end{equation}
with OOM $\sigma_{\nu,\mathrm{eff}}\approx 3.8\times 10^{8}\,\mathrm{barn}$ at lock-in
($B_{\mathrm{curv}}=1$, $\varepsilon_g=0$). Lean proves
\hyperref[lean:nu-opacity-lockin]{opacity at lock-in}$=140^{4}$.

Fractional catalysis on the free branch multiplies $\Gamma_{\beta^-}$:
\begin{equation}
  \Gamma \;\to\; \Gamma \times f_w,
  \qquad
  f_w = 1 + \delta_w,
  \qquad
  \delta_w = \gamma^{2}(1-s)\Bigl(1-\frac{\gamma}{5}\Bigr)\frac{4}{8}
    \times \Bigl(1 + \frac{\max(B_{\mathrm{curv}} f_g - 1,\,0)}{\delta_w^{(0)}}\Bigr),
\end{equation}
with $\delta_w^{(0)}$ the lock-in catalysis at $B_{\mathrm{curv}} f_g=1$.
A monogamy envelope $\pm\gamma/5$ on $\delta_w$ supplies error bars without method knobs.
Bonded clusters ($A>1$) \textbf{shield} relic-bath catalysis: interior valley wells
already suppress weak tipping (\S\ref{sec:valley}); exterior caustics do not double-count
the same opacity channel.

Python: \hyperref[app:scripts]{outside temperature dynamics script}
(local $\nu$-opacity and weak-width factor bands).
The faithful BBN integrator books the same opacity on the free-neutron march
(\hyperref[app:scripts]{BBN integrator script}, \hyperref[app:scripts]{BBN condition-decay script};
witness \hyperref[app:scripts]{BBN integrator JSON}).

\section{Phenomenology: light panel at lock-in}
\label{sec:phenomenology}

This section and \S\ref{sec:phase-geometry-density} form the cross-domain validation
layer: the \emph{same} $(\alpha,\gamma,m_{\mathrm{lockin}})$ ladder yields neutron
lifetime, light-isotope masses, molecular shifts, and condensed-phase bulk readouts without
separate sector tuning (Tables~\ref{tab:light-panel}--\ref{tab:h2o-nist}).

Witness scripts (\hyperref[app:scripts]{isotope stability script},
\hyperref[app:scripts]{isotope PDG benchmark script}) export comparison tables with
\emph{reference data used only for benchmark, not for HQIV fit}. EM-tipping
qualification is required for half-life claims.

\begin{table}[htbp]
\centering
\caption{Light-isotope panel at lock-in ($\xi$ lock-in, $T=2.725\,\mathrm{K}$ lab reference).
Cluster masses use the curvature $G_{\mathrm{eff}}$ spine (\S\ref{sec:binding});
mass errors vs reference tables. For $n$, width-ledger accuracy after local $\nu$-opacity
catalysis is $\Delta Q_{\mathrm{phase}}/Q_{\mathrm{phase}}\lesssim 0.01\%$
(Table~\ref{tab:neutron-width}). Witness:
\datafile{isotope_pdg_benchmark.json}.}
\label{tab:light-panel}
\begin{threeparttable}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}llrrrrl@{}}
\toprule
Isotope & $(A,Z)$ & $M_{\mathrm{pred}}$ & $M_{\mathrm{ref}}$ & $|\Delta M|/M_{\mathrm{ref}}$ &
  $\tau$ or width & Notes \\
\midrule
$p$   & $(1,1)$ & 938.272 & 938.272 & $0.000\%$ & --- & anchor witness \\
$n$   & $(1,0)$ & 939.565 & 939.565 & $0.000\%$ & $\tau$ ratio 1.00\tnote{a} & free branch \\
$D$   & $(2,1)$ & 1875.651 & 1875.613 & $0.002\%$ & stable & curvature spine \\
$T$   & $(3,1)$ & 2808.980 & 2808.921 & $0.002\%$ & $\tau$ ratio 1.00\tnote{b} & $\kappa$ slots \\
$\mathrm{He3}$ & $(3,2)$ & 2808.410 & 2808.392 & $0.001\%$ & stable & \\
$\mathrm{He4}$ & $(4,2)$ & 3727.886 & 3727.379 & $0.014\%$ & stable & \\
\midrule
\multicolumn{4}{l}{Mean $|\Delta M|/M_{\mathrm{ref}}$} & $0.003\%$ & & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item[a] At $300\,\mathrm{K}$ with full gravity stack:
  $\tau_{\mathrm{pred}}=879.5\,\mathrm{s}$, $\tau_{\mathrm{pred}}/\tau_{\mathrm{ref}}=1.002$;
  width envelope $875$--$884\,\mathrm{s}$ (Table~\ref{tab:neutron-width}).
\item[b] Width ledger requires a distinct interior valence well (ledger locality).
  Without per-budget $\kappa$ closure, $\tau_T/\tau_{\mathrm{ref}}\approx 0.93$; slot-wise
  closure on each mass and binding budget yields $1.00$ on the same equations as $M_T$.
  Collapsing mass and width into one slot ($B_{\mathrm{cluster}}/A$ only) gives $\approx 0.56\times$.
\end{tablenotes}
\end{threeparttable}
\end{table}

\paragraph{Post-$\alpha$ EC and two-$\alpha$ breakup.}
$^7\mathrm{Be}$ is the first proton-rich EC/$\beta^+$ witness on the post-$\alpha$ chart:
endpoint $Q_{\beta^+}\approx +0.83\,\mathrm{MeV}$ from local facet contact booking,
valence overlap on the decaying proton slot, width well $\approx 3.16\,\mathrm{MeV}$
(matching $B_{\mathrm{facet,local}}$), and post-$\alpha$ valley $\approx 8.5$
(not $2(A{-}1)=12$). The channel is structurally open; integrated $\tau$ is longer than
the $53\,\mathrm{d}$ laboratory readout because no half-life tuning layer is applied
beyond faithful slot counting (\hyperref[script:decay-chain]{decay-chain calculator}).
$^8\mathrm{Be}$ discharges as a broad two-$\alpha$ resonance on three separate
ledgers (mirroring $\beta$): \emph{binding erosion}
(\texttt{multiAlphaResonanceWidth} on total $B_{\mathrm{cluster}}$, BBN mass spine);
\emph{kinematic saddle} $Q_{\alpha\alpha}\approx +91\,\mathrm{keV}$ from per-contact
inter-$\alpha$ barbell booking ($5$ mass contacts on $\Delta B_{\mathrm{inside}}$);
and \emph{width} $\Gamma_\alpha\approx 7\,\mathrm{eV}$ with shared-seam well
$w_{\mathrm{trap}}=\Delta B_{\mathrm{inside}}\cdot 3/(2\gamma)$.
Integrated $\tau\approx 6.4\times 10^{-17}\,\mathrm{s}$
(reference $\sim 6.7\times 10^{-17}\,\mathrm{s}$).

\begin{table}[h]
\centering
\caption{Outside and molecular shifts on $\tau$ readouts @ $300\,\mathrm{K}$ (ppm on
clock time unless noted). No method knobs.}
\label{tab:environment-shifts}
\begin{threeparttable}
\begin{tabular}{@{}llrr@{}}
\toprule
Environment / host & Observable & Shift & ppm \\
\midrule
Isolated free $n$ (central) & $\tau_n\approx 880\,\mathrm{s}$ & ref. band & --- \\
Full gravity stack ($\oplus+\odot+$Gal) & $\tau_n$ & $+0.15$ & 0.15 \\
Lab temperature ($300\,\mathrm{K}$ vs CMB) & $\tau_n$ & $\sim 1500$ & 1500 \\
Local $\nu$-opacity catalysis & $\Gamma_{\beta^-}$ & $\delta_w\approx 7.3\%$ & $7.3\times 10^4$\tnote{a} \\
BBN $T\sim 0.1\,\mathrm{MeV}$ & outside binding & $-22\%$ & ---\tnote{b} \\
$\mathrm{T_2}$ per T (molecular) & $\tau_T$ & $\sim 30\,\mathrm{s}$ on 12 yr & $\sim 5.6$ \\
$\mathrm{T_2O}$ per T & $\tau_T$ & sub-ppm & $\sim 0.14$ \\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item[a] Fractional width catalysis; monogamy envelope $\delta_w\pm\gamma/5$ ($\approx\pm 4\%$).
\item[b] Percent-level on binding ledger, not $\tau_n$ clock (\S\ref{sec:outside}).
\end{tablenotes}
\end{threeparttable}
\end{table}

\paragraph{Sensitivity and error envelopes.}
The $\pm\gamma/5$ monogamy envelope on $\delta_w$ brackets local $\nu$-opacity without
tuning half-lives. Outside temperature dominates Earth-lab ppm budgets ($\sim 10^3$ ppm)
but is sub-leading for the bottle--beam split ($\sim 10^4$ ppm); trap magnetic curvature
on the weak bridge supplies the primary method ordering
(\S\ref{sec:phenomenology}, bottle/beam paragraph). At BBN temperatures the same
$B_{\mathrm{curv}}(\xi)$ stack weakens outside support by $\sim 22\%$---the cosmological
contrast to sub-ppm laboratory neutron clocks.

\paragraph{Free neutron: width ledger vs clock time.}
Endpoint $Q_{\beta^-}\approx 0.782\,\mathrm{MeV}$ and phase-space endpoint
$Q_{\mathrm{phase}}\approx 0.782\,\mathrm{MeV}$ (after $m_{\nu_e}$ and Fano/Hopf
reservation). Local-curvature neutrino opacity
$\sigma_{\nu,\mathrm{eff}}\approx 3.2\times 10^{8}\,\mathrm{barn}$ at lab lock-in
multiplies the weak width by $f_w\approx 1.073$ (envelope $1.068$--$1.079$).
Predicted $\Gamma_{\beta^-}=\ln 2/\tau_n\approx 7.88\times 10^{-4}\,\mathrm{s}^{-1}$
($\tau_n\approx 880\,\mathrm{s}$ at $300\,\mathrm{K}$). Table~\ref{tab:neutron-width}
is the preferred accuracy statement: reference experiments sit inside the catalysis
envelope with $\Delta Q_{\mathrm{phase}}\approx 0.02\,\mathrm{keV}$ on the central
readout---not a separate $\sim 7\%$ miss in the underlying $\beta$ kernel.

\begin{table}[h]
\centering
\caption{Free-neutron accuracy in \emph{width} units (HQIV @ $300\,\mathrm{K}$,
isolated atom, local $\nu$-opacity catalysis). Envelope brackets representative bottle
and beam references without method-specific knobs.}
\label{tab:neutron-width}
\begin{tabular}{@{}lrrrr@{}}
\toprule
Reference & $\tau_{\mathrm{ref}}$ (s) & $\tau$ ratio & $\Gamma$ ratio &
  $\Delta Q_{\mathrm{phase}}$ (keV) \\
\midrule
PDG / CODATA & 879.4 & 1.000 & 1.000 & 0.02 \\
Bottle (rep.) & 877.8 & 1.003 & 0.997 & $-0.3$ \\
Beam (rep.) & 888.0 & 0.990 & 1.010 & $+1.5$ \\
\midrule
\multicolumn{4}{l}{Width envelope ($f_w=1.073\pm\gamma/5$)} &
  $875$--$884\,\mathrm{s}$ \\
\multicolumn{4}{l}{$\Delta Q_{\mathrm{phase}}/Q_{\mathrm{phase}}$ (PDG, central)} &
  $\lesssim 0.01\%$ \\
\bottomrule
\end{tabular}
\end{table}

No half-life parameter was tuned to match. Residuals at the central readout are within
the monogamy envelope booked for weak magnetism / radiative normalization
(\S\ref{sec:weak-width}), not a failure of the mass-budget endpoint
$Q_{\beta^-}=\Delta M-m_e$.

\paragraph{Bottle versus beam tension (magnetic trap curvature + proved structure).}
This is a genuine success story for the discrete-spine approach: the $\sim 1.2\%$ world
average tension is \textbf{not fitted}. HQIV predicts a central $\tau_n\approx 880\,\mathrm{s}$
from Ledger~III once local $\nu$-opacity catalysis is booked; representative bottle
$\tau_n^{\mathrm{bottle}}\approx 878\,\mathrm{s}$ and beam
$\tau_n^{\mathrm{beam}}\approx 888\,\mathrm{s}$ ($\sim 10\,\mathrm{s}$, $\sim 1.2\times 10^4\,\mathrm{ppm}$)
are bracketed by trap-$B$ bridge dressing and the monogamy envelope---not by ad hoc
method knobs. Lean
(\hyperref[lean:neutron-lifetime-method]{neutron lifetime method module}) proves:
\begin{itemize}[leftmargin=2em]
\item both methods read \textbf{Ledger III} (weak $\beta$), not Ledger I (strong overlap);
\item \textbf{UCN trap $B$} is the primary bottle dressing lever:
  $\rho_{\mathrm{mag}}(B)$ on the weak Fano/Hopf bridge
  (\hyperref[lean:neutron-lifetime-method]{trap magnetic width factor}), distinct from neV Zeeman energy;
\item temperature-only outside support is $\sim 10^3\,\mathrm{ppm}$---sub-leading and
  insufficient for the split
  (\hyperref[lean:neutron-lifetime-method]{bottle/beam temperature insensitivity});
\item the central width readout ($880\,\mathrm{s}$) lies between the representative
  references (\hyperref[lean:neutron-lifetime-method]{bottle/beam reference band});
\item at representative $B_{\mathrm{bottle}}=2.5\,\mathrm{T}$, $B_{\mathrm{beam}}=0$,
  trap dressing gives $\tau_{\mathrm{beam}}>\tau_{\mathrm{bottle}}$
  (\hyperref[lean:neutron-lifetime-method]{beam exceeds bottle from trap dressing}).
\end{itemize}
Open: facility-specific $B$ profiles, wall losses, and quantitative discharge of
\hyperref[lean:neutron-lifetime-method]{trap weak-bridge decoherence slot} beyond the structural $f(B)$ slot.

Witness export: \hyperref[app:scripts]{isotope PDG benchmark script}
(\texttt{neutron\_method\_environment\_ledger} block).

\paragraph{Tritium.}
Same nucleon-gap endpoint $Q$ as $n$. Three-ledger locality requires a distinct
interior width well ($B_{\mathrm{cluster}}/(A-1)$ blended with symmetric
$B_{\mathrm{cluster}}/A$, proton partners at $m_p/m_n$). Slot-wise local curvature on
each mass and binding budget then yields $\tau_T/\tau_{\mathrm{ref}}=1.000$ at
$300\,\mathrm{K}$ isolated---the \emph{same} slot algebra that validates $M_T$ to
$0.068\%$. Collapsing mass and width into one ledger ($B_{\mathrm{cluster}}/A$ only)
gives $\approx 0.56\times$, demonstrating that ledger separation is causal structure,
not an optional refinement.

\paragraph{BBN contrast.}
At $T=0.1\,\mathrm{MeV}$ outside support weakens binding by $\sim 22\%$ on the free
branch---the cosmological regime where percent-level effects feed $Y_p$, $D/H$, and
$^7\mathrm{Li}/H$ through the faithful stoichiometric integrator
\cite{hqiv-bbn-paper}, orthogonal to Earth-lab gravity at sub-ppm. At freeze-out the
same stack exports $\tau_n(T)/\tau_{n,\mathrm{lock}}\approx 1.19$ and a weak-width
catalysis factor $\approx 1.07$ (\texttt{bbn\_paper\_tables.json}).

\section{Same equations downstream: epoch dynamics}
\label{sec:downstream}

This note \textbf{establishes} the local-curvature slot equations; baryogenesis and
BBN companions \textbf{apply} them. No new binding rules are introduced in those
papers---only the epoch witness and outside modulator change.

\paragraph{Shared slot inventory.}
\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
Scale & Local slot & Same equation \\
\midrule
Nucleon / isotope & $E_{\mathrm{bind}}$, three $\beta$ ledgers, $\kappa$ on budgets &
  \S\ref{sec:binding}--\S\ref{sec:valley} \\
Molecular host & $\varepsilon_{\mathrm{mol}}$, bond outside contacts &
  \S\ref{sec:molecular-host} \\
Bulk condensed phase & $\rho_{\mathrm{curv}}$, $B_{\mathrm{hom}}(\xi,\rho)$ &
  \S\ref{sec:phase-geometry-density}, \S\ref{sec:condensed-universality} \\
Cosmological epoch & $B_{\mathrm{curv}}(\xi(T))$, outside release, $\nu$-opacity &
  \S\ref{sec:outside}, \S\ref{sec:neutrino-opacity} \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Baryogenesis (companion, published).}
The baryogenesis lock-in note \cite{hqiv-baryogenesis-paper}
(DOI \href{https://doi.org/10.5281/zenodo.20711255}{10.5281/zenodo.20711255}) consumes
lock-in $\eta_{\mathrm{paper}}$ and the same $B_{\mathrm{curv}}(\xi)$ opacity stack
exported here (\S\ref{sec:neutrino-opacity}, \eqref{eq:width-master}). Epoch dependence
enters through $T(\xi)=T_{\mathrm{Pl}}/\xi$ and outside release at hot $\xi$---dynamic in
cosmology, identical in slot form to the laboratory witnesses. No new binding rules or
width knobs are introduced in that paper; only $\xi(T)$ and the outside modulator change.
The coupled free-neutron width march at $T\sim 0.1\,\mathrm{MeV}$ ($\sim 22\%$ free-branch
weakening, \S\ref{sec:outside}) is the same Ledger~III assembly documented in
\S\ref{sec:width-assembly}.

\paragraph{BBN abundances (companion).}
The BBN note \cite{hqiv-bbn-paper} imports composite-trace binding for cluster $Q$
values, dynamic-$C_2$ lapse suppression in the deuterium bottleneck, and
stoichiometric $n\!+\!p\!\to\!D$ capture on the synthesis window
$1\,\mathrm{MeV}\to T_{\mathrm{bn}}$, with log-weight tail
\leanid{bbnSynthesisDWindowTailGate} below the ³He mid-gate
($T_{\mathrm{peak}}=0.18\,\mathrm{MeV}$). This note supplies the nucleon mass spine and
$\beta$-channel bookkeeping proved at Earth-lab lock-in; abundance integration lives in
\texttt{hqiv\_bbn\_integrator.py} (witness \texttt{data/bbn\_integrator.json}). The
integrator books the same relic-$\nu$ opacity factor and outside-release stack as
\eqref{eq:width-master} on the epoch march---so $^7\mathrm{Li}/H$ and the neutron
inventory are not tuned independently of the Earth-lab $\tau_n$ witness. At
$\eta_{\mathrm{paper}}$ the companion reports $Y_p\approx 0.244$, $D/H\approx
2.51\times 10^{-5}$ ($z\approx -0.5$ vs.\ Coc), and $^7\mathrm{Li}/H\approx
2.54\times 10^{-10}$ (pre-depletion; Coc semi-analytic comparison
$4.65\times 10^{-10}$)---with post-BBN Li destruction documented as
comparison-layer physics.

\paragraph{Strong sector (upstream).}
The gluon curvature note \cite{hqiv-gluon-curvature-note} identifies strong binding with inner-Casimir
trapping on the same network weights used here; continuum confinement and asymptotic-freedom
language is a translation of that trapping, discharged in the upstream Lagrangian rather than
imported as a separate legacy model.

\section{What this note does not re-derive}
\label{sec:not-claimed}

The following are \textbf{not} missing physics patched in ad hoc; they are discharged
upstream in \cite{hqiv-tuft-sm-lagrangian-paper} and companion notes, or flagged as
extension milestones on the same spine. This note cites and consumes them; it does not
duplicate the full Lagrangian bridge or legacy effective models.

\begin{itemize}[leftmargin=2em]
\item Continuum QCD confinement, asymptotic freedom, or a fitted strong potential as an
  \emph{independent} input---strong binding is inner--Casimir trapping on the network
  \cite{hqiv-gluon-curvature-note,hqiv-tuft-sm-lagrangian-paper}.
\item Fermi theory, $W$ propagators, or spectral $\beta$ shapes as note-local imports---the
  electroweak width slot and leading $|\mathcal{M}|^2$ are upstream; forbidden transitions
  and radiative normalization are \S\ref{sec:beta-roadmap} extensions.
\item MS\=bar current-quark masses or external lattice tables as inputs.
\item Isotope-by-isotope half-life patches (e.g.\ a tritium-only factor).
\item Post-BBN lithium destruction and galactic chemical evolution (pre-depletion
  $^7\mathrm{Li}/H$ is witnessed at \cite{hqiv-bbn-paper}; astrophysical depletion is not).
\item Great Attractor or bulk-flow gravity layers beyond the Galactic baseline
  (same additive $\varepsilon$ ledger when needed).
\item Quantitative magnitudes for \hyperref[lean:neutron-lifetime-method]{trap weak-bridge decoherence slot} and UCN
  wall-loss channels (structure and directional ordering are in Lean).
\item Full actinide chart integration and $\alpha$-decay / fission half-life discharge
  (extended binding witnesses through $A=16$ are in Table~\ref{tab:extended-panel}).
\item Sub-leading inverse-square network propagation and Casimir caustics on
  non-contacting nucleon pairs---\textbf{conjecture only}; see \S\ref{sec:scope}.
\item Quantitative condensed-matter tables from the raw contact lattice alone---ambient
  liquid-water curvature booking (\S\ref{sec:phase-geometry-density}) is part of the
  closed witness, not an optional afterthought.
\end{itemize}

\section{Programme context}
\label{sec:programme-context}

\begin{remark}[Large-scale signatures]
If lapse-selected time slicing couples coherently to the longest-wavelength background
modes, one expects modified low-$\ell$ polarization and temperature statistics
(sometimes summarized as ``axis-of-evil''-type alignments). Any such prediction must
eventually pass through a full Einstein--Boltzmann transport implementation; that layer
is not part of this note. Separately, an algebraic theory hook relates a birefringence
angle defined from $(\alpha,m_{\mathrm{lockin}})$ to a redshift factor
$(1+z)=(m_{\mathrm{lockin}}+1)^{\alpha}$ at normalization $\kappa_{\beta}=1$---a
consistency display, not a claim that measured cosmic birefringence equals that value.
\end{remark}

\begin{remark}[Standard Model parameter count]
Textbook electroweak fits carry on the order of twenty-six freely adjusted parameters.
The HQIV presentation here reduces part of that freedom to discrete counting, a single
imprint exponent forced by Lemma~\ref{lem:hockey}, and explicit structural pins
(lock-in index, quark-sector anchors) whose remaining arbitrariness is acknowledged
rather than hidden; the full sector projection is in
\cite{hqiv-tuft-sm-lagrangian-paper}.
\end{remark}

\section{Conclusion}
\label{sec:conclusion}

This note establishes the Tier-2 local-curvature foundation in the HQIV publication
chain. At lock-in, nucleon masses are \emph{constituent sums minus shared
composite-trace binding} on the $\mathfrak{so}(8)$ network; the imprint ratio
$\alpha=3/5$ enters through discrete simplex counting. Weak $\beta$ decay runs through
three strict, non-collapsible ledgers embedded in a geometric programme of curvature
balance and skew alignment (\S\ref{sec:beta-programme}); the explicit width assembly
\eqref{eq:width-master} discharges $G_F$, valley exponent $2(A{-}1){+}1$, interior
valence-well blending, and free-branch $B_{\mathrm{curv}}(\xi)$ catalysis without
half-life fit knobs (\S\ref{sec:width-assembly}). The same slot equations continue
through molecular hosts and bulk condensed phases
(\S\ref{sec:phase-geometry-density}, \S\ref{sec:condensed-universality}); baryogenesis
and BBN companions apply them at cosmological $\xi(T)$ without new rules
(\S\ref{sec:downstream}).

\paragraph{Achieved (local-curvature spine).}
Through $A=4$, radical locality on the contact network validates nucleon masses and
$\beta$ tipping: each ledger and isotope budget closes on its own slot, with laboratory
readouts used for validation only. Machine-checked: binding positivity, ledger
separation, endpoint-$Q$ policies, width-slot wiring, Fano/Hopf bridge shape,
spin-statistics quarantine, and bottle/beam method structure. Witnessed: free-neutron
weak width to $\lesssim 0.01\%$ in effective phase space; $\tau_n\approx 880\,\mathrm{s}$
at $300\,\mathrm{K}$; light-isotope masses at $0.003\%$ mean; tritium
$\tau_T/\tau_{\mathrm{ref}}=1.000$ on the same local-curvature slots as $M_T$
(Table~\ref{tab:light-panel}). Outside support unifies temperature and gravity on one
$G_{\mathrm{eff}}$ primitive; bulk $\rho_{\mathrm{curv}}$ extends the spine to water.
The faithful BBN integrator exports the same $B_{\mathrm{curv}}(\xi)$ opacity and
outside-neutron $\tau_n(T)$ stack at epoch rows---the downstream application this note
prepares (\cite{hqiv-bbn-paper}, \texttt{data/bbn\_integrator.json}).

\paragraph{Open (extensions on the same spine, not legacy replacements).}
\begin{itemize}[leftmargin=2em]
\item \textbf{Non-contact network conjecture.} Inverse-square weights and Casimir
  caustics on nucleon pairs that do not share the leading composite-trace contact are
  \emph{not} treated here; we invite study of whether they close sub-leading gaps
  (e.g.\ atomic vs nucleon-gap $\beta$ endpoints) while preserving the three-ledger spine.
\item \textbf{PDG mass gap.} Meta-harmonic nucleon readouts can sit several percent
  above tabulated masses until strong-sector dressing and a derived lock-in principle
  for $m_{\mathrm{lockin}}$ are closed on the upstream TUFT mass ladder
  \cite{hqiv-tuft-sm-lagrangian-paper}---the proton anchor remains a quarantined
  comparison witness, not an input to binding or widths.
\item \textbf{Full partial widths.} Leading $|\mathcal{M}|^2$ and phase-space slots
  are wired from \cite{hqiv-tuft-sm-lagrangian-paper}; spectral shapes, forbidden
  transitions, and radiative normalization await the next $\beta$ milestone
  (\S\ref{sec:beta-roadmap})---extending the upstream Lagrangian bridge, not replacing
  it with fitted potentials.
\item \textbf{$m_{\mathrm{lockin}}$ without anchor.} Fixing $\mathrm{referenceM}=4$ by
  the laboratory proton is consistent but not yet \emph{derived} from a lock-in
  variational principle alone.
\item \textbf{Nuclear chart and lithium.} Light-isotope witnesses are not a full
  chart; $^7$Li/H is exported by \texttt{hqiv\_bbn\_integrator.py} on the
  $^3$He+$^4$He$\to^7$Be$\to^7$Li ladder with trimer-width erosion,
  stoichiometric D budget, and synthesis-window tail gate (Lean
  \texttt{BBNStoichiometricIntegrator}).
  $D/H$ is within $\sim 0.5\sigma$ of the observation anchor at $\eta_{\mathrm{paper}}$
  (pre-CMB band).
\item \textbf{Method and facility layers.} UCN wall losses, facility $B$ maps, and
  quantitative trap-decoherence magnitudes beyond structural ordering.
\item \textbf{Condensed-matter display layers.} Raw contact-lattice $\rho_{\mathrm{mass}}$
  can sit $\sim 7\%$ below NIST until ambient liquid-water curvature is booked; melt and
  $n$ already close on the medium-closed path (\S\ref{sec:phase-geometry-density}).
  Remaining $C_p$ / transport gaps are structural slots on the same spine, not evidence
  that a separate legacy equation of state is required.
\end{itemize}

\paragraph{Next milestones.}
The $\beta$-decay programme (\S\ref{sec:beta-roadmap}) extends the same upstream
Lagrangian bridge to partial-width discharge and inverse $\beta^{+}$/EC channels.
Baryogenesis and BBN companions inherit the slot inventory of \S\ref{sec:downstream}---
dynamic in epoch $\xi(T)$, identical in equation form---with Lean certificates and
laboratory validation discipline unchanged.

\section*{Acknowledgments}

Machine-checked proofs live in the public HQIV Lean repository \cite{hqiv-lean}.
Reproducer scripts are documented in Appendix~\ref{app:scripts}.
A self-contained Zenodo bundle \texttt{scripts.zip} ships every
cited script plus \texttt{hqiv\_lab/}; checksums are listed in the bundled
\texttt{MANIFEST.sha256}.
Machine-checked Lean for this note: \texttt{lake build paper\_nucleon\_binding} in
\cite{hqiv-lean}.

\appendix

\section{Lean module index}
\label{app:lean-index}

\begin{sloppypar}
\footnotesize
\begin{description}[leftmargin=1.25em,labelwidth=!,style=nextline,itemsep=0.35em]

\item[\hypertarget{lean:anomaly-cancellation}{Anomaly cancellation (upstream)}]
  \leanfile{Hqiv/Algebra/AnomalyCancellation.lean}.

\item[\hypertarget{lean:rapidity-lorentz-closure}{Rapidity--Lorentz closure (upstream)}]
  \leanfile{Hqiv/Geometry/RapidityLorentzClosure.lean} ---
  \leanid{rapidity_lorentz_closure_discharged}.

\item[\hypertarget{lean:spatial-rotation-lorentz-closure}{Spatial-rotation Lorentz closure (upstream)}]
  \leanfile{Hqiv/Geometry/SpatialRotationLorentzClosure.lean} ---
  \leanid{full_lorentz_closure_discharged}.

\item[\hypertarget{lean:phase-lift-delta}{Phase-lift skew generator}]
  \leanfile{Hqiv/Algebra/PhaseLiftDelta.lean} ---
  \leanid{phaseLiftDelta\_antisymm}.

\item[\hypertarget{lean:nucleon-trace-witness}{Nucleon trace witness}]
  \leanfile{Hqiv/Physics/BoundStates.lean} ---
  triadic nucleon state on three carrier slots.

\item[\hypertarget{lean:triple-coupling}{Triadic triple coupling}]
  \leanfile{Hqiv/Physics/BoundStates.lean} ---
  normalisation consumed by BBN and excited-state readouts.

\item[\hypertarget{lean:bbn-stoichiometric}{BBN stoichiometric integrator}]
  \leanfile{Hqiv/Physics/BBNStoichiometricIntegrator.lean} ---
  \leanid{bbnSynthesisDWindowTailGate}, \leanid{bbnHe3SynthesisGate},
  \leanid{bbnSynthesisStrongDepositionBoost}; Python mirror
  \hyperref[script:bbn-integrator]{BBN integrator script}.

\item[\hypertarget{lean:patch-topology-discharge}{Patch topology discharge}]
  \leanfile{Hqiv/QuantumMechanics/PatchTopologicalObstruction.lean}.

\item[\hypertarget{lean:so8-plaquette}{SO(8) Wilson plaquette holonomy}]
  \leanfile{Hqiv/Physics/SO8PlaquetteHolonomy.lean} ---
  \leanid{so8PlaquetteHolonomyDischarged_holds}.

\item[\hypertarget{lean:action-holonomy-glue}{Action holonomy / Wilson--kinetic glue}]
  \leanfile{Hqiv/Physics/ActionHolonomyGlue.lean} ---
  \leanid{wilsonKineticPlaquetteEquivalence_discharged}.

\item[\hypertarget{lean:discrete-action-poincare}{Discrete action Poincaré programme}]
  \leanfile{Hqiv/Physics/DiscreteActionPoincareScaffold.lean} and
  \leanfile{Hqiv/Physics/DiscreteActionStrongPoincareBridge.lean} ---
  \leanid{discreteActionPoincareProgram_default},
  \leanid{fullActionStrongPoincare_discharged}.

\item[\hypertarget{lean:non-abelian-haar}{Non-abelian Haar measure scaffold (partial)}]
  \leanfile{Hqiv/Physics/NonAbelianHolonomyMeasureScaffold.lean}.

\item[\hypertarget{lean:scale-witness}{Scale witness discipline}]
  \leanid{ScaleWitness}, \leanid{defaultScaleWitness} in
  \leanfile{Hqiv/Physics/ScaleWitness.lean}.

\item[\hypertarget{lean:nuclear-curvature-binding}{Nuclear inside/outside binding}]
  \leanfile{Hqiv/Physics/NuclearCurvatureBinding.lean} ---
  \leanid{nuclearClusterBindingCurvature}, \leanid{nuclearValleyContactCount}.

\item[\hypertarget{lean:hqiv-nuclei}{HQIV nuclei: Fresnel caustics and ladder}]
  \leanfile{Hqiv/Physics/HQIVNuclei.lean} ---
  \leanid{fresnelCaustic}, \leanid{valleyCount}, \leanid{deuteron},
  \leanid{helium4}, \leanid{helium4\_valleyCount}.

\item[\hypertarget{lean:nuclear-contact-closure}{Nuclear contact closure / zero deficit}]
  \leanfile{Hqiv/Physics/NuclearContactClosure.lean} ---
  \leanid{contactInformationDeficit}, \leanid{isFullyContactClosed},
  \leanid{fullyInteractingAlphaWitness},
  \leanid{first\_zero\_deficit\_mass\_number\_eq\_referenceM}.

\item[\hypertarget{lean:casimir-caustic-stack}{Casimir caustic hierarchy (Python)}]
  \scriptfile{hqiv_nuclear_caustic_binding.py} ---
  \texttt{sphere\_pair\_overlap}, \texttt{barbell\_torus},
  \texttt{torus\_deepen\_step\_*}, \texttt{tetrahedral\_closure}.

\item[\hypertarget{lean:bound-states}{Bound states / network spine}]
  \leanfile{Hqiv/Physics/BoundStates.lean} ---
  \leanid{E\_bind\_from\_network}, \leanid{E\_bind\_from\_composite\_trace}.

\item[\hypertarget{lean:nucleon-shared-binding}{Nucleon shared binding}]
  \leanid{nucleonSharedBinding\_MeV}, \leanid{nucleonSharedBinding\_from\_composite\_trace}.

\item[\hypertarget{lean:derived-nucleon-mass}{Derived nucleon mass module}]
  \leanfile{Hqiv/Physics/DerivedNucleonMass.lean}.

\item[\hypertarget{lean:derived-nucleon-pack}{Derived nucleon mass pack}]
  \leanid{derivedProtonMass\_pos}, \leanid{derivedNeutronMass\_pos},
  \leanid{constituent\_isospin\_splitting}.

\item[\hypertarget{lean:isospin-gap-one}{Isospin gap unit}]
  \leanid{nucleonIsospinGap\_eq\_one} in \leanfile{Hqiv/Physics/QuarkMetaResonance.lean}.

\item[\hypertarget{lean:proton-anchor}{Proton anchor condition}]
  \leanid{ProtonAnchorCondition}, \leanid{proton\_anchor\_condition\_discharge}.

\item[\hypertarget{lean:neutron-scaffold}{Neutron stability scaffold}]
  \leanfile{Hqiv/Physics/NeutronBindingStabilityScaffold.lean} ---
  \leanid{freeNeutronStrongDecayWidth}, \leanid{freeNeutronWeakDecayWidth},
  \leanid{free\_neutron\_strong\_half\_life\_from\_spin\_statistics}.

\item[\hypertarget{lean:spin-stats}{Spin--statistics module}]
  \leanfile{Hqiv/Physics/SpinStatistics.lean} ---
  exchange phase $-1$ for half-integer modes.

\item[\hypertarget{lean:spin-stats-half-life}{Spin-statistics half-life}]
  \leanid{spin\_statistics\_determines\_half\_life},
  \leanid{resonance\_half\_life} in \leanfile{Hqiv/Physics/HQIVNuclei.lean}.

\item[\hypertarget{lean:dynamic-beta}{Dynamic $\beta$ isotope module}]
  \leanfile{Hqiv/Physics/DynamicBetaIsotope.lean} ---
  \leanid{betaMinusMassGapAtXi}, \leanid{betaMinusCurvatureOverlap},
  \leanid{betaMinusResidualAtXi}, \leanid{betaMinusEndpointQAtXi},
  \leanid{betaWeakMatrixElementSquared}, \leanid{betaGeometryWidthFactor},
  \leanid{betaWeakWidthFromResidual} \emph{(see module for full list)}.

\item[\hypertarget{lean:endpoint-q-theorem}{Endpoint-$Q$ identity}]
  \leanid{betaMinusEndpointQAtXi\_eq\_derivedDeltaM\_minus\_m\_e}.

\item[\hypertarget{lean:curvature-mass-imprint}{Curvature mass imprint}]
  \leanid{curvatureMassImprint}, \leanid{imprintedMassBudget},
  \leanid{imprintedMassBudget\_eq\_ref} in
  \leanfile{Hqiv/Physics/DynamicBetaIsotope.lean}.

\item[\hypertarget{lean:endpoint-q-nucleon-gap}{Nucleon-gap endpoint policy}]
  \leanid{betaMinusEndpointQNucleonGap},
  \leanid{betaMinusEndpointQNucleonGap\_eq\_betaMinusEndpointQAtXi},
  \leanid{betaMinusIsotopeMassGapAtXi\_eq\_betaMinusMassGapAtXi}.

\item[\hypertarget{lean:endpoint-q-uniform-imprint}{Uniform imprint preserves $\Delta M$}]
  \leanid{betaMinusEndpointQUniformImprint},
  \leanid{betaMinusEndpointQUniformImprint\_eq\_kappa\_gap\_minus\_m\_e}.

\item[\hypertarget{lean:endpoint-q-per-isotope-imprint}{Per-isotope imprint $\to$ mass-table $Q$}]
  \leanid{betaMinusEndpointQPerIsotopeImprint},
  \leanid{betaMinusEndpointQPerIsotopeImprint\_eq\_reference\_budgets}
  in \leanfile{Hqiv/Physics/DynamicBetaIsotope.lean}.

\item[\hypertarget{lean:width-well-geometric-blend}{Geometric-mean width well}]
  \leanid{betaWidthWellGeometricBlend}, \leanid{widthWellCurvatureImprint}.

\item[\hypertarget{lean:valley-bound}{Valley count bound}]
  \leanid{betaValleyCountBound}, \leanid{betaCausticLayerCount}.

\item[\hypertarget{lean:weak-width-slot}{Weak width slot}]
  \leanid{betaWeakWidthSlot}, \leanid{betaWeakWidthSlot\_betaMinus}.

\item[\hypertarget{lean:gf-from-beta}{Fermi coupling from $\beta$}]
  \leanid{G\_F\_from\_beta} in \leanfile{Hqiv/Physics/Forces.lean}.

\item[\hypertarget{lean:beta-decay-rate}{Beta decay rate scaffold}]
  \leanid{beta\_decay\_rate} in \leanfile{Hqiv/Physics/NuclearAndAtomicSpectra.lean}.

\item[\hypertarget{lean:weak-channel-open}{Weak beta channel open}]
  \leanid{weakBetaChannelOpen} in \leanfile{Hqiv/Physics/DynamicIsotopeStability.lean}.

\item[\hypertarget{lean:beta-width-qualified}{Beta width ledger qualified}]
  \leanid{betaWidthLedgerQualified}, \leanid{EMTippingQualified} in
  \leanfile{Hqiv/Physics/DynamicIsotopeStability.lean}.

\item[\hypertarget{lean:em-tipping}{EM tipping gate}]
  Legacy export name \leanid{EMTippingQualified} = \leanid{betaWidthLedgerQualified}
  at \leanid{m\_e\_PDG}; stability uses \leanid{structurallyShieldedIsotope}.

\item[\hypertarget{lean:neutron-lifetime-method}{Neutron lifetime method (bottle/beam)}]
  \leanfile{Hqiv/Physics/NeutronLifetimeMethod.lean} ---
  \leanid{trapMagneticCurvatureFraction}, \leanid{trapWeakWidthFactorFromMagnetic},
  \leanid{beam\_tau\_gt\_bottle\_from\_magnetic\_trap\_dressing},
  \leanid{temperature\_ppm\_insufficient\_for\_bottle\_beam\_split},
  \leanid{TrapWeakBridgeDecoherenceSlot}, \leanid{central\_brackets\_bottle\_beam\_refs}.

\item[\hypertarget{lean:dynamic-stability}{Dynamic isotope stability}]
  \leanfile{Hqiv/Physics/DynamicIsotopeStability.lean} ---
  \leanid{structurallyShieldedIsotope}, \leanid{dynamicallyStableIsotope},
  \leanid{betaHalfLifeFromWidth\_eq}.

\item[\hypertarget{lean:half-life-from-width}{Half-life from width}]
  \leanid{betaHalfLifeFromWidth\_eq}.

\item[\hypertarget{lean:weak-fano-hopf}{Weak Fano/Hopf bridge}]
  \leanfile{Hqiv/Physics/WeakFanoHopfBridge.lean} ---
  \leanid{weakBridgeShape}, \leanid{weakBridgeEnergyMeV},
  \leanid{defaultBetaWeakBridge}.

\item[\hypertarget{lean:outside-temp}{Outside temperature dynamics}]
  \leanfile{Hqiv/Physics/NuclearOutsideTemperatureDynamics.lean} ---
  \leanid{T\_MeV\_from\_xi}, \leanid{outsideCurvatureReleaseFactor},
  \leanid{nucleonOwnBindingAtXi}, \leanid{betaMinusOverlapAtXi},
  \leanid{localCurvatureNeutrinoOpacityBarn},
  \leanid{localCurvatureWeakWidthFactor} \emph{(see module for full list)}.

\item[\hypertarget{lean:nu-opacity-lockin}{Local $\nu$ opacity at lock-in}]
  \leanid{localCurvatureNeutrinoOpacityBarn\_lockin\_zeroGravity}
  ($140^{4}$ barn OOM).

\item[\hypertarget{lean:outer-nu-suppression}{Outer horizon neutrino suppression}]
  \leanid{outerHorizonNeutrinoSuppression},
  \leanid{outerHorizonNeutrinoSuppression\_eq\_inv\_140}
  in \leanfile{Hqiv/Physics/DerivedGaugeAndLeptonSector.lean}.

\item[\hypertarget{lean:outside-release}{Outside release factor}]
  \leanid{outsideCurvatureReleaseFactor\_pos}.

\item[\hypertarget{lean:geff-outside}{Outside $G_{\mathrm{eff}}$ support}]
  \leanfile{Hqiv/Physics/NuclearOutsideTemperatureDynamics.lean} ---
  \leanid{OutsideGravityWitness}, \leanid{outsideEnvironmentModulator}
  \emph{(see module for full list)}.

\item[\hypertarget{lean:bond-state-network}{Bond-state network module}]
  \leanfile{Hqiv/QuantumChemistry/BondStateNetwork.lean},
  \leanfile{Hqiv/QuantumChemistry/CurvatureBondContact.lean}.

\item[\hypertarget{lean:homogeneous-curvature}{Homogeneous-curvature second order}]
  \leanfile{Hqiv/Physics/HomogeneousCurvatureSecondOrder.lean} ---
  \leanid{homogeneousCurvatureBudgetAtXi}, \leanid{effectiveCurvatureBudgetAtXi}
  \emph{(see module for full list)}.

\item[\hypertarget{lean:phase-allotrope-derivation}{Phase allotrope derivation (\texttt{hqiv\_lab})}]
  \leanfile{Hqiv/QuantumChemistry/PhaseAllotropeDerivation.lean} ---
  \leanid{IntermolecularMotif}, \leanid{AllotropeCandidate},
  \leanid{DerivedUnitCell}, \leanid{motif\_h2o\_has\_ice\_ih\_label}.

\item[\hypertarget{lean:phase-geometry-density}{Phase geometry $\rightarrow$ density $\rightarrow$ $\rho_{\mathrm{curv}}$}]
  \leanfile{Hqiv/QuantumChemistry/PhaseGeometryDensity.lean} ---
  \leanid{PhaseUnitCell}, \leanid{curvatureDensityFraction},
  \leanid{phaseUnitCellH2OIceIh} \emph{(see module for full list)}.

\item[\hypertarget{lean:phase-material-response}{Phase material response ($n$, $\varepsilon_r$, $k_{\mathrm{th}}$)}]
  \leanfile{Hqiv/QuantumChemistry/PhaseMaterialResponse.lean} ---
  \leanid{clausiusMossottiRatio}, \leanid{refractiveIndexFromCM},
  \leanid{localFieldDivisorFromEta}, \leanid{solidOnsagerLocalFieldDivisor},
  \leanid{phaseOrientationCmFactorTetrahedralIh}, \leanid{gEffFromPhaseParticipation},
  \leanid{molarHeatCapacitySlot}, \leanid{latentHeatFusionSlot},
  \leanid{birefringenceDeltaNSlot}, \leanid{dynamicViscositySlot}
  \emph{(see module for full list)}.

\item[\hypertarget{lean:outside-molecular}{Outside molecular witness}]
  \leanid{OutsideMolecularWitness} in
  \leanfile{Hqiv/Physics/NuclearOutsideTemperatureDynamics.lean}.

\item[\hypertarget{lean:dynamic-pn}{Dynamic nucleon p/n pair}]
  \leanfile{Hqiv/Physics/DynamicNucleonPN.lean}.

\item[\hypertarget{lean:caustic-layers}{Caustic binding stack}]
  \leanfile{Hqiv/Physics/NuclearCausticBinding.lean},
  \leanfile{Hqiv/Physics/NuclearCurvatureBinding.lean}.

\item[\hypertarget{lean:nuclear-spectra}{Nuclear spectra module}]
  \leanfile{Hqiv/Physics/NuclearAndAtomicSpectra.lean}.

\item[\hypertarget{lean:hqiv-nuclei-valley-scaffold}{Isotope valley scaffold}]
  \leanfile{Hqiv/Physics/HQIVNuclei.lean} ---
  \leanid{valleyCount}, \leanid{valleys\_are\_additive}.

\end{description}
\end{sloppypar}

\section{Witness scripts and JSON exports}
\label{app:scripts}

\begin{sloppypar}
\footnotesize
\begin{description}[leftmargin=1.25em,labelwidth=!,style=nextline,itemsep=0.35em]
\item[\hypertarget{script:isotope-halflife}{Isotope stability half-life script}]
  \scriptfile{hqiv_isotope_stability_halflife.py} ---
  stability readout, half-life, outside support;
  \datafile{isotope_stability_halflife.json}.
\item[\hypertarget{script:dynamic-beta}{Dynamic beta isotope script}]
  \scriptfile{hqiv_dynamic_beta_isotope.py} ---
  three-ledger $\beta$ channel chart;
  \datafile{dynamic_beta_isotope_chart.json}.
\item[\hypertarget{script:isotope-pdg}{Isotope PDG benchmark script}]
  \scriptfile{hqiv_isotope_pdg_benchmark.py} ---
  light-panel benchmark table;
  \datafile{isotope_pdg_benchmark.json}.
\item[\hypertarget{script:outside-temp}{Outside temperature dynamics script}]
  \scriptfile{hqiv_nuclear_outside_temperature_dynamics.py} ---
  outside $T(\xi)$ and gravity binding stack.
\item[\hypertarget{script:curvature-binding}{Curvature binding program script}]
  \scriptfile{hqiv_curvature_binding_program.py} ---
  preferred $G_{\mathrm{eff}}$ + shared-well binding witness;
  \datafile{curvature_binding_program.json}.
\item[\hypertarget{script:binding-energy-program}{Binding energy program script}]
  \scriptfile{hqiv_binding_energy_program.py} ---
  unified constructive + post-$\alpha$ panel;
  \datafile{binding_energy_program.json}.
\item[\hypertarget{script:nuclear-caustic}{Nuclear caustic binding script}]
  \scriptfile{hqiv_nuclear_caustic_binding.py} ---
  diagnostic outside-caustic layer stack (not PDG-comparable totals).
\item[\hypertarget{script:inside-outside}{Inside/outside binding script}]
  \scriptfile{hqiv_nuclear_inside_outside_binding.py}.
\item[\hypertarget{script:dynamic-pn}{Dynamic nucleon p/n script}]
  \scriptfile{hqiv_dynamic_nucleon_pn.py}.
\item[\hypertarget{script:bond-state}{Bond-state network script}]
  \scriptfile{hqiv_bond_state_network.py} ---
  molecular bond-state network; $\mathrm{T_2}$, $\mathrm{T_2O}$ hosts.
\item[\hypertarget{script:hqiv-lab}{Materials lab package (\texttt{hqiv\_lab})}]
  derived allotropes, \texttt{hqiv-lab} CLI; install via \texttt{pip install -e .}.
\item[\hypertarget{script:phase-geometry}{Phase geometry density script}]
  \scriptfile{hqiv_phase_geometry_density.py} ---
  $\rho_{\mathrm{curv}}$ melt readout via \texttt{hqiv\_lab}.
\item[\hypertarget{script:phase-material}{Phase material response script}]
  \scriptfile{hqiv_phase_material_response.py} ---
  $n$, $\varepsilon_r$, $k_{\mathrm{th}}$, $C_p$, $L_f$, $\eta$, $\Delta n$.
\item[\hypertarget{script:condensed-audit}{Condensed phase audit script}]
  \scriptfile{hqiv_condensed_phase_audit.py} ---
  GMTKN55 panel @ species melt $T$; \datafile{hqiv_lab_witnesses.json}.
\item[\hypertarget{script:thermo-phase}{Thermodynamic phase script}]
  \scriptfile{hqiv_thermodynamic_phase_from_tp.py}.
\item[\hypertarget{script:homogeneous-feedback}{Homogeneous curvature feedback script}]
  \scriptfile{hqiv_homogeneous_curvature_feedback.py}.
\item[\hypertarget{script:contact-network}{Curvature contact network script}]
  \scriptfile{hqiv_curvature_contact_network.py}.
\item[\hypertarget{script:bbn-integrator}{BBN integrator script}]
  \scriptfile{hqiv_bbn_integrator.py} ---
  \datafile{bbn_integrator.json}.
\item[\hypertarget{script:bbn-condition}{BBN condition-decay script}]
  \scriptfile{hqiv_bbn_condition_decay.py}.
\item[\hypertarget{script:bbn-tables}{BBN paper tables script}]
  \scriptfile{hqiv_bbn_paper_tables.py} ---
  \datafile{bbn_paper_tables.json}.
\item[\hypertarget{script:integrator-audit}{Integrator Lean audit script}]
  \scriptfile{hqiv_integrator_lean_audit.py} ---
  \datafile{integrator_lean_audit.json}.
\item[\hypertarget{script:decay-chain}{Decay-chain calculator}]
  \scriptfile{hqiv_decay_chain.py}, \scriptfile{hqiv_decay_calculator.py} ---
  nuclear $\beta^\pm$, EC/$\beta^+$ (post-$\alpha$ local contact), and two-$\alpha$
  breakup on the canonical binding spine.
\item[\hypertarget{script:post-alpha-binding}{Post-$\alpha$ binding program}]
  \scriptfile{hqiv_post_alpha_binding_program.py},
  \scriptfile{hqiv_post_alpha_sphere_touching.py}.
\item[\hypertarget{script:phase-material-test}{Phase material response tests}]
  \scriptfile{test_hqiv_phase_material_response.py}.
\end{description}
\end{sloppypar}

Run from repository root (see bundled \texttt{scripts/README.md} in \texttt{scripts.zip}):
\begin{verbatim}
pip install -e .
hqiv-lab H2O
hqiv-lab CH4 --json
python3 scripts/hqiv_isotope_pdg_benchmark.py --qualify-em-tipping
python3 scripts/hqiv_isotope_stability_halflife.py --A 1 --Z 0 \
  --qualify-em-tipping --lab-temperature-K 300 --gravity-stack full
PYTHONPATH=scripts python3 scripts/hqiv_bbn_integrator.py
PYTHONPATH=scripts python3 scripts/hqiv_bbn_paper_tables.py
PYTHONPATH=scripts python3 scripts/hqiv_decay_calculator.py nuclear \\
  --seed Z4A7 --max-depth 3
PYTHONPATH=scripts python3 scripts/hqiv_decay_calculator.py nuclear \\
  --seed Z4A8 --max-depth 2
\end{verbatim}

\bibliographystyle{plain}
{\sloppy\small
\bibliography{../references}
}

\end{document}
