\documentclass[11pt]{amsart}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\input{../include/patch_theory_messaging.tex}
\input{../include/referenceM_programme_pin.tex}

\title{Flavor Mixing, Rare Decays, and SMEFT Coefficients from the HQIV Discrete Octonion Carrier: A Machine-Checked Proof Ladder}

\author{Steven Ettinger Jr}

\begin{document}
\maketitle

\begin{abstract}
We discharge the flavor-mixing layer of the HQIV programme as a self-reinforcing,
machine-checked proof ladder on the discrete octonion carrier at \texttt{referenceM=4}.
The spine proves a finite $3\times3$ CKM magnitude ledger with row and column unitarity,
CP-odd holonomy skew $J_{\rm ledger}=3/80$, PDG-style Jarlskog factorization
$\sin\delta\cdot\sqrt{|V_{us}|^2|V_{cb}|^2}\cdot\cos$ envelopes, PMNS angles from T10
admissible-cycle overlaps with normal ordering, rare-decay and angular observables
(including $P_5'$ and $B\to K^*\mu\mu$ forward-backward skeleton), discrete SMEFT
coefficients, PDF moments, patch 2$\to$2 unitarity, and a Sievers glueball bridge
($\kappa_7/\kappa_1=|V_{cb}|^2/|V_{us}|^2$).
Capstone: \texttt{proofSpineInventory\_holds} in \texttt{Hqiv.Physics.ProofSpineInventory}.
Comparison PDG numerals remain outside proof hypotheses.
\end{abstract}

\section{Introduction and position in the HQIV chain}
HQIV derives flavor structure from informational monogamy on the discrete null lattice
and octonionic carrier holonomy, not from fitted Wolfenstein parameters.
This paper packages the flavor spine as seven proof phases plus a strong-sector bridge,
citing the core axioms (\texttt{OctonionicLightCone}, $\gamma=2/5$, $\alpha=3/5$) and
the inventory map \texttt{docs/PROOF\_SPINE\_MAP.md}.
It complements the TUFT electroweak readout and HEP decay ledger while standing alone
as the Tier-3 flavor target.

\section{The self-reinforcing proof ladder}
Each phase aggregates Lean certificates; Table~\ref{tab:lean-module-map} lists modules.
\input{generated/lean_module_map.tex}

\section{CKM ledger, $\delta_{\rm CP}$, and Jarlskog holonomy}
Slot squares $|V_{us}|^2=\gamma/8$, $|V_{cd}|^2=\gamma/16$, $|V_{cb}|^2=\gamma/32$
assemble the charge-conjugate unitary ledger in \texttt{CkmHolonomyReadout.lean}.
The CP phase is $\delta_{\rm CP}=3\pi/32$; the ledger Jarlskog equals the CP-odd Fano skew.
The PDG parameterization satisfies the proved factorization
\texttt{ckmJarlskogFromAngles\_eq\_sin\_delta\_sin\_cos\_envelope}; the correction ratio
$J_{\rm param}/J_{\rm ledger}\approx 0.021$ is \emph{not} $\mathcal{O}(1)$---ledger skew
remains the primary HQIV CP-odd invariant.
\input{generated/ckm_holonomy.tex}

\section{PMNS with normal ordering}
T10 overlap yields $\sin^2\theta_{12}=1/4$, $\sin^2\theta_{23}=1/3$,
$\delta_{\rm CP}^{\rm PMNS}=\pi/5$, with machine-checked normal ordering
$m_0<m_1<m_2$ and $\Delta m^2_{21}>0$ (\texttt{PMNSHolonomyReadout.lean}).
Full PMNS Jarlskog closure and sub-percent angle comparison remain lighter than the CKM
discharge; the programme synthesis paper flags this explicitly.

\section{Rare decays, CP/LFV, and differential observables}
Phase~4 routes $K\to\pi\nu\nu$, $B_s\to\mu\mu$, and $b\to s\gamma$ slots and extends
to angular distributions (\texttt{FlavorDifferentialReadout.lean}):
$P_5'=11/400$, $A_{\rm FB}(B\to K^*\mu\mu)=8800/160121$.
\input{generated/differential_readout.tex}

\section{Higgs self-couplings and discrete SMEFT exporter}
\texttt{HiggsSelfCouplingReadout.lean} certifies $\lambda_3$ and $\kappa$ modifiers;
\texttt{SMEFTDiscreteExporter.lean} emits dimension-6 holonomy-skew coefficients.

\section{Collider and cosmology witnesses}
\texttt{PartonCarrierEvolution.lean} and \texttt{PatchScatteringUnitarity.lean}
cover PDF moments and patch 2$\to$2 optical theorem;
\texttt{CosmologicalPerturbationReadout.lean} supplies the lighter Friedmann capstone.

\section{Longitudinal and modified-inertia signatures}
\texttt{FlavorLongitudinalAmplitudeBridge.lean} shows rare-decay and 2$\to$2 amplitudes
pick up a positive correction $1+\gamma|\Delta\phi_{\rm imprint}|$ from the O-Maxwell
phase-gradient slot at lock-in, with inertia screen unity at $a=1$.
At \texttt{referenceM=4}, the imprint increment yields $\sim 8\%$ shifts on the
proved $P_5'$ and $A_{\rm FB}$ skeleton (Table~\ref{tab:phenomenology})---illustrative
until form-factor insertion is proved.
\input{generated/phenomenology_edges.tex}

\section{Sievers glueball convergence (strong sector)}
Sievers~\cite{Sievers2026Glueball} models glueballs via $S^7\hookrightarrow S^{15}\to S^8$
with fiber drag $\kappa_7=3/2$ and singlet stiffness $\kappa_1=6$.
\texttt{GluonSieversHopfBridge.lean} maps the combined backreaction $1/4$ to the CKM
ledger ratio $|V_{cb}|^2/|V_{us}|^2$, linking octonionic Hopf fine structure to the
HQIV discrete carrier without injecting lattice masses.

\section{Conclusions and roadmap}
The flavor paper is anchored on \texttt{proofSpineInventory\_holds}.
The programme synthesis (\texttt{papers/proof\_spine\_synthesis/}) positions this Tier-3
readout within the full HQIV spine for a wider audience.
Open items: PMNS depth matching CKM, form-factor-level longitudinal insertion, and
$q^2$-dependent differential distributions.

\section*{Reproducibility}
\textbf{Lean:} \texttt{lake build paper\_flavor\_mixing} (3383 jobs).
\textbf{Witness:} \texttt{python3 scripts/hqiv\_flavor\_mixing\_export.py}.
\textbf{Bundle:} \texttt{papers/flavor\_mixing\_readout/scripts.zip}.

\bibliographystyle{amsplain}
\bibliography{../references}

\end{document}
