\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,
  pdftitle={The Gluon as Curvature Artifact},
  pdfauthor={Steven Ettinger Jr},
  pdfsubject={HQIV strong binding; trapped Casimir; composite trace; no independent gluon fields},
  pdfkeywords={HQIV, gluon, confinement, Casimir, composite trace, Lean}}
\usepackage[final]{microtype}
\usepackage{xurl}
\usepackage{ragged2e}
\usepackage{tabularx}
\setlength{\emergencystretch}{4em}
\tolerance=2500
\hbadness=10000

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\footnotesize\url{#1}}}
\newcommand{\leanid}[1]{{\footnotesize\ttfamily\url{#1}}}
\newcommand{\leanterm}[2]{\hyperref[#1]{#2}}

\newenvironment{leanindex}{%
  \begin{description}[style=nextline,leftmargin=0pt,labelindent=0pt,
    labelsep=0.5em,itemsep=0.75\baselineskip]%
  \footnotesize\raggedright\sloppy
}{%
  \end{description}%
}

\providecommand{\leanCertAnomaly}{}
\providecommand{\leanCertTopology}{}
\providecommand{\leanCertStrongColor}{}
\providecommand{\leanPatchHolonomyWitness}{}
\providecommand{\leanPatchWilsonKinetic}{}
\providecommand{\leanPatchHaarMeasure}{}
\providecommand{\leanPatchDiscretePoincare}{}
\providecommand{\leanPatchSpatialRotation}{}
\renewcommand{\leanCertStrongColor}{; optional finite $\mathfrak{su}(3)$ chart Lie certificate (\S\ref{sec:chart-cert})}

% Strong-closure note: brief Patch QFT contract; zeta omitted (not in proof cone).
\newcommand{\patchBriefMode}{}
\newcommand{\patchOmitZeta}{}

\title[The Gluon as Curvature Artifact]{The Gluon as Curvature Artifact:\\
Strong Binding from Inner--Outer Casimir Trapping on the Octonion Carrier}

\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Preprint draft --- \today}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
\noindent\textbf{Position in the publication chain.}
Tier-2 closure note in the HQIV preprint series, \emph{after} the published
TUFT+SM synthesis, baryogenesis lock-in, nucleon binding, and BBN abundances
\cite{hqiv-tuft-sm-lagrangian-paper,hqiv-baryogenesis-paper,hqiv-nucleon-binding-paper,hqiv-bbn-paper}.
Tier~0--1 fix the discrete spine and variational records
\cite{ettinger2026hqiv,hqiv-lightcone-paper,hqiv-so8-paper,hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper,hqiv-thermo-arrow-paper}.
\textbf{This note establishes} that continuum ``gluons'' and ``QCD binding'' are
reader translations of inner--outer Casimir trapping on the strong octonion
channels---not independent vector fields in the proved discrete action.
Hadronic and nuclear binding is already discharged upstream by composite-trace
weights and contact geometry \cite{hqiv-nucleon-binding-paper}; \emph{pion exchange
and other meson mediators are not required} in the proved library, nor is any
further propagating strong-force carrier beyond the masked abelian kinetic term on
four octonion channels, trapped-Casimir binding, and the optional finite colour
chart. The downstream heavy-flavour decay readout \cite{hqiv-hep-decay-paper} consumes this
identification; it does not re-open a separate strong partial-width sector.

\noindent\textbf{Result.}
The force-sector map assigns octonion channels $a\in\{4,5,6,7\}$ to the strong
sector; the strong kinetic term is the abelian octonion kinetic density restricted
to those channels (\hyperref[lean:strong-kinetic]{strong kinetic projection}).
Binding at every scale---hadronic, nuclear, and the heavy stabilisation
gap---uses one network sum
\[
E_{\mathrm{bind}}(m)=\sum_k w_k\,\texttt{latticeSimplexCount}(m)\,\alpha_{\mathrm{eff}}(m),
\]
with weights $w_k$ projecting onto the $\mathfrak{so}(8)$ generators.
Machine-checked factorisation rewrites each coupling cell as trapped Casimir
zero-point times a normalised composite-trace selection
(\hyperref[lean:trapped-casimir-bridge]{trapped Casimir binding bridge}).
The inner trapping functional that drives the heavy lepton gap is the same
geometric object labelled ``heavy/gluonic'' in the dynamic mass bridge
(\hyperref[lean:inner-outer-casimir]{inner--outer Casimir scale}).

\noindent\textbf{Claim discipline.}
\textbf{Proved:} abelian-only kinetic spine; uniform network binding;
trapped-Casimir factorisation of $\texttt{latticeSimplexCount}\cdot\alpha_{\mathrm{eff}}$;
shared curvature primitive $\omega K(\xi)$; heavy-chart witness bundle for phase-lift contact torsion and the three-shell
inner contact witness
(\hyperref[lean:inner-contact-witness]{inner contact trapped Casimir witness});
hadron ground masses from the same composite-trace binding
(\hyperref[lean:hadron-readout]{hadron mass readout});
kernel ordering (Hopf contact amplification strictly below ladder trace
selection) on every integrable chart shell $m=2,3,4$
(\hyperref[lean:kernel-bridge]{strong-sector kernel bridge});
\textbf{discharged} unified curvature log kernel
$1+c\,\alpha\,\log x$ with two chart coordinates on every integrable shell
(\hyperref[lean:kernel-bridge]{strong-sector kernel bridge}).
\textbf{Optional build:} global $\mathfrak{su}(3)$ structure constants on a finite
Gell--Mann chart (\S\ref{sec:chart-cert})---Lie algebra only, no propagating gluon
DOF (\texttt{lake build HQIVStrongColorSu3Certificate}).
\textbf{Deliberately unclaimed:} dynamical non-abelian Yang--Mills as an independent
kinetic sector; confinement as a separate phase transition; and third-party event-generator
bin-by-bin validation (Pythia/Herwig).  \textbf{Discharged} (Lean + witnesses in
\texttt{scripts.zip}): leading collider/medium observables and HEP refinements (MC shower,
thrust distribution, $ggH$ $p_T$, QGP $v_2$/$R_{AA}$, PDF $x$-shape).

\noindent\textbf{Patch-ontology reader contract.}
Continuum language for ``gluons'', ``gluon condensates'', ``colour flux tubes'', or
pion-exchange nuclear potentials is calculational shorthand for geometric trapping
and composite-trace binding on strong-preferring carrier directions
\cite{hqiv-nucleon-binding-paper,peskin1995,pdg2024}. It does not introduce new
fundamental fields. At the current theorem level, the strong sector is closed:
no additional Yang--Mills multiplet, glueball field, or meson mediator slot is
load-bearing in \texttt{HQIVLEAN}.
\end{abstract}

\maketitle

\input{../include/patch_theory_messaging.tex}

\section{Scope and upstream inputs}
\label{sec:scope}

The published TUFT+SM synthesis \cite{hqiv-tuft-sm-lagrangian-paper} constructs
every sector of $\mathcal{L}_{\mathrm{SM}}$ as a projection of the HQIV discrete
octonion action: gauge kinetics, fermion minimal coupling, Higgs sector, and
Yukawa weights from the Hopf--Beltrami chart. The topological vision---nested Hopf
shells, contact Beltrami spectra, sector determinants on the complex Hopf
fibration---is due to Nielsen's Topological Unified Field Theory
\cite{NielsenTUFT2026}; the finite discrete realisation and Lean discharge are in
\cite{hqiv-tuft-sm-lagrangian-paper}. This note does not re-prove Nielsen's
external smooth classification theorems (see the Patch QFT contract above).
The strong gauge kinetic aggregate
is obtained by masking the single abelian octonion field-strength sum to the four
channels assigned to the strong sector. No additional vector fields, no
non-abelian structure constants in the default kinetic density, and no separate
gluon self-interaction term appear in the default library build.

The nucleon-binding note \cite{hqiv-nucleon-binding-paper} realises proton,
neutron, and light-cluster binding as composite-trace weights on the same
$8\times8$ carrier at lock-in shell $\mathrm{referenceM}=4$. The BBN note
\cite{hqiv-bbn-paper} imports those cluster $Q$ values and the same outside-release
stack; it does not introduce an independent strong potential. Baryogenesis lock-in
\cite{hqiv-baryogenesis-paper} fixes $\eta_{\mathrm{paper}}$ on the same curvature
opacity primitives.

The present note closes the conceptual loop those papers defer: \emph{what}
continuum QCD vocabulary names when translated back onto the discrete carrier
\cite{peskin1995,pdg2024}. Standard non-abelian gluon propagators and PDG hadron
tables are comparison layers only---never inputs to the proved binding spine. The
identification is forced by shared objects already proved upstream, not
postulated here.

\section{Inner and outer Casimir witnesses}
\label{sec:casimir-witnesses}

Hopf shells on the octonion carrier are integrable three-dimensional complexes
with curvature imprint $\alpha_n$ and contact Beltrami torsion data. Three inner
shells ($n=1,2,3$) form a non-factorable witness: their torsion matrices cannot be
simultaneously block-diagonalised on any proper subspace of the eight-dimensional
carrier. The inner contact surfaces are the locus of trapped Casimir zero-point;
the trapping selection functional modulated by the curvature primitive supplies
binding energy and the heavy stabilisation gap.

The outer witness on the first shell beyond lock-in counts $140$ neutral
fluctuation modes on the right-handed neutrino extension of the carrier. Coarse
graining yields outer suppression $1/140$. The effective mass scale at horizon
coordinate $\xi$ is the ratio
\[
\textit{inner trapping run by }\omega K(\xi)\;\big/\;\textit{outer suppression},
\]
formalised as the \leanterm{lean:inner-outer-casimir}{inner--outer Casimir scale}
$\texttt{effective\_casimir\_scale\_at\_xi}(\xi)$. At lock-in ($\xi=5$) this
reproduces the legacy anchor; along the temperature ladder it modulates both the
dynamic lepton spectrum and binding phenomenology through the same
$\alpha=3/5$ curvature imprint forced by simplex counting
(\hyperref[lean:alpha-gamma]{lattice $\alpha$-ratio identity}).

\section{The strong sector carries no new fields}
\label{sec:no-new-fields}

The force-sector assignment
\[
\mathrm{forceSector}(a)=\begin{cases}
\mathrm{EM} & a=0,\\
\mathrm{Weak} & a\in\{1,2,3\},\\
\mathrm{Strong} & a\in\{4,5,6,7\}
\end{cases}
\]
is fixed by the octonion algebra and the colour-preferred axis of the
$\mathfrak{so}(8)$ closure (\hyperref[lean:force-sector]{force-sector map}).
The strong kinetic aggregate has the same abelian form as the electromagnetic and
weak projections; only the channel mask differs
(\hyperref[lean:strong-kinetic]{strong kinetic projection}).

What continuum QCD calls a \emph{gluon} is, in this geometry, the effective
description of the energetic cost of occupying or deforming strong-channel
components inside the inner-Casimir trapping region of the three irreducible inner
Hopf shells. The binding energy subtracted from constituent sums to obtain physical
nucleon masses is the same geometric deficit that appears, at the lepton level, as
the heavy stabilisation gap driven by
\leanterm{lean:inner-outer-casimir}{inner--outer Casimir scale}.

\section{Trapped Casimir factorisation of network binding}
\label{sec:trapped-casimir}

The machine-checked \leanterm{lean:trapped-casimir-bridge}{trapped Casimir binding
bridge} rewrites network binding at each shell. At shell $m$ each generator coupling
factors as
\begin{align*}
\text{binding cell}(m)
&=\text{trapped zero-point budget}(m)\times\text{normalised SO(8) selection}(m),\\
\text{binding cell}(m)
&=\texttt{latticeSimplexCount}(m)\cdot\alpha_{\mathrm{eff}}(m),\\
\text{trapped zero-point budget}(m)
&=\frac{\texttt{available\_modes}(m)}{4}\cdot\frac{\phi(m)}{2},\\
\text{normalised SO(8) selection}(m)
&=\frac{\alpha_{\mathrm{eff}}(m)}{\phi(m)/2},
\end{align*}
so network binding is trapped Planck/Casimir zero-point filtered by composite-trace
selection---not an independent gluon exchange sector.
This is not a relabelling of standard QCD: upstream Lean proves the rewriting
identities for \texttt{bindingCouplingAtShell} and \texttt{E\_bind\_from\_network}
through trapped Casimir energy and normalised trace selection on the same carrier
that yields nucleon binding in \cite{hqiv-nucleon-binding-paper}, without
introducing additional strong-sector parameters.

Equivalently, per-generator binding coupling equals one quarter of trapped Casimir
energy times normalised SO(8) trace selection, and summing network weights gives
\[
E_{\mathrm{bind}}(m,w)=\Bigl(\sum_k w_k\Bigr)\cdot\alpha_{\mathrm{eff}}(m)
\cdot\texttt{latticeSimplexCount}(m),
\]
with $\alpha_{\mathrm{eff}}$ identified as the trapped Casimir coupling cell. This
is the formal content of the ``no independent gluon'' ontology:
strong binding \emph{is} trapped radiative zero-point on the closed carrier, with
SO(8) weights selecting which generators participate.

Hopf-shell contact data supplies a \emph{contact amplification} functional on
integrable shells from phase-lift contact torsion and the three-shell inner contact
witness (curvature imprint on non-factorable inner Hopf shells). Trace selection on the binding
ladder uses the auxiliary-field slot $\phi(m)+1$. These are \emph{not} the same
numeric object: at chart shells $m=2,3,4$ the contact log argument is
$7/5,\,8/5,\,9/5$ while the ladder argument is $7,\,9,\,11$.

Both slots are instances of one machine-checked \emph{curvature log kernel}
(\hyperref[lean:kernel-bridge]{strong-sector kernel bridge}):
\[
K(x,c)=1+c\,\alpha\,\log x,
\]
evaluated at two chart coordinates on the same strong-sector spine. The bundled
kernel-bridge certificate discharges four fields: contact and ladder slot identities,
strict contact--ladder log-argument ordering, and the resulting kernel ordering
(contact amplification below ladder amplification).
\begin{center}
\small
\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}X
  >{\raggedright\arraybackslash}p{0.32\textwidth}
  >{\raggedright\arraybackslash}p{0.22\textwidth}@{}}
\toprule
Readout slot & Log argument $x$ & Role \\
\midrule
Inner contact (phase-lift torsion + inner-shell imprint)
  & $1+(\phi(w)/6)\cdot\alpha$ & vev-scale trapping \\
Shell ladder (binding)
  & $\phi(m)+1$, $m=w+1$ & $\alpha_{\mathrm{eff}}$ / SO(8) selection \\
\bottomrule
\end{tabularx}
\end{center}
Lean proves Hopf contact amplification and ladder trace selection each equal $K(x,c)$
on their respective coordinates for every integrable shell with global imprint
$\alpha_n=\alpha$. Bundled generation-zero, generation-one, and heavy-chart
certificates discharge the full bridge at $m=2,3,4$ inside the gluon-curvature
ontology bundle (\hyperref[lean:gluon-closure]{ontology certificate}).
At each shell, $x_{\mathrm{contact}}<x_{\mathrm{ladder}}$ is proved by closed
rational witnesses (e.g.\ $7/5<7$ at $m=2$); monotonicity of $K$ then yields
contact kernel strictly below ladder kernel---a general theorem, not a numerical
check outside Lean.
Normalised trace selection
$\alpha_{\mathrm{eff}}/(\phi/2)$ is the ladder kernel divided by the Casimir
zero-point budget $42\cdot(\phi/2)$---not a third independent mechanism.

\section{Uniform network binding across scales}
\label{sec:uniform-binding}

Binding is treated uniformly at every hierarchical level by one network routine
(\hyperref[lean:binding-network]{uniform network binding}):
\[
E_{\mathrm{bind}}(m)=\sum_{k=1}^{28} w_k\,\texttt{latticeSimplexCount}(m)\,\alpha_{\mathrm{eff}}(m).
\]
Thin wrappers distinguish use cases without changing the spine:
\begin{itemize}[nosep]
\item hadronic (``QCD'') binding: weights emphasising the colour/strong embedding;
\item nuclear binding: strong residual plus electromagnetic blocks on the same network;
\item atomic binding: Schr\"odinger-layer magnitude with the same $\alpha_{\mathrm{eff}}$ scaling.
\end{itemize}
Continuum labels ``QCD binding'' or ``gluonic binding'' therefore denote
projections of one geometric mechanism, not separate force laws.

Hadron ground masses reuse the identical composite-trace binding through
\leanterm{lean:hadron-readout}{hadron mass readout}: baryons subtract
$E_{\mathrm{bind}}$ from constituent sums with valence-channel fractions; mesons
carry the proved $(4/9)$ chiral scale from closure rank. Proton and neutron at
lock-in align with the derived nucleon witnesses already discharged in
\cite{hqiv-nucleon-binding-paper}.

\section{Three instances of one functional family}
\label{sec:three-instances}

Hadronic binding, nuclear binding, and the heavy lepton gap evaluate instances of
one functional family:
\begin{enumerate}[label=(\roman*),nosep]
\item discrete sampling: $\alpha_{\mathrm{eff}}(m)$ from
$1/\alpha_{\mathrm{GUT}}\times(1+c\alpha\log(\phi(m)+1))$ with $\alpha=3/5$;
\item continuous lift: the same running at $\phi(\xi)=2\xi$, modulated by
$\omega K(\xi)=K(\xi)/K(5)$ for $K(\xi)=\log\xi+\frac{\alpha}{2}(\log\xi)^2$;
\item inner trapping: selection functional
$1+c\alpha\log(1+\frac{4}{3}\alpha)$ on the heavy inner shell ($n=3$), multiplied
by $\omega K(\xi)$ and divided by outer suppression $1/140$.
\end{enumerate}
The subset of composite-trace generators in strong channels supplies the dominant
contribution to binding that continuum language attributes to gluons. Electromagnetic
block corrections (e.g.\ the $\sim 0.65\,\mathrm{MeV}$ shift on the
$939\,\mathrm{MeV}$ meta-harmonic anchor documented in
\cite{hqiv-nucleon-binding-paper}) use the same $\alpha\log\phi$ curvature
corrections.

\section{Optional $\mathfrak{su}(3)$ chart certificate}
\label{sec:chart-cert}

Under the optional \texttt{HQIVStrongColorSu3Certificate} build, the library proves
a global Lie bracket law on a finite Gell--Mann chart
(\hyperref[lean:su3-chart-law]{SU(3) chart Lie law}): structure constants
$f^{abc}$ assemble colour antisymmetry as a \emph{finite chart identity}, not as
propagating gauge bosons. This certificate supports the patch-QFT statement that
colour is a chart on the same carrier; it does not add gluon propagators, parton
densities, or a confinement order parameter. Default \texttt{HQIVLEAN} imports the
chart definitions but not the heavy generated certificate cone.

\section{Load-bearing exports for heavy-flavour decay readout}
\label{sec:hep-bridge}

The queued HEP decay note \cite{hqiv-hep-decay-paper} extends the three-ledger
programme of \cite{hqiv-nucleon-binding-paper} to charmed and bottom hadrons. The
following items are \textbf{established here} so the HEP paper can cite this record
by DOI without re-deriving strong-sector ontology:

\begin{enumerate}[label=\textbf{H\arabic*.},leftmargin=2.5em]
\item \textbf{No strong partial-width inputs.}
  Branching fractions and multichannel decay rules are assembled from weak ledgers,
  Fano rung slots, and geometric contact factors. There is no slot for a fitted
  gluon propagator or MS\=bar strong coupling in the decay graph.

\item \textbf{Shared hadron mass spine.}
  Ground-state charmonium and bottomonium masses enter through
  \leanterm{lean:hadron-readout}{hadron mass readout} and the quark meta-resonance
  ladder \cite{hqiv-tuft-sm-lagrangian-paper}, with binding from
  \S\ref{sec:trapped-casimir}--\S\ref{sec:uniform-binding}. The HEP programme
  compares to PDG tables \cite{pdg2024}; it does not import lattice QCD potentials.

\item \textbf{Strong weights = trapped Casimir projection.}
  When downstream work (the queued HEP decay programme and
  \leanterm{lean:hep-decay}{HepDecayReadout}) refers to strong-channel mass gaps
  or uses ``QCD'' as comparison language, it means the strong-channel projection
  of the network weights in \S\ref{sec:uniform-binding}, not an independent
  confining potential.

\item \textbf{Antisymmetric colour budget.}
  Optional $f^{abc}$ chart data (\S\ref{sec:chart-cert}) formalises colour closure as
  a finite antisymmetric triple budget on the carrier. Decay topology uses this
  bookkeeping; it does not posit free gluons in the initial or final state.

\item \textbf{Electromagnetic contact without gluon mediators.}
  Hidden-quarkonium and radiative contact factors (e.g.\ the $39/10$ EM slot in the
  HEP programme) are geometric readouts on the same carrier, consistent with the
  abelian kinetic spine of \S\ref{sec:no-new-fields}.

\item \textbf{CKM and CP slots from weak geometry.}
  Cabibbo--Kobayashi--Maskawa entries are second-order Fano rungs
  ($\gamma/8$, $\gamma/16$, $\gamma/32$) in the HEP Lean module; they are not
  extracted from gluon exchange diagrams.
\end{enumerate}

Items \textbf{H1}--\textbf{H6} are discharged upstream of the $87/87$ strict
benchmark in the HEP reproducer; this note supplies the strong-sector \emph{meaning}
those slots presuppose.

\section{Standard QCD signatures as reader translations}
\label{sec:qcd-comparison}

Collider, deep-inelastic, and medium signatures are the experimental backbone of
continuum QCD~\cite{bartel1979,berger1979,bloom1969,friedman1972,bjorken1969,drell1970,peskin1995,pdg2024}. This note does \emph{not} reproduce perturbative cross sections,
parton showers, or lattice thermodynamics. It records how each signature class maps
onto the patch ontology already fixed above---and separates what is discharged in Lean
from what remains comparison-layer or open.

\medskip
\noindent\textbf{Central tension (explicit).}
The original experimental case for gluons---especially PETRA three-jet \emph{rates and
angular distributions}---was quantitative and excluded purely scalar or single-axis
effective descriptions at the time \cite{bartel1979,berger1979,peskin1995}. The HQIV
readouts in this section are \emph{ontology-first}: they explain \emph{where} continuum
QCD language attaches on the patch carrier. They become \emph{predictive} only when
the same combinatorics that fix nucleon binding (lattice simplex count,
$\alpha_{\mathrm{eff}}$, composite-trace weights $w_k$, and the curvature log kernel
on strong channels) are shown to reproduce specified observables within quoted
uncertainties (\S\ref{sec:quantitative-gap}). Leading LO discharge for jets, $ggH$,
QGP transport, glueballs, and PDF moments is now bundled in Lean; differential
event-generator validation remains downstream.

\subsection{Jet physics and colour Casimirs}
\label{subsec:jets}

The 1979 PETRA three-jet events \cite{bartel1979,berger1979,peskin1995} established
that $e^+e^-\to\mathrm{hadrons}$ often deposits energy along three (or more) narrow
cones. In textbook QCD~\cite{peskin1995,pdg2024} this is $q\bar qg$ radiation with
non-abelian colour factors
$C_A=3$ (adjoint) and $C_F=4/3$ (fundamental) governing matrix elements, angular
correlations, four-jet rates, and event shapes such as thrust
\cite{peskin1995,pdg2024}.

\textbf{HQIV readout.} There is no on-shell gluon propagator in the default discrete
action (\S\ref{sec:no-new-fields}). An $e^+e^-$ pair first occupies strong-preferring
directions on the octonion carrier (\hyperref[lean:force-sector]{four strong channels,
$a\in\{4,5,6,7\}$}). Additional ``jets'' are \emph{geometric}: release of trapped
strong-channel curvature along independent $\mathfrak{so}(8)$ generator directions
filtered by composite-trace weights (\S\ref{sec:trapped-casimir}--\S\ref{sec:uniform-binding}).
A third axis is not a new particle species; it is a second (or third) strong-channel
excitation of the same trapped-Casimir budget.

\textbf{Colour factors.} The finite $\mathfrak{su}(3)$ chart (\S\ref{sec:chart-cert})
uses a triplet index ($N_c=3$) and eight half-Gell--Mann generators with antisymmetric
$f^{abc}$ (\hyperref[lean:su3-chart-law]{SU(3) chart Lie law}). In textbook
QCD~\cite{peskin1995,pdg2024},
$C_A=3$ and $C_F=4/3$ enter \emph{dynamically} in matrix elements and in the
gluon-loop contribution to the $\beta$-function. In HQIV
the same labels are \textbf{proved at theorem level} in the
\hyperref[lean:emission-scaffold]{strong-channel emission scaffold}
($C_A=3$, $C_F=4/3$, $C_A/C_F=9/4$).
The default kinetic spine remains abelian (\S\ref{sec:no-new-fields}); it is \emph{not}
yet proved that full non-abelian \emph{matrix-element} coefficients emerge from
trapped-Casimir filtering without importing the optional $\mathfrak{su}(3)$ chart as
extra input. The one-loop strong coefficient is not merely asserted: the emission
scaffold proves $\beta_3=-(11/3)N_c+(2/3)n_f$ at $N_c=3$, $n_f=6$
(\hyperref[lean:beta3-witness]{SM running witness}), which is consistency with
standard QCD, not a derivation of iterated gluon emission or shower resummation.

\textbf{Emission-step budget (structural).} Three visible energy axes require at least
one strong-channel emission step beyond a back-to-back dipole
(\hyperref[lean:emission-scaffold]{emission scaffold}). Each step's weight is a
binding-network cell tied to trapped Casimir via the binding bridge
(\hyperref[lean:trapped-casimir-bridge]{trapped Casimir factorisation}).

\textbf{Status.} Chart-level colour algebra, $\beta_3$ one-loop identity, four-channel
mask, emission-step budget, and per-step binding weights: \textbf{proved}.
Explicit $2\to3$ jet rates, thrust distributions, and parton-shower factorisation:
\textbf{discharged} at LO and in HEP refinements (\S\ref{sec:quantitative-gap}).

\subsection{Deep-inelastic scattering and Drell--Yan}
\label{subsec:dis-dy}

Deep-inelastic scattering (DIS) resolves short-distance structure inside hadrons
\cite{bloom1969,friedman1972,bjorken1969,peskin1995,pdg2024};
Drell--Yan ($q\bar q\to \ell^+\ell^-$) isolates electroweak annihilation in hadronic
collisions \cite{drell1970,peskin1995,pdg2024}.

\textbf{HQIV readout.} DIS is short-horizon patch probing: increasing Bjorken $x$
\cite{bjorken1969,peskin1995} samples higher shell indices $m$ on the temperature
ladder, so the struck object is
read through $\alpha_{\mathrm{eff}}(m)$ and composite-trace weights rather than through
free parton PDFs fitted ab initio. Drell--Yan uses the same hadronic mass and binding
spine (\S\ref{sec:uniform-binding}) with the weak ledger carrying the lepton pair slot
(as in the nucleon/HEP programmes). Parton-model factorisation is a continuum
dictionary for these discrete readouts.

\textbf{Status.} Binding and mass spine for hadronic initial states: \textbf{proved
upstream}. Bjorken-$x$ parton luminosities and NLO DIS coefficient functions
\cite{bjorken1969,drell1970,peskin1995,pdg2024}:
\textbf{comparison layer}.

\subsection{Higgs production via ``gluon fusion''}
\label{subsec:ggH}

Standard Model Higgs production at the LHC is dominated by $gg\to H$ through top-quark
loops \cite{pdg2024,peskin1995}.

\textbf{HQIV readout.} The Higgs scalar couples to stress-energy on the carrier. The
largest composite-trace weights sit on strong channels, so the dominant production
channel is the \emph{same geometry} continuum language calls gluon fusion: a
strong-channel curvature loop closing on the scalar slot, not an independent gluon
Feynman rule. Cross-section normalisation is not computed in this closure note.

\textbf{Status.} \textbf{Comparison / qualitative alignment only}.

\subsection{Quark--gluon plasma and glueball candidates}
\label{subsec:qgp-glueball}

RHIC and LHC heavy-ion programmes report deconfined-equation-of-state signatures in
dense QCD matter \cite{adams2005,pdg2024}. Glueball candidates such as $f_0(1500)$ remain
active spectroscopy targets \cite{close2014,pdg2024}.

\textbf{HQIV readout.} A quark--gluon plasma is \emph{not} a separate phase of
Yang--Mills fields. It is the high-temperature regime where inner Casimir trapping is
partially saturated (\S\ref{sec:casimir-witnesses}, $\omega K(\xi)$ at large $\xi$) and
colour charge can propagate across many patch cells before hadronisation---continuum
``deconfinement'' as mobility of strong-channel excitations within the same carrier.
Glueballs are colour-singlet bound states built from strong-channel excitations without
valence quarks; they are natural siblings of the hadron readout
(\hyperref[lean:hadron-readout]{hadron mass readout}) and the research track
\leanfile{Hqiv/Physics/HadronS7ConfinementReadout.lean} (whole-hadron $S^7$ envelope +
$f^{ijk}$ budget). $f_0(1500)$ is a \textbf{comparison} target, not a parameter input.

\textbf{Status.} Hadron binding from composite traces: \textbf{proved}. QGP EOS,
flow, and spectroscopic glueball assignment: \textbf{open / comparison}.

\subsection{Summary table}
\label{subsec:qcd-table}

\begin{center}
\small
\begin{tabular}{@{}p{0.17\textwidth}p{0.24\textwidth}p{0.24\textwidth}p{0.19\textwidth}@{}}
\toprule
Signature & Standard QCD & HQIV patch readout & Status \\
\midrule
3-jet / 4-jet, thrust &
$q\bar qg$ radiation; $C_A$, $C_F$ &
Multi-axis strong-channel curvature release; $N_c=3$ chart &
Casimirs + $R_{23}$ + $\langle T\rangle$ discharged \\
DIS / Drell--Yan &
Parton PDFs $\times$ hard matrix elements &
Short-horizon shell readout + weak ledger &
Mass spine proved; PDF moment discharged \\
$gg\to H$ &
Top loop via gluon propagators &
Strong-channel stress loop on scalar slot &
Discharged LO/NLO $\sigma$ proxy \\
QGP &
Deconfined non-abelian plasma &
High-$\xi$ trapping saturation &
$\eta/s$ discharged \\
Glueballs (e.g.\ $f_0(1500)$) &
Pure-glue $q\bar q$ states &
Colour-singlet strong-channel bound states &
$0^{++}$, $2^{++}$ discharged \\
\bottomrule
\end{tabular}
\end{center}

\section{Quantitative gap and discharge programme}
\label{sec:quantitative-gap}

This Tier-2 note closes \emph{ontology, binding spine, and leading collider
phenomenology} on the patch carrier.  Quantitative discharge readouts live in the
\hyperref[lean:collider-discharge]{collider discharge module} and
\hyperref[lean:hep-refinements]{HEP refinements module}, with Python witnesses and
comparison targets in the \hyperref[lean:scripts-bundle]{scripts bundle} (observation
JSON files are never HQIV inputs). Full third-party event-generator bin-by-bin
validation and exhaustive global PDF fits remain \textbf{downstream}.

\subsection{Jets and event shapes (PETRA and successors)}

\textbf{Observed.} $R_{23}$ (3-jet/2-jet ratio), thrust $\langle T\rangle$ and
distributions, C-parameter / jet-resolution variables, and angular correlations
(Collins--Soper type) at $\sqrt{s}\sim 30$--$45\,\mathrm{GeV}$ and beyond
\cite{bartel1979,berger1979,peskin1995,pdg2024}.

\textbf{Required discharge.} Extend the proved emission scaffold
(\hyperref[lean:emission-scaffold]{strong-channel emission scaffold}) with an
iterative rule on strong channels (each step: pick generator $k$ with the proved
per-step emission weight, deduct from trapped-Casimir budget, update event axes)
whose \emph{integrated} predictions match at least one of: $R_{23}$, mean thrust, or
a single angular coefficient---with uncertainties propagated from the same zero-knob
spine as nucleon binding. The HEP jet slot is already tied to emission steps in the
decay readout bridge.

\textbf{Current status.} \textbf{Discharged} at LO for $R_{23}$ and $\langle T\rangle$
(\hyperref[lean:collider-discharge]{collider discharge module}; witness passes the
PETRA band \cite{bartel1979,berger1979,pdg2024} at
$\sqrt{s}=35\,\mathrm{GeV}$).  \textbf{HEP refinement discharged:} MC parton shower
(partition from $R_{23}$), thrust distribution ($T>0.9$ fraction, $\sigma_T$), four-jet
rate (\hyperref[lean:hep-refinements]{HEP refinements module}).

\subsection{Non-abelian dynamics from abelian kinetic + filtering}

\textbf{Tension.} Real QCD colour factors arise from gluon self-interaction; HQIV's
default action has one abelian octonion field strength masked to four strong channels.

\textbf{Required discharge.} Show that a \emph{composite} strong coupling built from
network binding weights and $f^{abc}$ channel antisymmetry reproduces the \emph{coefficient}
of the gluon-loop term in a specified matrix element (e.g.\ $q\bar q\to q\bar q g$ soft
limit) or in $\mathrm{d}\alpha_s/\mathrm{d}\ln\mu$ beyond the existing $\beta_3=-7$
witness---without declaring a second kinetic term.

\textbf{Current status.} \textbf{Discharged}: adjoint/fundamental Casimir ratio
$C_A/C_F=9/4$ from the colour chart filter (soft splitting ratio); $\beta_3$ running
to PETRA scale via one-loop QCD identity
(\hyperref[lean:emission-scaffold]{emission scaffold}).  Full $2\to2\to3$ matrix-element
differential distributions remain \textbf{HEP refinements}.

\subsection{Higgs gluon fusion and loop observables}

\textbf{Observed.} LHC $ggH$ inclusive and differential cross sections ($p_T$,
rapidity, fiducial bins) at NLO+NNLO accuracy \cite{pdg2024,peskin1995}.

\textbf{Required discharge.} An explicit patch loop on strong channels coupling to the
Higgs stress slot, with normalisation fixed by the same scale witness as the proton
---predicting $\sigma(ggH)$ or a leading differential moment, not merely assigning
``largest composite weight'' to the channel.

\textbf{Current status.} \textbf{Discharged}: inclusive LO and NLO $ggH$ cross sections
(\hyperref[lean:collider-discharge]{collider discharge module}); \textbf{refinement}:
mean $p_T$ and $p_T>30\,\mathrm{GeV}$ fraction
(\hyperref[lean:hep-refinements]{HEP refinements module}).

\subsection{QGP transport}

\textbf{Observed.} $v_2$, jet quenching $R_{AA}(p_T)$, $\eta/s$, speed of sound
\cite{adams2005,pdg2024}.

\textbf{Required discharge.} A finite-patch transport readout from high-$\xi$
trapping saturation (patch cell mobility $\times$ curvature opacity), producing at
least one benchmark curve (e.g.\ $\eta/s$ order-of-magnitude or a single $R_{AA}$
slope) without fitted transport coefficients.

\textbf{Current status.} \textbf{Discharged}: $\eta/s$
(\hyperref[lean:collider-discharge]{collider discharge module});
\textbf{refinement}: $v_2$ and $R_{AA}(5\,\mathrm{GeV})$
(\hyperref[lean:hep-refinements]{HEP refinements module}).

\subsection{Glueballs and pure strong-channel spectra}

\textbf{Observed.} Lattice QCD glueball masses; candidates such as $f_0(1500)$
\cite{close2014,pdg2024}.

\textbf{Required discharge.} Masses and widths from uniform network binding and
whole-hadron $S^7$ confinement readout on colour-singlet strong-channel networks (no
valence quarks), compared to lattice bands---sharp test because no PDG mass input is
used.

\textbf{Current status.} \textbf{Discharged}: $0^{++}$ and $2^{++}$ glueball mass
witnesses from triple-budget $\times\gamma$ and whole-hadron $S^7$ dressing
(\hyperref[lean:collider-discharge]{collider discharge module},
\hyperref[lean:s7-glueball-track]{glueball research track}); witness matches lattice
bands within $2\sigma$.

\subsection{PDFs and global DIS fits}

\textbf{Observed.} Thousands of DIS/DY/collider points constraining $f(x,Q^2)$
\cite{pdg2024,peskin1995}.

\textbf{Required discharge.} Derive $x$- and $Q^2$-dependence of composite-trace
shell weights $\alpha_{\mathrm{eff}}(m(\xi))$ and show it reproduces a subset of
global PDF moments (e.g.\ $\int_0^1 x f_g(x,Q_0^2)\,\mathrm{d}x$) or a single
structure-function ratio---not merely label PDFs as a continuum dictionary.

\textbf{Current status.} \textbf{Discharged} for first moment and $x$-shape
(\hyperref[lean:collider-discharge]{collider discharge module} and
\hyperref[lean:hep-refinements]{HEP refinements module}: $xf_g$ at $x=0.1$, ratio
$f(0.1)/f(0.01)$).

\subsection{What this note \emph{does} discharge (scope boundary)}

Readers should not conflate the following \textbf{discharged} items with optional
\textbf{future} HEP extensions (e.g.\ full Pythia comparison, global PDF fits to
thousands of DIS points):
\begin{itemize}[nosep]
\item trapped-Casimir factorisation and uniform $E_{\mathrm{bind}}$ (nucleon/HEP mass spine);
\item unified curvature log kernel $K(x,c)=1+c\alpha\log x$ on Hopf/ladder coordinates;
\item abelian strong kinetic + four-channel mask + optional $\mathfrak{su}(3)$ chart Lie law;
\item $\beta_3=-7$ and standard one-loop QCD identity at $N_c=3$, $n_f=6$;
\item colour Casimirs $C_A=3$, $C_F=4/3$, $C_A/C_F=9/4$ and PETRA three-jet
  \cite{bartel1979,berger1979} emission-step
  budget (\hyperref[lean:emission-scaffold]{emission scaffold});
\item per-step emission weights summing to $E_{\mathrm{bind}}$; bundled phenomenology certificate
  (\hyperref[lean:gluon-closure]{ontology certificate});
\item quantitative collider discharge (\hyperref[lean:collider-discharge]{collider module}):
  $R_{23}$, $\langle T\rangle$, $C_A/C_F$, $\sigma(ggH)$, $\eta/s$, glueball masses,
  PDF gluon moment (witness in \hyperref[lean:scripts-bundle]{scripts bundle});
\item HEP refinements (\hyperref[lean:hep-refinements]{refinements module}): MC shower,
  thrust distribution, $ggH$ $p_T$, QGP $v_2$/$R_{AA}$, PDF $x$-shape (witness in
  \hyperref[lean:scripts-bundle]{scripts bundle}).
\end{itemize}
Those results fix \emph{what ``gluon'' means} on the carrier, \emph{how binding
works at lock-in}, and \emph{leading collider/medium observables} without importing
PDG mass inputs into the prediction spine.  Closing differential bins and event
generators is explicitly \textbf{downstream} in the HEP programme.

\section{Implications and claim boundaries}
\label{sec:boundaries}

The framework therefore contains: one abelian gauge theory on the octonion-indexed
carrier; one geometric binding mechanism (trapped Casimir $\times$ composite-trace
selection on $\mathfrak{so}(8)$); and one curvature primitive running every scale
with cosmic temperature. The three Standard-Model gauge factors arise as natural
algebra projections; the ``gluon'' is the continuum shadow of strong-channel binding
geometry.

\textbf{Deliberately unclaimed:} dynamical confinement as a phase transition in an
independent non-abelian gauge theory; a gluon condensate order parameter; asymptotic
freedom derived from a separate $\beta$-function; full PETRA differential cross
sections and angular bins \cite{bartel1979,berger1979,peskin1995},
parton showers, DIS coefficient functions, or QGP transport coefficients (see
\S\ref{sec:qcd-comparison}). Confinement, in the present ontology,
means that promoting a strong-channel excitation above the inner trapping depth set by
$\omega K(\xi)$ is geometrically disfavoured---analogous to outer-Casimir suppression
of the neutral singlet extension.

The last independent dimensionful strong-sector input beyond discrete lattice
combinatorics and the ``now'' slice on the temperature ladder has therefore been
removed \emph{within the proved ontology spine summarised here and in the upstream
trapped-Casimir and nucleon readout records}. Leading collider and medium observables
(\S\ref{sec:quantitative-gap}) are \textbf{discharged} at the structural/LO level;
third-party event-generator bin-by-bin validation remains \textbf{downstream}. Continuum
``gluon'' vocabulary names strong-channel trapping geometry on the proved spine;
electroweak and gravitational readouts on the same patch carrier are established in
sibling notes, not re-derived here.

\appendix

\section{Lean module index}
\label{app:lean-index}

\begin{leanindex}

\item[\phantomsection\label{lean:force-sector}Force-sector map.]
  \leanfile{Hqiv/Physics/Forces.lean}\\
  \leanid{O_component_to_sector}; \leanid{ForceSector.Strong}.

\item[\phantomsection\label{lean:strong-kinetic}Strong kinetic projection.]
  \leanfile{Hqiv/Physics/Forces.lean};
  SM Lagrangian bridge in \leanfile{Hqiv/Physics/DerivedGaugeAndLeptonSector.lean}.

\item[\phantomsection\label{lean:inner-outer-casimir}Inner--outer Casimir scale.]
  \leanfile{Hqiv/Physics/HopfShellBeltramiMassBridge.lean}\\
  \leanid{effective_casimir_scale_at_xi};
  uses \leanid{omegaK_xi} from \leanfile{Hqiv/Physics/ContinuousXiPath.lean}.

\item[\phantomsection\label{lean:curvature-primitive}Curvature primitive.]
  \leanfile{Hqiv/Physics/ContinuousXiCoupling.lean}\\
  \leanid{continuousCurvaturePrimitive}, \leanid{omegaKContinuous}.\\
  Discrete sampling: \leanid{alphaEffAtShell}, \leanid{oneOverAlphaEffAtShell} in
  \leanfile{Hqiv/Physics/SM_GR_Unification.lean}.

\item[\phantomsection\label{lean:alpha-gamma}Lattice $\alpha$-ratio identity.]
  \leanfile{Hqiv/Geometry/OctonionicLightCone.lean},
  \leanfile{Hqiv/Geometry/AlphaGammaForcedByLattice.lean}:
  $\alpha=3/5$, $\gamma=2/5$.

\item[\phantomsection\label{lean:trapped-casimir-bridge}Trapped Casimir binding bridge.]
  \leanfile{Hqiv/Physics/TrappedCasimirBindingBridge.lean}\\
  \leanid{trappedCasimirCouplingCell_eq_alphaEffAtShell}\\
  \leanid{bindingCouplingAtShell_eq_trappedEnergy_quarter_normalizedSelection}\\
  \leanid{E_bind_from_network_eq_sum_trappedCasimirCells}.

\item[\phantomsection\label{lean:binding-network}Uniform network binding.]
  \leanfile{Hqiv/Physics/BoundStates.lean}\\
  \leanid{E_bind_from_network}, \leanid{E_bind_QCD_from_network},
  \leanid{E_bind_nuclear_from_network}, \leanid{E_bind_from_composite_trace}.\\
  Witness numerics: \leanfile{Hqiv/Physics/DerivedNucleonMass.lean},
  \leanfile{Hqiv/Physics/QuarkMetaResonance.lean}.

\item[\phantomsection\label{lean:inner-contact-witness}Inner contact trapped Casimir witness.]
  Phase-lift contact torsion and three-shell inner contact witness
  (\leanfile{Hqiv/Physics/TrappedCasimirBindingBridge.lean}):\\
  \leanid{hopfTrappedSelectionFromShell};\\
  heavy-chart bundle discharged in
  \hyperref[lean:gluon-closure]{gluon curvature ontology certificate}.

\item[\phantomsection\label{lean:hadron-readout}Hadron mass readout.]
  \leanfile{Hqiv/Physics/HadronMassReadout.lean}\\
  \leanid{E_bind_from_composite_trace};
  \leanid{hadronIntrinsicScale_meson_eq_four_ninths_squared}.

\item[\phantomsection\label{lean:kernel-bridge}Strong-sector curvature log kernel.]
  \leanfile{Hqiv/Physics/GluonCurvatureArtifact.lean}\\
  \leanid{curvatureLogAmplification};
  \leanid{hopfContactLogInnerArgument}, \leanid{ladderLogArgumentAtChartShell};\\
  structure \leanid{StrongSectorCurvatureKernelBridge};\\
  bundled \leanid{strongSectorKernelBridge_generation_zero},
  \leanid{strongSectorKernelBridge_generation_one},
  \leanid{strongSectorCurvatureKernelBridge_heavy_chart};\\
  monotonicity \leanid{hopfKernel_lt_ladderKernel_of_contact_lt_ladder}.

\item[\phantomsection\label{lean:collider-discharge}Collider phenomenology discharge.]
  \leanfile{Hqiv/Physics/StrongSectorColliderDischarge.lean}\\
  \leanid{alphaStrongRunningLO}, \leanid{nonAbelianSplittingFromFilter},
  \leanid{petraR23Discharge}, \leanid{meanThrustDischarge},
  \leanid{partonShowerStepWeight}, \leanid{ggHSigmaPbLODischarge},
  \leanid{ggHSigmaPbNLODischarge}, \leanid{qgpEtaOverSDischarge},
  \leanid{glueball0ppMassMeV}, \leanid{glueball2ppMassMeV},
  \leanid{pdfGluonFirstMomentDischarge};\\
  bundled \leanid{gluonCurvatureFullDischarge}.\\
  Witness: \leanid{python3 scripts/hqiv_strong_sector_collider_discharge.py --strict}.

\item[\phantomsection\label{lean:hep-refinements}HEP collider refinements.]
  \leanfile{Hqiv/Physics/HepColliderRefinements.lean}\\
  \leanid{thrustDistributionWidth}, \leanid{partonShowerEmissionProbability},
  \leanid{partonShowerStepDamp}, \leanid{ggHpTFalloffGeV},
  \leanid{qgpV2Discharge}, \leanid{qgpRAAWeightAtPT}, \leanid{pdfGluonShapeAtX};\\
  bundled \leanid{hepColliderRefinementsDischarged}.\\
  Witness: \leanid{python3 scripts/hqiv_hep_collider_refinements.py --strict}.

\item[\phantomsection\label{lean:scripts-bundle}Reproducer bundle.]
  \leanid{papers/gluon_curvature_artifact/scripts.zip}\\
  Refresh: \leanid{python3 scripts/bundle_gluon_curvature_scripts.py}.\\
  Checksums: \leanid{scripts/MANIFEST.sha256}.

\item[\phantomsection\label{lean:emission-scaffold}Strong-channel emission scaffold.]
  \leanfile{Hqiv/Physics/StrongChannelEmissionScaffold.lean}\\
  \leanid{colourCasimirFundamental_three},
  \leanid{colourCasimirAdjoint_over_fundamental_three};\\
  \leanid{strongOctonionComponents_card} ($4$ channels);
  \leanid{beta_3_eq_standardQcd_oneLoop_nc3_nf6};\\
  \leanid{petraThreeJet_requires_emissionStep};
  \leanid{singleGeneratorEmissionWeight_eq_weight_times_lattice_trappedCell};\\
  \leanid{hepJetSlotFromEmissionSteps};\\
  bundled \leanid{strongSectorEmissionScaffoldDischarged},
  \leanid{gluonCurvaturePhenomenologyDischarged}.

\item[\phantomsection\label{lean:gluon-closure}Gluon curvature ontology certificate.]
  \leanfile{Hqiv/Physics/GluonCurvatureArtifact.lean}\\
  \leanid{gluonCurvatureOntologyDischarged};\\
  heavy-chart contact--ladder ordering certificate.\\
  Build: \leanid{lake build paper_gluon_curvature}.

\item[\phantomsection\label{lean:su3-chart-law}SU(3) chart Lie law (optional build).]
  \leanfile{Hqiv/Physics/StrongColorSu3LieChartLaw.lean}\\
  \leanid{colorHalfGellMannFull_lieBracket_eq_I_smul_f_sum}.\\
  Build: \leanid{lake build HQIVStrongColorSu3Certificate}.

\item[\phantomsection\label{lean:beta3-witness}Strong $\beta_3$ witness.]
  \leanfile{Hqiv/Physics/SM_GR_Unification.lean}: \leanid{beta_3 = -7}.\\
  One-loop QCD identity: \leanfile{Hqiv/Physics/StrongChannelEmissionScaffold.lean}\\
  \leanid{beta_3_eq_standardQcd_oneLoop_nc3_nf6}.

\item[\phantomsection\label{lean:s7-glueball-track}Whole-hadron $S^7$ + glueball research track.]
  \leanfile{Hqiv/Physics/HadronS7ConfinementReadout.lean}\\
  \leanid{hadronIjkSortedTripleBudget}; \leanid{hadronWholeS7IjkDressing}.

\item[\phantomsection\label{lean:hep-decay}HEP decay readout (downstream).]
  \leanfile{Hqiv/Physics/HepDecayReadout.lean}:
  CKM slot squares, chiral factors, weak-width reuse.

\end{leanindex}

\clearpage
\bibliographystyle{plain}
\bibliography{bibliography-abbrev}

\end{document}
