\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\usepackage[nameinlink,capitalise]{cleveref}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{xurl}
\urlstyle{same}

\newcommand{\fitbox}[1]{\begin{adjustbox}{max width=\linewidth,center}#1\end{adjustbox}}

\newcommand{\etap}{\ensuremath{\eta_{\mathrm{paper}}}}
\newcommand{\etaBBN}{\ensuremath{\eta_{\mathrm{BBN}}}}
\newcommand{\HQIVbox}[1]{%
  \medskip
  \noindent\setlength{\fboxsep}{6pt}%
  \fbox{\begin{minipage}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
  \sloppy\small\textbf{HQIV in five sentences.}\ #1
  \end{minipage}}%
  \medskip}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\footnotesize\url{#1}}}
\newcommand{\leanid}[1]{\nolinkurl{#1}}

\title[BBN light elements from network weights]{Big-Bang Nucleosynthesis from HQIV Network Weights:\\
Lock-In $\eta$, Epoch Ladder, and Integrated Witnesses in the Lean Library}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Published v1 --- Zenodo \href{https://doi.org/10.5281/zenodo.20723606}{10.5281/zenodo.20723606} (2026-06-16)}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
\noindent\textbf{Position in the publication chain.}
Published Tier-2 \#8 (Zenodo
DOI \href{https://doi.org/10.5281/zenodo.20723606}{10.5281/zenodo.20723606}).
This is the eighth paper in the HQIV preprint series and the
fifth Tier-2 closure note. Tier~0--1 fix the discrete spine
\cite{ettinger2026hqiv,hqiv-lightcone-paper,hqiv-so8-paper,hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper,hqiv-thermo-arrow-paper};
Tier~2 supplies the published mass lock-in synthesis, baryogenesis normalisation, and nucleon
composite-trace binding \cite{hqiv-tuft-sm-lagrangian-paper,hqiv-baryogenesis-paper,hqiv-nucleon-binding-paper}.
The present note states what the BBN Lean stack \emph{actually proves} about
primordial light-element readouts when lock-in $\eta$, nucleon masses, and
composite-trace binding weights are already fixed upstream.

\noindent\textbf{Result.}
At lock-in the library defines an $8\times8$ network readout for
$Y_p$, $D/H$, $^3\mathrm{He}/H$, and $^7\mathrm{Li}/H$ from valley
counts, Boltzmann weights, and cluster binding $Q$ values tied to
\cite{hqiv-tuft-sm-lagrangian-paper}. A separate epoch module places the BBN
window on the temperature shell ladder between QCD lock-in and the CMB
today; an integrated witness certifies $Y_p$ in the observed band and
within $1\%$ of the freeze-out partition value at the same $\eta$.
Machine-checked vital bundles aggregate positivity, ladder ordering, and
$\eta$ consistency (\hyperref[lean:bbn-full-vital]{full BBN vital readout}).
Reader-friendly anchors with clickable Lean links are in
Appendix~\ref{app:lean-index}.

\noindent\textbf{Single absolute anchor (strictly isolated).}
Dimensionful nucleon masses enter only through the proton lock-in witness
at $\mathrm{referenceM}=4$; lock-in \etap{} is imported from the baryogenesis
witness layer, not refit here. Coc et al.\ semi-analytic
fits \cite{coc2015} are labelled \emph{comparison-only} in Lean and are
never HQIV inputs.

\noindent\textbf{Patch-ontology reader contract.}
Per the patch-theory reader contract below, BBN is a discrete-epoch
readout on the shell--temperature map, not a continuum-limit completion
programme. Full PRIMAT-style reaction networks, Li-problem resolution, and
sphaleron/chemical-potential transport remain comparison-layer or future
work.

\noindent\textbf{Status.}
Proved: network vital readout, epoch vital readout, integrated $Y_p$
witness pack, umbrella \texttt{bbn\_full\_vital\_readout}, and the
stoichiometric integrator spine (\hyperref[lean:bbn-stoichiometric]{BBN
stoichiometric integrator}). Python reproduction at lock-in
$\eta_{\mathrm{paper}}$ yields $Y_p\approx 0.244$, $D/H\approx
2.51\times 10^{-5}$ ($z\approx -0.5$ vs.\ Coc \cite{coc2015}), $^3\mathrm{He}/H\approx
1.01\times 10^{-5}$, and $^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$
(pre-astrophysical-depletion; $\approx 45\%$ below Coc semi-analytic
$4.65\times 10^{-10}$ at the same $\eta$; within the Spite-plateau envelope
$(1.6$--$4.5)\times 10^{-10}$; \texttt{data/bbn\_integrator.json}). Not claimed: derivation of
\etap{} from first principles in this note.
\end{abstract}
\maketitle

\input{../include/patch_theory_messaging.tex}

\HQIVbox{%
(1)~Cosmology is encoded on a discrete null-shell ladder
$m(T)=\lfloor T_{\mathrm{Pl}}/T\rfloor-1$, not as a continuum mesh awaiting a
limit. (2)~Matter and gauge couplings read from an inner--outer Casimir chart
$\kappa_6(\xi)=\eta_{\mathrm{local}}\,\gamma\,C_2(\xi)$ tied to the Hopf--Beltrami
mass bridge \cite{hqiv-tuft-sm-lagrangian-paper}. (3)~Nucleon masses and cluster
binding $Q$ follow the eight-dimensional composite trace with a single proton
anchor at $\mathrm{referenceM}=4$. (4)~Baryon asymmetry \etap{} is fixed upstream
by curvature lock-in \cite{hqiv-baryogenesis-paper}, with
$\etaBBN=\etap$ proved in Lean. (5)~BBN abundances are \emph{weighted network
readouts} on that ladder---this note certifies what is already theorem-level and
documents the faithful Python integrator that closes $Y_p$, $D/H$, $^3$He/H, and
$^7$Li/H at \etap.}

\paragraph{Notation.}
We write \etap{} for the lock-in baryon-to-photon ratio imported from
baryogenesis; Lean proves $\etaBBN=\etap$ (no separate BBN fit).
Deuteron, helium, and lithium binding enter as $Q_D$, $Q_{^4\mathrm{He}}$,
$Q_{^3\mathrm{He}}$, $Q_{^7\mathrm{Be}}$, $Q_{^7\mathrm{Li}}$ from the
composite-trace network unless a temperature-dependent release suffix
$Q^{\mathrm{eff}}(T)$ is explicit. Strong-channel weight
$w_{\mathrm{strong}}=4/8$ and horizon ratio $\gamma=2/5$ are HQIV constants,
not BBN fit parameters.

\section{Scope and claim discipline}
\label{sec:scope}

The HQIV programme ties discrete null-shell counting, horizon ratios, and
octonionic-sector packaging \cite{ettinger2026hqiv,hqiv-so8-paper}. The
baryogenesis note \cite{hqiv-baryogenesis-paper} fixes how
$\eta(n;N)$ shares the same normalisation as $\Omega_k(n;N)$ at lock-in.
\textbf{This note answers: given that normalisation, what light-element
statements are already theorems in Lean, and what remains interpretive or
comparison-layer physics?}

\begin{enumerate}[leftmargin=2em]
\item \textbf{Proved in Lean (network spine).}
  Valley counts for $A=2,3,4$; cluster binding $Q$ for deuterium and
  $^4\mathrm{He}$ aligned with composite-trace nucleon binding; neutron
  fraction and $Y_p=2x_n/(1+x_n)$ in $(0,1)$; $\eta$ at BBN equals
  $\eta_{\mathrm{paper}}$ (the \hyperref[lean:eta-bbn-eq]{BBN--$\eta$ identity});
  positivity of $D/H$, $^3\mathrm{He}/H$, $^7\mathrm{Li}/H$ readouts at
  lock-in (the \hyperref[lean:bbn-network-vital]{network vital readout}).
\item \textbf{Proved in Lean (epoch spine).}
  Shell index $m(T)=\lfloor T_{\mathrm{Pl,MeV}}/T\rfloor-1$; BBN mid-epoch
  shell strictly above lock-in and below today (the
  \hyperref[lean:bbn-mid-shell-order]{mid-epoch shell ordering});
  epoch readout positivity at $T\approx 0.1\,\mathrm{MeV}$ (the
  \hyperref[lean:bbn-epoch-vital]{epoch vital readout}).
\item \textbf{Proved in Lean (integrated witness).}
  Freeze-out $Y_p$ in the Cooke-style observed band; integrated $Y_p$
  within $1\%$ of freeze-out at exported numerics (the
  \hyperref[lean:bbn-integrated-vital]{integrated witness vital readout}).
\item \textbf{Comparison layer only.}
  Coc et al.\ \cite{coc2015} semi-analytic $Y_p(\eta,N_\nu,\tau_n)$ and
  $D/H(\eta)$ in the \hyperref[lean:bbn-umbrella]{BBN umbrella module}---not
  derived from HQIV weights.
\item \textbf{Not claimed.}
  Full 400-reaction PRIMAT network in Lean; Li-problem solution; independent
  refit of $\eta$ or nucleon masses; continuum transport of weak rates.
\end{enumerate}

\section{Upstream inputs consumed at lock-in}
\label{sec:inputs}

The BBN stack does not introduce new dimensionful anchors. It consumes:

\begin{itemize}[leftmargin=2em]
\item \textbf{Baryon asymmetry.}
  $\eta_{\mathrm{paper}}$ from the baryogenesis witness chain
  \cite{hqiv-baryogenesis-paper}, with
  $\eta_{\mathrm{BBN}}=\eta_{\mathrm{paper}}$ proved in the
  \hyperref[lean:bbn-network]{network-from-weights module}.
\item \textbf{Nucleon masses.}
  $\texttt{derivedProtonMass}$ and $\texttt{derivedDeltaM}$ (neutron--proton
  gap) from the mass-spectrum lock-in discipline
  \cite{hqiv-tuft-sm-lagrangian-paper,hqiv-kirchhoff-paper}; binding shell
  $\texttt{bbnBindingShell}=\mathrm{referenceM}$.
\item \textbf{Composite-trace binding.}
  Deuterium and $^4\mathrm{He}$ reaction $Q$ values identified with
  cluster binding from the eight-dimensional composite trace (the
  \hyperref[lean:bbn-cluster-binding]{cluster-binding alignment theorems}).
\end{itemize}

Laboratory abundances and semi-analytic fits are \emph{comparison targets}
for scripts and narrative; they are not simultaneous fit rows in the proof
graph.

\section{Network readout from valley weights}
\label{sec:network}

The \hyperref[lean:bbn-network]{network-from-weights module} encodes light
elements as a weighted readout on the same Fano/network picture used for
nucleon binding \cite{hqiv-tuft-sm-lagrangian-paper}:

\begin{itemize}[leftmargin=2em]
\item \textbf{Valley counts.}
  $A=2\mapsto 2$, $A=3\mapsto 4$, $A=4\mapsto 6$ (helium-4 valley).
\item \textbf{Boltzmann factor.}
  $\exp(-Q/T)$ at temperature $T$ (MeV) for each channel $Q$.
\item \textbf{Neutron sector.}
  $x_n=n/p$ from $\Delta m$ and $T$; $Y_p=2x_n/(1+x_n)$.
\item \textbf{Light ions.}
  $D/H\propto \eta_{10}^{-\alpha_D}\times$ thermal sink $\times$
  Boltzmann weight; analogous exponent slots for $^3\mathrm{He}$ and
  $^7\mathrm{Li}$.
\end{itemize}

At lock-in temperature the readout packages
$(\eta,Y_p,D/H,{}^3\mathrm{He}/H,{}^7\mathrm{Li}/H)$ with
$\eta=\eta_{\mathrm{paper}}$. The
\hyperref[lean:bbn-network-vital]{network vital readout} theorem bundles
positivity and the $\eta$ identity---this is the certificate-level content
of ``BBN from network weights,'' not a hand-tuned abundance table.

\begin{remark}[Epoch dependence]
Partition formulas evaluated at lock-in temperature are \emph{not} the same
as relic abundances after integration over the BBN window; Section~\ref{sec:epoch}
and Section~\ref{sec:integrated} separate partition readouts from
cooling-integrated witnesses.
\end{remark}

\section{BBN as a cosmological epoch on the temperature ladder}
\label{sec:epoch}

The \hyperref[lean:bbn-epoch-evolution]{epoch evolution module} maps
temperature to shell index
\[
  m(T)+1 = T_{\mathrm{Pl,MeV}}/T,
\]
with a QCD onset shell, lock-in shell $m_{\mathrm{lockin}}=\mathrm{referenceM}$,
BBN window $T\in[0.01,1]\,\mathrm{MeV}$, and CMB today as the cold end of
the ladder. Proved ordering:
\[
  m_{\mathrm{lockin}} < m_{\mathrm{BBN,mid}} < m_{\mathrm{now}},
\]
where $m_{\mathrm{BBN,mid}}$ corresponds to $T\approx 0.1\,\mathrm{MeV}$.
Freeze-out internal temperature is defined consistently with neutron--proton
equilibrium break (the \hyperref[lean:bbn-freezeout-temp]{freeze-out
temperature identity}).

\begin{remark}[Discrete shells and freeze-out sharpness]
Because $m(T)$ uses a floor map, the temperature ladder is piecewise constant
in shell index: a cooling step crosses a shell boundary when $T$ drops below
$T_{\mathrm{Pl}}/(m+2)$.  Freeze-out is therefore a \emph{shell predicate}
($\tau_{n\!\leftrightarrow\! p}\sim t_{\mathrm{exp}}$ at the matching shell),
not a smooth continuum crossover.  The integrator uses log-spaced sub-steps
inside the BBN window to approximate the cooling history; residual
$\mathcal{O}(\Delta m)$ artifacts near boundaries are absorbed into the
documented $D/H$ residual rather than hidden by a continuum transport code.
\end{remark}

The \hyperref[lean:bbn-epoch-network]{epoch network module} supplies a
rate scaffold (weak, $np$, $dd$, $^3\mathrm{He}$ branches) and Hubble
dilution---schematic reaction bookkeeping for integration witnesses, not a
published PRIMAT clone in Lean.

\subsection{Dynamic $C_2$ lapse clock (BBN lab)}
\label{sec:dynamic-c2}

\paragraph{Physical picture.}
In standard BBN one often parametrizes departures from freeze-out equilibrium
via $H(T)$, $\tau_n$, or effective reaction rates.  HQIV instead advances
species by a \textbf{shell reaction opportunity}
$\mathcal{O}\propto \Delta\log\xi\times B_{\mathrm{curv}}(\xi)$, where
$\xi=T_{\mathrm{Pl}}/T$ is the cosmological shell coordinate
\cite{hqiv-baryogenesis-paper}.  The Hopf--Beltrami mass bridge packages the
matter slot as $\kappa_6(\xi)=\eta_{\mathrm{local}}\,B_{\mathrm{curv}}(\xi)\,
\gamma\,C_2(\xi)$ (\hyperref[lean:bbn-dynamic-c2]{Hopf--Beltrami mass bridge}).
On the hot BBN chart $B_{\mathrm{curv}}$ is nearly flat while lapse
concentration $C_2(\xi)$ grows with $\xi$; the anchor-normalized ratio
$\kappa_{6,\mathrm{ref}}/\kappa_6(T)$ therefore acts as a \emph{local lapse
reaction clock}: it is \textbf{not} a variable speed of light in the metric
sense, but an effective rescaling of strong-channel opportunity relative to the
freeze-out anchor temperature $T_{\mathrm{freeze}}(\eta)$.  In Friedmann
language, HQIV does not import $H(T)$ as the integration clock; the
dictionary is ``comoving temperature step $\Delta\log T$'' $\leftrightarrow$
``shell opportunity $\Delta\log\xi$'' modulated by $(\kappa_{6,\mathrm{ref}}/
\kappa_6)^{w_{C_2}(T)}$ in the deuterium bottleneck.  This is closest to a
\textbf{temperature-dependent effective rate} in the bottleneck window
$T\le T_{\mathrm{bn}}(\eta)$, with exponent tied to composite-trace binding
release rather than a fitted MeV cutoff.

\paragraph{Definitions.}
For $T\le T_{\mathrm{bn}}(\eta)=\gamma\,w_{\mathrm{strong}}\,T_{\mathrm{freeze}}(\eta)$,
\[
\begin{aligned}
w_{C_2}(T) &= \gamma\,w_{\mathrm{strong}}\,\frac{Q_D^{\mathrm{eff}}(T)}{Q_{np}}, \\
f_{C_2}(T) &= \left(\frac{\kappa_{6,\mathrm{ref}}}{\kappa_6(T)}\right)^{w_{C_2}(T)}, \\
\kappa_6(T) &= \etap\,B_{\mathrm{curv}}(\xi(T))\,\gamma\,C_2(\xi(T)).
\end{aligned}
\]
Lean names: \leanid{bbnDynamicC2BottleneckT\_MeV},
\leanid{bbnDynamicC2LapseExponent},
\leanid{bbnDynamicC2OpportunitySuppression}
(\hyperref[lean:bbn-dynamic-c2]{DynamicBBNBaryogenesis}).
Multiplying opportunity by raw $\kappa_6$ at low $T$ \emph{over-suppresses}
$D/H$; the inverse lapse ratio is the certified dynamic-$C_2$ readout.

The legacy dynamic-bulk integrator (\texttt{hqiv\_dynamic\_bulk\_bbn.py})
implements this opportunity clock without the synthesis-window stoichiometry of
\Cref{sec:faithful-integrator}.
\paragraph{Provider decomposition at \etap.}
\Cref{tab:provider-decomposition} quantifies how each layer raises $D/H$ toward
the faithful readout (exported by \texttt{hqiv\_bbn\_paper\_tables.py}).
Stoichiometric $n\!+\!p\!\to\!D$ capture and the synthesis window
($1\,\mathrm{MeV}\to T_{\mathrm{bn}}$) dominate ($4.4\times10^{-6}\to
1.84\times10^{-5}$); coupled inventory with outside-neutron curvature supplies
the final increment to $2.51\times10^{-5}$ and sets $Y_p$ via the
post-synthesis $x_n$ survivor.  Turning off $C_2$ lapse alone leaves the
faithful $D/H$ unchanged at this precision---$f_{C_2}$ is booked primarily on
the $n\!+\!p$ and $D\!+\!D$ branches inside the bottleneck, not as a separate
post-hoc multiplier.

\begin{table}[ht]
\centering
\caption{Layered $D/H$ providers at $\etap=6.10\times10^{-10}$
(\texttt{data/bbn\_paper\_tables.json}).}
\label{tab:provider-decomposition}
\begin{tabular}{lcc}
\toprule
Provider layer & $D/H$ & $Y_p$ \\
\midrule
Lock-in $Q_D$ at $T_{\mathrm{freeze}}$ & $\lesssim 10^{-30}$ & --- \\
Hybrid $Q^{\mathrm{eff}}(T)$, partition window & $4.4\times 10^{-6}$ & $0.258$ \\
+ stoichiometric + $C_2$, full window & $8.1\times 10^{-6}$ & $0.247$ \\
+ synthesis window (uncoupled) & $1.84\times 10^{-5}$ & $0.247$ \\
Faithful (coupled inventory + curvature) & $2.51\times 10^{-5}$ & $0.244$ \\
\midrule
Coc et al.\ comparison \cite{coc2015} & $(2.53\pm0.04)\times10^{-5}$ & $0.244$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Faithful stoichiometric integrator}
\label{sec:faithful-integrator}

{\sloppy
The recommended reproduction path is \Cref{app:minimal-repro}
(\texttt{hqiv\_bbn\_integrator.py}), mirrored in Lean by the
\hyperref[lean:bbn-stoichiometric]{stoichiometric integrator module}
and \texttt{data/bbn\_integrator.json}.
At $\eta_{\mathrm{paper}}=6.10\times 10^{-10}$ it couples:
}

\begin{itemize}[leftmargin=2em]
\item \textbf{Synthesis-window $D/H$.}
  Log-weighted integration from $1\,\mathrm{MeV}$ through the $C_2$
  bottleneck ($T_{\mathrm{bn}}\approx 0.14\,\mathrm{MeV}$), not the full
  cold tail to $0.01\,\mathrm{MeV}$.  Below the $^3\mathrm{He}$ synthesis mid-gate
  ($T_{\mathrm{peak}}=0.18\,\mathrm{MeV}$, Lean
  \leanid{bbnHe3SynthesisMidMeV}) the log weights taper linearly
  $T_{\mathrm{bn}}\!\to\!T_{\mathrm{peak}}$ via
  \leanid{bbnSynthesisDWindowTailGate}---a thermal-crush tail, not a hard
  shut-off at mid-MeV.  The numerator adds an explicit
  $n\!+\!p\!\to\!D$ capture term (\texttt{branch\_np\_to\_d}) and a strong-channel
  deposition boost $\propto \gamma\,w_{\mathrm{strong}}(1+g_{np}+c_2)$
  (\leanid{bbnSynthesisStrongDepositionBoost}).
\item \textbf{Coupled neutron inventory.}
  Outside-curvature and $\nu$-opacity set $\tau_n(T)$; the same $x_n(T)$
  ladder scales $n\!+\!p\!\to\!D$ and $D\!+\!D\!\to\!{}^4\mathrm{He}$.
  $Y_p$ is read from the \emph{post-synthesis} survivor
  $x_n(T_{\mathrm{low}})$, not the instant freeze-out partition.
\item \textbf{$^3\mathrm{He}/H$ window.}
  Dedicated $0.45\!\to\!0.05\,\mathrm{MeV}$ ladder with trimer-width
  erosion on the strong channel only; a narrower feed gate routes $^3\mathrm{He}$
  into the $^7\mathrm{Be}$ tail without inflating the $^3\mathrm{He}$ readout.
\item \textbf{$^7\mathrm{Li}/H$ tail.}
  $^7\mathrm{Be}$ production and electron capture on
  $0.25\!\to\!0.05\,\mathrm{MeV}$;   witness $^7\mathrm{Li}/H\approx
  2.54\times 10^{-10}$ (pre-depletion HQIV readout).
\end{itemize}

Exported readout: $Y_p\approx 0.244$, $D/H\approx
2.51\times 10^{-5}$ ($z\approx -0.5$ vs.\ $(2.53\pm 0.04)\times 10^{-5}$),
$^3\mathrm{He}/H\approx 1.01\times 10^{-5}$, $^7\mathrm{Li}/H\approx
2.54\times 10^{-10}$ ($\approx 45\%$ below Coc $4.65\times 10^{-10}$).  Curvature relative to bare equilibrium \emph{lowers}
$Y_p$ by $\sim 3\times 10^{-3}$ while \emph{raising} $D/H$ slightly---the
intended unified inventory behaviour, not a Y$_p$-only tweak.

\section{Integrated witnesses and Python export}
\label{sec:integrated}

The \hyperref[lean:bbn-witness]{BBN witness module} pins numeric exports
from \texttt{data/bbn\_witnesses.json} (generated by
\texttt{scripts/hqiv\_bbn\_abundances.py} and
\texttt{scripts/hqiv\_bbn\_epoch\_network.py}; Appendix~\ref{app:scripts}).
At the exported lock-in $\eta=6.10\times 10^{-10}$ (partition readout in
\texttt{data/bbn\_witnesses.json}):

\begin{itemize}[leftmargin=2em]
\item Lock-in cluster $Q$ values from the composite trace:
  $Q_D\approx 2.34\,\mathrm{MeV}$, $Q_{^4\mathrm{He}}\approx 7.03\,\mathrm{MeV}$,
  $Q_{^3\mathrm{He}}\approx 4.39\,\mathrm{MeV}$; post-$\alpha$ wells
  $Q_{^7\mathrm{Be}}\approx 14.9\,\mathrm{MeV}$,
  $Q_{^7\mathrm{Li}}\approx 14.4\,\mathrm{MeV}$ ($^7\mathrm{Be}$ deeper than $^7\mathrm{Li}$).
\item Integrated window $Y_p\approx 0.247$ lies in the Cooke et al.\ band
  checked in Lean (\hyperref[lean:bbn-yp-band]{freeze-out $Y_p$ band witness}).
\item Integrated $Y_p$ is within $1\%$ of the freeze-out partition value
  (\hyperref[lean:bbn-yp-near-freeze]{integrated--freeze-out proximity
  witness}).
\item Faithful stoichiometric integrator at $\eta_{\mathrm{paper}}$:
  $Y_p\approx 0.244$, $D/H\approx 2.51\times 10^{-5}$,
  $^3\mathrm{He}/H\approx 1.0\times 10^{-5}$,
  $^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$
  (\texttt{data/bbn\_integrator.json}; \S\ref{sec:faithful-integrator}).
\item Legacy dynamic-bulk row (\texttt{legacy\_dynamic\_bulk\_row} in
  \texttt{integrator\_lean\_audit.json}): pre-synthesis-window scaffold only;
  superseded by \texttt{faithful\_bbn\_integrator}.
\end{itemize}

\textbf{Honesty on lithium and deuterium.}
The faithful integrator exports $^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$
(pre-astrophysical-depletion) at lock-in $\eta$, $\approx 45\%$ below the Coc
semi-analytic comparison $4.65\times 10^{-10}$ but inside the broad Spite
envelope $(1.6$--$4.5)\times 10^{-10}$, without refitting nuclear $Q$ values.
$D/H$ at
$\eta_{\mathrm{paper}}$ sits within $\sim 0.5\sigma$ of the Coc anchor once
the synthesis-window tail gate (\leanid{bbnSynthesisDWindowTailGate}) is
booked---documented as a structural readout, not a comparison-layer fit.
Mid-epoch partition witnesses in
\texttt{bbn\_witnesses.json} can still diverge on Li; treat those as
epoch-scaffold exports distinct from the stoichiometric integrator.

\section{Relation to standard BBN physics}
\label{sec:standard-bbn}

\paragraph{Neutron lifetime and weak rates.}
Laboratory $\tau_n\simeq 879.4\,\mathrm{s}$ enters as a \textbf{reference
scale} in \hyperref[lean:bbn-outside-neutron]{NuclearOutsideTemperatureDynamics}
and the Python outside-neutron module; it is \textbf{not refit} at BBN.
Instead HQIV books a temperature-dependent ratio
$\tau_n(T)/\tau_{n,\mathrm{lock}}$ from (i)~outside-curvature support on the
free neutron branch and (ii)~relic-neutrino opacity catalysis on the weak width
(same stack as nucleon $\beta$ tipping \cite{hqiv-nucleon-binding-paper}).
At $T_{\mathrm{freeze}}$ the witness exports
$\tau_n(T)/\tau_{n,\mathrm{lock}}\approx 1.19$ and a weak-width factor
$\approx 1.07$ (\texttt{data/bbn\_paper\_tables.json}).  This lengthens the
effective freeze-out epoch slightly and feeds the coupled $x_n(T)$ march that
sets both $Y_p$ and $D/H$.

\paragraph{Effective neutrino species $N_\nu$.}
HQIV does not introduce an explicit $N_\nu$ fit parameter in the proof graph.
The epoch scaffold assumes the Standard Model relativistic neutrino content
implicitly through the thermal $g_\ast$ bookkeeping used in comparison-layer
Coc fits \cite{coc2015}.  A future extension would route $N_\nu$ through the
same relic-$\nu$ opacity slot already present in the outside-neutron module.

\paragraph{Baryon density \etap{} vs.\ Planck/CMB.}
$\etap=6.10\times10^{-10}$ ($\eta_{10}=6.10$) is the Lean lock-in anchor
imported from baryogenesis \cite{hqiv-baryogenesis-paper}, matching the
BBN+$C$MB comparison target $6.10\pm0.06$ recorded there---not a simultaneous
abundance fit.  The dynamic bulk integrator independently derives
$\eta_{10}\approx 6.20$ from $\Omega_b$ and curvature closure; the faithful
BBN readout is evaluated at the proved identity $\etaBBN=\etap$.

\paragraph{The lithium problem.}
Standard SBBN with unmodified rates typically \emph{overpredicts}
$^7\mathrm{Li}/H$ by factors $\sim 2$--$3$ relative to the Spite plateau
once post-BBN destruction is accounted for.  At \etap{} Coc semi-analytics
\cite{coc2015} give $^7\mathrm{Li}/H\approx 4.65\times10^{-10}$; the HQIV
faithful witness is $^7\mathrm{Li}/H\approx 2.54\times10^{-10}$
(pre-astrophysical-depletion), $\approx 45\%$ below that comparison anchor
but inside the broad $(1.6$--$4.5)\times10^{-10}$ envelope, without
refitting $Q$.
\Cref{tab:li7-gates} shows the mechanism: unstratified $^3\mathrm{He}$ routing
inflates $^7\mathrm{Be}$ production by orders of magnitude; the combination of
\textbf{trimer-width erosion} on the $^3\mathrm{He}$ readout and a
\textbf{narrow $^3\mathrm{He}\!\to\!{}^7\mathrm{Be}$ feed gate}
($0.05$--$0.35\,\mathrm{MeV}$) brings the ladder into band.  Composite-trace
$Q_{^7\mathrm{Be}}$ sets the Boltzmann weight; $C_2$ lapse affects $D/H$ more
strongly than $^7\mathrm{Li}$ in the final witness.

\begin{table}[ht]
\centering
\caption{$^7\mathrm{Li}/H$ sensitivity to $^3\mathrm{He}$ stratification gates
at \etap{} (\texttt{bbn\_paper\_tables.json}).}
\label{tab:li7-gates}
\begin{tabular}{lcc}
\toprule
$^3\mathrm{He}$ stratified & synthesis window & $^7\mathrm{Li}/H$ \\
\midrule
off & off & $3.1\times 10^{-7}$ \\
off & on  & $4.4\times 10^{-10}$ \\
on  & off & $1.8\times 10^{-9}$ \\
on  & on  & $2.54\times 10^{-10}$ \\
\bottomrule
\end{tabular}
\end{table}

\section{Quantitative validation and uncertainties}
\label{sec:validation}

\paragraph{$Y_p$ band used in witnesses.}
Lean \hyperref[lean:bbn-yp-band]{freeze-out $Y_p$ band witness} checks against
center $0.244\pm0.004$ (Cooke-style helium band; \cite{coc2015}), \textbf{not}
the wider $0.244$--$0.256$ envelope sometimes quoted from heterogeneous
datasets.  The faithful integrator readout $Y_p\approx 0.2442$ has
$z\approx +0.05$ against that witness; the partition freeze-out witness
$0.2469$ remains inside the same band.

\paragraph{$D/H$ and observational context.}
The model $D/H\approx 2.51\times10^{-5}$ lies within $\sim 0.5\sigma$ of the Coc
semi-analytic anchor $(2.53\pm0.04)\times10^{-5}$ and the high-precision
deuterium measurements of Cooke et al.\ \cite{cooke2014,cooke2018}.  The
log-weight tail below $T_{\mathrm{peak}}=0.18\,\mathrm{MeV}$ (shared with the
$^3\mathrm{He}$ synthesis mid-gate) closes the prior $\sim 6\%$ undershoot that arose from
treating the synthesis window as a hard floor at $0.15\,\mathrm{MeV}$.  Recent
``high-$D$'' interstellar sightlines and galactic chemical evolution (GCE)
depletion corrections remain an active comparison layer: HQIV documents offsets
at \etap{} rather than tuning $\gamma$, $Q_D$, or $\eta$ to match Coc.  At
$\eta_{10}=5.8$ the same integrator reaches $D/H\approx 2.79\times10^{-5}$,
bracketing the Coc value from above.

\paragraph{Sensitivity to HQIV constants.}
\Cref{tab:sensitivity} summarizes one-at-a-time variations around the proved
constants $\gamma=2/5$ and $w_{\mathrm{strong}}=4/8$.  $D/H$ scales roughly
linearly with $\gamma$ and $w_{\mathrm{strong}}$ in the synthesis window;
$Y_p$ is comparatively stable because it is anchored by the post-synthesis
$x_n$ march.

\begin{table}[ht]
\centering
\caption{Sensitivity of faithful readout to $\gamma$ and $w_{\mathrm{strong}}$
(one-at-a-time; \texttt{bbn\_paper\_tables.json}).}
\label{tab:sensitivity}
\begin{tabular}{lccc}
\toprule
Parameter & value & $D/H$ & $Y_p$ \\
\midrule
$\gamma$ & $0.36$ & $1.95\times10^{-5}$ & $0.2447$ \\
         & $0.40$ & $2.51\times10^{-5}$ & $0.2442$ \\
         & $0.44$ & $3.23\times10^{-5}$ & $0.2437$ \\
$w_{\mathrm{strong}}$ & $0.45$ & $1.95\times10^{-5}$ & $0.2447$ \\
                      & $0.50$ & $2.51\times10^{-5}$ & $0.2442$ \\
                      & $0.55$ & $3.23\times10^{-5}$ & $0.2437$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{$\eta$ sweep vs.\ comparison layer.}
\Cref{tab:eta-sweep} lists the faithful integrator across
$\eta_{10}\in[5.8,6.4]$.  Coc semi-analytics \cite{coc2015} are not inputs;
the table is for external calibration only.

\begin{table}[ht]
\centering
\caption{Faithful integrator abundances vs.\ $\eta_{10}$
(\texttt{bbn\_paper\_tables.json}).}
\label{tab:eta-sweep}
\fitbox{%
\small
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lcccc@{}}
\toprule
$\eta_{10}$ & $Y_p$ & $D/H$ & $^3\mathrm{He}/H$ & $^7\mathrm{Li}/H$ \\
\midrule
$5.8$ & $0.254$ & $2.79\times10^{-5}$ & $1.08\times10^{-5}$ & $2.83\times10^{-10}$ \\
$5.9$ & $0.251$ & $2.69\times10^{-5}$ & $1.05\times10^{-5}$ & $2.73\times10^{-10}$ \\
$6.0$ & $0.247$ & $2.60\times10^{-5}$ & $1.03\times10^{-5}$ & $2.63\times10^{-10}$ \\
$\mathbf{6.1}$ & $\mathbf{0.244}$ & $\mathbf{2.51\times10^{-5}}$ & $\mathbf{1.01\times10^{-5}}$ & $\mathbf{2.54\times10^{-10}}$ \\
$6.2$ & $0.241$ & $2.43\times10^{-5}$ & $9.9\times10^{-6}$ & $2.45\times10^{-10}$ \\
$6.3$ & $0.238$ & $2.35\times10^{-5}$ & $9.7\times10^{-6}$ & $2.36\times10^{-10}$ \\
$6.4$ & $0.235$ & $2.27\times10^{-5}$ & $9.5\times10^{-6}$ & $2.28\times10^{-10}$ \\
\bottomrule
\end{tabular}}
\end{table}

\section{Comparison layer: Coc et al.\ semi-analytics}
\label{sec:comparison}

The \hyperref[lean:bbn-umbrella]{BBN umbrella module} records Coc et al.\
(2015) semi-analytic $Y_p(\eta,N_\nu,\tau_n)$ and $D/H(\eta)$ for
side-by-side plots. These definitions are explicitly \textbf{not HQIV
inputs}; they exist so reviewers can see how close the weight readout sits
to standard BBN phenomenology without contaminating the proof graph.
Cooke et al.\ deuterium and helium measurements \cite{cooke2014,cooke2018}
supply independent observational anchors for the same comparison layer.

\section{Umbrella certificate}
\label{sec:umbrella}

The \hyperref[lean:bbn-umbrella]{BBN umbrella module} conjoins network,
epoch, and integrated vital readouts into
\texttt{bbn\_full\_vital\_readout}. This is the single top-level Prop a
downstream cosmology or CMB paper can cite once lock-in $\eta$ and nucleon
witnesses are accepted from earlier tiers.

\section{What this note does not claim}
\label{sec:not-claimed}

\begin{itemize}[leftmargin=2em]
\item Derivation of $\eta_{\mathrm{paper}}$ from associator phases or
  sphaleron yields (baryogenesis interpretive layer).
\item Replacement of PRIMAT/MESA with the HQIV rate scaffold.
\item Full Lean certificate that integrated $D/H$ matches Coc et al.\ within
  stated uncertainty at every $\eta$ (Python witness is within $\sim 0.5\sigma$
  at $\eta_{\mathrm{paper}}$; see \S\ref{sec:faithful-integrator}).
\item Post-BBN Li destruction and galactic chemical evolution (pre-depletion
  $^7\mathrm{Li}/H$ only).
\item Any continuum-limit or patch-refinement programme.
\end{itemize}

\section{Conclusion}

Big-bang nucleosynthesis in HQIV is formalised as a \emph{weighted network
readout} at lock-in, lifted to a \emph{cosmological epoch} on the
temperature--shell ladder, and closed by \emph{integrated abundance witnesses}
tied to Python exports. The proof graph is sorry-free at the vital-bundle
level; observational agreement is claimed only where the witness module
states it ($Y_p$ in the fixed-$\eta$ bundle; light-ion band agreement in the
faithful stoichiometric witness of \S\ref{sec:faithful-integrator}, including
$D/H$ within $\sim 0.5\sigma$ of Coc at \etap{}). Companion reproduction is documented in
Appendix~\ref{app:scripts}.

\section*{Acknowledgments}

Machine-checked proofs live in the public HQIV Lean repository
\cite{hqiv-lean}. BBN witness JSON is regenerated from the scripts named in
Appendix~\ref{app:scripts}.

\appendix

\section{Lean module index}
\label{app:lean-index}

Each entry lists the repository path (audit) and the reader-facing label used
in the body. All modules are in the \texttt{Hqiv.Physics} namespace unless
noted.

\begin{description}[leftmargin=2.5em,style=nextline]
\footnotesize\sloppy
\item \phantomsection\label{lean:bbn-umbrella}\textbf{BBN umbrella}
  \leanfile{Hqiv/Physics/BigBangNucleosynthesis.lean} ---
  Coc comparison defs; \leanid{bbn\_full\_vital\_readout}.
\item \phantomsection\label{lean:bbn-network}\textbf{Network from weights}
  \leanfile{Hqiv/Physics/BBNNetworkFromWeights.lean} ---
  Valley counts, $Y_p$, $D/H$, $\eta$ identity;
  \leanid{bbn\_network\_vital\_readout\_holds}.
\item \phantomsection\label{lean:bbn-epoch-evolution}\textbf{Epoch evolution}
  \leanfile{Hqiv/Physics/BBNEpochEvolution.lean} ---
  Shell ladder, epoch readout;
  \leanid{bbn\_epoch\_vital\_readout\_holds}.
\item \phantomsection\label{lean:bbn-epoch-network}\textbf{Epoch network}
  \leanfile{Hqiv/Physics/BBNEpochNetwork.lean} ---
  Rate scaffold, cooling integration state.
\item \phantomsection\label{lean:bbn-witness}\textbf{BBN witness}
  \leanfile{Hqiv/Physics/BBNWitness.lean} ---
  JSON-linked $Y_p$ witnesses;
  \leanid{bbn\_integrated\_witness\_vital\_holds}.
\item \phantomsection\label{lean:bbn-network-vital}\textbf{Network vital readout}
  \leanid{bbn\_network\_vital\_readout} in
  \leanfile{Hqiv/Physics/BBNNetworkFromWeights.lean}.
\item \phantomsection\label{lean:bbn-epoch-vital}\textbf{Epoch vital readout}
  \leanid{bbn\_epoch\_vital\_readout} in
  \leanfile{Hqiv/Physics/BBNEpochEvolution.lean}.
\item \phantomsection\label{lean:bbn-integrated-vital}\textbf{Integrated witness vital}
  \leanid{bbn\_integrated\_witness\_vital} in
  \leanfile{Hqiv/Physics/BBNWitness.lean}.
\item \phantomsection\label{lean:bbn-full-vital}\textbf{Full BBN vital readout}
  \leanid{bbn\_full\_vital\_readout} in
  \leanfile{Hqiv/Physics/BigBangNucleosynthesis.lean}.
\item \phantomsection\label{lean:eta-bbn-eq}\textbf{BBN--$\eta$ identity}
  \leanid{eta\_bbn\_eq\_eta\_paper} in
  \leanfile{Hqiv/Physics/BBNNetworkFromWeights.lean}.
\item \phantomsection\label{lean:bbn-mid-shell-order}\textbf{Mid-epoch shell ordering}
  \leanid{bbnMidEpochShell\_gt\_lockinShell},
  \leanid{bbnMidEpochShell\_lt\_nowShell} in
  \leanfile{Hqiv/Physics/BBNEpochEvolution.lean}.
\item \phantomsection\label{lean:bbn-freezeout-temp}\textbf{Freeze-out temperature identity}
  \leanid{bbnInternalTemperatureMeV\_eq\_freezeout} in
  \leanfile{Hqiv/Physics/BBNEpochEvolution.lean}.
\item \phantomsection\label{lean:bbn-cluster-binding}\textbf{Cluster-binding alignment}
  \leanid{bbnDeuteronBindingQ\_eq\_clusterBinding},
  \leanid{bbnHelium4BindingQ\_eq\_clusterBinding} in
  \leanfile{Hqiv/Physics/BBNNetworkFromWeights.lean}.
\item \phantomsection\label{lean:bbn-yp-band}\textbf{Freeze-out $Y_p$ band witness}
  \leanid{bbnYpFreezeoutWitness\_in\_observed\_band} in
  \leanfile{Hqiv/Physics/BBNWitness.lean}.
\item \phantomsection\label{lean:bbn-yp-near-freeze}\textbf{Integrated--freeze-out proximity}
  \leanid{bbnIntegratedYp\_near\_freezeout\_witness} in
  \leanfile{Hqiv/Physics/BBNWitness.lean}.
\item \phantomsection\label{lean:nuclei}\textbf{Nuclear valley scaffold}
  \leanfile{Hqiv/Physics/HQIVNuclei.lean} ---
  Isotope valley counts feeding BBN weights.
\item \phantomsection\label{lean:eta-paper}\textbf{Lock-in $\eta$ paper module}
  \leanfile{Hqiv/Physics/BaryogenesisEtaPaper.lean} (upstream).
\item \phantomsection\label{lean:derived-mass}\textbf{Derived nucleon mass}
  \leanfile{Hqiv/Physics/DerivedNucleonMass.lean} (upstream).
\item \phantomsection\label{lean:bbn-dynamic-c2}\textbf{BBN dynamic-$C_2$ pack}
  \leanfile{Hqiv/Physics/DynamicBBNBaryogenesis.lean} ---
  \leanid{bbnDynamicC2OpportunitySuppression},
  \leanid{bbnShellReactionOpportunity\_dynamic\_integrator};
  \leanfile{Hqiv/Physics/HopfShellBeltramiMassBridge.lean} ---
  \leanid{tuftLapseConcentrationAtXi}, \leanid{tuftHopfKappa6AtXi}.
\item \phantomsection\label{lean:bbn-stoichiometric}\textbf{BBN stoichiometric integrator}
  \leanfile{Hqiv/Physics/BBNStoichiometricIntegrator.lean} ---
  synthesis-window $D/H$ with \leanid{bbnSynthesisDWindowTailGate},
  \leanid{bbnHe3SynthesisGate}, \leanid{bbnSynthesisStrongDepositionBoost},
  coupled neutron inventory; Python mirror \texttt{hqiv\_bbn\_integrator.py}.
\item \phantomsection\label{lean:bbn-outside-neutron}\textbf{Outside neutron dynamics}
  \leanfile{Hqiv/Physics/NuclearOutsideTemperatureDynamics.lean} ---
  free-neutron $\tau_n(T)$ modulator feeding the faithful integrator.
\end{description}

\section{Stoichiometric integrator flow}
\label{app:flow}

Reader-facing pseudocode for \texttt{hqiv\_bbn\_integrator.py} (Lean mirror:
\hyperref[lean:bbn-stoichiometric]{stoichiometric integrator module}):

\begin{verbatim}
INPUT: eta_paper, lock-in Q from composite trace, gamma, w_strong
BUILD temperature ladder: T = 1 MeV -> T_bn (C2 bottleneck)
MARCH coupled neutron inventory x_n(T):
    tau_n(T) from outside curvature + relic-nu weak-width factor
    beta-decay survival freeze -> T_bn
FOR each T on ladder (log weights w):
    tail_gate(T) = bbnSynthesisDWindowTailGate(T)  # unity >= T_peak=0.18 MeV
    w_eff = w * tail_gate(T)
    Q_D^eff(T), Q_3^eff(T) from binding release + trimer width
    f_C2(T) = (kappa6_ref/kappa6(T))^w_C2(T)  if T <= T_bn else 1
    D/H from stoichiometric branches:
        n+p->D capture (branch_np_to_d * f_C2)
        D+D->4He, D photodissociation, D+p->3He (width-suppressed)
        strong deposition boost ~ gamma * w_strong * (1 + g_np + c2)
    integrate D/H with sum(D/H(T) * w_eff) / sum(w_eff)
    3He/H on 0.45 -> 0.05 MeV window (separate ladder)
    7Be -> 7Li tail on 0.25 -> 0.05 MeV
OUTPUT Y_p = 2*x_n(T_low)/(1+x_n(T_low))   # post-synthesis survivor
        D/H, 3He/H, 7Li/H integrated abundances
\end{verbatim}

\section{Minimal reproduction and sample witness}
\label{app:minimal-repro}

From the repository root, a single command regenerates the primary witness:
\begin{verbatim}
PYTHONPATH=scripts python3 scripts/hqiv_bbn_integrator.py
\end{verbatim}
Paper tables (\Cref{tab:provider-decomposition}, \Cref{tab:eta-sweep}, \Cref{tab:sensitivity}):
\begin{verbatim}
PYTHONPATH=scripts python3 scripts/hqiv_bbn_paper_tables.py
\end{verbatim}
Excerpt from \texttt{data/bbn\_integrator.json} at \etap:
\begin{verbatim}
"abundance_readout": {
  "Y_p": 0.24422,
  "D_over_H": 2.5119e-05,
  "He3_over_H": 1.0122e-05,
  "Li7_over_H": 2.5362e-10,
  "synthesis_d_T_peak_MeV": 0.18,
  "x_n_at_T_low": 0.13909,
  "delta_Y_p_curvature": -0.00269
}
\end{verbatim}

\section{Reproducer scripts}
\label{app:scripts}

From the repository root:
\begin{verbatim}
PYTHONPATH=scripts python3 scripts/hqiv_bbn_integrator.py
PYTHONPATH=scripts python3 scripts/hqiv_bbn_paper_tables.py
python3 scripts/hqiv_bbn_abundances.py
python3 scripts/hqiv_bbn_epoch_network.py
python3 scripts/hqiv_dynamic_bulk_bbn.py
python3 scripts/hqiv_integrator_lean_audit.py
python3 -m unittest scripts.test_hqiv_bbn_integrator
python3 -m unittest scripts.test_hqiv_dynamic_c2_bbn
\end{verbatim}
\paragraph{Outputs.}
\begin{itemize}[leftmargin=2em]
\item \texttt{data/bbn\_integrator.json} --- primary abundance witness
\item \texttt{data/bbn\_paper\_tables.json}
\item \texttt{data/bbn\_witnesses.json}
\item \texttt{data/dynamic\_bulk\_bbn\_v2.json}
\item \texttt{data/integrator\_lean\_audit.json}
\end{itemize}
See \texttt{papers/bbn/scripts/README.md} for the patch-ontology disclaimer and
input list.

\bibliographystyle{plain}
{\sloppy\small
\bibliography{../references}
}

\end{document}
