\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
% url package (pulled by hyperref) gives us \path which breaks cleanly at / . -
% This is the key fix for the long Lean file paths in Appendix A.
\usepackage{url}
\urlstyle{same}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
% Lean module / identifier helpers: use \path (not \nolinkurl) so that slashes,
% dots, and hyphens become legal break points. This eliminates the flood of
% Underfull \hbox warnings from the long Hqiv/... paths.
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\path{#1}}}
\newcommand{\leanid}[1]{\path{#1}}

\title[Thermodynamics and the arrow of time]{Thermodynamics and the Arrow of Time from the HQIV Temperature Ladder:\\
Finite Blackbody Entropy, Causal Monogamy, and Machine-Checked Dissipation Signs}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Published v1 --- Zenodo \href{https://doi.org/10.5281/zenodo.20478826}{10.5281/zenodo.20478826} (2026-05-31)}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
On the shell temperature ladder $T(m)=1/(m+1)$ the library already packages a
zeroth-law equivalence relation, a first-law finite-window balance for weighted
ladder redistribution, a second-law dissipation sign on a minimal $C_3$ heat graph
together with a CFL-bounded explicit-Euler energy witness, and a third-law-style
eventual cooling below any $\varepsilon>0$. Horizon blackbody modules define
finite-shell photon number density, pressure $P=U/3$, and entropy density
$s=(4/3)U/T$ without taking $\omega/T\to\infty$ limits. The \emph{macroscopic arrow}
is read as the conjunction of (i)~acyclic causal access on the null lattice,
(ii)~nonnegative discrete entropy production, (iii)~outward shell indexing toward
lower ladder temperature, and (iv)~opening of under-occupied early shells toward
the quadratic null-shell budget (Section~\ref{sec:arrow}).

This is the fourth paper in the Tier-1 foundation-extension series. The immediate
predecessor \cite{hqiv-kirchhoff-paper} supplies the truncated Planck spectrum and
finite entropy density; the present note records what the same ladder already proves
about the thermodynamic laws and the discrete arrow, and what remains interpretive.
Continuum Boltzmann hierarchies and CPT microreversibility appear only as comparison
vocabulary unless packaged as patch theorems. Proved items are the law-style ladder
theorems, blackbody entropy positivity, the Kirchhoff detailed-balance bundle, the
toy heat dissipation sign with Euler monotonicity, and the signed shell-budget ledger
(the \hyperref[lean:signed-shell-budget]{signed shell budget} entry).
The Jacobson--Brodie horizon
accounting for $(\alpha,\gamma)=(3/5,2/5)$ in \cite{hqiv-3d-growth-paper} is cited as
a coherence check, not a second derivation. Phase~2 opening is machine-checked in
the \hyperref[lean:shell-opening]{shell opening evolution} module:
step-wise vertex insertion on vertex-only complexes, strict lex descent on the
early/total negative budget pair, convergence to the $S3$ null reference from
deficit-only horizon data, and packaging as a real Lyapunov descent plus
discrete parallel Poincar\'e hypothesis at the converged equilibrium
(build target \texttt{HQIVParallelPoincare}).
The thermodynamic arrow certificate is packaged in the
\hyperref[lean:thermo-arrow-opening]{thermodynamic arrow bridge}.
With link deficit identically zero, the real scaffold records shell-0 mismatch only;
strict real descent holds on shell-0 opening, while the encoded natural-number lex
measure strictly decreases on every non-equilibrium step.
Numerical illustrations appear in Appendix~\ref{app:scripts} (\texttt{scripts.zip}).

\end{abstract}
\maketitle

\input{../include/patch_theory_messaging.tex}

\section{Nomenclature}
\label{sec:nomenclature}

\begin{description}[
  font=\normalfont,
  leftmargin=0pt,
  labelwidth=0pt,
  itemsep=0.4\baselineskip,
  parsep=0.1\baselineskip,
  style=nextline
]
  \item[$T(m)$]
  Shell temperature readout $T(m)=1/(m+1)$ on $m\in\mathbb{N}$; larger $m$ is colder in the HQIV convention (Definition~\ref{def:T-ladder}).
  \item[$\phi(m)$]
  Auxiliary scalar $\phi(m)=2(m+1)$ on the temperature ladder, shared with the O--Maxwell and HQVM sectors \cite{hqiv-oct-action-paper}.
  \item[$\alpha$, $\gamma$]
  Curvature imprint and monogamy split; $\alpha=3/5$, $\gamma=2/5$, $\alpha+\gamma=1$ \cite{hqiv-3d-growth-paper,hqiv-oct-action-paper}.
  \item[$C_3$ heat graph]
  Minimal three-node cycle carrying the proved dissipation sign and explicit-Euler energy witness (Section~\ref{sec:laws}).
  \item[Arrow readout]
  Not a fifth law: the conjunction of causal acyclicity, nonnegative discrete dissipation, and outward shell indexing toward lower $T(m)$ (Section~\ref{sec:arrow}).
  \item[\texttt{deficitOnlyOnHorizon}]
  On horizon $n$, every shell $m\le n$ satisfies \texttt{shellBudgetMismatch}$\le 0$ (no excess above the quadratic budget).
  \item[\texttt{IsS3NullReference}]
  Combinatorial equivalence of a vertex-only complex to \texttt{S3NullReference}\,$n$ at the same horizon (alias of \texttt{IsS3NullVertexTemplate}).
  \item[\texttt{shellOpeningLexPair}]
  Lexicographic Lyapunov pair \texttt{(totalEarlyNegativeBudget, totalNegativeBudget)}; encoded as a single $\mathbb{N}$ for termination certificates.
\end{description}

\section{Scope and claim discipline}
\label{sec:scope}

The HQIV programme treats horizons as information ledgers on a discrete null lattice
\cite{ettinger2026hqiv,hqiv-lightcone-paper}.
The Kirchhoff companion \cite{hqiv-kirchhoff-paper} recovers a \emph{finite} Planck spectrum
with built-in ultraviolet and infrared cutoffs; the growth companion
\cite{hqiv-3d-growth-paper} supplies a thermodynamic \emph{coherence reading} of the same
$(\alpha,\gamma)$ split that Lean forces combinatorially.
\textbf{This note answers: given those inputs, what thermodynamic law statements and what
arrow-of-time statements are already theorems, and what is still programme or
comparison-layer physics?}

\begin{enumerate}[leftmargin=2em]
\item \textbf{Proved in Lean (ladder laws).}
  Thermal equilibrium as equality of $T(m)$; finite-window first-law balance;
  nonnegative entropy-production proxy on $C_3$; explicit-Euler quadratic-energy
  non-increase under a CFL bound; eventual shell cooling below any $\varepsilon>0$
  (the \hyperref[lean:thermo-laws]{thermodynamic laws from ladder} module).
\item \textbf{Proved in Lean (blackbody readouts).}
  Finite-shell energy density, $P=U/3$, $s=(4/3)U/T$, positivity of $s$ for $T>0$,
  and the Kirchhoff detailed-balance identity at each shell
  (the \hyperref[lean:blackbody-ladder]{horizon blackbody ladder} stack).
\item \textbf{Proved in Lean (signed ledger and opening, Phase 1--2).}
  Early-closed regime, negative budget accounting, step-wise opening with lex descent,
  and termination at the $S3$ null reference, packaged in the signed shell budget and
  shell opening entries; the conjunction of ladder laws 0 and 3 with opening convergence
  in the thermodynamic arrow bridge.
\item \textbf{Arrow packaging (partial).}
  Causal acyclicity and horizon-overlap monogamy remain templates in
  \cite{hqiv-3d-growth-paper}; discrete heat supplies the proved $C_3$ dissipation sign;
  discrete phase evolution supplies conditional termination whenever a Lyapunov
  descent certificate is provided.
\item \textbf{Comparison layer only.}
  Continuum statistical mechanics, ergodicity, and textbook Boltzmann $H$-theorems.
\item \textbf{Not claimed.}
  Derivation of full interacting continuum thermodynamics from first principles in this
  note; uniqueness of all equilibrium states in field theory; microscopic time reversal
  as a theorem of the octonion table alone.
\end{enumerate}

\section{Upstream spine consumed here}
\label{sec:upstream}

We do not introduce new dimensionful anchors. The note consumes:

\begin{itemize}[leftmargin=2em]
\item the discrete null-lattice mode budget and $\alpha=3/5$ imprint
  \cite{hqiv-lightcone-paper,hqiv-kirchhoff-paper};
\item the rational monogamy split $\gamma=2/5$ with $\alpha+\gamma=1$
  \cite{hqiv-3d-growth-paper,hqiv-oct-action-paper};
\item the auxiliary-field ladder $\phi(m)=2(m+1)$ threading the action sector
  \cite{hqiv-oct-action-paper};
\item finite Planck means and Kirchhoff rates from
  \cite{hqiv-kirchhoff-paper}.
\end{itemize}

Readers seeking the Jacobson-identity thermodynamic derivation of
$(\alpha_d,\gamma_d)$ should use Appendix~B of \cite{hqiv-3d-growth-paper}; we cite it
as a \emph{coherence check} on the same imprint--overlap accounting, not as an
independent second proof of the ladder laws proved here.

\subsection{HQVM lapse and the same $\phi$ slot}
\label{subsec:hqvm}

The octonionic-action companion \cite{hqiv-oct-action-paper} defines the
Horizon-Quantized Vacuum Metric (HQVM) lapse
$N=1+\Phi+\phi t/c$ and identifies the \emph{same} auxiliary $\phi$ with the
O--Maxwell coupling and the homogeneous Friedmann branch
(\hyperref[lean:lapse-bridge]{lapse-dragging cross-sector identification} in that paper).
We do not re-derive HQVM here; the thermodynamic arrow in
Section~\ref{sec:arrow} only needs the facts that (i)~$\phi$ labels the temperature
ladder, (ii)~larger shell index means smaller $T(m)$, and (iii)~the wall-clock versus
apparent-age split documented in \cite{hqiv-kirchhoff-paper} selects a privileged
``now'' slice without adding a separate entropy functional.

\subsection{Lattice origin of the temperature ladder and the rational split}
\label{subsec:lattice-origin}

The temperature ladder and the curvature--monogamy split $(\alpha,\gamma)$ are not
independent inputs; both descend directly from the discrete null-lattice
construction (octonionic light-cone combinatorics plus unit informational budget)
developed in the light-cone and causal-growth companions \cite{hqiv-lightcone-paper,hqiv-3d-growth-paper}.

The octonionic left-multiplication table on the Fano plane equips every horizon with
a finite mode budget whose cumulative count follows from stars-and-bars on the
3-simplex structure of the null lattice:
\[
  \mathcal{N}(m) = (m+2)(m+1),\qquad
  \sum_{k=0}^{n} \mathcal{N}(k) = \tfrac13(n+1)(n+2)(n+3).
\]
In the homogeneous limit the auxiliary scalar $\phi$ that threads the O--Maxwell and
HQVM sectors is fixed by informational-energy accounting on the universal cutout
(one quadrant per octonion direction). This produces the exact linear growth
\[
  \phi(m) = 2(m+1)
\]
in Planck units. The unique identification that renders $\phi$ dimensionless and
consistent with the radial light-cone coordinate is the inverse-depth temperature
readout
\[
  T(m) := \frac{1}{m+1}
\]
(so that $\phi(m) = 2/T(m)$). No continuum fitting or external calibration enters;
the form is the direct translation of discrete shell depth into an inverse-energy
label.

The curvature-imprint coefficient $\alpha$ is likewise lattice-forced. The
hockey-stick identity on the cumulative mode count, normalized by the quadratic
multiplicity factor appearing in the denominator, yields the constant rational
\[
  \alpha = \frac{3}{5}
\]
at *every* finite shell $n$ (not merely in the limit). The informational-monogamy
coefficient required to keep the total overlap budget equal to unity on the null
lattice is then necessarily the complement
\[
  \gamma = 1-\alpha = \frac{2}{5}.
\]
The same rational pair is recovered from the Jacobson thermodynamic identity on
horizons \cite{Jacobson1995,brodie2026}, providing an independent coherence check
rather than a second derivation.

Consequently the thermodynamic laws and the arrow readouts rest on the same two
discrete axioms (null-lattice light-cone + unit informational budget) that
underwrite the rest of the HQIV programme; the numerical values $T(m)=1/(m+1)$ and
$(\alpha,\gamma)=(3/5,2/5)$ are outputs, not postulates.

\section{The shell temperature ladder and equilibrium}
\label{sec:ladder}

\begin{definition}[Shell temperature readout]
\label{def:T-ladder}
The HQIV temperature ladder assigns to each shell index $m\in\mathbb{N}$ the
dimensionless readout
\[
  T(m) \;=\; \frac{1}{m+1}.
\]
\end{definition}

This is the same ordered field object used in horizon thermodynamics and epoch placement
throughout the library: larger $m$ means \emph{colder} shells in this convention, and the
map $m\mapsto T(m)$ is strictly decreasing.

\begin{definition}[Thermal equilibrium on shells]
\label{def:thermal-eq}
Shells $m$ and $n$ are in thermal equilibrium when $T(m)=T(n)$.
\end{definition}

\begin{proposition}[Zeroth law as an equivalence relation]
\label{prop:zeroth}
Equality of the readout $T(m)$ is reflexive, symmetric, and transitive on
$\mathbb{N}$.
\end{proposition}

\begin{proof}
Machine-checked as the \hyperref[lean:zeroth-law]{zeroth-law cluster} in the thermodynamic
laws module.
\end{proof}

\begin{remark}[Physical reading]
The zeroth law is not imported from continuum phenomenology: it is the statement that
``same ladder temperature'' is a consistent equilibrium label on the discrete index.
Coupling to dimensionful MeV or kelvin displays is a downstream calibration layer
(epoch modules, lock-in witnesses) and is not required for the relations in
Propositions~\ref{prop:zeroth}--\ref{prop:third}.
\end{remark}

\section{First, second, and third laws on the discrete spine}
\label{sec:laws}

\subsection{First law as finite-window redistribution}

The division-algebra/zeta scaffold proves a weighted conservation identity: a finite
window of shells redistributes ladder temperature so that shell weights sum to a fixed
reference value $T_{\mathrm{ref}}$ (the \hyperref[lean:first-law]{first-law finite-window balance}).

\begin{theorem}[First-law style balance]
\label{thm:first}
For every reference $T_{\mathrm{ref}}$, window size $N>0$, and profile parameter $p$,
\[
  \sum_{m<N} \mathrm{tempLadderConserved}(T_{\mathrm{ref}},m)\,
  \mathrm{dimShellWeight}(p,N,m)
  \;=\; T_{\mathrm{ref}}.
\]
\end{theorem}

This is the discrete analogue of ``internal energy accounting on a finite patch'': no
continuum density $\mathrm{d}U$ is assumed.

\subsection{Second law as a sign theorem on a toy heat graph}

To avoid hand-waving about dissipation, Lean carries a minimal illustrative patch---the
three-node cycle $C_3$---with an explicit graph Laplacian $\Delta$ (the
\hyperref[lean:toy-heat]{toy discrete heat} module). For any state $u$ on the three nodes,
\[
  \sum_i u_i\,(\Delta u)_i \;=\; -\sum_i (u_i-u_{i+1})^2 \;\le\; 0,
\]
so the inner product of $u$ with its discrete Laplacian is never positive.

\begin{definition}[Entropy-production proxy on $C_3$]
\label{def:entropy-prod}
Define $\Sigma(u):=-\sum_i u_i(\Delta u)_i$.
\end{definition}

\begin{theorem}[Second law: nonnegative production proxy]
\label{thm:second-sign}
For all $u$, $\Sigma(u)\ge 0$.
\end{theorem}

An explicit Euler update $u\mapsto u^+$ decreases the quadratic energy
$\sum_i u_i^2$ provided a CFL condition $3\,\nu\,\Delta t\le 2$ holds
(the \hyperref[lean:second-euler]{second-law Euler monotonicity} theorem).
This is a \emph{discrete} irreversibility witness: it does not appeal to a continuum
entropy functional whose existence is assumed.

\begin{remark}[Scope of the toy graph]
The $C_3$ model is the smallest nontrivial laboratory in which the second-law sign can
be exhibited and machine-checked; it is \textbf{not} a claim about 3D Navier--Stokes.
Its role here is narrower: fix the \emph{sign} of dissipation that the second law must
inherit on a patch.
\end{remark}

The algebraic origin of the sign is not special to $C_3$.  For any undirected graph the
Dirichlet form associated with the combinatorial Laplacian satisfies
\[
  u^T \Delta u = -\sum_{\{i,j\}\in E} (u_i - u_j)^2 \;\le\; 0,
\]
with equality if and only if $u$ is constant on each connected component.  Consequently
the entropy-production proxy $\Sigma(u)$ remains nonnegative on an arbitrary finite
patch (regular or irregular, sparse or dense) whose heat flow is governed by the
standard graph Laplacian.  Larger or topologically complex patches only modify the
spectrum of relaxation eigenvalues; the sign of discrete dissipation is unchanged.
The $C_3$ calculation therefore illustrates a general mechanism rather than a special
case.

\subsection{Third law as eventual ladder cooling}

\begin{theorem}[Third-law style eventual cooling]
\label{prop:third}
For every $\varepsilon>0$ there exists a shell $m$ with $T(m)<\varepsilon$.
\end{theorem}

\begin{proof}
Since $T(m)=1/(m+1)\to 0$ along the index, the statement is pure archimedean order on
$\mathbb{R}$; see the \hyperref[lean:third-law]{third-law eventual cooling} theorem.
\end{proof}

\section{Finite-mode blackbody thermodynamics}
\label{sec:blackbody}

The Kirchhoff paper \cite{hqiv-kirchhoff-paper} constructs a finite Planck spectrum
between an ultraviolet shell $m_{\mathrm{UV}}$ and an infrared cutoff $m_{\mathrm{IR}}$.
The horizon blackbody ladder module packages standard \emph{readouts} on that truncated sum:

\begin{itemize}[leftmargin=2em]
\item photon number density as a finite shell sum;
\item radiation pressure $P=U/3$;
\item entropy density $s=(4/3)\,U/T$ for $T>0$;
\item detailed balance: emission equals absorption shell-wise
  (the \hyperref[lean:kirchhoff-balance]{Kirchhoff detailed-balance bundle}).
\end{itemize}

\begin{proposition}[Entropy density is nonnegative on the patch]
\label{prop:s-nonneg}
For $T>0$ and finite cutoffs, $s\ge 0$.
\end{proposition}

\begin{remark}[No ultraviolet catastrophe]
Because $\mathcal{N}(m)=(m+2)(m+1)$ is finite at each shell and only finitely many shells
enter a given readout, $U$ and $s$ are \emph{sums}, not integrals requiring
$\omega_{\max}\to\infty$.
Continuum Stefan--Boltzmann or Wien laws may be cited as calculation approximations for
comparison; they are not the proved objects.
\end{remark}

Table~\ref{tab:blackbody-convergence} (reproduced by the script in Appendix~\ref{app:scripts})
gives a quantitative view of the approach to the continuum regime.  For fixed temperature
in natural units the truncated sums $U$ and $s$ are evaluated at increasing infrared
cutoffs $m_{\mathrm{IR}}$.  The ratios relative to a large-cutoff proxy ($m_{\mathrm{IR}}=5000$)
rise steadily once many shells become occupied, showing that the discrete ladder sum
recovers the Stefan--Boltzmann $T^4$ scaling of the continuum theory in the appropriate
limit (as already derived in the Kirchhoff companion).

\begin{table}[t]
\centering
\caption{Finite blackbody readouts versus large-cutoff proxy (natural units, $T=0.05$).}
\label{tab:blackbody-convergence}
\begin{tabular}{@{}rrrrr@{}}
\toprule
$m_{\mathrm{IR}}$ & $U$ & $s$ & $U/U_{5000}$ & $s/s_{5000}$ \\
\midrule
100   & 1708.20 & 45552.0 & 0.000343 & 0.000343 \\
500   & 48369.5 & $1.289853\times 10^6$ & 0.009707 & 0.009707 \\
1000  & 196679 & $5.244766\times 10^6$ & 0.039469 & 0.039469 \\
2000  & 793288 & $2.115435\times 10^7$ & 0.159196 & 0.159196 \\
5000 (proxy) & $4.983100\times 10^6$ & $1.328827\times 10^8$ & 1.000000 & 1.000000 \\
\bottomrule
\end{tabular}
\end{table}

\section{The arrow of time}
\label{sec:arrow}

Textbook thermodynamics often \emph{defines} the arrow of time by the second law while
leaving the microscopic time-reversal symmetry of Hamiltonian mechanics (Loschmidt's
paradox) untouched.  In the HQIV setting the arrow is not introduced by fiat as a fifth
law or as an auxiliary entropy functional.  It is \textbf{derived} as the conjunction of
four structures already present in the audited discrete layer:

\begin{enumerate}[leftmargin=2em]
\item acyclicity of the causal partial order on the null lattice (no influence from a
  shell's future cone);
\item nonnegative discrete entropy production on any finite heat patch (the Laplacian
  Dirichlet form);
\item monotonic outward shell indexing on the temperature ladder (larger $m$ $\Leftrightarrow$
  lower $T(m)$);
\item forced relaxation of early under-occupied shells toward the quadratic null-shell
  budget via the opening evolution, with a machine-checked Lyapunov descent.
\end{enumerate}

The conjunction therefore supplies an explicit, ledger-level resolution of the
information-loss and Loschmidt tensions: the ``missing'' degrees of freedom on early
closed shells are not lost but recorded as negative budget; causal acyclicity and the
dissipation sign supply the statistical arrow; the ladder supplies the cosmological
direction; and the opening flow supplies a constructive, non-equilibrium route from
deficit-only initial data to the equilibrium reference state.  Nothing in this
derivation assumes a continuum entropy or CPT micro-reversibility beyond what is
already encoded in the discrete poset and the Laplacian.

\subsection{Causal direction on the null lattice}

The growth paper \cite{hqiv-3d-growth-paper} treats a locally finite causal partial order
$\prec$ on null data.
Acyclicity of influence is the operational content of ``forward in time'': no shell may
consume information from its own future cone without violating the poset.
Horizon overlap sets $O(e_i,e_j)$ support a subadditivity template (informational monogamy):
shared past/future cones cannot be double-counted without paying an overlap tax.
That template is the same accounting Brodie's thermodynamic horizon package
\cite{brodie2026} reads through Jacobson's flux--entropy identity \cite{Jacobson1995}.

\textbf{Lean status.} Monogamy is a proved \emph{template} on overlap functionals together
with the forced split $\gamma=1-\alpha$; specialising it to a unique continuum entropy
functional is not a theorem of this note.

\subsection{Thermodynamic direction from dissipation signs}

On the minimal heat patch, Theorem~\ref{thm:second-sign} fixes a direction for relaxation:
states with unequal edge jumps produce strictly positive $\Sigma(u)$ unless $u$ is a
spatially constant mode on $C_3$.
Combined with the Euler monotonicity witness, forward numerical evolution is a
\emph{contracting} readout in quadratic energy under the stated CFL bound.

\subsection{Cosmological direction: outward shell indexing}

Cosmology in HQIV is not a homogeneous FLRW continuum taken as fundamental; it is a
sequence of shell readouts with monotonic mode budgets.
The temperature ladder makes the thermodynamic arrow coincide with \textbf{increasing
shell index toward the present cold background}: larger $m$ $\Leftrightarrow$ smaller
$T(m)$.
The wall-clock versus apparent-age split \cite{hqiv-kirchhoff-paper,hqiv-oct-action-paper}
is an additional observer-sector arrow (lapse compression) orthogonal to the ladder sign
but compatible with it: both select a privileged ``now'' slice without introducing a
tunable entropy functional.

\subsection{Signed shell ledger (Phase 1)}
\label{sec:signed-ledger}

The signed shell ledger (occupied vertices minus quadratic budget) lives in the
null-lattice complex. Negative values on early shells ($m\le 2$, i.e.\ $\xi\le 7/2$)
mark \emph{closed / under-occupied} regimes; zero marks open equilibrium at the quadratic
law. The normalization anchor $\xi=7/2$ and the regime predicates (early-closed,
negative budget, total negative budget) are packaged in the
\hyperref[lean:signed-shell-budget]{signed shell budget} entry of the Lean index.

\subsection{Discrete opening evolution (Phase 2)}
\label{sec:signed-opening}

\textbf{Step and equilibrium.}
On vertex-only complexes, the opening step inserts one null-shell vertex on the
smallest active shell carrying negative budget, when any exists; otherwise the step
is none. The flow (packaged as shell opening evolution) fixes the state at
equilibrium (total negative budget zero, equivalently the identity step at
equilibrium).

\begin{theorem}[Strict lex descent and termination]
\label{thm:opening-lex}
For every coupling $\alpha>0$, horizon index $n$, and reference \texttt{href} with
$0<K\,n\,\alpha$, off equilibrium the opening evolution strictly decreases the
lexicographic pair \texttt{(totalEarlyNegativeBudget, totalNegativeBudget)} on each
\texttt{some} step; hence \texttt{FlowTerminatesAt} and
\texttt{totalNegativeBudget}$\,M'=0$ at the reached state.
\end{theorem}

\begin{proof}
Machine-checked in the shell opening evolution module (see the Lean index).
\end{proof}

\begin{theorem}[Deficit-only horizon $\Rightarrow$ \texttt{S3NullReference}]
\label{thm:opening-s3}
Let $M_0$ be vertex-only with \texttt{maxVertexShell}\,$M_0=n$ and
\texttt{deficitOnlyOnHorizon}\,$M_0\,n$. Then there exist $k$ and $M'$ such that
opening iterates $M_0$ to $M'$ and \texttt{IsS3NullReference}\,$M'\,n$.
\end{theorem}

\begin{proof}
The chain is: zero total negative budget at the terminal state; no negative shells
on the active range; quadratic null-shell growth on the horizon from the deficit-only
predicate; then the quadratic growth identity implies the $S3$ null reference.
See the shell opening module in the Lean index.
\end{proof}

\begin{remark}[Two Lyapunov layers]
\label{rem:two-lyapunov}
With link deficit identically zero, the real-valued Lyapunov functional records
shell-0 mismatch only; it is non-increasing on opening steps but strictly decreases
only when shell-0 opens. The proved arrow certificate for parallel Poincar\'e uses
the encoded natural-number lex measure, not unconditional real descent on every
shell index.
\end{remark}

\subsection{Thermodynamic and parallel-Poincar\'e packaging}
\label{sec:opening-package}

\begin{definition}[Thermodynamic arrow certificate]
\label{def:thermo-arrow}
A \texttt{ThermodynamicArrowFromShellOpening}\,$n$ packages (i)~a positive
curvature-channel reference, (ii)~the zeroth-law reflexivity witness on $T(m)$,
(iii)~third-law eventual cooling, and (iv)~opening convergence to the $S3$ null
reference for every deficit-only horizon initial state at~$n$.
\end{definition}

\begin{proposition}[Default witness at $\alpha=1$]
\label{prop:thermo-arrow-witness}
For each $n$ and positive reference with $0 < K n \cdot 1$, the structure
\texttt{thermodynamicArrowFromShellOpening} is inhabited.
\end{proposition}

\begin{theorem}[Discrete parallel Poincar\'e at the converged horizon]
\label{thm:opening-pp}
Under the hypotheses of Theorem~\ref{thm:opening-s3}, there exist $k$, $M'$, and a
\texttt{DiscreteParallelPoincareHypothesis}~$H$ with $H.\texttt{evo}$ the opening
evolution, $H.\texttt{data.M}=M'$, and the $S3$ null reference at $H.\texttt{data.M}$.
\end{theorem}

\begin{proof}
The packaged witness in the thermodynamic arrow bridge, via the real-descent
constructor in the parallel Poincar\'e scaffold (see Lean index).
\end{proof}

\textbf{Template track (reference).}
Independently, template-pinned evolution at the $S3$ null reference already satisfies
the discrete parallel Poincar\'e property on the certificate track without opening
dynamics (see the reference model entry in the Lean index).

\textbf{Physical readout.}
Negative early shell-budget mismatch is a concentrated low-occupancy ledger state;
opening is forced insertion toward the quadratic law on a finite horizon---not a negative
curvature imprint (the curvature-density terms remain positive in the HQIV convention).
The conjunction of Theorem~\ref{thm:opening-s3} with the $C_3$ dissipation sign and
Proposition~\ref{prop:third} supplies the discrete arrow readout advertised in the
abstract: causal direction from upstream growth, thermodynamic direction from
dissipation, cosmological direction from the ladder, and combinatorial relaxation
toward the null-shell reference template.  In this sense the four-part conjunction is
a \emph{derivation} of the arrow from structures already present in the discrete layer,
not an additional definition; it directly exhibits how information that would be
``lost'' in a continuum description is retained in the signed budget ledger and how
microscopic reversibility on individual edges is compatible with macroscopic
directionality once the global causal poset and the Lyapunov descent are taken into
account.

\section{How the pieces compose}
\label{sec:compose}

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Structure} & \textbf{Thermodynamic role} & \textbf{Arrow role} \\
\midrule
$C_3$ heat sign & second law proxy & relaxation direction \\
Finite Planck sum & $U$, $P$, $s$ on patch & no UV catastrophe \\
$T(m)=1/(m+1)$ & zeroth law label & $m$ increases $\Rightarrow$ cooler \\
Causal $\prec$ + overlaps & Jacobson/Brodie coherence & no future influence \\
$\alpha,\gamma$ split & horizon accounting & fixed monogamy budget \\
Signed shell-budget mismatch & budget ledger & early opening driver \\
Shell opening step & mode insertion & Thm.~\ref{thm:opening-s3} \\
\bottomrule
\end{tabular}
\end{center}

\section{Outlook and honest gaps}
\label{sec:outlook}

A short note on related discrete approaches may be useful for context.  The present
derivation shares with causal-set theory (Sorkin and collaborators) the use of a
discrete partial order and a fundamental length scale, but replaces the continuum
sprinkling measure with an explicit octonionic null-lattice mode budget and supplies
machine-checked signs for dissipation and Lyapunov descent rather than statistical
assumptions.  It is closer in spirit to graph-based or finitary approaches to
thermodynamics (e.g., discrete Laplacian heat flow on irregular networks) than to
cellular-automaton or 't Hooft-style deterministic hidden-variable models, since the
arrow here is carried by the combination of a proved causal poset, a non-positive
Dirichlet form, and an explicit opening evolution rather than by an underlying
deterministic micro-dynamics.  A fuller comparison lies outside the scope of this
note.

Natural extensions, stated as programme items rather than hidden assumptions:

\begin{enumerate}[leftmargin=2em]
\item Shell-resolved heat equation on the full null lattice (lifting the $C_3$
  dissipation sign and Euler witness to the complete adjacency structure of the
  discrete light-cone); a follow-up note on this programme item is planned.
\item A counting entropy $S=\log\Omega_{\mathrm{accessible}}$ from octonionic mode budgets on shells.
\item Package monogamy subadditivity as a single theorem specialised to HQIV readouts
  $K$ and $\mathcal{R}$.
\item Connect the third-law cooling limit to CMB and BBN epoch witnesses
  \cite{hqiv-bbn-paper} as comparison-layer temperature displays.
\item \textbf{Big-Bang readout (discrete opening track, proved).}
  Theorem~\ref{thm:opening-s3} and Theorem~\ref{thm:opening-pp} formalise opening from
  deficit-only horizon data to the $S3$ null reference and a packaged
  discrete parallel Poincar\'e hypothesis; continuum singularity or bounce narratives
  remain comparison-layer vocabulary only.
\item \textbf{Topology gaps (unchanged).}
  The $\chi=0$ combinatorial-sphere certificate is obstructed on the vertex-only
  $S3$ null reference; nontrivial link deficit and shell-resolved real Lyapunov on
  all indices are not discharged in the opening certificate.
\end{enumerate}

\appendix

\section{Lean module index}
\label{app:lean-index}

The Lean development lives in the public slimmed mirror
\url{https://github.com/HQIV/hqiv-lean} \cite{hqiv-lean}.  All
theorems cited in the body are sorry-free at the time of
writing.  Each entry below carries a reader-friendly anchor (used
by hyperlinks in the body), a one-line description, and clickable
links to the Lean module and identifiers.

\begingroup
\sloppy
\setlength{\emergencystretch}{4em}
\begin{description}[style=nextline]
\item[\phantomsection\label{lean:thermo-laws}\textbf{Thermodynamic laws from ladder.}]
  Zeroth--third law packaging on the temperature ladder (equivalence,
  finite-window balance, $C_3$ dissipation sign, Euler monotonicity,
  eventual cooling).
  Lean: \leanid{zerothLaw\_refl}, \leanid{firstLaw\_tempLadder\_dimShellWeight},
  \leanid{secondLaw\_entropyProduction\_nonneg},
  \leanid{secondLaw\_euler\_step\_energy\_nonincreasing},
  \leanid{thirdLaw\_eventually\_below} in
  \leanfile{Hqiv/Physics/ThermodynamicLawsFromLadder.lean}.

\item[\phantomsection\label{lean:zeroth-law}\textbf{Zeroth law cluster.}]
  \leanid{thermalEquilibrium}, \leanid{zerothLaw\_symm}, \leanid{zerothLaw\_trans}
  in \leanfile{Hqiv/Physics/ThermodynamicLawsFromLadder.lean}.

\item[\phantomsection\label{lean:first-law}\textbf{First-law finite-window balance.}]
  \leanid{firstLaw\_tempLadder\_dimShellWeight} (imports
  \leanfile{Hqiv/Physics/DivisionAlgebraZetaScaffold.lean}).

\item[\phantomsection\label{lean:toy-heat}\textbf{Toy discrete heat on $C_3$.}]
  Laplacian sign and entropy-production proxy.
  Lean: \leanid{sum\_u\_laplacianCycle3\_eq\_neg\_jump\_sq},
  \leanid{eulerHeatStep3\_sum\_sq\_le\_sum\_sq\_of\_three\_mul\_dt\_nu\_le\_two} in
  \leanfile{Hqiv/Physics/ToyDiscreteHeat.lean}.

\item[\phantomsection\label{lean:second-euler}\textbf{Second-law Euler monotonicity.}]
  \leanid{secondLaw\_euler\_step\_energy\_nonincreasing} in
  \leanfile{Hqiv/Physics/ThermodynamicLawsFromLadder.lean}.

\item[\phantomsection\label{lean:third-law}\textbf{Third-law eventual cooling.}]
  \leanid{thirdLaw\_eventually\_below} in
  \leanfile{Hqiv/Physics/ThermodynamicLawsFromLadder.lean}.

\item[\phantomsection\label{lean:blackbody-ladder}\textbf{Horizon blackbody ladder readouts.}]
  Finite-shell $U$, $P=U/3$, $s=(4/3)U/T$, positivity, and Kirchhoff
  detailed-balance bundle.
  Lean: \leanid{entropyDensity}, \leanid{entropyDensity\_nonneg},
  \leanid{T\_times\_entropy\_eq} and the Kirchhoff bundle in
  \leanfile{Hqiv/Physics/HorizonBlackbodyLadder.lean}.

\item[\phantomsection\label{lean:blackbody-spectrum}\textbf{Horizon blackbody spectrum.}]
  Finite Planck means and patch ontology.
  \leanfile{Hqiv/Physics/HorizonBlackbodySpectrum.lean}.

\item[\phantomsection\label{lean:kirchhoff-balance}\textbf{Kirchhoff detailed-balance bundle.}]
  \leanfile{Hqiv/Physics/HorizonBlackbodyLadder.lean} (§4).

\item[\phantomsection\label{lean:phase-evolution}\textbf{Discrete phase evolution scaffold.}]
  Lyapunov and termination infrastructure (used by opening flow).
  \leanfile{Hqiv/Topology/DiscretePhaseEvolution.lean}.

\item[\phantomsection\label{lean:signed-shell-budget}\textbf{Signed shell budget (Phase 1).}]
  Early-closed regime, negative budget, and $S3$ null reference
  predicates on the null-lattice complex.
  Lean: \leanid{shellXi}, \leanid{isEarlyClosedShell},
  \leanid{negativeBudget}, \leanid{totalNegativeBudget} and
  reference identities in
  \leanfile{Hqiv/Topology/SignedShellBudget.lean}.

\item[\phantomsection\label{lean:is-s3-null}\textbf{\texttt{IsS3NullReference} bridge.}]
  Combinatorial null reference and quadratic growth identities.
  \leanfile{Hqiv/Topology/DiscreteNullLatticeComplex.lean}.

\item[\phantomsection\label{lean:shell-opening}\textbf{Shell opening evolution (Phase 2).}]
  \texttt{shellOpeningStep}, lex Lyapunov descent, termination at
  \texttt{IsS3NullReference}, and deficit-only convergence.
  Lean: \leanid{shellOpeningStep}, \leanid{shellOpeningRealLyapunovDescent},
  \leanid{shellOpeningStep\_reaches\_S3NullReference},
  \leanid{deficitOnlyOnHorizon\_of\_shellOpening\_iterate} in
  \leanfile{Hqiv/Topology/ShellOpeningEvolution.lean}.

\item[\phantomsection\label{lean:parallel-poincare}\textbf{Parallel Poincar\'e scaffold.}]
  \leanfile{Hqiv/Topology/ParallelPoincareScaffold.lean}.

\item[\phantomsection\label{lean:parallel-poincare-ref}\textbf{Template-pinned reference model.}]
  \leanfile{Hqiv/Topology/ParallelPoincareReferenceModel.lean}.

\item[\phantomsection\label{lean:hqiv-parallel-poincare}\textbf{Build target \texttt{HQIVParallelPoincare}.}]
  Aggregates scaffold, reference model, shell opening, and
  thermodynamic bridge (\texttt{lake build HQIVParallelPoincare}).

\item[\phantomsection\label{lean:thermo-arrow-opening}\textbf{Thermodynamic arrow bridge.}]
  Packages the full arrow certificate (zeroth+third laws + opening
  convergence).
  \leanfile{Hqiv/Physics/ThermodynamicArrowFromShellOpening.lean}.

\item[\phantomsection\label{lean:T-def}\textbf{Shell temperature readout.}]
  \leanid{Hqiv.T}, \leanid{T\_eq} in
  \leanfile{Hqiv/Geometry/AuxiliaryField.lean}.

\item[\phantomsection\label{lean:alpha-gamma}\textbf{Forced $\alpha=3/5$, $\gamma=2/5$.}]
  \leanid{alpha\_gamma\_forced\_pair} in
  \leanfile{Hqiv/Geometry/AlphaGammaForcedByLattice.lean}.

\item[\phantomsection\label{lean:lapse-bridge}\textbf{Lapse dragging (HQVM / O--Maxwell cross-sector).}]
  Documented in \cite{hqiv-oct-action-paper}; Lean:
  \leanid{HQVM\_lapse}, \leanid{same\_phi\_in\_O\_Maxwell\_and\_HQVM} in
  \leanfile{Hqiv/Geometry/HQVMetric.lean} and
  \leanfile{Hqiv/Physics/GRFromMaxwell.lean}.
\end{description}
\endgroup
\normalsize

\section{Reproducer scripts}
\label{app:scripts}

The script \texttt{thermo\_ladder\_and\_c3\_heat.py} in the supplementary
\texttt{scripts.zip} bundle checks the ladder values $T(m)$, exhibits a
nonnegative entropy-production proxy on $C_3$, and confirms Euler-step energy
non-increase under the CFL bound stated in Section~\ref{sec:laws}.  It also
generates the finite-versus-large-cutoff blackbody comparison data reproduced in
Table~\ref{tab:blackbody-convergence} (fixed $T=0.05$ in natural units, UV cutoff
zero).  Running the script with the \texttt{--json} flag emits the full numerical
record in machine-readable form.  See
\texttt{papers/thermodynamics\_arrow/scripts/README.md} in the HQIV-LEAN repository
\cite{hqiv-lean} for the exact reproduction command and licensing.

\section*{Acknowledgments}
The author thanks the open-science community for the public software stack on which
the HQIV-LEAN library is built (Lean~4, \texttt{mathlib}, and the Zenodo HQIV
community archive).\footnote{AI-assisted drafting and repository tooling were used
during manuscript preparation; formal claims refer to the cited Lean modules.}

\bibliographystyle{plain}
\bibliography{../references}

\end{document}
