% Shared paper-facing statement: patch QFT + discrete numerics + gauge consistency.
% From papers/*.tex:     \input{../include/patch_theory_messaging.tex}
% Papers with a Lean module index may define \leanCertAnomaly, \leanCertTopology,
% and \leanCertStrongColor before this include; otherwise they default to empty.
% Strong-closure notes may define \newcommand{\patchBriefMode}{} before this include
% for a shorter gauge bullet list. Define \newcommand{\patchOmitZeta}{} to drop the
% zeta pointer / dictionary (appropriate when zeta is not in the note's proof cone).

\providecommand{\leanCertAnomaly}{}
\providecommand{\leanCertTopology}{}
\providecommand{\leanCertStrongColor}{}
\providecommand{\leanPatchHolonomyWitness}{(\leanfile{Hqiv/Physics/SO8PlaquetteHolonomy.lean},
witness \leanid{so8PlaquetteHolonomyDischarged_holds})}
\providecommand{\leanPatchWilsonKinetic}{(\leanfile{Hqiv/Physics/ActionHolonomyGlue.lean},
witness \leanid{wilsonKineticPlaquetteEquivalence_discharged})}
\providecommand{\leanPatchHaarMeasure}{(\leanfile{Hqiv/Physics/NonAbelianHolonomyMeasureScaffold.lean})}
\providecommand{\leanPatchDiscretePoincare}{(\leanfile{Hqiv/Physics/DiscreteActionStrongPoincareBridge.lean},
\leanfile{Hqiv/Physics/DiscreteActionPoincareScaffold.lean})}
\providecommand{\leanPatchSpatialRotation}{(\leanfile{Hqiv/Geometry/SpatialRotationLorentzClosure.lean},
witness \leanid{spatial_rotation_lorentz_closure_discharged})}
\providecommand{\leanfile}[1]{\nolinkurl{#1}}
\providecommand{\leanid}[1]{\nolinkurl{#1}}

\paragraph{Patch QFT.}
The quantum field theory used here is the HQIV \emph{patch QFT}: fields, actions,
charges, amplitudes, and readouts live on finite horizon patches labelled by shell
indices $m\in\mathbb{N}$, finite spacetime index types such as $\mathrm{Fin}\,4$,
and discrete carrier charts.  Each patch is the \emph{entire accessible universe} in
theory terms---the full causally accessible patch net, not a mesh tile waiting for a
fundamental continuum limit.  This is the operative QFT of the model.  Its empirical content is
the finite-patch calculation itself: mass and coupling ladders, curvature ratios,
shell thermodynamics, anomaly traces, branching ledgers, and rational witnesses
are computed on the patch carrier and then compared with laboratory observables.
Continuum notation, when used, is a dictionary for readers trained in smooth
field theory; it is not the definition of the theory.

\paragraph{Gauge and topology on the patch carrier.}
The patch QFT has no hidden continuum gauge sector whose consistency must be
assumed.  The relevant gauge and topological statements are finite statements on
the same carrier that produces the numerical readouts:
\ifdefined\patchBriefMode
\begin{itemize}[nosep]
\item \textbf{Anomaly cancellation.} For one Standard-Model generation
($Q_L$, $u^c$, $d^c$, $L$, $e^c$, $\nu^c$), the patch trace identities give
zero $U(1)_Y^3$, gravitational--$U(1)_Y$, $SU(3)_c^2U(1)_Y$,
$SU(2)_L^2U(1)_Y$, and $SU(3)_c^3$ anomaly coefficients; the $SU(2)_L^3$
coefficient vanishes because the weak doublet slot is pseudoreal.  Three
generations are a finite $\mathrm{Fin}\,3$ label sum of the same cancelled
one-generation block, not a fitted multiplicity\leanCertAnomaly.
\item \textbf{Topological sector.} The patch abelian gauge datum has a single
topological sector.  Instanton number, Pontryagin number, first Chern number, and
$U(1)$ winding number vanish identically on patch data; the $\theta$ contribution
is therefore $\theta$-independent; and abelian commutator obstructions vanish\leanCertTopology.
\item \textbf{Non-abelian charts and holonomy.} Light-cone $\mathfrak{so}(8)$ Lie
closure, G$_2\cup\{\Delta\}$ data, admissible holonomy on integrable Hopf shells,
cyclic plaquette flatness, and weak/colour commutator certificates are discharged on
the octonion carrier.  Matrix Wilson transport and Hopf-shell plaquette witnesses
\leanPatchHolonomyWitness; Wilson--kinetic equivalence
\leanPatchWilsonKinetic.  Partial / open: Haar measure on rotated charts
\leanPatchHaarMeasure; full strong-interacting discrete-action Poincaré
\leanPatchDiscretePoincare.  Optional strong-colour build: global Gell-Mann bracket
law on the finite $\mathfrak{su}(3)$ chart\leanCertStrongColor.
\item \textbf{Rapidity phase and Lorentz invariance.} The null-lattice carrier
charts into $1{+}1$ Minkowski space by rapidity $\Lambda(\eta)(1,1)$; rapidity
addition is equivariant under $\mathrm{boostMatrix11}$, the Minkowski pairing is
bilinearly invariant, and the polar-angle
\ifdefined\patchOmitZeta
phase
\else
/ \texttt{zetaHQIVTerm} phase
\fi
is unchanged
when $\phi$ transforms as a Lorentz scalar.
\item \textbf{Spatial rotations and directional readouts.} Orthogonal $O(3)$
rotations on $\mathrm{Fin}\,3$ preserve Euclidean inner products,
$\|{\mathbf r}\times{\mathbf v}\|^2$, and the spatial block of \texttt{minkowskiSq4}
\leanPatchSpatialRotation.  Full patch Lorentz is rapidity plus this spatial bundle
(witness \leanid{full_lorentz_closure_discharged}).  Continuum interacting QFT /
Wightman theorems are not claimed.
\end{itemize}
\else
\begin{itemize}[nosep]
\item \textbf{Anomaly cancellation.} For one Standard-Model generation
($Q_L$, $u^c$, $d^c$, $L$, $e^c$, $\nu^c$), the patch trace identities give
zero $U(1)_Y^3$, gravitational--$U(1)_Y$, $SU(3)_c^2U(1)_Y$,
$SU(2)_L^2U(1)_Y$, and $SU(3)_c^3$ anomaly coefficients; the $SU(2)_L^3$
coefficient vanishes because the weak doublet slot is pseudoreal.  Three
generations are a finite $\mathrm{Fin}\,3$ label sum of the same cancelled
one-generation block, not a fitted multiplicity\leanCertAnomaly.
\item \textbf{Topological sector.} The patch abelian gauge datum has a single
topological sector.  Instanton number, Pontryagin number, first Chern number, and
$U(1)$ winding number vanish identically on patch data; the $\theta$ contribution
is therefore $\theta$-independent; and abelian commutator obstructions vanish\leanCertTopology.
\item \textbf{Non-abelian charts and holonomy.} Light-cone $\mathfrak{so}(8)$ Lie
closure, G$_2\cup\{\Delta\}$ data, admissible holonomy on integrable Hopf shells,
cyclic plaquette flatness, and weak/colour commutator certificates are discharged on
the octonion carrier (see the sector zeta and holonomy discharge bundle in the Lean
catalog).  **Matrix Wilson layer:** full four-edge SO(8) generator transport and
Hopf-shell $T^4$ torsion plaquettes with pinned non-abelian witness
\leanPatchHolonomyWitness.  **Wilson--kinetic equivalence**
\leanPatchWilsonKinetic.  **Partial / open:** Haar
measure on rotated charts
\leanPatchHaarMeasure; general flat HQVM kinetic
boost and full strong-interacting discrete-action Poincaré
\leanPatchDiscretePoincare.  Where the optional
strong-colour certificate is invoked, the global Gell-Mann bracket law is proved on
the finite $\mathfrak{su}(3)$ chart\leanCertStrongColor.
\item \textbf{Rapidity phase and Lorentz invariance.} The null-lattice carrier
$\texttt{NullLatticeEvent}$ charts into $1{+}1$ Minkowski space by the classical
rapidity parametrization $\Lambda(\eta)(1,1)$; rapidity addition on the carrier is
equivariant under $\mathrm{boostMatrix11}$, the Minkowski pairing is bilinearly
invariant, and the polar-angle / \texttt{zetaHQIVTerm} phase is unchanged when
$\phi$ transforms as a Lorentz scalar.  The honest $3{+}1$ extension preserves
\texttt{minkowskiSq4} on the embedded $(t,x^1,0,0)$ plane.
\item \textbf{Spatial rotations and directional readouts.} Orthogonal $O(3)$
rotations on $\mathrm{Fin}\,3$ spatial vectors preserve Euclidean inner products,
$\|{\mathbf r}\times{\mathbf v}\|^2$, axis-projection fractions, and the spatial
block of \texttt{minkowskiSq4} when time is fixed
\leanPatchSpatialRotation.  Applied discharge:
CMB sky rotation shifts the Stokes angle additively while preserving
$\cos^2+\sin^2=1$; O-Maxwell fluid sources rotate covariantly with gradients;
coronal axis projections are joint rotation invariants.  Full patch Lorentz on the
flat chart is the rapidity bundle plus this spatial bundle
(witness \leanid{full_lorentz_closure_discharged}).  Continuum interacting QFT /
Wightman theorems are not claimed.
\end{itemize}
\fi

\paragraph{External continuum theorems not required.}
Several textbook continuum constructions are \emph{not} load-bearing prerequisites
because the patch carrier already fixes the traces, holonomy transport, topological
charges, and
\ifdefined\patchOmitZeta
finite spectral readout bodies
\else
sector zeta bodies
\fi
used for predictions: BPST instanton classification
on a separate smooth four-manifold, Nielsen-style principal-bundle completeness on
an external $C^\infty$ base, and path-integral measure choice on that base.  HQIV
does not claim to re-prove those external theorems; it proves the finite patch
statements that supply the same consistency checks for the readout pipeline.  This
is why familiar QFT language can appear in the comparison layer while the
predictions remain patch-level statements with fewer adjustable inputs than a
continuum effective fit.

\ifdefined\patchOmitZeta
% (zeta paragraph omitted --- not load-bearing for this note's proof cone)
\else
\ifdefined\patchBriefMode
\paragraph{Sector zeta and classical $\zeta$ (pointer).}
Sector zeta bodies and \texttt{zetaHQIVTerm} phase on Hopf shells are discharged on
the patch carrier in the TUFT synthesis \cite{hqiv-tuft-sm-lagrangian-paper};
classical $\zeta(s)$ as an SO(4) projection readout is developed separately in
\cite{hqiv-s3-zeta-so4-paper}.
\else
\paragraph{Projection geometry and the classical $\zeta$ dictionary (2026-06).}
HQIV uses \emph{zeta} in two distinct, now formally separated roles:
\begin{itemize}[nosep]
\item \textbf{Sector zeta on the patch carrier} (TUFT+SM synthesis): fibre-torsion
subleading bodies, \texttt{zetaHQIVTerm} phase on Hopf shells, and holonomy
discharge bundles on the octonion light-cone carrier---finite spectral readouts
indexed by shell and Fano sector, not the Riemann function.
\item \textbf{Classical $\zeta(s)$ as an SO(4) projection readout}
(\cite{hqiv-s3-zeta-so4-paper}): the Mathlib Riemann zeta packaged as a
\emph{regional} closed form on $\mathbb{C}$---Dirichlet/Euler for $\Re s>1$,
sin--cos--$\Gamma$ assembly on the strip, Bernoulli slots---with the
functional-equation pair $(\sigma,1-\sigma)$ interpreted as a $45^\circ$
quaternionic projection.  Machine-checked on this spine: harmonic divergence
forces $\Delta$ and $\mathfrak{so}(4)$ closure; shell sum $=\zeta(s)$ for
$\Re s>1$; polar decoupling $n^{-s}=n^{-\sigma}\cdot e^{-it\log n}$ (modulus
$\sigma$-only, phase $t$-only); prime-log Diophantine independence; Goldbach
circle $pq+(q-N)^2=N^2$ as additive curvature in the log channel; and a large
family of RH-\emph{equivalent} locators (twiddle, spectral weights, polar
collapse, transformer adjoint law, S$^7$ torsion cancellation) with
\emph{zero slack}---not a discharge of RH or Goldbach parity.
\end{itemize}
Continuum heat-flow and de~Bruijn--Newman notation in the same story spine are
\emph{comparison ports} ($\texttt{dbnHeatFamily}$, lattice dominance hypotheses);
they are not load-bearing continuum extensions of the patch.  Formal mathematical
notes may include only the readout-dictionary paragraph
(\texttt{readout\_dictionary\_messaging.tex}) rather than this full contract when
the manuscript is projection geometry rather than patch phenomenology.
\fi
\fi
