\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\nolinkurl{#1}}}
\newcommand{\leanid}[1]{\nolinkurl{#1}}

\title[Coronal heating from longitudinal HQIV stress]{A Longitudinal HQIV/O-Maxwell Channel for Coronal Heating:\\
Boundary-Form Stress on Photosphere-to-Corona Flux Tubes}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Preprint v1 --- \today}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
\noindent\textbf{Position in the publication chain.}
This is the tenth paper in the HQIV preprint series and the
second Tier-3 applied note, written back-to-back with the axial wire-force
companion \cite{ettinger2026longitudinal}. Tier~0--1 supply the O-Maxwell spine
\cite{ettinger2026hqiv,hqiv-lightcone-paper,hqiv-so8-paper,hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper};
Tier~2 closes cosmological and nucleon readouts
\cite{hqiv-tuft-sm-lagrangian-paper,hqiv-baryogenesis-paper,hqiv-nucleon-binding-paper,hqiv-bbn-paper}.

\noindent\textbf{Result.}
A coronal magnetic loop is the domain in which the wire-paper's three
preconditions coincide: an axial current channel, a photosphere--corona
$\varphi$ jump on the HQIV shell ladder, and asymmetric footpoint
``contacts.'' The \hyperref[lean:coronal-longitudinal]{coronal longitudinal
stress module} packages the boundary-form integrated force
\[
\Delta F_\parallel
  = A\,nq\,E_\ast\,\frac{3}{20\pi}\,\Lambda_s\,
    \bigl[\varphi_{\mathrm{cor}}-\varphi_{\mathrm{photo}}\bigr],
\]
the heating-flux analogue $Q_\parallel/A$, shell-gap scaling
$\Delta\varphi=2(m_{\mathrm{cor}}-m_{\mathrm{photo}})$, and even/odd-in-current
discriminants---all as sorry-free $\mathsf{Prop}$ certificates.

\noindent\textbf{Honesty.}
This note does \emph{not} derive a full stellar-atmosphere $\varphi(s)$ profile,
match observed coronal energy budgets in SI, or displace Alfv\'en-wave or
nanoflare candidates. It fixes the algebraic spine for a falsifiable \emph{third}
steady-state channel when longitudinal phase-gradient stress is nonzero.

\noindent\textbf{Patch-ontology reader contract.}
Coronal heating here is a discrete shell-ladder boundary readout, not a
continuum-limit completion programme.
\end{abstract}
\maketitle

\input{../include/patch_theory_messaging.tex}

\section{Scope and claim discipline}
\label{sec:scope}

The longitudinal wire note \cite{ettinger2026longitudinal} derives an axial
O-Maxwell source term from $\alpha\Lambda_s\partial_s\varphi$ and lists
laboratory discriminants (current reversal, contact asymmetry, length scans).
\textbf{This note instantiates the same mechanism on a stellar flux tube} and
states what Lean already proves about boundary-form heating.

\begin{enumerate}[leftmargin=2em]
\item \textbf{Proved in Lean.}
  Axial HQIV field with $\alpha=3/5$ rewrite
  (\hyperref[lean:hqiv-field-alpha]{HQIV field $\alpha$ form});
  vanishing when $\partial_s\varphi=0$
  (\hyperref[lean:flat-limit]{flat-limit collapse});
  boundary force and heating-flux boundary forms
  (\hyperref[lean:boundary-force]{boundary force},
  \hyperref[lean:heating-flux]{heating-flux boundary});
  shell-ladder $\varphi$ jump $\Delta\varphi=2(m_{\mathrm{cor}}-m_{\mathrm{photo}})$
  (\hyperref[lean:phi-jump]{shell $\varphi$ jump});
  fluid axial momentum source $g_{{\rm vac},s}=-\frac{1}{15}(\cdots)$
  (\hyperref[lean:vacuum-momentum]{vacuum momentum source});
  even/odd current decomposition for $I^2$ vs $I,I^3$ channels
  (\hyperref[lean:even-odd]{even/odd current split}).
\item \textbf{Interpretive (not Lean).}
  Mapping $(m_{\mathrm{photo}},m_{\mathrm{corona}})$ to observed atmospheric
  heights; absolute SI heating rates; network-wide energetics.
\item \textbf{Not claimed.}
  Replacement of wave heating or reconnection; peer-reviewed wire-force
  confirmation \cite{graneau2026}; horizon-shell derivation of solar $\varphi(s)$.
\end{enumerate}

\section{Why a coronal loop satisfies the wire preconditions}
\label{sec:geometry}

A force-free coronal flux tube carries axial current along $\mathbf B$.
In HQIV readout:

\begin{itemize}[leftmargin=2em]
\item \textbf{Axial channel.}
  $J_\parallel$ along the loop supplies the Ohmic background
  $E_{\mathrm{Ohm}}=J/\sigma$; in the corona $\sigma$ is enormous, so Ohmic
  heating is suppressed while the HQIV channel does not scale as $1/\sigma$
  (\hyperref[lean:ohmic-suppression]{high-conductivity Ohmic limit}).
\item \textbf{Nonzero $\partial_s\varphi$.}
  The discrete ladder gives $\varphi(m)=2(m+1)$ at shell $m$
  (\hyperref[lean:phi-of-shell]{shell $\varphi$ ladder}). A column with
  $m_{\mathrm{photo}}<m_{\mathrm{corona}}$ has positive
  $\varphi_{\mathrm{cor}}-\varphi_{\mathrm{photo}}$
  (\hyperref[lean:phi-jump-pos]{positive $\varphi$ jump}).
\item \textbf{Asymmetric contacts.}
  Opposite-polarity convecting footpoints play the role of oriented wire
  contacts, enabling odd-in-current components in the wire-paper test polynomial
  (\hyperref[lean:test-poly-odd]{odd part of force test polynomial}).
\end{itemize}

\section{Axial HQIV field and boundary force}
\label{sec:field}

Along the loop axis $s$, the HQIV electric channel is
\[
E_{\mathrm{HQIV}}
  = E_\ast\,\frac{\alpha}{4\pi}\,\Lambda_s\,\partial_s\varphi
  = E_\ast\,\frac{3}{20\pi}\,\Lambda_s\,\partial_s\varphi,
\qquad \alpha=\frac35,
\]
with $E_\ast$ the dimensionless-to-SI readout factor as in
\cite{ettinger2026longitudinal}. The effective axial field is
$E_{\mathrm{eff}}=E_{\mathrm{Ohm}}+E_{\mathrm{HQIV}}$; when $\partial_s\varphi=0$,
$E_{\mathrm{eff}}$ reduces to pure Ohmic
(\hyperref[lean:classical-effective]{classical effective-field limit}).

Integrating over a column of cross-section $A$ with approximately constant
$\Lambda_s$ gives the boundary form proved in Lean:
\[
\Delta F_\parallel
  = A\,nq\,E_\ast\,\frac{3}{20\pi}\,\Lambda_s\,
    \bigl[\varphi_{\mathrm{cor}}-\varphi_{\mathrm{photo}}\bigr].
\]
Equal footpoint $\varphi$ values imply $\Delta F_\parallel=0$
(\hyperref[lean:boundary-zero]{zero boundary force at equal $\varphi$}).
The \hyperref[lean:column-witness]{column boundary witness} packages the
integration step the wire paper performs explicitly.

\section{Heating-rate density and flux boundary}
\label{sec:heating}

Force density $f_\parallel=nq\,E_{\mathrm{eff}}$ yields heating-rate density
$\dot q=f_\parallel v_\parallel$. In the flat limit ($\partial_s\varphi=0$),
$\dot q$ collapses to Joule-style $nq(J/\sigma)v_\parallel$
(\hyperref[lean:heating-flat]{classical heating-rate limit}). With Ohmic channel
zeroed, the HQIV-only contribution is linear in $\partial_s\varphi$
(\hyperref[lean:heating-hqiv-only]{HQIV-only heating rate}).

The area-normalised heating-flux boundary form is
\[
\frac{Q_\parallel}{A}
  = nq\,v_\parallel\,E_\ast\,\frac{3}{20\pi}\,\Lambda_s\,
    \bigl[\varphi_{\mathrm{cor}}-\varphi_{\mathrm{photo}}\bigr],
\]
registered as \hyperref[lean:heating-flux]{heating-flux boundary} in Lean.
The end-to-end \hyperref[lean:heating-witness]{coronal HQIV heating witness}
records $\dot q$ for caller-supplied plasma parameters.

\section{Shell-ladder anchoring and length independence}
\label{sec:shell}

Given a shell pair $m_{\mathrm{photo}}\le m_{\mathrm{corona}}$,
\[
\Delta\varphi
  = \varphi(m_{\mathrm{corona}})-\varphi(m_{\mathrm{photo}})
  = 2\,(m_{\mathrm{corona}}-m_{\mathrm{photo}}),
\]
so the integrated boundary force scales with the \emph{shell index gap}, not
with bulk loop length $L$
(\hyperref[lean:shell-gap-force]{boundary force shell-gap form},
\hyperref[lean:shell-gap-flux]{heating flux shell-gap form}).
This is the coronal analogue of the wire-paper prediction that symmetric bulk
with oriented boundaries can show a boundary-dominated signal.

\section{Fluid momentum companion}
\label{sec:fluid}

The wire note's mechanical channel
\[
g_{{\rm vac},s}
  =
  -\frac{\gamma}{6}
  \bigl(\varphi\,\partial_s\dot{\delta\theta'}
        +\dot{\delta\theta'}\,\partial_s\varphi\bigr),
\qquad \gamma=\frac25,
\]
appears in Lean as \hyperref[lean:vacuum-momentum]{coronal axial vacuum
momentum source}, identified with component $0$ of the three-vector
$\mathbf g_{\rm vac}$ from the \hyperref[lean:fluid-scaffold]{fluid closure
scaffold}. With $\gamma=2/5$ the prefactor is $-1/15$
(\hyperref[lean:minus-fifteenth]{$-1/15$ prefactor identity}).

\section{Solar discriminants}
\label{sec:discriminants}

Translating the wire-paper tests to loops:

\begin{enumerate}[leftmargin=2em]
\item \textbf{Footpoint polarity / network asymmetry.}
  Odd-in-current components isolate oriented $\partial_s\varphi$ or
  $\partial_s\dot{\delta\theta'}$; opposite-polarity footpoints are the solar
  realisation of asymmetric contacts
  (\hyperref[lean:even-odd]{even/odd decomposition}).
\item \textbf{Length independence at fixed footpoints.}
  If heating flux depends only on $\varphi_{\mathrm{cor}}-\varphi_{\mathrm{photo}}$,
  lengthening the loop while holding footpoint anchoring fixed should modulate
  standard wave/reconnection channels differently from the HQIV boundary term.
\item \textbf{Sub-Alfv\'enic timing.}
  Phase-equilibration signatures can precede thermal diffusion along the loop,
  analogous to fast electrical pulses in wires.
\item \textbf{Quiet-Sun vs active-region networks.}
  Footpoint complexity and $\Lambda_s$ readout should correlate with odd-channel
  strength if the HQIV boundary term is physical.
\end{enumerate}

Use the wire-paper test polynomial with even part $a_2 I^2+a_4 I^4+F_0$ and odd
part $b_1 I+b_3 I^3$; Lean proves the decomposition
(\hyperref[lean:test-poly-even]{even part},
\hyperref[lean:test-poly-odd]{odd part}).

\section{What this note does not claim}
\label{sec:not-claimed}

\begin{itemize}[leftmargin=2em]
\item A derived $(m_{\mathrm{photo}},m_{\mathrm{corona}})$ pair from horizon
  thermodynamics alone.
\item Agreement with $10^3\,\mathrm{W\,m^{-2}}$ coronal flux budgets without an
  $E_\ast$ and plasma-parameter witness layer.
\item Displacement of Alfv\'en-wave dissipation or nanoflare reconnection.
\item Any continuum-limit or patch-refinement programme.
\end{itemize}

\section{Conclusion}

Coronal loops are the natural stellar laboratory for the longitudinal HQIV
stress mechanism: enormous conductivity suppresses Ohmic heating while a
photosphere--corona $\varphi$ jump on the shell ladder supplies a boundary-form
HQIV channel. The Lean module fixes every algebraic identity needed for a
falsifiable third heating channel; stellar profile derivation and SI energetics
remain programme targets. The wire companion \cite{ettinger2026longitudinal}
gives the laboratory protocol; this note gives the flux-tube instantiation.

\section*{Acknowledgments}

Machine-checked proofs live in the public HQIV Lean repository \cite{hqiv-lean}.

\appendix

\section{Lean module index}
\label{app:lean-index}

\begin{description}[leftmargin=2.5em,style=nextline]
\item[\hypertarget{lean:coronal-longitudinal}{Coronal longitudinal stress module}]
  \leanfile{Hqiv/Physics/CoronalLongitudinalStress.lean} --- umbrella for this paper.
\item[\hypertarget{lean:hqiv-field-alpha}{HQIV field $\alpha$ form}]
  \leanid{coronalLongitudinalHQIVField\_alpha\_3\_5}.
\item[\hypertarget{lean:flat-limit}{Flat-limit collapse}]
  \leanid{coronalLongitudinalHQIVField\_zero\_of\_dphi\_zero}.
\item[\hypertarget{lean:boundary-force}{Boundary force}]
  \leanid{coronalLongitudinalForceBoundary},
  \leanid{coronalLongitudinalForceBoundary\_eq\_alpha\_form}.
\item[\hypertarget{lean:boundary-zero}{Zero boundary force at equal $\varphi$}]
  \leanid{coronalLongitudinalForceBoundary\_zero\_of\_phi\_equal}.
\item[\hypertarget{lean:column-witness}{Column boundary witness}]
  \leanid{CoronalColumnBoundaryWitness}.
\item[\hypertarget{lean:ohmic-suppression}{High-conductivity Ohmic limit}]
  \leanid{ohmicAxialField\_eq\_zero\_of\_J\_zero}; HQIV channel not $1/\sigma$-suppressed
  (\leanid{coronalHeatingRateDensity\_hqiv\_only}).
\item[\hypertarget{lean:classical-effective}{Classical effective-field limit}]
  \leanid{coronalEffectiveAxialField\_classical\_limit}.
\item[\hypertarget{lean:heating-flat}{Classical heating-rate limit}]
  \leanid{coronalHeatingRateDensity\_classical\_flat\_limit}.
\item[\hypertarget{lean:heating-hqiv-only}{HQIV-only heating rate}]
  \leanid{coronalHeatingRateDensity\_hqiv\_only}.
\item[\hypertarget{lean:heating-flux}{Heating-flux boundary}]
  \leanid{coronalHeatingFluxBoundary},
  \leanid{coronalHeatingFluxBoundary\_eq\_alpha\_form}.
\item[\hypertarget{lean:heating-witness}{Coronal HQIV heating witness}]
  \leanid{CoronalHQIVHeatingWitness}.
\item[\hypertarget{lean:phi-of-shell}{Shell $\varphi$ ladder}]
  \leanid{phi\_of\_shell}, \leanid{phi\_of\_shell\_closed\_form} (imported in module).
\item[\hypertarget{lean:phi-jump}{Shell $\varphi$ jump}]
  \leanid{coronalPhiJump}, \leanid{coronalPhiJump\_closed\_form}.
\item[\hypertarget{lean:phi-jump-pos}{Positive $\varphi$ jump}]
  \leanid{coronalPhiJump\_pos\_of\_lt}.
\item[\hypertarget{lean:shell-gap-force}{Boundary force shell-gap form}]
  \leanid{coronalLongitudinalForceBoundaryShells\_eq\_two\_shellGap}.
\item[\hypertarget{lean:shell-gap-flux}{Heating flux shell-gap form}]
  \leanid{coronalHeatingFluxBoundaryShells\_eq\_two\_shellGap}.
\item[\hypertarget{lean:vacuum-momentum}{Vacuum momentum source}]
  \leanid{coronalAxialVacuumMomentumSource}.
\item[\hypertarget{lean:minus-fifteenth}{$-1/15$ prefactor identity}]
  \leanid{coronalAxialVacuumMomentumSource\_eq\_minus\_one\_fifteenth}.
\item[\hypertarget{lean:fluid-scaffold}{Fluid closure scaffold}]
  \leanfile{Hqiv/Physics/HQIVFluidClosureScaffold.lean}.
\item[\hypertarget{lean:even-odd}{Even/odd current split}]
  \leanid{evenInCurrent}, \leanid{oddInCurrent},
  \leanid{evenInCurrent\_pow\_two}, \leanid{oddInCurrent\_id}.
\item[\hypertarget{lean:test-poly-even}{Even part of force test polynomial}]
  \leanid{coronalForceTestPoly\_even\_part}.
\item[\hypertarget{lean:test-poly-odd}{Odd part of force test polynomial}]
  \leanid{coronalForceTestPoly\_odd\_part}.
\item[\hypertarget{lean:o-maxwell-upstream}{O-Maxwell upstream}]
  \leanfile{Hqiv/Physics/OMaxwellAlgebraSeed.lean},
  \leanfile{Hqiv/Physics/ModifiedMaxwell.lean}.
\end{description}

\bibliographystyle{plain}
\bibliography{../references}

\end{document}
