\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,booktabs,array,tabularx}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\usepackage[final]{microtype}
\setlength{\emergencystretch}{3em}

\newcommand{\leanrepo}{https://github.com/HQIV/hqiv-lean/blob/main}
\newcommand{\leanrepourl}{https://github.com/HQIV/hqiv-lean}
\newcommand{\programmeurl}{https://disregardfiat.tech}
\newcommand{\scriptrepo}{https://github.com/HQIV/hqiv-lean/blob/main/scripts}
\newcommand{\leanfile}[1]{\href{\leanrepo/#1}{\nolinkurl{#1}}}
\newcommand{\scriptfile}[1]{\href{\scriptrepo/#1}{\nolinkurl{scripts/#1}}}
\newcommand{\leanid}[1]{\nolinkurl{#1}}
\newcommand{\LeanEng}[2]{\hyperref[#1]{#2}}
\newcommand{\datafile}[1]{\texttt{#1}}

\newenvironment{summarybox}{%
  \medskip\begin{quote}\small
}{%
  \end{quote}\medskip
}

\title[Curvature-ratio lock-in on the HQIV spine]{Curvature-Ratio Lock-In for Baryon Asymmetry:\\
Demonstrated on the HQIV Discrete Spine (Lean Formalisation)}
\author{Steven Ettinger Jr}
\address{Independent Researcher}
\email{steven@disregardfiat.tech}
\date{Preprint v3 --- \today}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\begin{document}

\begin{abstract}
This note---the seventh HQIV preprint---is a capability demonstration on
the shared discrete curvature-ratio spine. Baryon-asymmetry bookkeeping is
already a theorem-level consequence of the same lock-in calibration,
monotonicity results, and triality scaffolds used for hadron structure,
BBN, and other applications in the library. At lock-in shell $m=4$,
horizon ratios $\Omega_k(n;N)=I(n)/I(N)$ are proved monotone with
$\Omega_k(M;M)=1$, and the baryon readout $\eta(n;N)$ shares that
normalisation; a single quarantined anchor $\eta_{\mathrm{paper}}$
supplies the comparison-layer scale. The primary artefacts are the Lean
formalisation and the Python reproduction bundle
(\S\ref{sec:repro}). Programme context, witness panels, and the
proved/witnessed/deferred split are in \S\ref{sec:intro}; Lean module
and identifier names are in Appendix~\ref{app:lean-index}.
\end{abstract}
\maketitle

\input{../include/patch_theory_messaging.tex}

\section{Introduction}
\label{sec:intro}

\subsection{Programme context and doorway}
This note is a \textbf{Tier-2 checkpoint}, not the programme's goal.
The HQIV objective is a discrete curvature-ratio ledger on the null
lattice---with sector projections to hadronic binding, electroweak
synthesis, chemistry, compact objects, and cosmology---not a standalone
theory of baryogenesis. Baryon-asymmetry bookkeeping is one cosmological
\emph{application}: it records how the same lock-in calibration,
$\Omega_k$ monotonicity, and quarantined $\eta_{\mathrm{paper}}$ anchor
answer a single ledger question at shell $m=4$.

Geometry and ratios live in the core formal modules
(\LeanEng{lean:lightcone}{octonionic lightcone},
\LeanEng{lean:baryogenesis-core}{baryogenesis core},
\LeanEng{lean:baryogenesis-witness}{baryogenesis witness};
Appendix~\ref{app:lean-index}); $\eta$ bookkeeping is how that spine
answers one comparison-layer question. Readers should treat this note as
an on-ramp to the larger corpus, not as an isolated claim about observed
$\eta$ alone.

\begin{itemize}[leftmargin=2em,nosep]
\item \textbf{Plain-language tour:} \href{\programmeurl}{disregardfiat.tech}
  (clickable module anchors).
\item \textbf{Formal library:} \href{\leanrepourl}{HQIV-LEAN}
  \cite{hqiv-lean}.
\item \textbf{Reproduction bundle:} \texttt{scripts.zip} in this folder
  (\S\ref{sec:repro}).
\item \textbf{Companion BBN stack:} \cite{hqiv-bbn-paper}.
\end{itemize}

\subsection{Publication chain}
Tier~0 fixes the framework
\cite{ettinger2026hqiv,hqiv-lightcone-paper,hqiv-so8-paper}; Tier~1
supplies the variational and lattice spine
\cite{hqiv-3d-growth-paper,hqiv-oct-action-paper,hqiv-kirchhoff-paper,hqiv-thermo-arrow-paper};
Tier~2 adds the published Lagrangian bridge
\cite{hqiv-tuft-sm-lagrangian-paper}. The present note states the
strongest \emph{ratio-level} baryogenesis theorem currently supported:
baryogenesis is not an independent knob---it is the same partial
curvature imprint as the growth channel, triality CP-bias scaffolds, and
the HQVM Friedmann compatibility package.

The same spine already shows machine-checked readouts across applied
scales---nucleon binding \cite{hqiv-nucleon-binding-paper}, TUFT+SM
synthesis \cite{hqiv-tuft-sm-lagrangian-paper}, chemistry and phase
geometry, compact-object witnesses, and the BBN integrator
\cite{hqiv-bbn-paper}. This note records how the baryon-asymmetry ratio
slots into that spine at lock-in.

\subsection{Formal results (summary)}
Discrete shell combinatorics define a cumulative curvature integral
$I(n)$; horizon ratios $\Omega_k(n;N)=I(n)/I(N)$ are machine-checked
with proved monotonicity at fixed $N$; the lock-in shell
$m_{\mathrm{lockin}}=M=4$ calibrates $\Omega_k=1$ at the lock-in
horizon; and $\eta(n;N)$ is \emph{defined} to share that same
normalisation---the convention the companion BBN note
\cite{hqiv-bbn-paper} takes at lock-in. Light-element abundances are
\emph{not} claimed here.

\subsection{Quarantined anchor and reader contract}
The observed-order scale $\eta_{\mathrm{paper}}$ appears in exactly one
formal module (\LeanEng{lean:eta-paper}{paper $\eta$ anchor module}) and
is wired only through the witness layer. Ratio laws and lock-in
calibration are proved without it. Under proton lock-in witness discipline
\cite{hqiv-tuft-sm-lagrangian-paper,hqiv-kirchhoff-paper}, laboratory
$\eta$ is a comparison target, not a simultaneous fit row alongside other
PDG inputs.

Per the patch-theory reader contract below, the discrete patch layer is
the ontology. Smooth-manifold transport, washout, and sphaleron yield are
comparison-layer targets---not foundational requirements.

\subsection{Dynamic bulk and BBN witnesses (v3)}
A per-shell matter integrator derives comparison-layer $\Omega_b$ and
$\eta$ without importing $\eta_{\mathrm{paper}}$ as an input. Central
readout: $\eta_{10}=6.20$, $\Omega_b=0.0499$ at lock-in
(\datafile{data/dynamic\_bulk\_bbn\_v2.json}; \S\ref{sec:repro}). At
$\eta_{\mathrm{paper}}=6.10\times 10^{-10}$, the faithful stoichiometric
BBN integrator yields $Y_p=0.244$, $D/H=2.51\times 10^{-5}$,
$^3\mathrm{He}/H\approx 1.01\times 10^{-5}$, and
$^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$ (pre-depletion;
$\approx 45\%$ below Coc semi-analytic $4.65\times 10^{-10}$ at the same
$\eta$ \cite{coc2015}; $z(D/H)\approx -0.45$). Legitimate $\eta$ stages
bracket $6.0$--$6.4$ (Tables~\ref{tab:eta-methods}--\ref{tab:bbn-methods}).
Full abundance certificates remain in \cite{hqiv-bbn-paper}.

\subsection{Proved, witnessed, and deferred}
\begin{itemize}[leftmargin=2em,nosep]
\item \textbf{Proved (Lean):} monotonicity of $I$ and $\Omega_k$; lock-in
  calibration; parallel $\eta$ laws; triality tilt averaging; GR/O--Maxwell
  compatibility; dynamic-bulk ratio packaging
  (\LeanEng{lean:dynamic-bulk-vital}{dynamic bulk vital bundle}).
\item \textbf{Witnessed (Python):} dynamic bulk $\eta_{10}\approx 6.20$,
  $\Omega_b\approx 0.050$; faithful BBN panel at $\eta_{\mathrm{paper}}$.
\item \textbf{Deferred:} associator-derived $\eta_{\mathrm{paper}}$; SM
  washout/sphaleron yield; closed-form $\eta$ without comparison-layer
  $H_0$, $T_{\mathrm{CMB}}$, or $\Omega_b$.
\end{itemize}

All numerical tables are regenerated from the standalone Python bundle
(\S\ref{sec:repro}); no Lean installation is required to verify witnesses.

\section*{Imported spine (at a glance)}
\label{sec:imported-spine}

This note proves ratio-level baryogenesis bookkeeping on the discrete shell
ladder. The following inputs are \emph{imported} from earlier Tier-0--2
records; they are listed here so a reader need not open those manuscripts
to see what is assumed versus what is derived here. Programme context is
in \S\ref{sec:intro}.

\begin{summarybox}
\noindent\textbf{What this paper adds (new relative to upstream Tier-0--2).}
\begin{itemize}[leftmargin=1.5em,nosep]
\item Horizon-ratio theorems $\Omega_k(n;N)=I(n)/I(N)$ with strict
  monotonicity and lock-in calibration $\Omega_k(M;M)=1$.
\item Parallel $\eta(n;N)$ law sharing the identical normalisation as
  $\Omega_k$ (\LeanEng{lean:eta-omega-same}{shared-normalisation theorem}).
\item Triality CP-bias scaffold with proved three-representation averaging
  (\LeanEng{lean:rapidity-cp-bias}{rapidity CP-bias},
  Appendix~\ref{app:lean-index}).
\item Lean packaging of dynamic-bulk colour-singlet fraction, $\Omega_b$
  channel, and comparison-layer $\eta$ bridge
  (\LeanEng{lean:dynamic-bulk-vital}{dynamic bulk vital bundle}).
\item v3 Python witnesses: per-shell $\eta_{10}\approx 6.20$,
  $\Omega_b\approx 0.050$, and faithful BBN panel at $\eta_{\mathrm{paper}}$
  (Tables~\ref{tab:eta-methods}--\ref{tab:bbn-methods}).
\end{itemize}
\end{summarybox}

\begin{table}[ht]
\centering
\caption{Imported quantities from upstream HQIV papers (not refit here).
Lean file paths: Appendix~\ref{app:lean-index}.}
\label{tab:imported-spine}
\footnotesize
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash}p{0.19\linewidth}>{\raggedright\arraybackslash}p{0.28\linewidth}X@{}}
\toprule
Quantity & Upstream source & Role in this note \\
\midrule
$\alpha=3/5$, $\gamma=2/5$, shell shape $\sigma(m)$ &
  Tier-0--1: \cite{hqiv-so8-paper,hqiv-3d-growth-paper,hqiv-oct-action-paper} &
  Curvature imprint ladder; $I(n)$, $\Omega_k$ \\
$\mathcal{N}_{\mathrm{curv}}=6^7\sqrt{3}$, $\delta_E(m)$ &
  \cite{hqiv-so8-paper,hqiv-lightcone-paper} &
  Per-shell imprint; octonion/quaternion contrast \\
$m_{\mathrm{lockin}}=4$, $\xi_{\mathrm{lock}}=5$ &
  \cite{hqiv-kirchhoff-paper,hqiv-tuft-sm-lagrangian-paper,hqiv-so8-paper} &
  Lock-in horizon; $\Omega_k(M;M)=1$. First shell with full
  $G_2\cup\{\Delta\}\Rightarrow\mathfrak{so}(8)$ sector spine
  ($N_{\mathrm{new}}(4)=40$; octonionic lightcone module,
  App.~\ref{app:lean-index}). \\
$\kappa_6=\eta_{\mathrm{local}}\,\gamma\,C_2(\xi)$, $C_2=56/45$ at lock-in &
  \cite{hqiv-tuft-sm-lagrangian-paper} &
  Matter slot; dynamic-$C_2$ BBN clock \\
Lock-in nuclear $Q$'s ($Q_D$, $Q_{np}$, cluster traces) &
  \cite{hqiv-nucleon-binding-paper,hqiv-tuft-sm-lagrangian-paper} &
  Binding-release factor $f_{\mathrm{bind}}(T)$; BBN network \\
$\eta_{\mathrm{paper}}=6.1\times 10^{-10}$ (quarantined) &
  Paper $\eta$ anchor module (App.~\ref{app:lean-index}) &
  Single absolute anchor; parallel to $\Omega_k$ \\
Comparison $H_0$, $T_{\mathrm{CMB}}$, $\Omega_b$ &
  \cite{planck2018cosmo,pdg2024} (comparison only) &
  Late-universe $\eta$ conversion layer \\
Observed BBN band ($Y_p$, $D/H$; Coc $^7\mathrm{Li}/H\approx 4.65\times 10^{-10}$) &
  \cite{coc2015} (comparison only) &
  Audit targets; HQIV faithful $^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$
  is pre-depletion model output \\
\bottomrule
\end{tabularx}
\end{table}

\begin{summarybox}
\noindent\textbf{Derived in this note (not imported).}
Horizon ratios $\Omega_k(n;N)=I(n)/I(N)$ with strict monotonicity;
lock-in calibration $\Omega_k(M;M)=1$; parallel $\eta(n;N)$ laws sharing
the same normalisation; triality CP-bias scaffold; GR/O--Maxwell $(\alpha,\phi)$
compatibility \cite{hqiv-oct-action-paper}; dynamic-bulk comparison-layer
$\eta_{10}\approx 6.20$ and $\Omega_b\approx 0.050$ without importing
$\eta_{\mathrm{paper}}$ (\S\ref{sec:dynamic-bulk}, Table~\ref{tab:eta-methods}).
\end{summarybox}

\section{Scope and claim discipline}
\label{sec:scope}

The HQIV programme ties discrete null-shell counting, curvature imprint
exponent $\alpha=3/5$ \cite{brodie2026}, and an eight-dimensional
octonionic carrier with machine-checked
$G_2\cup\{\Delta\}\Rightarrow\mathfrak{so}(8)$ data
\cite{hqiv-so8-paper,hqiv-3d-growth-paper}. See \S\ref{sec:intro} for
programme position and on-ramps.

\paragraph{Lock-in shell $m=4$ and the $G_2$ sector spine.}
The lock-in index $M=4$ is not an arbitrary fit row: on
the discrete null ladder it is the \textbf{first shell that carries enough
degrees of freedom} to support the full octonion-sector readout used
downstream. At $m=4$ the library proves $N_{\mathrm{new}}(4)=40$ new
lattice modes on the carrier
(\LeanEng{lean:lightcone}{octonionic lightcone module}), while the
algebraic closure $G_2\cup\{\Delta\}\Rightarrow\mathfrak{so}(8)$ supplies
fourteen $G_2$ generators plus the phase-lift harmonic $\Delta$ closing to
the full $28$-dimensional gauge-sector carrier
\cite{hqiv-so8-paper}. Proton lock-in, $\Omega_k(M;M)=1$, and the
baryon-asymmetry ratio normalisation are all evaluated at this same row
($\xi_{\mathrm{lock}}=m+1=5$). Earlier shells are pre-sector or
pre-closure rungs on the same ladder; later shells extend the chart but do
not replace the calibration pin.

\paragraph{Baryogenesis as an application.}
This note is \textbf{not} a standalone theory of baryogenesis; it is one
checkpoint on a spine that already shows promising results at every applied
scale exercised in the repository so far---hadronic binding, electroweak
synthesis, molecular phase geometry, compact-object witnesses, and BBN
abundance integration. The formal objective remains the discrete
curvature-ratio ledger and its sector projections; $\eta$ bookkeeping is
how that ledger answers a cosmological question at lock-in.

This note answers: \textbf{what baryogenesis statement is already a
theorem, and what is still interpretive physics?} Geometry and ratios
live in the \LeanEng{lean:baryogenesis-core}{baryogenesis core module}
and the \LeanEng{lean:lightcone}{octonionic lightcone module}; the
$\eta_{\mathrm{paper}}$ anchor is isolated in the
\LeanEng{lean:eta-paper}{paper $\eta$ anchor module} and wired
only through the \LeanEng{lean:baryogenesis-witness}{baryogenesis witness
module} (Appendix~\ref{app:lean-index}).

\begin{enumerate}[leftmargin=2em]
\item \textbf{Proved in Lean (formal spine).} Shell indices, ladder
  temperatures, combinatorial curvature norm $\mathcal{N}_{\mathrm{curv}}$,
  per-shell imprint $\delta_E(m)$, cumulative integral $I(n)$ with strict
  monotonicity (\LeanEng{lean:curvature-integral-mono}{curvature-integral
  strict-monotonicity theorem}), horizon ratios $\Omega_k(n;N)$ with
  monotonicity, strict monotonicity, and the bound $\Omega_k\le 1$ for
  $n\le N$ (\LeanEng{lean:omega-k-mono}{horizon-ratio monotonicity
  pack}), lock-in calibration $\Omega_k(M;M)=1$
  (\LeanEng{lean:omega-k-lockin}{lock-in $\Omega_k$ calibration}), parallel
  $\eta(n;N)$ with matching monotonicity lemmas
  (\LeanEng{lean:eta-mono}{$\eta$ monotonicity pack}), and the identity
  $\eta=\eta_{\mathrm{paper}}\cdot\Omega_k$ at fixed horizon
  (\LeanEng{lean:eta-omega-same}{shared-normalisation theorem}).
\item \textbf{Proved dynamic-bulk packaging (v3).}
  Colour-singlet projector $f_{\mathrm{color}}=1/3$, baryon channel
  $(4/8)\cdot(1/3)=1/6$, dimensionless binding-curvature correction linked
  to MeV cluster binding, dynamic $\eta$ at lock-in with dimensionless
  feedback, and comparison-layer $\eta$ bridge
  (\LeanEng{lean:dynamic-bulk-vital}{dynamic bulk vital bundle};
  Appendix~\ref{app:lean-index}).
\item \textbf{Proved algebraic packaging.} Triality-sensitive increments
  from the \LeanEng{lean:rapidity-cp-bias}{rapidity CP-bias} in the
  \LeanEng{lean:triality-well}{triality--rapidity-well module}; affine
  tilt averaging (\LeanEng{lean:triality-tilt-avg}{triality tilt
  average identity}); three-representation averaging back to the rapidity
  well (\LeanEng{lean:triality-inc-avg}{triality increment average
  identity}).
\item \textbf{Structural compatibility (not transport).} The
  \LeanEng{lean:gr-from-maxwell}{GR-from-Maxwell module}: shared
  $\phi$, shared $\alpha$, and homogeneous HQVM Friedmann equivalence---a
  compatibility bridge, not a washout-limited yield theorem.
\item \textbf{Companion: BBN \cite{hqiv-bbn-paper}.}
  The BBN formal stack---umbrella certificate, network-from-weights readout,
  epoch ladder, and integrated witness (Appendix~\ref{app:lean-index})---is
  documented in the companion note; it consumes lock-in $\eta$ from this note
  and is \emph{not} part of this note's proof graph. Coc et al.\ semi-analytic
  fits in the library are labelled comparison-only, not HQIV inputs.
\item \textbf{Yield-level SM physics (not Lean).} Associator/vorticity
  sources, QCD-window maximisation, and sphaleron rates remain narrative
  targets beyond both this note and \cite{hqiv-bbn-paper}. No
  continuum-spacetime limit is invoked or required.
\end{enumerate}

\section{Discrete shells, temperatures, and the baryogenesis window}

Fix the HQIV discrete shell index $m\in\mathbb{N}$ as in
\cite{hqiv-so8-paper,hqiv-3d-growth-paper}. The library defines a QCD
onset shell index $m_{\mathrm{QCD}}$ and a lock-in shell
\[
m_{\mathrm{lockin}} \;:=\; m_{\mathrm{QCD}}+k_{\mathrm{QCD}\to\mathrm{lock}},
\]
with fixed step count $k_{\mathrm{QCD}\to\mathrm{lock}}$ (Appendix~\ref{app:lean-index}),
aligned with $M=4$ (the same lock-in convention used
throughout the formalisation and pinned by the proton lock-in witness in
\cite{hqiv-tuft-sm-lagrangian-paper,hqiv-kirchhoff-paper}). Temperatures on the
ladder are $T(m)=1/(m+1)$ in natural units
(\LeanEng{lean:T-QCD-lockin}{QCD and lock-in temperature closures} in the
baryogenesis core module).

The finite shell set used for discrete bookkeeping is the explicit
\LeanEng{lean:baryogenesis-shells}{baryogenesis shell window}: a finite
interval from $m_{\mathrm{QCD}}$ through a few shells after
lock-in. This is the honest discrete substitute for a continuum ``GeV
window'': it is \emph{not} proved from washout maximisation in Lean; it is
the packaged combinatorial stage on which curvature ratios are evaluated.

\section{Curvature imprint, \(\mathcal{N}_{\mathrm{curv}}\), and \(\delta_E\)}

\begin{summarybox}
\noindent\textbf{Imported from Tier-0--1 (summary).}
The shell-shape factor $\sigma(m)$ and curvature exponent $\alpha=3/5$ enter
from the growth trilogy \cite{hqiv-3d-growth-paper,hqiv-so8-paper}; the
octonionic-action note \cite{hqiv-oct-action-paper} documents the same
$(\alpha,\phi)$ slots in O--Maxwell stationarity. The combinatorial norm
$\mathcal{N}_{\mathrm{curv}}$ and per-shell imprint $\delta_E(m)$ are proved
in the lightcone module---not re-derived here. See
Table~\ref{tab:imported-spine}.
\end{summarybox}

Let $\sigma(m)$ denote the shell-shape factor (the
\LeanEng{lean:shell-shape}{shell-shape factor}), built from the same
$\alpha$-weighted logarithmic curvature density used in the growth trilogy.
The combinatorial curvature norm is the exact real scalar
$\mathcal{N}_{\mathrm{curv}} = 6^{7}\sqrt{3}$, proved in the lightcone
module from the structural ``six cube directions, seven imaginary octonion
directions, half-diagonal $\sqrt{3}$'' bookkeeping (the
\LeanEng{lean:curvature-norm}{curvature-norm structural theorem} and
related lemmas). The per-shell imprint is
$\delta_E(m)=\mathcal{N}_{\mathrm{curv}}\cdot\sigma(m)$ (the
\LeanEng{lean:deltaE-QCD}{QCD-shell imprint} specialises at
$m_{\mathrm{QCD}}$).

\medskip
\noindent\textbf{Structural rigidity (octonionic vs.\ quaternionic comparison).}
Lean also proves that a natural quaternionic Maxwell comparison norm
undershoots the HQIV octonionic norm by a rigid factor $1296$ shellwise
(the \LeanEng{lean:deltaE-1296}{octonion--quaternion imprint ratio
theorem}). This supports the narrative that $\mathcal{N}_{\mathrm{curv}}$
is not an arbitrary normalisation: it is the imprint constant selected by
the octonionic lift inside the same eight-dimensional carrier used for
gauge closure.

\section{Horizon ratios \(\Omega_k\) and lock-in calibration}

Define the cumulative partial integral
\[
I(n) \;:=\; \sum_{m=0}^{n-1} |\sigma(m)|
\]
(the \LeanEng{lean:curvature-integral}{curvature integral}). For horizons
$N$ with $I(N)>0$, the library defines
$\Omega_k(n;N):=I(n)/I(N)$. At the lock-in shell, Lean proves positivity
$I(m_{\mathrm{lockin}})>0$ and the calibration
$\Omega_k(m_{\mathrm{lockin}};m_{\mathrm{lockin}})=1$ (the
\LeanEng{lean:omega-k-lockin}{lock-in $\Omega_k$ calibration}). This is
the precise sense in which ``spatial curvature lock-in'' is a
\textbf{theorem}: normalisation of the discrete imprint integral at the
declared reference horizon, not an experimental input.

\begin{proposition}[Monotonicity and approach to lock-in, Lean]
\label{prop:mono-lockin}
Assume $I(N)>0$. Then for all $n_1\le n_2$,
$\Omega_k(n_1;N)\le \Omega_k(n_2;N)$, with strict inequality when
$n_1<n_2$ (the \LeanEng{lean:omega-k-mono}{horizon-ratio monotonicity
pack}). If additionally $n\le N$, then $\Omega_k(n;N)\le 1$. In
particular, along shells $n\le m_{\mathrm{lockin}}$ with horizon
$N=m_{\mathrm{lockin}}$, the ratio $\Omega_k(n;m_{\mathrm{lockin}})$
increases strictly toward $1$ as $n$ approaches $m_{\mathrm{lockin}}$
from below.
\end{proposition}

\section{Baryon asymmetry \(\eta\) as the same normalization (optional absolute anchor)}

\paragraph{Without importing any experimental $\eta$.}
Everything in the baryogenesis core module is logically independent of
PDG-scale baryon asymmetry.

\paragraph{With a single absolute anchor (quarantined).}
The constant $\eta_{\mathrm{paper}}$ appears \emph{only} in the
\LeanEng{lean:eta-paper}{baryogenesis $\eta$ paper module}. The witness
module defines $\eta(n;N):=\eta_{\mathrm{paper}}\cdot I(n)/I(N)$ when
$I(N)>0$ (and a conservative fallback otherwise), and proves the parallel
identities to the $\Omega_k$ ratio (the
\LeanEng{lean:eta-at-horizon}{horizon $\eta$ readout} and
\LeanEng{lean:eta-self}{self-horizon $\eta$ identity} at $n=N$).

The conceptual punchline is already a theorem:
\[
\eta(n;N) \;=\; \eta_{\mathrm{paper}}\cdot \Omega_k(n;N)
\]
(the \LeanEng{lean:eta-omega-same}{shared-normalisation theorem}). At
fixed horizon, baryogenesis is not a second independent imprint law: it is
the same partial-integral ratio as $\Omega_k$, scaled by one overall anchor.

\medskip
\noindent\textbf{Same monotonicity and bounds as $\Omega_k$.}
Lean also proves $\eta(n_1;N)\le \eta(n_2;N)$ for $n_1\le n_2$ (the
\LeanEng{lean:eta-mono}{$\eta$ monotonicity pack}), and
$\eta(n;N)\le \eta_{\mathrm{paper}}$ whenever $n\le N$. Thus
$\eta_{\mathrm{paper}}$ is both the absolute scale at lock-in and an
upper envelope for readouts inside the horizon---a direct consequence of
shared monotonicity with $\Omega_k$.

\paragraph{Honesty boundary (first principles vs.\ calibration).}
Lean does not currently derive $\eta_{\mathrm{paper}}$ from associator
phases or SM chemical potentials. What \emph{is} first-principles inside
the formalisation is the shared ratio law and the lock-in calibration of
$\Omega_k$. Attaching the observed $\eta$ scale is a \emph{single-point
calibration} quarantined to one module---analogous to how the mass and
Kirchhoff papers treat amplitude normalisation while keeping imprint
\emph{shape} combinatorially rigid.

\section{Dynamic bulk matter and method range for \texorpdfstring{$\eta$}{eta}}
\label{sec:dynamic-bulk}

The ratio spine of \S\ref{sec:scope}--\S5 fixes how $\eta(n;N)$ is
\emph{normalized} once an absolute anchor is declared. A separate Python
witness chain now asks what $\eta$ is when that anchor is \emph{not}
imported: baryon budget is accumulated shell-by-shell from curvature
imprint, Casimir inside/outside balance, and composite-trace binding
feedback, then converted to $\eta$ only at a late comparison layer using
present-day $H_0$, $T_{\mathrm{CMB}}$, and $\Omega_b$.

\paragraph{Per-shell integrator (v3).}
The dynamic bulk witness (\S\ref{sec:repro}) evolves
shells $m_{\mathrm{QCD}}\le m\le m_{\mathrm{lock}}$ with emergent
hadronic lock-in at $m=4$ ($\xi_{\mathrm{lock}}=5$). At each step it
reads shell shape $\sigma(m)$, inner/outer curvature proxies, the
per-shell curvature budget $B_{\mathrm{curv}}(m)$ (early matter--antimatter /
radiation seed relaxing to $1$ at lock-in; excess imprint outside the baryon track),
a temperature-relative vev factor from the inner/outer Casimir balance
(described below), and dynamic proton mass and nuclear $Q$ values from
the $8\times8$ composite-trace network. Matter increments
\[
\Delta S_m = \sigma(m)\,f_{\mathrm{vev}}(\xi_m)
  + w_{\mathrm{strong}}\,f_{\mathrm{bind}}(m,\xi_m)\,B_{\mathrm{curv}}(\xi_m)
\]
accumulate into a cumulative source $C=\sum_m \Delta S_m$. Baryon and
total matter fractions follow
\[
\Omega_m = \gamma\, C/(C+\rho_{\mathrm{rad}}),
\qquad
\Omega_b = \Omega_m \cdot w_{\mathrm{strong}} \cdot f_{\mathrm{color}},
\]
with HQIV structural weights $\gamma=2/5$, $w_{\mathrm{strong}}=4/8$,
and a colour-singlet projector $f_{\mathrm{color}}=1/3$ (trace-derived
sweep below). The inner/outer Casimir scale uses a dynamic \emph{outer}
suppression $\omega K(\xi)/140$ (not a fixed $1/140$ at all epochs;
the integer $140$ counts discrete fluctuation modes on the outer
neutral horizon shell at lock-in, while the amplitude runs with
$\omega K(\xi)$ like the inner trapped-Casimir factor). The witness
reports $C=3.80$, $\Omega_b=0.0499$, $\Omega_m=0.299$ at $m=4$
(\datafile{dynamic\_bulk\_bbn\_v2.json}, lock-in row).

\paragraph{What ``inside'' and ``outside'' mean (reader-facing).}
HQIV treats the octonionic carrier as having two competing Casimir-like
surfaces on the \emph{same} geometric object, not two unrelated constants.
\begin{itemize}[leftmargin=2em]
\item \textbf{Inside (contact shells).}
  On the three inner Hopf contact shells (the heavy-sector winding
  ladder), modes are ``trapped'' by phase-lift and Beltrami torsion.
  This inner surface sets the scale of binding, the heavy-lepton gap,
  and the hadronic chart.  Its strength grows with the curvature
  readout $\omega K(\xi)$ on the temperature ladder.
\item \textbf{Outside (neutral horizon shell).}
  On the first shell \emph{beyond} proton lock-in, fluctuations on the
  right-handed neutrino/neutral channel of the same carrier are
  coarse-grained into a finite number of discrete modes.  At lock-in
  that count is $140$, giving the familiar factor $1/140$ when the
  curvature amplitude is normalized to unity.  Away from lock-in the
  outer suppression is $\omega K(\xi)/140$, parallel to the inner side.
\end{itemize}
The overall mass scale and vev factor are the ratio inner/outer.  The
formal library records this as the \emph{outer neutral-shell fluctuation
witness} (Appendix~\ref{app:lean-index})---not a standard physics
abbreviation and not a particle label.

\paragraph{Comparison-layer $\eta$.}
Only after the shell ladder is integrated does the witness form
\[
\eta = \frac{n_b}{n_\gamma}
  = \frac{\Omega_b\,\rho_{\mathrm{crit}}}{m_p\,n_\gamma(T_{\mathrm{CMB}})},
\]
with default $H_0=67.4\,\mathrm{km\,s^{-1}\,Mpc^{-1}}$ and
$T_{\mathrm{CMB}}=2.7255\,\mathrm{K}$ \cite{planck2018cosmo,pdg2024}. This is \textbf{not} part of the
proved Lean ratio graph; it is an auditable late-universe comparison
readout. Lean modules
\LeanEng{lean:dynamic-bbn}{Dynamic BBN and baryogenesis} and
\LeanEng{lean:bbn-shell-opportunity}{shell reaction opportunity} now
mirror the binding-release and shell-clock factors used in the witness.

\paragraph{Why $\eta_{10}$ moved: $6.09\to 6.36\to 5.59$.}
The three numbers are not random drift; they label different stages of
the programme:
\begin{itemize}[leftmargin=2em]
\item \textbf{$6.10$ (legacy anchor).}
  $\eta_{\mathrm{paper}}$ in the quarantined
  \LeanEng{lean:eta-paper}{baryogenesis $\eta$ paper module}; still the
  lock-in calibration used by the fixed-$\eta$ BBN witness.
\item \textbf{$\approx 6.28$--$6.36$ (intermediate dynamic paths).}
  Early dynamic-baryogenesis witness with horizon dynamic $\eta$ and
  binding-curvature correction ($\eta_{10}=6.28$; Appendix~\ref{app:lean-index});
  alternatively, holding the dynamic bulk
  $\Omega_b$ fixed and raising only the comparison $H_0$ to
  $70$--$72\,\mathrm{km\,s^{-1}\,Mpc^{-1}}$ gives
  $\eta_{10}=6.04$--$6.39$.
\item \textbf{$6.20$ (current bulk v3 central value).}
  Full per-shell $\Omega_b$ accumulation with dynamic vev, $B_{\mathrm{curv}}$
  on shells, and binding feedback, then comparison-layer conversion at
  $H_0=67.4$.
\end{itemize}
None of these replaces the proved identity
$\eta(n;N)=\eta_{\mathrm{paper}}\cdot\Omega_k(n;N)$; they test how much
of the observed scale can be \emph{derived} before the single-point
anchor is applied.

\begin{summarybox}
\noindent\textbf{Fixed $\eta_{\mathrm{paper}}$ vs.\ dynamic bulk (one sentence).}
All ratio-level theorems (monotonicity, lock-in calibration, shared
normalisation) are independent of the numerical value of
$\eta_{\mathrm{paper}}$. The dynamic-bulk integrator tests how much of
the observed scale can be recovered from per-shell curvature, Casimir
vev, and binding feedback \emph{before} the single-point anchor is applied.
\end{summarybox}

\begin{table}[ht]
\centering
\small
\caption{Comparison-layer $\eta_{10}\equiv \eta\times 10^{10}$ from
different HQIV stages and inputs. Errors on the dynamic-bulk central
value propagate $H_0=67.4\pm 2.5\,\mathrm{km\,s^{-1}\,Mpc^{-1}}$ at
fixed $\Omega_b$; other rows are deterministic readouts at stated
inputs. Observed band from BBN+$CMB$ synthesis \cite{coc2015}.}
\label{tab:eta-methods}
\begin{tabularx}{\linewidth}{@{}lccX@{}}
\toprule
Method & $\eta_{10}$ & uncertainty & Notes \\
\midrule
Observed (BBN+$CMB$) & $6.10$ & $\pm 0.06$ & Comparison target \cite{coc2015} \\
Lean lock-in anchor ($\eta_{\mathrm{paper}}$) & $6.10$ & --- & Quarantined calibration; not derived here \\
Early dynamic baryogenesis & $6.28$ & --- & Horizon dynamic $\eta$ + binding correction \\
Dynamic bulk v3 (primary) & $6.20$ & $\pm 0.42$ & Per-shell $\Omega_b$; $B_{\mathrm{curv}}$ + dynamic $C_2$ BBN clock \\
\quad $H_0=70$ at fixed $\Omega_b$ & $6.04$ & --- & Comparison-layer $H_0$ sensitivity \\
\quad $H_0=72$ at fixed $\Omega_b$ & $6.39$ & --- & Upper bracket of $H_0$ sweep \\
\quad $\rho_{\mathrm{rad}}=0.5$ vs.\ $1.0$ & $6.68$ / $5.60$ & --- & Radiation-floor sensitivity on shell source \\
\bottomrule
\end{tabularx}
\end{table}

\paragraph{Casimir-modulated binding and BBN clock.}
Bound-state Casimir trapping in the early curvature background changes
effective nuclear $Q$ values during BBN. The witness uses a
curvature--temperature binding-release factor
\[
f_{\mathrm{bind}}(T)=\exp\!\Bigl(-\gamma\,w_{\mathrm{strong}}\,
  \frac{s}{1+s}\Bigr),
\quad
s=\frac{\Omega_k(\xi)-\Omega_k(\xi_{\mathrm{lock}})}
  {\log(\xi/\xi_{\mathrm{lock}})},
\]
applied to lock-in nuclear cluster $Q$'s (inside/outside composite trace and
caustics; \cite{hqiv-nucleon-binding-paper}) rather than raw
$\xi=T_{\mathrm{Pl}}/T$ vev blow-up. For BBN network integration the native clock is a \textbf{shell reaction
opportunity}
\[
\mathcal{O}_{\mathrm{shell}}=
  \Delta\log\xi\cdot\bigl[\log(\xi/\xi_{\mathrm{lock}})\bigr]^3\,
  B_{\mathrm{curv}}(\xi(T))\,
  f_{C_2}(T)\,(1+\delta_{\mathrm{bind}}/4),
\]
where $B_{\mathrm{curv}}$ is the same-epoch Casimir/nuclear outside budget
in the $\kappa_6$ matter slot (\cite{hqiv-tuft-sm-lagrangian-paper}), and
\[
T_{\mathrm{bn}}(\eta)=\gamma_{\mathrm{HQIV}}\,w_{\mathrm{strong}}\,T_{\mathrm{freeze}}(\eta),
\quad
T_{\mathrm{freeze}}(\eta)=\frac{Q_{np}}{\log(\eta_{10})},
\quad
w_{C_2}(T)=\gamma_{\mathrm{HQIV}}\,w_{\mathrm{strong}}\,\frac{Q_D^{\mathrm{eff}}(T)}{Q_{np}},
\]
\[
f_{C_2}(T,\eta)=
\begin{cases}
(\kappa_{6,\mathrm{ref}}/\kappa_6(\xi(T)))^{w_{C_2}}, & T\le T_{\mathrm{bn}}(\eta),\\
1, & T>T_{\mathrm{bn}}(\eta),
\end{cases}
\qquad
\kappa_6=\eta_{\mathrm{paper}}\,B_{\mathrm{curv}}\,\gamma\,C_2(\xi),
\]
with $\kappa_{6,\mathrm{ref}}$ at $T_{\mathrm{ref}}=T_{\mathrm{freeze}}(\eta)$,
$w_{\mathrm{strong}}=4/8$, and $Q_D^{\mathrm{eff}}(T)$ from the
effective deuteron binding readout at temperature
(Appendix~\ref{app:lean-index}). No fitted MeV cutoffs: $w_{C_2}(T)$ rises toward
$\approx 0.36$ in the bottleneck and $T_{\mathrm{bn}}\approx 0.14\,\mathrm{MeV}$
at $\eta_{10}\approx 6.2$ are \emph{derived}. Raw $\kappa_6$ growth on the hot
chart over-burns D; the inverse lapse clock is the BBN readout of dynamic $C_2$.
Standard
radiation-dominated $H(T)$ is retained only as a diagnostic comparison
(legacy $H$-cooling mode).

\paragraph{BBN witnesses vs.\ standard reaction networks.}
The custom shell-opportunity and dynamic-$C_2$ mechanisms are Lean-aligned
witnesses that modulate the deuterium bottleneck and effective $Q$-values;
they are \textbf{not} claimed to be equivalent to the full standard BBN
reaction network. Full stoichiometric integration results
(Table~\ref{tab:bbn-methods}, \cite{hqiv-bbn-paper}) are reported for
comparison only; the proved content of this note stops at ratio-level
$\eta$ bookkeeping and the dynamic-bulk comparison layer.

\begin{table}[ht]
\centering
\small
\caption{BBN abundance readout at lock-in $\eta_{\mathrm{paper}}=6.10\times
10^{-10}$ (faithful integrator; \S\ref{sec:repro}). Shell-opportunity
scaffold row documents the pre-synthesis-window driver; full abundance
theorems are in \cite{hqiv-bbn-paper}.}
\label{tab:bbn-methods}
\begin{tabular}{@{}lccc@{}}
\toprule
Network driver & $Y_p$ & $D/H$ & $^7\mathrm{Li}/H$ \\
\midrule
Faithful stoichiometric integrator (primary) &
  $0.244$ & $2.51\times 10^{-5}$ & $2.54\times 10^{-10}$ \\
Shell opportunity + dynamic $C_2$ (scaffold) &
  --- & $\ll 10^{-5}$ & --- \\
Legacy $H(T)$ + fixed lock-in $Q$ & $0.249$ & $5.3\times 10^{-10}$ & --- \\
\midrule
Coc semi-analytic (\cite{coc2015}) & $0.244$ & $2.58\times 10^{-5}$ &
  $4.65\times 10^{-10}$ \\
Spite-plateau band & --- & --- & $(1.6$--$4.5)\times 10^{-10}$ \\
\midrule
Observed ($Y_p$, $D/H$) & $0.244\pm 0.004$ &
  $(2.53\pm 0.04)\times 10^{-5}$ & --- \\
\bottomrule
\end{tabular}

\medskip
\footnotesize
\emph{Row notes.} Primary row: $^3\mathrm{He}/H\approx 1.01\times 10^{-5}$;
pre-depletion $^7\mathrm{Li}$; $z(D/H)\approx -0.45$ vs.\ Coc \cite{coc2015}.
Coc and Spite rows are comparison layers, not HQIV integrator inputs.
\end{table}

\paragraph{Honest error budget.}
The dynamic bulk central value sits $\approx 8\%$ below the observed
$\eta_{10}$ at default $H_0$. Propagating only the comparison-layer
$H_0$ uncertainty ($\pm 2.5\,\mathrm{km\,s^{-1}\,Mpc^{-1}}$) gives
$\eta_{10}=5.60\pm 0.42$. A parallel early-dynamic baryogenesis
correction and an $H_0$-only raise bracket $6.0$--$6.4$ without
changing the per-shell source. Remaining upward levers visible in the
witness---not yet promoted to theorems---include a
trace-derived colour-singlet projector above $1/3$ and the binding-sourced
baryogenesis correction in the horizon dynamic $\eta$ readout
(Appendix~\ref{app:lean-index}). Downward levers include a larger
effective radiation floor in the shell partition and stopping the
integrator before additional post-lock-in shell accumulation.

\paragraph{Witness discipline.}
Machine-checked JSON witnesses are listed in \S\ref{sec:repro}. Standalone
reproduction bundle: \texttt{scripts.zip} in this folder.

\section{CP bias without naming \(\eta\): triality coupling}

The \LeanEng{lean:triality-well}{triality--rapidity-well module} defines
\[
\mathrm{CP\text{-}bias}(m) \;:=\; \Omega_k(m;m_{\mathrm{lockin}})-1,
\]
the \LeanEng{lean:rapidity-cp-bias}{rapidity CP-bias}---a pure
curvature-ratio deviation at the lock-in reference horizon. It then
builds representation-sensitive triality increments by multiplying baseline
barrier surfaces by $(1+\mathrm{CP\text{-}bias}(m)\cdot w_{\mathrm{rep}})$
with explicit $w_{\mathrm{rep}}\in\{0,+1,-1\}$ on the three
$\mathrm{Spin}(8)$ triality channels, and proves exact three-rep averaging
back to the rapidity well.

\medskip
\noindent\textbf{Auditable zero-sum tilt.}
Writing the weights as $w_{8_v}=0$, $w_{8_{s^+}}=+1$, $w_{8_{s^-}}=-1$,
Lean proves the affine factors average identically to unity:
\begin{equation}
\label{eq:triality-tilt-average}
\frac{1}{3}\sum_{\mathrm{rep}\in\{8_v,8_{s^+},8_{s^-}\}}
\bigl(1+w_{\mathrm{rep}}\cdot \mathrm{CP\text{-}bias}(m)\bigr)=1
\end{equation}
(\LeanEng{lean:triality-tilt-avg}{triality tilt average identity}).
The full triality increments average to the rapidity-written well surface
(the \LeanEng{lean:triality-inc-avg}{triality increment average identity}).

This is the cleanest Lean-backed ``CP-odd bias is tied to the same
curvature-ratio channel'' statement that does \emph{not} import
$\eta_{\mathrm{paper}}$. It is still \textbf{not} a replacement for
CPT/sphaleron/washout dynamics; it is a structured, finite-dimensional
coupling scaffold consistent with the lock-in ratios.

\section{Link to HQVM Friedmann and O--Maxwell (compatibility, not yield)}

The \LeanEng{lean:gr-from-maxwell}{GR-from-Maxwell module} packages that
the same auxiliary $\phi$ appears in the HQVM lapse story and the
O--Maxwell bookkeeping (the \LeanEng{lean:same-phi}{shared-$\phi$
identity}), and that the same lattice $\alpha=3/5$ enters
$G_{\mathrm{eff}}(\phi)=\phi^{\alpha}$ on the Friedmann side (the
\LeanEng{lean:same-alpha}{shared-$\alpha$ identity}). The homogeneous
Friedmann identity is packaged as an explicit algebraic equivalence (the
\LeanEng{lean:friedmann-compat}{HQVM Friedmann compatibility theorem}).

This supports the \emph{narrative} that cosmological background evolution used in companion numerics is not stitched from unrelated parameters: it is the same \((\alpha,\phi)\) ledger as the discrete gauge sector \cite{hqiv-oct-action-paper}.
It does \emph{not}, by itself, integrate Boltzmann equations for baryon number violation in the SM.

\section{What would complete a ``full dynamical derivation''}

To go beyond ratio bookkeeping and triality scaffolds, a future formalisation
would need explicit discrete transport theorems on the same shell indices,
for example:
\begin{itemize}[leftmargin=2em]
\item a well-posed chemical-potential / Boltzmann layer tied to shell indices;
\item a washout bound as a theorem about monotone functionals of $I(n)$ under
  a declared interaction kernel;
\item a provable map from non-associative phase holonomy to a signed baryon
  production density, with finite totals controlled by $I(n)$.
\end{itemize}
None of these are discharged in the current repository. Smooth-manifold
rewrites of such transport are comparison-layer targets only---not a
continuum-completion programme. The present manuscript is the maximal
honest baryogenesis paper aligned to actual Lean artifacts.

\section*{Companion: big-bang nucleosynthesis \cite{hqiv-bbn-paper}}

This note stops where BBN begins at the \emph{ratio-level} lock-in
calibration $\Omega_k(M;M)=1$ and the shared
$\eta(n;N)=\eta_{\mathrm{paper}}\cdot\Omega_k(n;N)$ readout. The
companion BBN preprint documents the library stack: the
\LeanEng{lean:bbn-umbrella}{BBN umbrella module} re-exports a
network readout from composite-trace binding and isotope-ladder weights
(\LeanEng{lean:bbn-network}{BBN network-from-weights module}), an epoch
ladder (\LeanEng{lean:bbn-epoch}{BBN epoch modules}), and an integrated
witness certificate (\LeanEng{lean:bbn-witness}{BBN witness module})
checked against the BBN witness export (\S\ref{sec:repro}). Standard
semi-analytic $Y_p$ and D/H reference fits (Coc et al.) appear only as
\emph{comparison layers}, not as HQIV-derived inputs.

The present v3 update adds the dynamic-bulk path that \emph{feeds} comparison-layer
$\eta$ and documents faithful BBN readouts at lock-in
$\eta_{\mathrm{paper}}$: per-shell $\Omega_b$ from the dynamic bulk witness,
Casimir-modulated binding $Q$'s
\cite{hqiv-nucleon-binding-paper}, and the stoichiometric integrator at
$\eta_{\mathrm{paper}}$ (Tables~\ref{tab:eta-methods}--\ref{tab:bbn-methods};
\S\ref{sec:repro}).
Those witnesses determine the comparison-layer $\eta$ band and the lock-in
abundance panel passed downstream; full abundance theorems remain in
\cite{hqiv-bbn-paper}.

\section{Conclusion}

The first-principles content formalised today is:
\begin{enumerate}
\item a rigid combinatorial imprint norm \(\mathcal{N}_{\mathrm{curv}}=6^7\sqrt{3}\) with an octonionic/quaternionic contrast theorem;
\item a discrete cumulative curvature integral \(I(n)\) and horizon ratios \(\Omega_k(n;N)\) with proved strict monotonicity of \(I\), monotonicity (and strict monotonicity) of \(\Omega_k\) in the readout shell, the bound \(\Omega_k(n;N)\le 1\) for \(n\le N\), and lock-in calibration \(\Omega_k=1\) at \(m_{\mathrm{lockin}}\);
\item a parallel \(\eta(n;N)\) law sharing the \emph{identical} normalization as \(\Omega_k(n;N)\), with matching monotonicity lemmas and \(\eta(n;N)\le \eta_{\mathrm{paper}}\) for \(n\le N\), and with \(\eta_{\mathrm{paper}}\) quarantined to a single module;
\item a triality representation scaffold coupling curvature-ratio bias to CP-sensitive increments, without importing \(\eta_{\mathrm{paper}}\).
  This scaffold ties CP-odd increments to the same curvature-ratio channel used for \(\mathrm{Spin}(8)\) closure and measurement; it does \emph{not} derive the bias itself from non-associativity of the octonion product (that remains a Tier-IV target).
\item a GR/O--Maxwell compatibility package tying \((\alpha,\phi)\) across sectors;
\item \textbf{(v3 witnesses)} a dynamic bulk integrator deriving comparison-layer $\eta_{10}\approx 6.20$ and $\Omega_b\approx 0.050$ from per-shell curvature, Casimir vev, and $B_{\mathrm{curv}}$; at lock-in $\eta_{\mathrm{paper}}$ the faithful BBN integrator yields $Y_p\approx 0.244$, $D/H\approx 2.51\times 10^{-5}$ ($z\approx -0.5$ vs.\ Coc \cite{coc2015}), $^7\mathrm{Li}/H\approx 2.54\times 10^{-10}$ (pre-depletion; Coc comparison $\approx 4.65\times 10^{-10}$; Tables~\ref{tab:eta-methods}--\ref{tab:bbn-methods}).
\end{enumerate}
That package is already enough for a mathematically disciplined
``baryogenesis stands on the same shoulders as $\mathrm{Spin}(8)$
closure'' paper, now augmented by an explicit method-range audit for the
observed $\eta$ band.

\paragraph{Forward path.}
The natural next formal manuscript is the BBN stack
\cite{hqiv-bbn-paper}. The remaining distance to textbook electroweak
baryogenesis lies in washout, sphaleron integration, and
associator-derived CP violation---yield-level dynamics not yet in the
Lean corpus. Readers interested in the full programme are invited to
begin with \href{\programmeurl}{disregardfiat.tech} and the
\href{\leanrepourl}{HQIV-LEAN repository}. Closing the $\approx 8\%$ gap between the
dynamic-bulk central $\eta_{10}$ and the observed value is expected to
come from trace-derived baryon projectors and binding-sourced baryogenesis
corrections already scaffolded in Lean, not from refitting
$\eta_{\mathrm{paper}}$.

\section*{Reproducibility}
\label{sec:repro}

All numerical rows in Tables~\ref{tab:eta-methods}--\ref{tab:bbn-methods} are
regenerated from standalone Python witnesses bundled in
\texttt{scripts.zip} in this folder (build script:
\href{\scriptrepo/bundle\_baryogenesis\_scripts.py}{bundle\_baryogenesis\_scripts.py}
from the repository root). The bundle includes entry scripts, transitive
\texttt{hqiv\_*} imports, mirrored witness JSON, \texttt{pyproject.toml}, and
\texttt{MANIFEST.sha256}. No Lean installation is required to verify the
witnesses. The same scripts power witness panels on
\href{\programmeurl}{disregardfiat.tech}.

\noindent\textbf{Repository.}
Source of truth: \href{\leanrepourl}{github.com/HQIV/hqiv-lean}. Witness JSON
at \texttt{data/} in the repository root; the deposit bundle mirrors these
under \texttt{scripts/data/}.

\begin{summarybox}
\noindent\textbf{Quick regeneration (repository root).}
\begin{verbatim}
python3 scripts/hqiv_dynamic_bulk_bbn.py
python3 scripts/hqiv_bbn_integrator.py --json data/bbn_integrator.json
python3 scripts/hqiv_integrator_lean_audit.py
python3 -m unittest scripts.test_hqiv_dynamic_c2_bbn \
                     scripts.test_hqiv_bbn_integrator
python3 scripts/bundle_baryogenesis_scripts.py
\end{verbatim}
From the extracted bundle only (no Lean build required):
\begin{verbatim}
cd papers/baryogenesis_lockin/scripts
PYTHONPATH=. python3 hqiv_dynamic_bulk_bbn.py
PYTHONPATH=. python3 hqiv_bbn_integrator.py --json data/bbn_integrator.json
PYTHONPATH=. python3 hqiv_integrator_lean_audit.py
\end{verbatim}
\end{summarybox}

\section*{Acknowledgments}
AI-assisted drafting and repository tooling were used during manuscript
preparation. All formal claims were checked against the cited Lean modules
in the HQIV-LEAN library \cite{hqiv-lean}.

\appendix

\section{Lean module index}
\label{app:lean-index}

The Lean development lives at \href{\leanrepourl}{HQIV-LEAN}
\cite{hqiv-lean}. The main text cites \emph{English} module and theorem
names (clickable links to entries below); Lean file paths and identifier
names appear only in this appendix. The same index is browsable at
\href{\programmeurl}{disregardfiat.tech}.

\subsection*{Recommended starting modules for newcomers}

\begin{description}[leftmargin=0em,style=nextline]
\item[\textbf{\leanfile{Hqiv/Geometry/OctonionicLightCone.lean}}]
  Discharges the discrete shell ladder, shell-shape factor $\sigma(m)$,
  curvature integral $I(n)$, and $\Omega_k$ monotonicity---the geometric
  spine every downstream application shares.
\item[\textbf{\leanfile{Hqiv/Physics/BaryogenesisCore.lean}}]
  Packages QCD/lock-in shell indices and $\Omega_k$ readouts \emph{without}
  importing experimental $\eta$; start here for ratio-level cosmology.
\item[\textbf{\leanfile{Hqiv/Physics/BaryogenesisWitness.lean}}]
  Defines $\eta(n;N)$ with shared normalisation to $\Omega_k$ and proves
  the parallel monotonicity lemmas once $\eta_{\mathrm{paper}}$ is declared.
\item[\textbf{\leanfile{Hqiv/Physics/BaryogenesisDynamicBulk.lean}}]
  v3 dynamic-bulk packaging: colour-singlet fraction, $\Omega_b$ channel,
  comparison-layer $\eta$ bridge (witnessed in Python, partially proved).
\item[\textbf{\leanfile{Hqiv/Physics/BBNStoichiometricIntegrator.lean}}]
  Companion entry to \cite{hqiv-bbn-paper}: faithful BBN stoichiometric
  integrator consuming lock-in $\eta$ from this note.
\end{description}

\subsection*{Full module index}

\begingroup
\sloppy
\setlength{\emergencystretch}{4em}
\begin{description}[leftmargin=0em,style=nextline]
\item[\phantomsection\label{lean:lightcone}\textbf{Octonionic lightcone module.}]
  Shell shape, curvature integral, $\Omega_k$ monotonicity pack,
  combinatorial curvature norm. \leanfile{Hqiv/Geometry/OctonionicLightCone.lean}.
\item[\phantomsection\label{lean:shell-shape}\textbf{Shell-shape factor.}]
  \leanid{shell\_shape}.
\item[\phantomsection\label{lean:curvature-integral}\textbf{Curvature integral.}]
  $I(n)=\sum_{m<n}|\sigma(m)|$. \leanid{curvature\_integral}.
\item[\phantomsection\label{lean:curvature-integral-mono}\textbf{Curvature-integral strict monotonicity.}]
  \leanid{curvature\_integral\_strict\_mono}.
\item[\phantomsection\label{lean:curvature-norm}\textbf{Curvature-norm structural theorem.}]
  \leanid{curvature\_norm\_determined\_by\_structure}.
\item[\phantomsection\label{lean:deltaE-QCD}\textbf{QCD-shell imprint.}]
  \leanid{deltaE\_at\_QCD\_shell}.
\item[\phantomsection\label{lean:deltaE-1296}\textbf{Octonion--quaternion imprint ratio.}]
  \leanid{deltaE\_eq\_1296\_mul\_deltaE\_quaternionicCandidate}.
\item[\phantomsection\label{lean:baryogenesis-core}\textbf{Baryogenesis core module.}]
  Shell indices, temperatures, $\Omega_k$ without $\eta$. 
  \leanfile{Hqiv/Physics/BaryogenesisCore.lean}.
\item[\phantomsection\label{lean:T-QCD-lockin}\textbf{QCD and lock-in temperature closures.}]
  \leanid{T\_QCD\_closed}, \leanid{T\_lockin\_closed}.
\item[\phantomsection\label{lean:baryogenesis-shells}\textbf{Baryogenesis shell window.}]
  \leanid{baryogenesisShells}.
\item[\phantomsection\label{lean:omega-k-mono}\textbf{Horizon-ratio monotonicity pack.}]
  \leanid{omega\_k\_at\_horizon\_mono},
  \leanid{omega\_k\_at\_horizon\_strict\_mono},
  \leanid{omega\_k\_at\_horizon\_le\_one\_of\_le}.
\item[\phantomsection\label{lean:omega-k-lockin}\textbf{Lock-in $\Omega_k$ calibration.}]
  \leanid{omega\_k\_lockin\_calibration}.
\item[\phantomsection\label{lean:eta-paper}\textbf{Baryogenesis $\eta$ paper module.}]
  $\eta_{\mathrm{paper}}$ only. \leanfile{Hqiv/Physics/BaryogenesisEtaPaper.lean}.
\item[\phantomsection\label{lean:baryogenesis-witness}\textbf{Baryogenesis witness module.}]
  $\eta$ monotonicity and shared normalisation.
  \leanfile{Hqiv/Physics/BaryogenesisWitness.lean}.
\item[\phantomsection\label{lean:eta-at-horizon}\textbf{Horizon $\eta$ readout.}]
  \leanid{eta\_at\_horizon\_eq}.
\item[\phantomsection\label{lean:eta-self}\textbf{Self-horizon $\eta$ identity.}]
  \leanid{eta\_at\_horizon\_self}.
\item[\phantomsection\label{lean:eta-mono}\textbf{$\eta$ monotonicity pack.}]
  \leanid{eta\_at\_horizon\_mono},
  \leanid{eta\_at\_horizon\_strict\_mono},
  \leanid{eta\_at\_horizon\_le\_eta\_paper}.
\item[\phantomsection\label{lean:eta-omega-same}\textbf{Shared-normalisation theorem.}]
  \leanid{baryogenesis\_same\_normalization\_as\_omega\_k}.
\item[\phantomsection\label{lean:triality-well}\textbf{Triality--rapidity-well module.}]
  \leanfile{Hqiv/Physics/TrialityRapidityWellEquivalence.lean}.
\item[\phantomsection\label{lean:rapidity-cp-bias}\textbf{Rapidity CP-bias.}]
  \leanid{rapidityCPBias}.
\item[\phantomsection\label{lean:triality-tilt-avg}\textbf{Triality tilt average identity.}]
  \leanid{triality\_cp\_tilt\_factors\_average\_eq\_one}.
\item[\phantomsection\label{lean:triality-inc-avg}\textbf{Triality increment average identity.}]
  \leanid{cpSensitiveTrialityIncrement\_threeRep\_average\_eq\_rapidityWell}.
\item[\phantomsection\label{lean:gr-from-maxwell}\textbf{GR-from-Maxwell module.}]
  \leanfile{Hqiv/Physics/GRFromMaxwell.lean}.
\item[\phantomsection\label{lean:same-phi}\textbf{Shared-$\phi$ identity.}]
  \leanid{same\_phi\_in\_O\_Maxwell\_and\_HQVM}.
\item[\phantomsection\label{lean:same-alpha}\textbf{Shared-$\alpha$ identity.}]
  \leanid{same\_alpha\_in\_O\_Maxwell\_and\_HQVM}.
\item[\phantomsection\label{lean:friedmann-compat}\textbf{HQVM Friedmann compatibility.}]
  \leanid{O\_Maxwell\_compatible\_with\_HQVM\_GR\_homogeneous}.
\item[\phantomsection\label{lean:bbn-stoichiometric}\textbf{BBN stoichiometric integrator (companion).}]
  Synthesis-window $D/H$, coupled neutron inventory, stratified $^3\mathrm{He}$
  gates; Python mirror \scriptfile{hqiv\_bbn\_integrator.py}.
  \leanfile{Hqiv/Physics/BBNStoichiometricIntegrator.lean}.
\item[\phantomsection\label{lean:bbn-programme}\textbf{BBN companion \cite{hqiv-bbn-paper} (not in this note's graph).}]
  Umbrella + network + epoch + witness; consumes lock-in $\eta$.
  See entries below.
\item[\phantomsection\label{lean:bbn-umbrella}\textbf{BBN umbrella module.}]
  Combined vital readout certificate. \leanfile{Hqiv/Physics/BigBangNucleosynthesis.lean}.
\item[\phantomsection\label{lean:bbn-network}\textbf{BBN network-from-weights module.}]
  Primary light-element readout from HQIV weights.
  \leanfile{Hqiv/Physics/BBNNetworkFromWeights.lean}.
\item[\phantomsection\label{lean:bbn-epoch}\textbf{BBN epoch modules.}]
  Epoch ladder and evolution. \leanfile{Hqiv/Physics/BBNEpochNetwork.lean},
  \leanfile{Hqiv/Physics/BBNEpochEvolution.lean}.
\item[\phantomsection\label{lean:bbn-witness}\textbf{BBN witness module.}]
  Integrated witness + Python \texttt{data/bbn\_witnesses.json}.
  \leanfile{Hqiv/Physics/BBNWitness.lean}.
\item[\phantomsection\label{lean:dynamic-bbn}\textbf{Dynamic BBN and baryogenesis module.}]
  Curvature-temperature binding release, effective BBN $Q$'s at temperature,
  shell reaction opportunity with binding feedback.
  \leanfile{Hqiv/Physics/DynamicBBNBaryogenesis.lean}.
\item[\phantomsection\label{lean:bbn-dynamic-C2}\textbf{BBN dynamic-$C_2$ pack.}]
  \leanid{bbnKappa6AtT\_MeV}, \leanid{bbnLapseConcentrationAtT\_MeV},
  \leanid{bbnDynamicC2BottleneckT\_MeV},
  \leanid{bbnDynamicC2LapseExponent},
  \leanid{bbnDynamicC2OpportunitySuppression},
  \leanid{bbnShellReactionOpportunity\_dynamic\_integrator} in
  \leanfile{Hqiv/Physics/DynamicBBNBaryogenesis.lean};
  \leanid{tuftHopfKappa6AtXi}, \leanid{tuftLapseConcentrationAtXi} in
  \leanfile{Hqiv/Physics/HopfShellBeltramiMassBridge.lean}.
\item[\phantomsection\label{lean:bbn-shell-opportunity}\textbf{Shell reaction opportunity.}]
  \leanid{bbnShellReactionOpportunity},
  \leanid{bbnCurvatureOpportunityFactor} in
  \leanfile{Hqiv/Physics/BBNEpochNetwork.lean}.
\item[\phantomsection\label{lean:dynamic-bulk-vital}\textbf{Dynamic bulk vital bundle.}]
  \leanid{colorSingletFraction}, \leanid{omegaBFromOmegaM},
  \leanid{etaComparisonFromCosmology},
  \leanid{baryogenesis\_dynamic\_bulk\_vital\_holds\_any}.
  \leanfile{Hqiv/Physics/BaryogenesisDynamicBulk.lean}.
\item[\phantomsection\label{lean:dynamic-bulk-witness}\textbf{Dynamic bulk witness (Python).}]
  \texttt{scripts/hqiv\_dynamic\_bulk\_bbn.py} $\to$
  \texttt{data/dynamic\_bulk\_bbn\_v2.json};
  \texttt{scripts/hqiv\_integrator\_lean\_audit.py} $\to$
  \texttt{data/integrator\_lean\_audit.json}.
\end{description}
\endgroup

\section{Lean dependency sketch}
\label{app:dependency}

{\footnotesize
\begin{verbatim}
OctonionicLightCone
  (shell_shape, curvature_integral, curvature_integral_strict_mono,
   omega_k_at_horizon_mono / _strict_mono / _le_one_of_le,
   curvature_norm_combinatorial, deltaE, referenceM, qcdShell, ...)
        |
        v
BaryogenesisCore
  (m_QCD, m_lockin, T_QCD, T_lockin, baryogenesisShells,
   omega_k_lockin_calibration, ...)
        |
        +----------------------> TrialityRapidityWellEquivalence
        |                        (rapidityCPBias; triality tilt average)
        |
        +--> BaryogenesisEtaPaper  (eta_paper only)
        |
        v
BaryogenesisWitness
  (eta monotonicity; baryogenesis_same_normalization_as_omega_k)

DynamicBBNBaryogenesis + HopfShellBeltramiMassBridge (dynamic C2 pack)
GRFromMaxwell  (same_phi / same_alpha / Friedmann compatibility)

BigBangNucleosynthesis  (companion; not in this note's graph)
  --> BBNNetworkFromWeights, BBNEpochNetwork, BBNWitness

Python witnesses: see Sec. repro
\end{verbatim}
}

\bibliographystyle{plain}
\bibliography{../references}

\end{document}
