% HQIV: O-Maxwell ↔ effective fluid chart hypothesis (F2) — Lean-aligned
% Build: cd paper && pdflatex HQIV_OMaxwell_fluid_chart.tex
\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,mathtools}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\newcommand{\lean}[1]{\path{#1}}
\usepackage[margin=1in]{geometry}

\title{HQIV chart hypothesis:\\
O-Maxwell scalar fields and the modified fluid vacuum source}
\author{Steven Ettinger Jr}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This note is the \textbf{paper companion} to the \textbf{F2 typed bundle} in
\lean{Hqiv/Physics/HQIVFluidClosureScaffold.lean}: at a fixed chart point~$c$ on
$\mathbb{R}^{1+3}$ (coordinates $\mathrm{Fin}\,4\to\mathbb{R}$), the fluid closure inputs
$(\varphi,\nabla\varphi,\dot\delta\theta',\nabla\dot\delta\theta')$ are identified with continuum
scalars $\varphi_F$, $\dot F$, an electric-energy proxy~$E'$, and spatial gradients from
\texttt{coordsGradientComponents}. \textbf{Not claimed:} a theorem that solutions of full MHD or
Navier--Stokes coincide with O-Maxwell; only \textbf{hypothesis bookkeeping} aligned with Lean.
See \texttt{AGENTS/FLUID\_OMAXWELL\_ROADMAP.md}~\S\,F2.
\end{abstract}

\input{../include/patch_theory_messaging.tex}

\paragraph{Index convention.}
$\nu=0$ is time; $\nu\in\{1,2,3\}$ are spatial. The fluid uses $\mathrm{Fin}\,3\to\mathbb{R}$ vectors;
\lean{spatialFin4} embeds $i\in\mathrm{Fin}\,3$ as $\nu=i+1$.

\paragraph{Modified fluid vacuum momentum (Python / Lean).}
The spatial source $\mathbf{g}_{\mathrm{vac}}$ in \texttt{pyhqiv.fluid} matches
\lean{hqivVacuumMomentumSource3}:
\begin{equation}
  \mathbf{g}_{\mathrm{vac}} = -\frac{\gamma}{6}\,\nabla(\varphi\,\dot\delta\theta')
  = -\frac{\gamma}{6}\bigl(\varphi\,\nabla\dot\delta\theta' + \dot\delta\theta'\,\nabla\varphi\bigr),
  \qquad \gamma = \gamma_{\mathrm{HQIV}} = \tfrac{2}{5}.
\end{equation}

\paragraph{F2 chart hypothesis (Lean name: \texttt{OMaxwellFluidChartHypothesis}).}
At a basepoint~$c$:
\begin{align}
  \varphi_{\mathrm{fluid}} &= \varphi_F(c), \\
  (\nabla\varphi)_{\mathrm{spatial}} &= \bigl(\partial_\nu \varphi_F(c)\bigr)_{\nu=1,2,3}
     &&\text{via \lean{chartSpatialPhiGradient}}, \\
  \dot\delta\theta' &= \delta\theta'(E') = \texttt{delta\_theta\_prime}(E')
     &&\text{(\lean{Hqiv.delta\_theta\_prime} in \lean{ModifiedMaxwell.lean})}, \\
  (\nabla\dot\delta\theta')_{\mathrm{spatial}} &= \bigl(\partial_\nu \dot F(c)\bigr)_{\nu=1,2,3}
     &&\text{via \lean{chartSpatialDotGradient} — user-chosen scalar $\dot F$ on the chart.}
\end{align}
Under these equalities, \lean{hqivVacuumMomentumSource3_of_OMaxwellFluidChartHypothesis} rewrites the
vacuum source to depend only on $(\varphi_F,\dot F,c,E')$.

\paragraph{MHD (ideal / resistive) as composition.}
\textbf{Ideal MHD} couples fluid velocity~$\mathbf{u}$, magnetic field~$\mathbf{B}$, and electric field
$\mathbf{E} = -\mathbf{u}\times\mathbf{B}$ (in suitable units). HQIV does not reprove MHD here; the
\textbf{closure alignment} is: the same $(\varphi,\dot\delta\theta')$ slots that enter
$\mathbf{g}_{\mathrm{vac}}$ and eddy viscosity (see \lean{hqivEddyViscosity}) are the ones whose
\textbf{chart gradients} are fixed by the F2 bundle when coupling to O-Maxwell. Resistive terms add
$\eta\,\mathbf{J}$; current $\mathbf{J}$ is still only \textbf{hypothesis-mapped} to stress (see F2 table,
``Plasma $J$'' row in \texttt{FLUID\_OMAXWELL\_ROADMAP.md}).

\paragraph{Repository.}
\lean{Hqiv/Physics/HQIVFluidClosureScaffold.lean},
\lean{Hqiv/Physics/ContinuumOmaxwellClosure.lean},
\lean{Hqiv/Physics/ModifiedMaxwell.lean}.

\end{document}
